### 2017-08-01 paper.tex
Project created

\documentclass{llncs}
% * <peter.christen@anu.edu.au> 2016-08-19T15:07:29.567Z:
%
% ^.
%\usepackage{llncsdoc}
\usepackage{epsfig}
\usepackage{graphicx}

%\usepackage{caption}
%\usepackage{subcaption}

\newtheorem{mydef}{Def.}
\newtheorem{myhyp}{Hypothesis}

\usepackage{url}
\usepackage{hyperref}

\pagestyle{empty}

% PAKDD 2017: maximum 12 pages
% Abstract max 200 words
% title, abstract: 1/2 page
% introduction: 1 1/2 page (2 pages so far)
% related work: 1 to 1 1/2 pages (3 pages so far)
% method: 4 pages (7 pages so far)
% experiments: 3 (10 pages so far)
% discussion /conclusion: 1 page
% citations 1

% ====================================================================

% Eamonn ICDM'10 tutorial slides
% - clear problem statement in abstract
% - To convince a reviewer, you must think like a reviewer

% Writing the paper:
% - Make a working title
% - Introduce the topic and define (informally at this stage)
%   terminology
% - Motivation: Emphasize why is the topic important
% - Relate to current knowledge: what’s been done
% - Indicate the gap: what need’s to be done?
% - Formally pose research questions
% - Explain any necessary background material.
% - Introduce formal definitions.
% - Introduce your novel algorithm/representation/data structure etc.
% - Describe experimental set-up, explain what the experiments will
%   show
% - Describe the datasets
% - Summarize results with figures/tables
% - Discuss results
% - Explain conflicting results, unexpected findings and discrepancies
%   with other research
% - State limitations of the study
% - State importance of findings
% - Announce directions for further research
% - Acknowledgements
% - References
%
% - Don’t make the reviewer of your paper think!
% - Reviewers make an initial impression on the first page and don’t
%   change 80% of the time
% - A good introduction with a good motivation is half your success
% - By the end of the introduction the reviewer mustknow.
%   - What is the problem?
%   - Why is it interesting and important?
%   - Why is it hard?why do naive approaches fail?
%   - Why hasn't it been solved before?(Or, what's wrong with previous
%     proposed solutions?)
%   - What are the key components of my approach and results?Also
%     include any specific limitations.
%   - A final paragraph or subsection: “Summary of Contributions”.
%     It should list the major contributions in bullet form,
%     mentioning in which sections they can be found. This material
%     doubles as an outline of the rest of the paper, saving space and
%     eliminating redundancy
% - Unjustified Choices (are bad)
% - Optimal: Does not mean `very good'
% - Proved: Does not mean `demonstrated'
% - Significant: There is a danger of confusing the informal statement
%   and the statistical claim
% - Use all the Space Available
% - Avoid Weak Language: aim, attempt, might, etc.
% - Use the Active Voice
% - ALWAYS put some variance estimate on performance measures (do
%   everything 10 times and give me the variance of whatever you are
%   reporting)

% Figures:
% - Don't cover the data with the labels!
% - Color helps -Direct labeling helps -Meaningful captions help
%
% Common problem with figures:
% 1.Too many patterns on bars
% 2.Use of both different symbols and different lines
% 3.Too many shades of gray on bars
% 4.Lines too thin (or thick)
% 5.Use of three-dimensional bars for only two variables
% 6.Lettering too small and font difficult to read
% 7.Symbols too small or difficult to distinguish
% 8.Redundant title printed on graph
% 9.Use of gray symbols or lines
% 10.Key outside the graph
% 11.Unnecessary numbers in the axis
% 12.Multiple colors map to the same shade of gray
% 13.Unnecessary shading in background
% 14.Using bitmap graphics (instead of vector graphics)
% 15.General carelessness

% ====================================================================

% TODO:
% - clearly point out novelty of this work
% - contribution compared to earlier work

\begin{document}

% ideas for a better shorter title?
\title{Efficient Cryptanalysis of Bloom Filters for
       Privacy-Preserving Record Linkage
  \thanks{The authors would like to thank the Isaac Newton Institute
  for Mathematical Sciences, Cambridge, for support and hospitality
  during the programme \emph{Data Linkage and Anonymisation} where
  this work was conducted (EPSRC grant EP/K032208/1). Peter Christen
  was also supported by a grant from the Simons Foundation. The work
  was also partially funded by the Australian Research Council
  (DP130101801).}}

\author{Peter Christen\inst{1} \and Rainer Schnell\inst{2}
        \and Dinusha Vatsalan\inst{1} \and Thilina
        Ranbaduge\inst{1}}
\institute{Research School of Computer Science,
           The Australian National University, \\
           Canberra, Australia.~
           Contact: \email{peter.christen@anu.edu.au}
           \and
           Methodology Research Group, %Institute for Sociology,
           University Duisburg-Essen, Duisburg, Germany.} %~
%           \email{rainer.schnell@uni-due.de}}

\maketitle

\begin{abstract}
Privacy-preserving record linkage (PPRL) is the process of identifying
records that represent the same entity across databases held by
different organizations without revealing any sensitive information
about these entities. A popular technique used in PPRL is Bloom
filter encoding, which has shown to be an efficient and effective way
to encode sensitive information into bit vectors while still enabling
approximate matching of attribute values. However, the encoded values
in Bloom filters are vulnerable to cryptanalysis attacks. Under
specific conditions, these attacks are successful in that some
frequent sensitive attribute values can be re-identified.
%
%While previous attacks have high computational requirements
In this paper we propose and evaluate on real databases a novel
efficient attack on Bloom filters. Our approach is based on the
construction principle of Bloom filters of hashing elements of sets
into bit positions. The attack is independent of the encoding
function and its parameters used, it can correctly re-identify
sensitive attribute values even when various recently proposed
hardening techniques have been applied, and it runs in a few seconds
instead of hours.
\end{abstract}

\keywords Privacy; re-identification; frequency analysis; data
          linkage.

% ====================================================================

\section{Introduction}
\label{sec-intro}

Integrating data from different sources with the aim to remove
duplicates, enrich data, and correct errors and inconsistencies is a
crucial data pre-processing task for many data mining and analytics
applications~\cite{Chr12}. Example applications include healthcare,
business analytics, national censuses, population informatics, fraud
detection, government services, and national security.

However, growing concerns about privacy and confidentiality
increasingly preclude the exchange or sharing of personal identifying
attributes, such as names, dates of birth, and addresses, which are
generally required for linking databases due to the non-existence of
common unique entity identifiers~\cite{Chr12,Vat13}. Work in
privacy-preserving record linkage (PPRL) aims to develop techniques
for identifying records that correspond to the same entity across
several databases while not compromising the privacy and
confidentiality of the entities~\cite{Vat13}.

PPRL is achieved by conducting linkage on the encoded (masked) values
of the identifying attributes of records across two or more databases.
%~\cite{Fie05}.
%
Several data encoding techniques for PPRL have been developed. These
can be categorized into cryptographic secure multi-party computation
(SMC) and perturbation-based techniques~\cite{Vat13}. The former are
accurate and provably secure, but they incur expensive computation
and communication costs.
%The latter category includes more efficient techniques which often
%provide a trade-off between accuracy and privacy.
%Since existing cryptographic-based techniques are not scalable for
%real applications,
Most PPRL techniques are therefore based on perturbation-based
techniques that provide adequate privacy protection while achieving
acceptable linkage quality~\cite{Ran16,Vat13}.

Bloom filter (BF) encoding is one such perturbation-based technique
that has successfully been used in several recent practical PPRL
applications~\cite{Boy15,Ran14}. A BF is a binary vector with bits
initially set to $0$. 
%with bits set to $0$ or $1$. As we describe in detail in
%Sect.~\ref{sec-overview},
A value can be encoded into a BF using a set of hash functions by
setting corresponding bits to $1$~\cite{Sch09}, and the approximate
similarity between two BFs can be calculated by counting the number
of positions where both BFs have $1$-bits in common.

% Peter: I don't think this next paragraph is needed for this paper
% as it is not needed to understand the paper
%The similarity (or distance) between a pair of values is preserved in
%the BF space such that the similarity calculated using their BF
%encodings is greater than or equal to the similarity calculated using
%their unencoded values. BF-based similarity calculation allows false
%positives due to collision of bits (where different values are
%hash-mapped into the same bit) which improves the privacy aspect
%while compromising accuracy (depending on the false positive rate).
%Therefore, it is crucial to set the BF related parameters ($k$, $l$,
%and number of elements to be hash-mapped into a BF) appropriately to
%balance the accuracy and privacy aspects of
%linkage~\cite{Sch09,Vat16}.

As we discuss in detail in the next section, BFs can be susceptible
to cryptanalysis attacks that aim to re-identify the encoded
sensitive attribute values~\cite{Kro15,Kuz11,Kuz13,Nie14}. Using 
frequency counts and patterns in a set of BFs, these attacks
iteratively map bit patterns to known attribute values. These existing
attacks are however not practical as they require knowledge of certain
parameters used during the BF encoding phase, and they have high
computational costs.

Our contribution in this paper is an efficient frequency-based
approach for attacking BFs that exploits the fundamental property of
how the elements of sets are hashed into BFs, as we describe in
Sect.~\ref{sec-overview}. In contrast to existing attack methods, our
novel approach does not require any assumption on the BF parameters
used when sensitive attribute values were encoded. It is also
significantly faster, making it a viable attack on large sets of
BFs to evaluate whether they provide adequate privacy protection. We
experimentally evaluate our attack method on two real-world data
sets, showing its efficiency and effectiveness.

Given BF encoding is now being employed in real-world PPRL
applications~\cite{Boy15,Ran14}, it is crucial to study possible
attacks on BFs to ensure their security and to make users of such
systems aware of the weaknesses of BF encoding. Our novel attack
method allows data custodians to identify such weaknesses of BF
encoding that otherwise could be exploited by an attacker.
%if appropriate BF encodings and parameters are not employed.

%Maybe add: Contributions:
%We propose a novel and efficient BF attack method and analyze its
%complexity

% PC 17/01/2107 until here **************************

% --------------------------------------------------------------------

\section{Prior Attacks on Bloom Filter based PPRL}
\label{sec-related}

%Bloom filters (BFs)~\cite{Blo70} are commonly used for the encoding of
%records in PPRL due to their capability of computing
%similarities~\cite{Dur14,Ran16,Sch09,Vat13}.
Schnell et al.~\cite{Sch09} were the first to introduce an
approximate matching approach for PPRL using Bloom filters (BFs), as
we describe in detail in Sect.~\ref{sec-overview}. A recent study by
Randall et al.~\cite{Ran14} has shown that PPRL based on BF encoding
can achieve similar linkage quality as can be achieved with
traditional linkage methods on the unencoded attribute values. As a
result, BF encoding has been the PPRL technique of choice in several
recent practical PPRL applications~\cite{Boy15,Ran14}.
%
%However, BFs are prone to different
%attacks~\cite{Kro15,Kuz11,Kuz13,Nie14} which will be described %shortly.
%
However, BFs are prone to different
attacks~\cite{Kro15,Kuz11,Kuz13,Nie14}. Therefore, a comprehensive
analysis of the weaknesses of BFs in the PPRL context is required. We
now describe the few studies that have been conducted on attacks on
BFs.

% From Dinusha's introduction:
%An attack using a constraint satisfaction solver was studied by
%Kuzu et al.~\cite{Kuz11,Kuz13}, where values are mapped to bit
%patterns that meet constraints such as frequency alignments. A
%successful attack on BFs using filtering and statistical analysis
%techniques was recently investigated by Niedermeyer et
%al.~\cite{Nie14}. Based on the linear combination of the values used
%to generate hash functions, this attack first generates a set of
%possible bit patterns (so called \emph{atoms}) that are then aligned
%according to their frequencies with known attribute
%values~\cite{Nie14}.

Kuzu et al.~\cite{Kuz11} formulated a cryptanalysis attack on BFs
as a constraint satisfaction problem (CSP). CSP is characterized by
a set of variables and a set of constraints on these variables. The
aim of this attack is to identify a set of values from a given domain
that can be assigned to each variable such that the constraints are
satisfied. This is achieved by a frequency analysis of the sensitive
attribute values and the BF encodings of the records in a database.
However, this attack requires that the attacker has access to a
global database where the encoded records are drawn from. This
is unlikely in practical applications.

In 2013, Kuzu et al.~\cite{Kuz13} investigated the accuracy of their
CSP attack with two real-world databases. The authors tried to
re-identify the personal details of patients in an encoded medical
database by using a frequency analysis of BF encodings of a public
voter registration database that has a different frequency
distribution to the medical database. Their results indicated that
although the CSP attack might be feasible in such situations, it is
less likely to be accurate in identifying original attribute
values and it requires more computational resources. The attack
re-identified four out of 20 frequent names correctly.

% Peter: results below are different to Table 1? *******************

Niedermeyer et al.~\cite{Nie14} more recently proposed an attack on
BFs built from German surnames. The attack was based on the
frequencies of sub-strings of length 2 extracted from frequent
surnames. Of $7,580$ surnames, the authors re-identified the $934$
most frequent ones (about $12\%$) before stopping the attack. In
contrast to the approach in~\cite{Kuz13}, this attack only 
%is not based on the assumption that the attacker has
%access to a population data set from which the surnames are drawn.
%However, the feasibility of this approach 
depends on the availability of a list of the most common surnames.
This work was extended by Kroll and Steinmetzer~\cite{Kro15} into a
cryptanalysis on BF encodings of several attributes, which was able
to re-identify $44\%$ of all attribute values correctly. However, both
attacks are based on the specific double hashing scheme used by
Schnell at al.~\cite{Sch09}.

Existing cryptanalysis attacks are feasible only for certain
settings and assumptions used in the BF encoding phase. They also
require excessive computational resources making them not practical
in real settings. Our novel attack method, described next, improves
on both these drawbacks of existing methods.

% --------------------------------------------------------------------

\begin{figure}[t]
  \centering
  \includegraphics[width=1.0\textwidth]{bf-attack-outline}
  \caption{Outline of the proposed cryptanalysis attack, which is
           based on a set of BFs and a set attribute values, both
           sorted according to their frequencies. In steps (1a) and
           (1b), the attack exploits the bit patterns in BFs to
           identify the sets of q-grams that are possible,
           $\mathbf{c}^+[p]$, and \emph{not} possible,
           $\mathbf{c}^-[p]$, respectively, for each bit position
           $p$. In step (2), the re-identification of a set of
           (sensitive) attribute values, $\mathbf{G}$, is conducted
           by intersecting the q-gram sets of values in $\mathbf{G}$
           with the sets $\mathbf{c}$ (illustrated using $\odot$).
           %as outlined in Sect.~\ref{sec-overview} and detailed in
           %Sect.~\ref{sec-bf-attack}.
           \label{fig:outline}}
\end{figure}

\section{Overview and Preliminaries}
\label{sec-overview}

We now provide an overview of our attack on BFs, as illustrated in
Fig.~\ref{fig:outline}. As with other attacks on
BFs~\cite{Kro15,Kuz11,Kuz13,Nie14}, our approach exploits the
frequency distribution of a set of BFs that were generated from a
large database. As for notation, we use bold letters for sets
(with upper-case bold letters for lists or sets of sets) and normal
type letters for integer or string values. We denote sets with curly
and lists with square brackets, where lists have an order while sets
do not.

We assume the attacker has access to a set of encoded BFs,
$\mathbf{B}$, and their frequencies, but he does not know anything
about the parameters used in the encoding process (such as the number
of hash functions used, or the actual hashing mechanism). We assume
these BFs represent a set of records that encode sensitive values
from one or a few attributes. Based on the frequency distribution of
the Hamming weights (number of $1$-bits) in $\mathbf{B}$, the attacker
can guess which attribute(s) have been encoded, because different
attributes (such as first name, surname, city name, or postcode) have
distinctive distributions of Hamming weights~\cite{Sch16}. The
distribution of Hamming weights is independent of the (unknown)
secret key. Therefore, an attacker can sample attribute values from
a publicly available population database (such as a telephone
directory) and select a set of frequent values, $\mathbf{V}$, from
an attribute that has a frequency distribution that is similar to
the distribution of the set to be attacked.
%matches with the distribution
%of the set of BFs $\mathbf{B}$ to be attacked.

In step (1), we first align BFs and attribute values according to
their frequencies, and consider the set of most frequent values in 
both. For each bit position $p$ in the BFs,
%of length $l$ (with $1 \le p \le l$),
for all corresponding attribute values that have this bit set to $1$
we add their q-grams (sub-strings of length $q$ generated from
attribute values) to the set $\mathbf{c}^+[p]$ of possible
q-grams for that position. The reasoning is that a $1$-bit means at
least one q-gram of an attribute value was hashed to this position.
For all attribute values with a value of $0$ at bit
position $p$ we add their q-grams to the set $\mathbf{c}^-[p]$ of
\emph{not} possible q-grams for that position, because a $0$-bit
means no q-gram of an attribute value could have been mapped to this
position.

At the end of step (1), for each position $p$ we obtain the set
$\mathbf{c}[p] = \mathbf{c}^+[p] \setminus \mathbf{c}^-[p]$ of
q-grams that potentially could have been hashed to position $p$.
Based on the list $\mathbf{C} = [\mathbf{c}[1], \ldots,
\mathbf{c}[l]]$, where $l$ is the length of the BFs, and a set
$\mathbf{G}$ of attribute values we aim to re-identify (i.e.\ learn
which BF possibly encodes which value in $\mathbf{G}$), in step (2)
we analyze each BF in $\mathbf{B}$ and remove those attribute values
from $\mathbf{G}$ that are not possible matches according to
$\mathbf{C}$ because they do not contain any q-grams that would have
been hashed to a certain $1$-bit.

For example, in Fig.~\ref{fig:outline}, for the most frequent BF$_1$,
`kate' is not a possible value because in order to obtain a $1$-bit
in position $p_1$, it would have to contain either the q-gram `en'
or `ry'; while `mary' is also not possible because it would need to
contain one of the q-grams `ka', `en', `at', or `te' in position
$p_3$.

Before we formalize and present our approach in detail in
Sect.~\ref{sec-bf-attack}, we first describe BF encoding, as well as
some recent approaches to harden them.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\smallskip

%\subsection{Bloom Filter Encoding}
%\label{sec-bf-encoding}

\begin{figure}[!t]
  \centering
  %\scalebox{1.0}[1.0]
  \includegraphics[width=0.65\textwidth]{bloomfilter}
  \caption{An example Dice coefficient similarity calculation of the
           two first names `peter' and `pete'  encoded in BFs, as
           described in Sect.~\ref{sec-overview}. The dark bit shows
           a hash collision.}
           \label{fig:bloomfilter}
\end{figure}

\textbf{Bloom Filter Encoding:}~
Proposed by Bloom~\cite{Blo70} in 1970 for the space and time
efficient representation of sets, a BF $\mathbf{b}$ is a bit vector
of length $l$ where all bits are initially set to $0$. $k$
independent hash functions, $h_1, \ldots, h_k$, each with range
$1, \ldots, l$, are used to map the elements $s$ in a set
$\mathbf{s}$ into the BF by setting the bit positions
$\mathbf{b}[h_j(s)]=1$, with $1 \le j \le k$.

For PPRL, the set $\mathbf{s}$ of q-grams generated from string
values~\cite{Sch09}, or neighboring values for numerical
values~\cite{Vat16}, can be hash-mapped into a BF. These BFs are then
either sent to a linkage unit (LU, an external party that conducts
the linkage) to calculate the similarity between BFs in order to
classify them as matches or non-matches~\cite{Sch09}, or they are
partially exchanged among the database owners to distributively
calculate the similarities between BFs~\cite{Vat14c}.

%Any set-based similarity function (such as Overlap, Jaccard, and Dice
%coefficient~\cite{Chr12}) can be used to calculate the similarity of
%pairs or sets of (multiple) BFs.

The Dice coefficient has been used for comparing BFs since it is
insensitive to many matching zeros
%(bit positions to which no elements are hash-mapped)
in long BFs~\cite{Sch09}. 
%
For two BFs, $\mathbf{b}_1$ and $\mathbf{b}_2$, the Dice coefficient
similarity is: $sim_D(\mathbf{b}_1, \mathbf{b}_2) = 2  c / (x_1 +
x_2)$,
%\begin{eqnarray}
%\label{eq:Dice_coefficient}
%sim_D(\mathbf{b}_1, \mathbf{b}_2) &=& \frac{2 \times c}{x_1 + x_2},
%\end{eqnarray}
where $c$ is the number of bit positions that are set to $1$ in both
BFs (common $1$-bits), and $x_1$ and $x_2$ are the number of bit
positions set to $1$ in $\mathbf{b}_1$ and $\mathbf{b}_2$,
respectively.
%
Figure~\ref{fig:bloomfilter} shows the encoding of bigrams ($q=2$)
of two string values into $l=14$ bits long BFs using $k=2$ hash
functions, and their Dice coefficient similarity calculation.

Different encoding methods have been proposed for BFs. Hashing
several attributes of a record into one BF is a method known as
cryptographic long term key (CLK). It is used to improve
privacy~\cite{Sch11}. Another record-level BF encoding (RBF) was
proposed to improve linkage quality~\cite{Dur14}. In RBF,
attribute values are first hashed into different BFs and then bits
are selected from each attribute-level BF into a RBF according to
attribute weights.

The initial proposal of BFs for PPRL used a double hashing
scheme~\cite{Sch09}, where the $k$ individual bit positions for an
element $s$ to be hashed are determined by the sum of the integer
representation of two independent hash functions that are mapped into
the range $1-l$. Random hashing has recently been proposed as an
improvement over double hashing to prevent against cryptanalysis
attacks, where $k$ random numbers are drawn for every element $s$ to
be hashed~\cite{Nie14,Sch16}.

%  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\smallskip

%\subsection{Bloom Filter Hardening}
%\label{sec-bf-harden}

\textbf{Bloom Filter Hardening:}~
Several BF hardening methods have been studied in recent times to
reduce the vulnerability of BFs against cryptanalysis
attacks~\cite{Sch16}. Compared to attribute-level BFs, record-level
BF encodings such as CLK and RBF reduce the risk of re-identification
by such attacks~\cite{Kuz11,Nie14}.

Exploiting the fact that data sets with (near) uniform Hamming
weight distribution of BFs are more difficult to attack (with
existing attack methods) than data sets with non-uniform
distributions, balancing BFs with constant Hamming weight has been
proposed~\cite{Sch16}. Balanced BFs can be constructed by
concatenating a BF of length $l$ with its negated copy (all bits
flipped) and then permuting the $2l$ bits. Another proposed approach
to harden BFs is XOR-folding, where a BF of length $l$ is split into
two halves of length $l/2$ each, and then bit-wise exclusive OR is
applied to combine the two shorter BFs~\cite{Sch16}.

While balancing and XOR-folding are easy and data independent
hardening techniques, salting with record-specific values has been
suggested as an alternative hardening method where an additional
(record-specific) value is concatenated with attribute values
before being hashed into the BF~\cite{Nie14}. A cryptanalysis attack
is unlikely to be successful without knowing the salting key, however
attributes suitable for salting might not be available in a database.
Other, more experimental hardening techniques, include random bits
and fake record injection~\cite{Dur14,Vat13}, as well as BLIP
(BLoom-and-flIP) which flips bits (noise addition) in a BF according 
to a differential privacy model~\cite{Sch16}. Many of these
hardening techniques improve security against attacks at the cost
of a reduction in linkage quality~\cite{Sch16}. In the experiments
in Sect.~\ref{sec-data} we will investigate if balancing and
XOR-folding make BFs more resistant to our proposed attack.

% --------------------------------------------------------------------

\section{Frequency-based Bloom Filter Cryptanalysis}
\label{sec-bf-attack}

% define problem? or define problem in sect 3?

% below explain what we aim to achieve

% I think we need to explain that our attack does better than simply
% sorting attribute values and BF by frequencies and aligning
% them! we do find which q-grams are at which position

We now describe our frequency-based attack on BFs in detail. As shown
in Fig.~\ref{fig:outline}, the attack consists of two main steps.
First, for each BF position we find its set of possible and
\emph{not} possible q-grams
%$p$, $1 \le p \le l$, we find its set of possible q-grams,
%$\mathbf{c}[p]$
(steps (1a) and (1b) in Fig.~\ref{fig:outline}).
%
Next, 
%using the list $\mathbf{C} = [\mathbf{c}[1], \ldots, \mathbf{c}
%[l]]$,
we re-identify for each BF  
%in $\mathbf{B}$
the set of attribute values
% $g_i$ from a set of values $\mathbf{G}$
that possibly were hashed into this BF (step (2) in
Fig.~\ref{fig:outline}).
%
Algorithms~1 and~2 show the details of steps (1a) and (1b), and (2),
respectively, which we describe in the next two sub-sections. We then
provide an analysis of our attack and discuss its limitations.

%  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Candidate Q-gram Set Generation}
\label{sec-cand-q-gram}

For our attack we require as input a set of BFs, $\mathbf{B}$, that
we assume to come from a sensitive database (the one from which we
aim to re-identify its sensitive values), and a set of attribute
values, $\mathbf{V}$, assumed to come from a public database. Each BF
$\mathbf{b}_i \in \mathbf{B}$ and each attribute value $v_i \in
\mathbf{V}$ has a frequency attached to it, denoted with
$\mathbf{b}_i.f$ and $v_i.f$, respectively. Unlike previous attacks on
BFs~\cite{Kuz11,Kuz13,Nie14}, we do not require any other information
about how the BFs were encoded.

\begin{figure}[t]
  \begin{center}
  \begin{scriptsize}
%    \begin{footnotesize}
  \begin{tabular}{ll} \hline
\multicolumn{2}{l}{\textbf{Algorithm 1: \emph{Candidate q-gram set
  generation}} -- Steps (1a) and (1b) } \\ \hline
\multicolumn{2}{l}{Input:} \\
\multicolumn{2}{l}{- $\mathbf{V}$: Set of attribute values and their
  frequencies from a public database} \\
\multicolumn{2}{l}{- $\mathbf{B}$: Set of BFs and their
  frequencies from the sensitive database} \\
\multicolumn{2}{l}{- $l$:~\, Length of Bloom filters} \\
\multicolumn{2}{l}{- $q$:~ Length of sub-strings to extract from
  attribute values} \\  
\multicolumn{2}{l}{- $m$: Minimum frequency for BFs and attribute
  values} \\
\noalign{\smallskip}
\multicolumn{2}{l}{Output:} \\
\multicolumn{2}{l}{- $\mathbf{C}$: List of possible q-grams for each
  BF position} \\
\noalign{\smallskip}
  1:  & $\mathbf{c}^+[p] = \{\}$, $\mathbf{c}^-[p] = \{\}$,
        $1 \le p \le l$ ~~~ // Initialize list of candidate q-gram
        sets \\
  2:  & $\mathbf{V}_F = \{v \in \mathbf{V}: v.f \ge m \}$
        ~~~~~~~~~~~~~~ // Attribute values with frequency of at
        least $m$ \\
  3:  & $\mathbf{B}_F = \{\mathbf{b} \in \mathbf{B}:
        \mathbf{b}.f \ge m \}$
        ~~~~~~~~~~~~~~\,// BFs with frequency of at least $m$ \\
  4:  & $revSort(\mathbf{B}_F)$, $revSort(\mathbf{V}_F)$ ~~~~~~~~~~ 
        // Sort according to frequencies, highest first \\
  5:  & $\mathbf{A} = [(\mathbf{b}_i, v_i) : 
        \mathbf{b}_i \in \mathbf{B}_F, v_i \in \mathbf{V}_F:
        \mathbf{b}_i.f > \mathbf{b}_j.f \land v_i.f > v_j.f:
        1 \le i < j \le min(|\mathbf{V}_F |, |\mathbf{B}_F|)]$ \\
  ~  &   ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ // Align $\mathbf{V}_F$
        and $\mathbf{B}_F$ as long as their frequencies are
        unique \\  
  6:  & \textbf{for} $(\mathbf{b}_i, v_i) \in \mathbf{A}$
        \textbf{do}: ~~~~~~~~~~~~~~~~~~~~ //  Step (1a): Get
        candidate sets of q-grams \\
  7:  & ~~~ $\mathbf{q}_i = genQGramSet(v_i, q)$ ~~~~~~~~\,//
        Convert attribute value into its q-gram set \\
  8:  & ~~~  \textbf{for} $1 \le p \le l$ \textbf{do}: 
        ~~~~~~~~~~~~~~~~~~~\, // Loop over all BF positions \\
  9:  & ~~~ ~~~ \textbf{if} $\mathbf{b}_i[p] == 1$ \textbf{then}:
        ~~~~~~~~~~~~\, // Bit at position p is 1 \\
  10: & ~~~ ~~~ ~~~ $\mathbf{c}^+[p] = \mathbf{c}^+[p] \cup
        \mathbf{q}_i$ ~~~~~~~~~~~\,// Add to set of possible
        q-grams \\
  11: & ~~~ ~~~ \textbf{else}: ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ //
        Bit at position $p$ is $0$ \\
  12: & ~~~ ~~~ ~~~ $\mathbf{c}^-[p] = \mathbf{c}^-[p] \cup
        \mathbf{q}_i$  ~~~~~~~~~~~// Add to set of \emph{not}
        possible q-grams \\
  13: & $\mathbf{C} = [\,]$ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\,
        // Step (1b): Initialize empty list of q-gram sets \\
  14: & \textbf{for} $1 \le p \le l$ \textbf{do}: 
        ~~~~~~~~~~~~~~~~~~~~~~~~\,// Loop over all BF positions to
        combine q-gram sets \\
  15: & ~~~ $\mathbf{C}.append(\mathbf{c}^+[p] \setminus
        \mathbf{c}^-[p])$ ~~~~~~~~~~~\,// Remove not possible from
        possible q-grams \\
%  15: & ~~~ $\mathbf{c}[p] = \mathbf{c}^+[p] \setminus
%        \mathbf{c}^-[p]$ ~~~~~~~~~~~~~~~~~ // Remove not possible
%        from possible q-grams \\
%  16: & ~~~ $\mathbf{C}.append(\mathbf{c}[p])$
%        ~~~~~~~~~~~~~~~~~~~~~~ // Add to list of candidates for all
%        BF positions\\
16: & \textbf{return} $\mathbf{C}$ \\
  \hline
  \end{tabular}
%  \end{footnotesize}
 \end{scriptsize}
  \end{center}
  \label{algo:step1}
\end{figure}

Algorithm~1 starts by initializing two empty sets for each BF
position $p$: $\mathbf{c}^+[p]$ of possible q-grams at that
position, and $\mathbf{c}^-[p]$ of \emph{not} possible q-grams at
that position. Next, in lines 2 and 3, we find all BFs and attribute
values that occur at least $m$ times. In line 4 we sort both the BFs
and attribute values according to their frequencies in reverse order
(most frequent first), and then we align a BF $\mathbf{b}_i$ and an
attribute value $v_i$ into the sorted list $\mathbf{A}$ of pairs
$(\mathbf{b}_i, v_i)$. We do this as long as both the $i$th BF
$\mathbf{b}_i$ and $i$th attribute value $v_i$ have a unique
frequency compared to the next, less frequent, BF or attribute
value, respectively. This stopping criterion ensures that we do not
have a BF that could correspond to two or more attribute values,
and vice versa, as this would lead to more uncertainty in the
mapping of q-grams into BF positions.

Step (1a) of our approach starts in line 6 and loops over pairs of
aligned BFs and attribute values, $(\mathbf{b}_i, v_i) \in
\mathbf{A}$. First, we convert $v_i$ into its set of q-grams,
$\mathbf{q}_i$, in line 7. Then we loop over all BF positions,
$1 \le p \le l$, in line 8, and if the bit at position $p$ in BF is
$1$  (i.e., $\mathbf{b}_i[p] = 1$), we add the q-gram set
$\mathbf{q}_i$ to the set $\mathbf{c}^+[p]$ of possible q-grams at
that position (line 10) because a $1$-bit means any q-gram from
$\mathbf{q}_i$ could have been hashed to that position. Conversely,
if the BF bit at position $p$ is $0$, then no q-gram from
$\mathbf{q}_i$ could have been hashed to that position and so we add
the q-grams in $\mathbf{q}_i$ to $\mathbf{c}^-[p]$ (line 12).

In step (1b), line 13 onwards in Algo.~1, we get the final set of
q-grams for each BF position $p$ as the set of possible q-grams
($\mathbf{c}^+[p]$) minus the set of not possible q-grams
($\mathbf{c}^-[p]$), and add these sets to the list %of q-gram sets
$\mathbf{C}$ in line 15.

%  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Attribute Value Re-identification}
\label{sec-reidentify}

\begin{figure}[t]
  \begin{center}
  \begin{scriptsize}
%    \begin{footnotesize}
  \begin{tabular}{ll} \hline
\multicolumn{2}{l}{\textbf{Algorithm 2: \emph{Attribute value
  re-identification}} -- Step (2) } \\ \hline
\multicolumn{2}{l}{Input:} \\
\multicolumn{2}{l}{- $\mathbf{B}$: Set of BFs from the sensitive
  database (same as in Algo.~1)} \\
\multicolumn{2}{l}{- $\mathbf{C}$: List of possible q-grams for
  each BF position (from Algo.~1)} \\
  \multicolumn{2}{l}{- $\mathbf{G}$: Set of attribute values we aim
  to re-identify in the set of BFs $\mathbf{B}$} \\
\noalign{\smallskip}
\multicolumn{2}{l}{Output:} \\
\multicolumn{2}{l}{- $\mathbf{R}$: List of possible attribute values
  re-identified for each BF in $\mathbf{B}$} \\
\noalign{\smallskip}
  1:  & $\mathbf{R} = [\,]$ ~~~~~~~~~~~~~~~~~~~~ // Initialize empty
        list of re-identified attribute values \\
%  1:  & $\mathbf{Q} = [\,]$ ~~~~ // Initialize empty list of q-gram
%        sets and attribute values \\
%  2:  & \textbf{for} $v_i \in \mathbf{G}$ \textbf{do}: ~~~~~~
%      // Loop
%      over all attribute values \\
%  3:  & ~~~ $\mathbf{q}_i = genQGramSet(v_i, q))$ ~ // Convert
%        attribute value into its q-gram set \\
%  4:  & ~~~ $\mathbf{Q}.append((\mathbf{q}_i, v_i))$ \\
  2:  & \textbf{for} $\mathbf{b}_i \in \mathbf{B}$
        \textbf{do}: ~~~~~~~~~ // Loop over all BFs \\
  3:  & ~~~ $\mathbf{g}_i = \mathbf{G}$ ~~~~~~~~~~~~~~~ //
        Initialize set of candidate attribute values as all
        possible values \\        
  4:  & ~~~  \textbf{for} $1 \le p \le l$ \textbf{do}: 
        ~~~~~~~~~~~~~~~~~~~~~~~\, // Loop over all BF positions \\
  5:  & ~~~ ~~~ \textbf{if} $\mathbf{b}_i[p] == 1$ \textbf{then}:
        ~~~~~~~~~~~~~~~~ // Bit at position $p$ is 1 \\
  6:  & ~~~ ~~~ ~~~ $\mathbf{c} = \mathbf{C}[p]$  ~~~~~~~~~~~~~~~~
        ~~~~~~~~~\, // Set of possible q-grams at this bit position
        \\
  7:  & ~~~ ~~~ ~~~ \textbf{for} $v_j \in \mathbf{g}_i$ \textbf{do}: 
        ~~~~~~~~~~~~~~~~~~~// Check all candidate attribute
        values \\
  8:  & ~~~ ~~~ ~~~ ~~~ \textbf{if} $(\forall q \in \textbf{c}:
        q \notin v_j)$ \textbf{then}: ~ // Check if no q-gram from
        $\mathbf{c}$ in attribute value \\
  9:  & ~~~ ~~~ ~~~ ~~~ ~~~ $\mathbf{g}_i = \mathbf{g}_i \setminus
        v_j$ ~~~~~~~~~~~~~~~\,// Value could not have been hashed
        into this BF \\ 
  10: & ~~~ $\mathbf{R}.append(\mathbf{g}_i)$ ~~~~~~~~~~~~~~~~~~~~~
        ~~~~~~ // Append possible attribute values for this BF \\
  11: & \textbf{return} $\mathbf{R}$ \\
  \hline
  \end{tabular}
%    \end{footnotesize}
  \end{scriptsize}
  \end{center}
  \label{algo:step2}
\end{figure}

The re-identification step, shown in Algo.~2, has as input the same
set of BFs, $\mathbf{B}$, as used in the first step, as well as the
list of candidate q-grams sets, $\mathbf{C}$, generated in Algo.~1.
A set of frequent attribute values, $\mathbf{G}$, also needs to be
provided. These are the values we aim to re-identify (guess) in
$\mathbf{B}$ using $\mathbf{C}$ (as shown in
Fig.~\ref{fig:outline}). The output of the algorithm is a set
of one or more re-identified attribute value(s), $\mathbf{g}_i
\subset \mathbf{G}$, for each BF $\mathbf{b}_i \in \mathbf{B}$,
collected in the result list $\mathbf{R}$.

The algorithm loops over all BFs $\mathbf{b}_i$ in $\mathbf{B}$ (line
2 onwards), and for each $\mathbf{b}_i$ it initializes the set of
possible candidate attribute values $\mathbf{g}_i$ (that potentially
have been hashed into this BF) as all values in $\mathbf{G}$ (line 3).
Next we loop over all BF positions $p$ (line  4), and for any position
that has a $1$-bit ($\mathbf{b}_i[p] = 1$) we retrieve the set of
possible q-grams $\mathbf{c} = \mathbf{C}[p]$ at that position
(line 6).

Using $\mathbf{c}$, we now check for each attribute value $v_j$ in
$\mathbf{g}_i$ if \emph{at least one of the q-grams} in $\mathbf{c}$
occurs in that value (lines 7 and 8). If $v_j$ does not contain at
least one q-gram from $\mathbf{c}$, it could not have generated the
$1$-bit at position $p$. If this is the case, in line 9 we remove
$v_j$ from the set of candidates $\mathbf{g}_i$. At the end of this
loop (line 10), the set $\mathbf{g}_i$ will contain all attribute
values from $\mathbf{G}$ that possibly have generated the BF
$\mathbf{b}_i$, and we append this set to the results list
$\mathbf{R}$. Note that positions with $0$-bits do not help us to
remove values $v_j \in \mathbf{g}_i$ because any q-gram in a value
$v_j$ is potentially hashed into other positions.
% Peter: not sure if this makes sense, we need to think further

%\textbf{\emph{In Algo.~2, we do not consider in a bit position $p$
%is $0$ (i.e. no else for the if in line 5). Are we missing something
%here? Would $0$-bits allow us to do more filtering? Can you please
%all have a think and let me know.. if there is nothing we can do with
%$0$-bits we should add a sentence why we cannot use them in the
%description of Algo.~2 below.}}

%% old below, PC 19/10 *****************************************

% Peter: Old and notes below,  maybe we can use somewhere:

%As with earlier attacks on BF-based PPRL approaches~\cite{...}, we
%assume that BFs are generated attribute-wise rather than record-wise,
%allowing us to analyze the frequency distribution of BFs as correlated
%with the frequency distribution of attribute values.

%Many attribute types commonly used in record linkage, such as first
%and last name, town or city names, postcodes or zipcodes, exhibit
%highly skewed frequency distributions of their values~\cite{?}. Often
%these distributions follow a Zipf-like distribution~\cite{}, where a
%few values occur very frequently (like surname `Smith' and `Miller' in
%English speaking countries), while the majority of values occur only
%rarely or once.

%As described in details below, we align the most frequent BFs from a
%set of BFs with the most frequent attribute values. We assume that
%an initial analysis of BF frequencies and attribute value frequencies
%(as shown in Fig.~\ref{xx}) has allowed us to identify which set of
%BFs corresponds to which attribute.

%Unlike other attack methods (which assume an attack knows certain
%parameters of the encoding approach used to generate the BFs~\cite{}),
%our attack does not require any such knowledge, making the attack more
%applicable and successful in situations where an attack only gains
%access to a set of BFs.

%Our attack exploits the fundamental properties of BFs: if a certain
%bit in a BF is not set to 1, then a corresponding element (q-gram in
%our case) of the set used to generate the BF cannot be in the %corresponding record.
%
%Using the alignment of frequent BFs and frequent attribute values,
%for each bit position of the BF we generate two sets: (1) the set of
%potential q-grams that \emph{cannot} be mapped to this position based
%on the q-grams in the frequent attribute values, and (2) the set of
%potential q-grams that \emph{can} be mapped to this position based on
%the q-grams in the frequent attribute values.

%- should we include how we find which BF correspond to which 
%  attributes? \\
% or is this an assumption? In section experiments we can show
% frequency plots of attributes and corresponding BF to show this
% step is easy
%
%- should we include the estimation of the number of hash functions k?
%  it is not too accurate at the moment, and somewhat independent of
%  our approach.

%  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Analysis and Limitations}

\label{sec-analysis}
We now discuss the complexity and limitations of our cryptanalysis
attack.
\smallskip

\textbf{Complexity:}
We assume $N = |\mathbf{B}|$ is the number of BFs coming from the
sensitive database, $Q$ is the average number of q-grams per value
in the set of attribute values $\textbf{V}$ from the public
database, $k$ is the number of hash functions used to hash q-grams
into BFs, and $l$ is the length of a BF.

In step (1) of our attack, we first find all frequent BFs and
attribute values that occur at least $m$ times. This requires a
linear scan through $\mathbf{B}$ and $\textbf{V}$, respectively,
which has a complexity of $O(N)$ and $O(|\textbf{V}|)$. In line 4 of
Algo.~1, the sorting function used on values in $\textbf{B}_F$ and
$\textbf{V}_F$ has a complexity of $O(M \cdot log\,M)$, where $M =
min(|\textbf{B}_F|, |\textbf{V}_F|)$.
%
In step (1a), each attribute value in $\textbf{A}$ is converted into
its set of q-grams which are then added to the sets $\mathbf{c}^+[p]$
and $\mathbf{c}^-[p]$ for all BF bit positions $1 \le b \le l$. This
step has a computational complexity of $O(M \cdot Q \cdot l)$. In
step (1b), the sets of possible q-grams are added to list
$\mathbf{C}$ by performing a set difference operation between
$\mathbf{c}^+[p]$ and $\mathbf{c}^-[p]$. Assuming $\mathbf{Q}$ is
the set of all unique q-grams generated from all values $v_i \in
\mathbf{V}$, then on average $t=|\mathbf{Q}|*k/l$ q-grams are hashed
into each bit position. Therefore each set difference is of $O(t)$,
and so step (1b) has an overall complexity of $O(t \cdot l)$ for $l$
bit positions.
%The memory requirements of our algorithm is of $O(\mathbf{C})$ at
%the end of step 1.

In step (2) of our attack, Algo.~2 iterates through each BF
$\mathbf{b}_i \in \mathbf{B}$ to re-identify the possible attribute
values $\mathbf{g}_i \in \mathbf{G}$ that map to $\mathbf{b}_i$. For
each bit position $p$ that is set to $1$, we check if any q-gram in
$\mathbf{C}[p]$ occurs in a value $v_j \in \mathbf{g}_i$. This leads
to a total complexity of $O(N \cdot |\mathbf{G}| \cdot l \cdot t)$,
assuming the average size of a $\mathbf{C}[p]$ is $t$. Finally, the
worst case space complexity of the list $\mathbf{R}$ returned by
Algo.~2 is $O(N \cdot |\mathbf{G}|)$ if every value in $\mathbf{G}$
is mapped to each $\mathbf{b}_i \in \mathbf{B}$.

%the set of possible attribute values $g_i$ are identified from
%$\mathbf{G}$ if any of their q-grams are available in the set of 
%candidate q-grams $C[p]$. Thus, step 2 of our attacks has a
%computational complexity of $O(N\cdot|G|\cdot l)$. Finally, a list
%$R$ of possible attribute values for $B$ is returned from Algorithm
%$2$ which requires a worst case space complexity of $O(G\cdot B)$ if
%all attribute values of $G$ are mapped to each $b\in B$.
\smallskip

%What are the limitations?
\textbf{Limitations:} Our frequency-based cryptanalysis attack on
BFs depends on several assumptions. First, we assume the attacker
has access to a publicly available population database from where
the set $\mathbf{V}$ of attribute values and their frequencies can
be extracted. We also assume that $\mathbf{B}$ does contain a sub-set
of BFs that occur several times, and that the frequency distribution
of BFs in $\mathbf{B}$ is similar to the frequency
%can be aligned by an attacker to the
distribution of a single or a sub-set of attribute values in
$\mathbf{V}$. Without such frequency information, that can be
aligned, our attack (like previous cryptanalysis attacks on
BFs~\cite{Kro15,Kuz11,Kuz13,Nie14}) would not be possible.

%$G$ which is a superset of the set of attribute values which are
%encoded in BFs in $B$. We also assume based on the frequency
%distribution of BFs, the attacker can identify the list of
%attributes that are encoded in $B$ which he used to extract the
%sets of frequent values from $G$.

% --------------------------------------------------------------------

\section{Experiments and Results}
\label{sec-data}

We conducted our experimental study using real data sets from two
domains.
%
The first are a pair of North Carolina Voter Registration (NCVR) data
sets (\url{ftp://alt.ncsbe.gov/data/})
%~\footnote{Available from \url{ftp://alt.ncsbe.gov/data/}}
collected in June 2014 (the database to be attacked) and October 2016
(the public database).
%, that have increasingly large differences in their time collection
%to see how this influences re-identification accuracy.
These data sets contain over five million records of voters including
their first names, surnames, and addresses. We present results of our
attack individually on the first name
%, last name, and address attributes, 
attribute, as well as on the concatenation of the first name and
surname attributes.

\begin{figure}[!t]
  \centering
%  \includegraphics[width=0.31\textwidth]
%  {reident-bf_len-ncvoter_20140619_1}
%  ~~
%  \includegraphics[width=0.31\textwidth]
%  {reident-hash_type-ncvoter_20140619_1}
%  ~~
%  \includegraphics[width=0.31\textwidth]
%  {reident-q-ncvoter_20140619_1}
  %\\ ~ \\
  %\includegraphics[width=0.31\textwidth]
  %{reident-bf_len-ncvoter_20140619_3}
  %~~
  %\includegraphics[width=0.31\textwidth]
  %{reident-hash_type-ncvoter_20140619_3}
  %~~
  %\includegraphics[width=0.31\textwidth]
  %{reident-q-ncvoter_20140619_3}
  %\\ ~ \\
  %\includegraphics[width=0.31\textwidth]
  %{reident-bf_len-ncvoter_20140619_6}
  %~~
  %\includegraphics[width=0.31\textwidth]
  %{reident-hash_type-ncvoter_20140619_6}
  %~~
  %\includegraphics[width=0.31\textwidth]
  %{reident-q-ncvoter_20140619_6}
%  \\ ~ \\
%\includegraphics[width=0.31\textwidth]
%  {reident-bf_len-ncvoter_20140619_1-3}
%  ~~
%  \includegraphics[width=0.31\textwidth]
%  {reident-hash_type-ncvoter_20140619_1-3}
%  ~~
%  \includegraphics[width=0.31\textwidth]
%  {reident-q-ncvoter_20140619_1-3}
%
% results with no hardening below
%~ \\~\\~\\
  \includegraphics[width=0.31\textwidth]
  {reident-bf_len-ncvoter_20140619_1}
  ~~
  \includegraphics[width=0.31\textwidth]
  {reident-hash_type-ncvoter_20140619_1}
  ~~
  \includegraphics[width=0.31\textwidth]
  {reident-q-ncvoter_20140619_1}
  \\ ~ \\
\includegraphics[width=0.31\textwidth]
  {reident-bf_len-ncvoter_20140619_1-3}
  ~~
  \includegraphics[width=0.31\textwidth]
  {reident-hash_type-ncvoter_20140619_1-3}
  ~~
  \includegraphics[width=0.31\textwidth]
  {reident-q-ncvoter_20140619_1-3}
\caption{Results for the NCVR data sets with first name (top)
    and combination of first and surname (bottom) for different BF
    lengths (left), different hashing methods (middle), and different
    values of $q$ (right). No BF hardening was applied.
   \label{fig:ncvr}}
\end{figure}

The second are a pair of census data sets (named UKCD) collected
from the years 1851 and 1901 (used as the sensitive and public
databases, respectively) for the town of Rawtenstall in
England~\cite{Fu14b}. These data sets contain around 50,000
records with personal details of individuals. We again use the
first name attribute, and the concatenation of the first name and
surname attributes.
%surname and address attributes.

The parameter settings we use in our experiments are $q = [2,3,4]$
(length of sub-strings used in BF encoding), BF length
$l = [250, 500, 1000,$ $2000]$, and either the double or random
hashing method~\cite{Sch16} as described in
Sect.~\ref{sec-overview}. We calculate the number of hash functions
$k$ based on $l$ and $q$ such that the false positive rate is
minimized~\cite{Vat14c}. We also apply the BF hardening techniques
balancing and XOR-folding~\cite{Sch16}.
% as described in Sect.~\ref{sec-overview}.
We use different numbers for the most frequent attribute values
(ranked by their frequencies) to be re-identified: $|\mathbf{G}| =
[10,20,50,100]$.
%most frequent values from the public data set to be guessed.

We evaluate the accuracy of our attack by calculating (1) the
percentage of correct guesses with 1-to-1 matching, (2) the
percentage of correct guesses with 1-to-m (many) matching, (3) the
percentage of wrong guesses, and (4) the percentage of no guesses,
where these four percentages sum to 100. These four categories are
labeled as `1-1 corr', `1-m corr', `Wrong', and `No' in the plots,
respectively. We evaluate the efficiency of our attack using run
time.

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.31\textwidth]
    {reident-bf_len-1851_allcleaned_1901_allcleaned_6}
  ~~
  \includegraphics[width=0.31\textwidth]
    {reident-hash_type-1851_allcleaned_1901_allcleaned_6}
  ~~
  \includegraphics[width=0.31\textwidth]
    {reident-bf_harden-1851_allcleaned_1901_allcleaned_6}
    \\ ~ \\
  \includegraphics[width=0.31\textwidth]
    {reident-bf_len-1851_allcleaned_1901_allcleaned_5-6}
  ~~
  \includegraphics[width=0.31\textwidth]
    {reident-hash_type-1851_allcleaned_1901_allcleaned_5-6}
  ~~
  \includegraphics[width=0.31\textwidth]
    {reident-bf_harden-1851_allcleaned_1901_allcleaned_5-6}
  \caption{Results for the UKCD data sets with first name (top) and
  combination of first and surname (bottom) for different BF lengths
  (left), different hashing methods (middle), and different BF
  hardening methods (right).
  \label{fig:ukcd}}
\end{figure}

We implemented our attack using Python 2.7 and ran all experiments on
a server with 64-bit Intel Xeon 2.4 GHz CPUs, 128 GBytes of memory
and running Ubuntu 14.04. The programs and data sets are available
from the authors.

%Toto: generate a similar figure for UKCS? add data set table
%(num rec, attributes use?, num matches? what do we need?
%discuss data sets, table with details - UKC, NCVR balanced,
%NVR 2011 to 2016 (select
%a few, give number of records, number of values in selected attributes)
%show plots with freq distributions of 3 selected attribute values
%(first name, last name, postcode, city?) and their corresponding BF distributions
%If room, show for different data sets.
%experimental setup
%also show if possible for pairs of attributes mapped into 1 BF

\smallskip

\textbf{Discussion:}~
In Fig.~\ref{fig:ncvr} we show the results for the NCVR data sets.
% for one individual attribute (first name) in the top row, and
%three individual attributes (first name, last name, and address in
%the first three rows, respectively), and
%combined attributes of first and surnames in the bottom row.
% BFs are hardened using the balanced method.
%
As can be seen, when the BF length $l$ increases (left column) the
percentage of correct re-identifications mostly increases, as with
larger $l$ the number of q-grams mapped to a certain bit position
decreases. All values can be correctly re-identified for individual
attributes
%nearly $90\%$ of values can be correctly re-identified
%%as 1-to-1 or 1-to-many matches
when the number of frequent values is 10.
%or 20. 
Around half of all
values can still be re-identified even when values from two
attributes are combined.
%
The middle column
%in Fig.~\ref{fig:ncvr}
shows that random hashing
(which supposedly improves privacy on BFs compared to double
hashing~\cite{Kro15,Nie14}) does not provide improved protection
against our attack, as a similar percentage of values can be
correctly re-identified for both hashing approaches. As for using
different values of $q$ (right column)
%in Fig.~\ref{fig:ncvr}),
the accuracy of an attack somewhat improves when $q$ is increased
because larger values of $q$ result in more unique q-grams.

The accuracy results for the UKCD data sets are shown in
Fig.~\ref{fig:ukcd} for the same attributes as for the NCVR data
sets. Similar re-identification patterns as with the NCVR data sets
can be seen for different BF lengths (left) and hashing methods
(middle). We also study how different BF hardening methods affect the
re-identification accuracy, as shown in the right column plots in
Fig.~\ref{fig:ukcd}. For the single attribute case (top right),
neither of the hardening techniques is capable of reducing the
re-identification accuracy, however for the combined attribute
case both hardening techniques improve the privacy of BF encoding.

\begin{table}[!t]
 %\addtolength{\tabcolsep}{-3pt}
  \caption{Comparison of re-identification results with existing BF
     attack methods.}
     \label{table:exist_approaches}
  \centering
  \begin{scriptsize}
    \begin{tabular}{llccc}
    \hline\noalign{\smallskip}
   Publication & Data set & Num BFs~ & ~1-1 corr~ & Run time
     \\ \noalign{\smallskip}\hline\noalign{\smallskip}
  Kuzu et al. (2011)~\cite{Kuz11} & NCVR first names & 3,500 & 400 &
    1,000 sec \\
  Kuzu et al. (2013)~\cite{Kuz13} & Patient names & 20 & 4 & 
    few sec \\
  ~~~~ " & ~~~~ " & 42 & 0 & $>$ week \\
  Niedermeyer et al. (2014)~\cite{Nie14}~ & German surnames & 7,580 &
    934 & $>$ days  \\
  Kroll and Steinmetzer (2015)~\cite{Kro15}~ & German names and
    locations & 100K & 44K & $>$ days \\
  Our approach & NCVR first names & 10--100 & 7--10 & 0.73--0.75 sec
    \\
  ~~~~ " & NCVR first and surnames & 10--100 & 3--6 & 1.5--1.9
  sec  \\
%  ~~~~ " & UKCD surname and address~ & 10 & 10 & 0.05 sec \\
  \noalign{\smallskip} \hline
  \end{tabular}
  \end{scriptsize}
\end{table}

% -------------------
%DV: Peter, for your question on what is the difference between 2 and
%3 in table 1:
%In that paper they first tried to re-identify 40 most frequent
%values ant it took more than a week without any results (0 
%re-identification percentage), and then they used 2 most frequent
%names and it results in 20% (i.e. 4 values are re-identified)
%accuracy in few seconds (they didn't mention the actual number of
%seconds).
%
%That's why I included both entries in the table to show the %scalability of our proposed approach.

% --------------

In Table~\ref{table:exist_approaches} we compare our attack method
with existing approaches in terms of accuracy and efficiency (with
results taken from the corresponding papers). As can be seen, our
method is both more efficient and effective in re-identification,
having both higher accuracy and reduced computational requirements.
% than existing approaches.

These results show the vulnerability of basic BF encoding to our
novel attack. They highlight the need for improved hardening
techniques to overcome such attacks. Our attack provides data
custodians with an efficient method to evaluate the privacy of their
BF encoded databases before using them for PPRL.

\smallskip
\textbf{Recommendations:}~
As a set of guidelines for the practical application of BF based PPRL
systems, to limit the vulnerability of such systems to known attack
methods we recommend to use record-level BF encoding (CLK or RBF),
apply advanced BF hardening methods~\cite{Sch16}, and reduce the
frequency of bit patterns (for example, by salting) to prevent any
frequency analysis.

% --------------------------------------------------------------------

\section{Conclusions and Future Work}
\label{sec-concl}

We have presented a novel efficient frequency-based attack on Bloom
filters (BFs) that contain encoded sensitive attribute values
intended for privacy-preserving record linkage (PPRL). Unlike earlier
attacks on BFs for PPRL, our approach only requires an attacker to
have access to a public database of attribute values, but no
information about the BF encoding used.
%Most importantly, it is independent of the kind of hash function
%used for BF encoding.
Our approach is faster than earlier attacks, making it feasible for
database owners to efficiently validate the security of their
encoded sensitive databases before they are being sent to other
parties for conducting PPRL. We believe our attack is an important 
component to making PPRL more secure for practical applications.

As future work we will study how our approach can be modified for
attacking composite and record-level BFs. We also plan to investigate
the risk of re-identification when advanced hardening techniques, such
as BLIP or BF salting, have been applied. Finally, our attack can be
accelerated by further analyzing the re-identified attribute values,
while correlations between $1$-bits and q-gram sets across BFs could
be identified using association rule mining
techniques~\cite{Ceg06}.

% --------------------------------------------------------------------

%\section*{Acknowledgments}
%
%Peter Christen was supported by a grant from the Simons Foundation.
%The authors would also like to thank the Isaac Newton Institute for
%Mathematical Sciences, Cambridge, for support and hospitality during
%the programme \emph{Data Linkage and Anonymisation} where this work
%was conducted (EPSRC grant EP/K032208/1). This work was partially
%funded by the Australian Research Council under Discovery Project
%DP130101801.

% --------------------------------------------------------------------

%\bibliographystyle{abbrv}
\bibliographystyle{splncs03}
\bibliography{paper}

\end{document}

% ====================================================================

### 2017-09-25 paper.tex
Update on Overleaf.

@@
 
-% PAKDD 2017: maximum 12 pages
+% PAKDD 2018: maximum 12 pages
 % Abstract max 200 words
@@
 
-% ideas for a better shorter title?
-\title{Efficient Cryptanalysis of Bloom Filters for
-       Privacy-Preserving Record Linkage
-  \thanks{The authors would like to thank the Isaac Newton Institute
-  for Mathematical Sciences, Cambridge, for support and hospitality
-  during the programme \emph{Data Linkage and Anonymisation} where
-  this work was conducted (EPSRC grant EP/K032208/1). Peter Christen
-  was also supported by a grant from the Simons Foundation. The work
-  was also partially funded by the Australian Research Council
-  (DP130101801).}}
+\title{Efficient and accurate record linkage using similarity spaces}
+%  \thanks{The authors would like to thank the Isaac Newton Institute
+%  for Mathematical Sciences, Cambridge, for support and hospitality
+%  during the programme \emph{Data Linkage and Anonymisation} where
+%  this work was conducted (EPSRC grant EP/K032208/1). Peter Christen
+%  was also supported by a grant from the Simons Foundation. The work
+%  was also partially funded by the Australian Research Council
+%  (DP130101801).}}
 
-\author{Peter Christen\inst{1} \and Rainer Schnell\inst{2}
-        \and Dinusha Vatsalan\inst{1} \and Thilina
-        Ranbaduge\inst{1}}
-\institute{Research School of Computer Science,
-           The Australian National University, \\
-           Canberra, Australia.~
-           Contact: \email{peter.christen@anu.edu.au}
-           \and
-           Methodology Research Group, %Institute for Sociology,
-           University Duisburg-Essen, Duisburg, Germany.} %~
-%           \email{rainer.schnell@uni-due.de}}
+\author{Submitted for double-blind review}
+%\author{Peter Christen\inst{1} \and Rainer Schnell\inst{2}
+%        \and Dinusha Vatsalan\inst{1} \and Thilina
+%        Ranbaduge\inst{1}}
+%\institute{Research School of Computer Science,
+%           The Australian National University, \\
+%           Canberra, Australia.~
+%           Contact: \email{peter.christen@anu.edu.au}
+%           \and
+%           Methodology Research Group, %Institute for Sociology,
+%           University Duisburg-Essen, Duisburg, Germany.} %~
+%%           \email{rainer.schnell@uni-due.de}}
 
@@
 \begin{abstract}
-Privacy-preserving record linkage (PPRL) is the process of identifying
-records that represent the same entity across databases held by
-different organizations without revealing any sensitive information
-about these entities. A popular technique used in PPRL is Bloom
-filter encoding, which has shown to be an efficient and effective way
-to encode sensitive information into bit vectors while still enabling
-approximate matching of attribute values. However, the encoded values
-in Bloom filters are vulnerable to cryptanalysis attacks. Under
-specific conditions, these attacks are successful in that some
-frequent sensitive attribute values can be re-identified.
-%
-%While previous attacks have high computational requirements
-In this paper we propose and evaluate on real databases a novel
-efficient attack on Bloom filters. Our approach is based on the
-construction principle of Bloom filters of hashing elements of sets
-into bit positions. The attack is independent of the encoding
-function and its parameters used, it can correctly re-identify
-sensitive attribute values even when various recently proposed
-hardening techniques have been applied, and it runs in a few seconds
-instead of hours.
+% usually about 200 words max- best to check
+
 \end{abstract}
 
-\keywords Privacy; re-identification; frequency analysis; data
-          linkage.
+\keywords Entity resolution; deduplication; metric spaces; etc.
 
 % ====================================================================
+
+\section{The Plan}
+
+We tried to develop a plan with PeterC. Here is the photo of the white-board.
+
+\includegraphics[width=\textwidth]{2017-08-01_12_08_43.jpg}
+
+\vspace{5mm}
+
+But, even better, here it is in a bulleted list.
+
+\begin{itemize}
+\item Design: Traditionally it has two stages.
+A blocking method needs to be developed separately from a similarity method.
+With similarity search based record linkage, there is only one method, and that is the similarity method.
+\item Efficiency: I am not convinced of this, but Peter thinks there may be efficiency benefits.
+The reasoning behind this is that with blocking-based methods we would have to calculate some sort of a similarity value first.
+This value will only be used for blocking, and later a similarity value between candidate pairs will need to be calculated again.
+\item Completeness: Similarity search based methods will be complete (with respect to the given similarity method) whereas blocking based methods (including LSH) are likely to be incomplete (there will be high-similarity pairs which are not places in the same block)
+\item Data sets
+\begin{itemize}
+\item Kilmarnock Birth-Death
+\item Skye Birth-Death
+\item CORA
+\item North Carolina Voter Registration DB
+\end{itemize}
+\item The following is a list of all methods we thought we would have to compare.
+\begin{itemize}
+\item Similarity search (M-Tree, others?)
+\item LSH-blocking
+\item Traditional blocking (e.g. Region, Surname)
+\item Sorted Neighbourhood blocking
+\end{itemize}
+\item Pairs-completeness and pairs-quality (w.r.t high similarity instead of true-link status). I will expand on this a bit more later.
+\item Comparison (for blocking based methods)
+		Precision/Recall/F1-Measure
+\item Which similarity methods are metric?
+\item Run time and memory usage comparisons
+\item Theorem. Precision might get better or worse with similarity search based methods, but recall should never get worse.
+\end{itemize}
 
@@
 
-Integrating data from different sources with the aim to remove
-duplicates, enrich data, and correct errors and inconsistencies is a
-crucial data pre-processing task for many data mining and analytics
-applications~\cite{Chr12}. Example applications include healthcare,
-business analytics, national censuses, population informatics, fraud
-detection, government services, and national security.
+something about record linkage and its importance, application examples
 
-However, growing concerns about privacy and confidentiality
-increasingly preclude the exchange or sharing of personal identifying
-attributes, such as names, dates of birth, and addresses, which are
-generally required for linking databases due to the non-existence of
-common unique entity identifiers~\cite{Chr12,Vat13}. Work in
-privacy-preserving record linkage (PPRL) aims to develop techniques
-for identifying records that correspond to the same entity across
-several databases while not compromising the privacy and
-confidentiality of the entities~\cite{Vat13}.
+drawback of existing approaches: blocking and comparisons/classification
+done separately, sometimes if bad linkage quality means to go back to redo
+blocking.
 
-PPRL is achieved by conducting linkage on the encoded (masked) values
-of the identifying attributes of records across two or more databases.
-%~\cite{Fie05}.
-%
-Several data encoding techniques for PPRL have been developed. These
-can be categorized into cryptographic secure multi-party computation
-(SMC) and perturbation-based techniques~\cite{Vat13}. The former are
-accurate and provably secure, but they incur expensive computation
-and communication costs.
-%The latter category includes more efficient techniques which often
-%provide a trade-off between accuracy and privacy.
-%Since existing cryptographic-based techniques are not scalable for
-%real applications,
-Most PPRL techniques are therefore based on perturbation-based
-techniques that provide adequate privacy protection while achieving
-acceptable linkage quality~\cite{Ran16,Vat13}.
+something about similarity spaces / metric spaces, M-trees and its
+more recent advanced/improved versions, and how they have been used in
+similarity search.
 
-Bloom filter (BF) encoding is one such perturbation-based technique
-that has successfully been used in several recent practical PPRL
-applications~\cite{Boy15,Ran14}. A BF is a binary vector with bits
-initially set to $0$. 
-%with bits set to $0$ or $1$. As we describe in detail in
-%Sect.~\ref{sec-overview},
-A value can be encoded into a BF using a set of hash functions by
-setting corresponding bits to $1$~\cite{Sch09}, and the approximate
-similarity between two BFs can be calculated by counting the number
-of positions where both BFs have $1$-bits in common.
-
-% Peter: I don't think this next paragraph is needed for this paper
-% as it is not needed to understand the paper
-%The similarity (or distance) between a pair of values is preserved in
-%the BF space such that the similarity calculated using their BF
-%encodings is greater than or equal to the similarity calculated using
-%their unencoded values. BF-based similarity calculation allows false
-%positives due to collision of bits (where different values are
-%hash-mapped into the same bit) which improves the privacy aspect
-%while compromising accuracy (depending on the false positive rate).
-%Therefore, it is crucial to set the BF related parameters ($k$, $l$,
-%and number of elements to be hash-mapped into a BF) appropriately to
-%balance the accuracy and privacy aspects of
-%linkage~\cite{Sch09,Vat16}.
-
-As we discuss in detail in the next section, BFs can be susceptible
-to cryptanalysis attacks that aim to re-identify the encoded
-sensitive attribute values~\cite{Kro15,Kuz11,Kuz13,Nie14}. Using 
-frequency counts and patterns in a set of BFs, these attacks
-iteratively map bit patterns to known attribute values. These existing
-attacks are however not practical as they require knowledge of certain
-parameters used during the BF encoding phase, and they have high
-computational costs.
-
-Our contribution in this paper is an efficient frequency-based
-approach for attacking BFs that exploits the fundamental property of
-how the elements of sets are hashed into BFs, as we describe in
-Sect.~\ref{sec-overview}. In contrast to existing attack methods, our
-novel approach does not require any assumption on the BF parameters
-used when sensitive attribute values were encoded. It is also
-significantly faster, making it a viable attack on large sets of
-BFs to evaluate whether they provide adequate privacy protection. We
-experimentally evaluate our attack method on two real-world data
-sets, showing its efficiency and effectiveness.
-
-Given BF encoding is now being employed in real-world PPRL
-applications~\cite{Boy15,Ran14}, it is crucial to study possible
-attacks on BFs to ensure their security and to make users of such
-systems aware of the weaknesses of BF encoding. Our novel attack
-method allows data custodians to identify such weaknesses of BF
-encoding that otherwise could be exploited by an attacker.
-%if appropriate BF encodings and parameters are not employed.
-
-%Maybe add: Contributions:
-%We propose a novel and efficient BF attack method and analyze its
-%complexity
-
-% PC 17/01/2107 until here **************************
+How our work is different from these other metric space works, and
+description of our contributions
 
@@
 
-\section{Prior Attacks on Bloom Filter based PPRL}
+\section{Related Work}
 \label{sec-related}
 
-%Bloom filters (BFs)~\cite{Blo70} are commonly used for the encoding of
-%records in PPRL due to their capability of computing
-%similarities~\cite{Dur14,Ran16,Sch09,Vat13}.
-Schnell et al.~\cite{Sch09} were the first to introduce an
-approximate matching approach for PPRL using Bloom filters (BFs), as
-we describe in detail in Sect.~\ref{sec-overview}. A recent study by
-Randall et al.~\cite{Ran14} has shown that PPRL based on BF encoding
-can achieve similar linkage quality as can be achieved with
-traditional linkage methods on the unencoded attribute values. As a
-result, BF encoding has been the PPRL technique of choice in several
-recent practical PPRL applications~\cite{Boy15,Ran14}.
-%
-%However, BFs are prone to different
-%attacks~\cite{Kro15,Kuz11,Kuz13,Nie14} which will be described %shortly.
-%
-However, BFs are prone to different
-attacks~\cite{Kro15,Kuz11,Kuz13,Nie14}. Therefore, a comprehensive
-analysis of the weaknesses of BFs in the PPRL context is required. We
-now describe the few studies that have been conducted on attacks on
-BFs.
+summarise related work in blocking/indexing
 
-% From Dinusha's introduction:
-%An attack using a constraint satisfaction solver was studied by
-%Kuzu et al.~\cite{Kuz11,Kuz13}, where values are mapped to bit
-%patterns that meet constraints such as frequency alignments. A
-%successful attack on BFs using filtering and statistical analysis
-%techniques was recently investigated by Niedermeyer et
-%al.~\cite{Nie14}. Based on the linear combination of the values used
-%to generate hash functions, this attack first generates a set of
-%possible bit patterns (so called \emph{atoms}) that are then aligned
-%according to their frequencies with known attribute
-%values~\cite{Nie14}.
+summarise related work in metric spaces as used in record linkage
 
-Kuzu et al.~\cite{Kuz11} formulated a cryptanalysis attack on BFs
-as a constraint satisfaction problem (CSP). CSP is characterized by
-a set of variables and a set of constraints on these variables. The
-aim of this attack is to identify a set of values from a given domain
-that can be assigned to each variable such that the constraints are
-satisfied. This is achieved by a frequency analysis of the sensitive
-attribute values and the BF encodings of the records in a database.
-However, this attack requires that the attacker has access to a
-global database where the encoded records are drawn from. This
-is unlikely in practical applications.
-
-In 2013, Kuzu et al.~\cite{Kuz13} investigated the accuracy of their
-CSP attack with two real-world databases. The authors tried to
-re-identify the personal details of patients in an encoded medical
-database by using a frequency analysis of BF encodings of a public
-voter registration database that has a different frequency
-distribution to the medical database. Their results indicated that
-although the CSP attack might be feasible in such situations, it is
-less likely to be accurate in identifying original attribute
-values and it requires more computational resources. The attack
-re-identified four out of 20 frequent names correctly.
-
-% Peter: results below are different to Table 1? *******************
-
-Niedermeyer et al.~\cite{Nie14} more recently proposed an attack on
-BFs built from German surnames. The attack was based on the
-frequencies of sub-strings of length 2 extracted from frequent
-surnames. Of $7,580$ surnames, the authors re-identified the $934$
-most frequent ones (about $12\%$) before stopping the attack. In
-contrast to the approach in~\cite{Kuz13}, this attack only 
-%is not based on the assumption that the attacker has
-%access to a population data set from which the surnames are drawn.
-%However, the feasibility of this approach 
-depends on the availability of a list of the most common surnames.
-This work was extended by Kroll and Steinmetzer~\cite{Kro15} into a
-cryptanalysis on BF encodings of several attributes, which was able
-to re-identify $44\%$ of all attribute values correctly. However, both
-attacks are based on the specific double hashing scheme used by
-Schnell at al.~\cite{Sch09}.
-
-Existing cryptanalysis attacks are feasible only for certain
-settings and assumptions used in the BF encoding phase. They also
-require excessive computational resources making them not practical
-in real settings. Our novel attack method, described next, improves
-on both these drawbacks of existing methods.
+other related works as relevant
 
@@
   \centering
-  \includegraphics[width=1.0\textwidth]{bf-attack-outline}
-  \caption{Outline of the proposed cryptanalysis attack, which is
-           based on a set of BFs and a set attribute values, both
-           sorted according to their frequencies. In steps (1a) and
-           (1b), the attack exploits the bit patterns in BFs to
-           identify the sets of q-grams that are possible,
-           $\mathbf{c}^+[p]$, and \emph{not} possible,
-           $\mathbf{c}^-[p]$, respectively, for each bit position
-           $p$. In step (2), the re-identification of a set of
-           (sensitive) attribute values, $\mathbf{G}$, is conducted
-           by intersecting the q-gram sets of values in $\mathbf{G}$
-           with the sets $\mathbf{c}$ (illustrated using $\odot$).
+%  \includegraphics[width=1.0\textwidth]{xyz.eps}
+  \caption{Descriptive caption
            %as outlined in Sect.~\ref{sec-overview} and detailed in
            %Sect.~\ref{sec-bf-attack}.
-           \label{fig:outline}}
+           \label{fig:xyz}}
 \end{figure}
 
-\section{Overview and Preliminaries}
+\section{approach section?}
 \label{sec-overview}
 
-We now provide an overview of our attack on BFs, as illustrated in
-Fig.~\ref{fig:outline}. As with other attacks on
-BFs~\cite{Kro15,Kuz11,Kuz13,Nie14}, our approach exploits the
-frequency distribution of a set of BFs that were generated from a
-large database. As for notation, we use bold letters for sets
-(with upper-case bold letters for lists or sets of sets) and normal
-type letters for integer or string values. We denote sets with curly
-and lists with square brackets, where lists have an order while sets
-do not.
+coverage: proportion of all rec pairs with a certain similarity
+(independent of their match status) that are being generated by
+a blocking  technique
 
-We assume the attacker has access to a set of encoded BFs,
-$\mathbf{B}$, and their frequencies, but he does not know anything
-about the parameters used in the encoding process (such as the number
-of hash functions used, or the actual hashing mechanism). We assume
-these BFs represent a set of records that encode sensitive values
-from one or a few attributes. Based on the frequency distribution of
-the Hamming weights (number of $1$-bits) in $\mathbf{B}$, the attacker
-can guess which attribute(s) have been encoded, because different
-attributes (such as first name, surname, city name, or postcode) have
-distinctive distributions of Hamming weights~\cite{Sch16}. The
-distribution of Hamming weights is independent of the (unknown)
-secret key. Therefore, an attacker can sample attribute values from
-a publicly available population database (such as a telephone
-directory) and select a set of frequent values, $\mathbf{V}$, from
-an attribute that has a frequency distribution that is similar to
-the distribution of the set to be attacked.
-%matches with the distribution
-%of the set of BFs $\mathbf{B}$ to be attacked.
+transitive closure is a problem if we have a blocking technique with overlapping blocks.
 
-In step (1), we first align BFs and attribute values according to
-their frequencies, and consider the set of most frequent values in 
-both. For each bit position $p$ in the BFs,
-%of length $l$ (with $1 \le p \le l$),
-for all corresponding attribute values that have this bit set to $1$
-we add their q-grams (sub-strings of length $q$ generated from
-attribute values) to the set $\mathbf{c}^+[p]$ of possible
-q-grams for that position. The reasoning is that a $1$-bit means at
-least one q-gram of an attribute value was hashed to this position.
-For all attribute values with a value of $0$ at bit
-position $p$ we add their q-grams to the set $\mathbf{c}^-[p]$ of
-\emph{not} possible q-grams for that position, because a $0$-bit
-means no q-gram of an attribute value could have been mapped to this
-position.
-
-At the end of step (1), for each position $p$ we obtain the set
-$\mathbf{c}[p] = \mathbf{c}^+[p] \setminus \mathbf{c}^-[p]$ of
-q-grams that potentially could have been hashed to position $p$.
-Based on the list $\mathbf{C} = [\mathbf{c}[1], \ldots,
-\mathbf{c}[l]]$, where $l$ is the length of the BFs, and a set
-$\mathbf{G}$ of attribute values we aim to re-identify (i.e.\ learn
-which BF possibly encodes which value in $\mathbf{G}$), in step (2)
-we analyze each BF in $\mathbf{B}$ and remove those attribute values
-from $\mathbf{G}$ that are not possible matches according to
-$\mathbf{C}$ because they do not contain any q-grams that would have
-been hashed to a certain $1$-bit.
-
-For example, in Fig.~\ref{fig:outline}, for the most frequent BF$_1$,
-`kate' is not a possible value because in order to obtain a $1$-bit
-in position $p_1$, it would have to contain either the q-gram `en'
-or `ry'; while `mary' is also not possible because it would need to
-contain one of the q-grams `ka', `en', `at', or `te' in position
-$p_3$.
-
-Before we formalize and present our approach in detail in
-Sect.~\ref{sec-bf-attack}, we first describe BF encoding, as well as
-some recent approaches to harden them.
+define blocking: only aimed at improving computational aspects by reducing the number of record pairs to be compared
 
@@
 
-\smallskip
-
-%\subsection{Bloom Filter Encoding}
-%\label{sec-bf-encoding}
-
-\begin{figure}[!t]
-  \centering
-  %\scalebox{1.0}[1.0]
-  \includegraphics[width=0.65\textwidth]{bloomfilter}
-  \caption{An example Dice coefficient similarity calculation of the
-           two first names `peter' and `pete'  encoded in BFs, as
-           described in Sect.~\ref{sec-overview}. The dark bit shows
-           a hash collision.}
-           \label{fig:bloomfilter}
-\end{figure}
-
-\textbf{Bloom Filter Encoding:}~
-Proposed by Bloom~\cite{Blo70} in 1970 for the space and time
-efficient representation of sets, a BF $\mathbf{b}$ is a bit vector
-of length $l$ where all bits are initially set to $0$. $k$
-independent hash functions, $h_1, \ldots, h_k$, each with range
-$1, \ldots, l$, are used to map the elements $s$ in a set
-$\mathbf{s}$ into the BF by setting the bit positions
-$\mathbf{b}[h_j(s)]=1$, with $1 \le j \le k$.
-
-For PPRL, the set $\mathbf{s}$ of q-grams generated from string
-values~\cite{Sch09}, or neighboring values for numerical
-values~\cite{Vat16}, can be hash-mapped into a BF. These BFs are then
-either sent to a linkage unit (LU, an external party that conducts
-the linkage) to calculate the similarity between BFs in order to
-classify them as matches or non-matches~\cite{Sch09}, or they are
-partially exchanged among the database owners to distributively
-calculate the similarities between BFs~\cite{Vat14c}.
-
-%Any set-based similarity function (such as Overlap, Jaccard, and Dice
-%coefficient~\cite{Chr12}) can be used to calculate the similarity of
-%pairs or sets of (multiple) BFs.
-
-The Dice coefficient has been used for comparing BFs since it is
-insensitive to many matching zeros
-%(bit positions to which no elements are hash-mapped)
-in long BFs~\cite{Sch09}. 
-%
-For two BFs, $\mathbf{b}_1$ and $\mathbf{b}_2$, the Dice coefficient
-similarity is: $sim_D(\mathbf{b}_1, \mathbf{b}_2) = 2  c / (x_1 +
-x_2)$,
-%\begin{eqnarray}
-%\label{eq:Dice_coefficient}
-%sim_D(\mathbf{b}_1, \mathbf{b}_2) &=& \frac{2 \times c}{x_1 + x_2},
-%\end{eqnarray}
-where $c$ is the number of bit positions that are set to $1$ in both
-BFs (common $1$-bits), and $x_1$ and $x_2$ are the number of bit
-positions set to $1$ in $\mathbf{b}_1$ and $\mathbf{b}_2$,
-respectively.
-%
-Figure~\ref{fig:bloomfilter} shows the encoding of bigrams ($q=2$)
-of two string values into $l=14$ bits long BFs using $k=2$ hash
-functions, and their Dice coefficient similarity calculation.
-
-Different encoding methods have been proposed for BFs. Hashing
-several attributes of a record into one BF is a method known as
-cryptographic long term key (CLK). It is used to improve
-privacy~\cite{Sch11}. Another record-level BF encoding (RBF) was
-proposed to improve linkage quality~\cite{Dur14}. In RBF,
-attribute values are first hashed into different BFs and then bits
-are selected from each attribute-level BF into a RBF according to
-attribute weights.
-
-The initial proposal of BFs for PPRL used a double hashing
-scheme~\cite{Sch09}, where the $k$ individual bit positions for an
-element $s$ to be hashed are determined by the sum of the integer
-representation of two independent hash functions that are mapped into
-the range $1-l$. Random hashing has recently been proposed as an
-improvement over double hashing to prevent against cryptanalysis
-attacks, where $k$ random numbers are drawn for every element $s$ to
-be hashed~\cite{Nie14,Sch16}.
+% subsection
 
@@
 
-\smallskip
-
-%\subsection{Bloom Filter Hardening}
-%\label{sec-bf-harden}
-
-\textbf{Bloom Filter Hardening:}~
-Several BF hardening methods have been studied in recent times to
-reduce the vulnerability of BFs against cryptanalysis
-attacks~\cite{Sch16}. Compared to attribute-level BFs, record-level
-BF encodings such as CLK and RBF reduce the risk of re-identification
-by such attacks~\cite{Kuz11,Nie14}.
-
-Exploiting the fact that data sets with (near) uniform Hamming
-weight distribution of BFs are more difficult to attack (with
-existing attack methods) than data sets with non-uniform
-distributions, balancing BFs with constant Hamming weight has been
-proposed~\cite{Sch16}. Balanced BFs can be constructed by
-concatenating a BF of length $l$ with its negated copy (all bits
-flipped) and then permuting the $2l$ bits. Another proposed approach
-to harden BFs is XOR-folding, where a BF of length $l$ is split into
-two halves of length $l/2$ each, and then bit-wise exclusive OR is
-applied to combine the two shorter BFs~\cite{Sch16}.
-
-While balancing and XOR-folding are easy and data independent
-hardening techniques, salting with record-specific values has been
-suggested as an alternative hardening method where an additional
-(record-specific) value is concatenated with attribute values
-before being hashed into the BF~\cite{Nie14}. A cryptanalysis attack
-is unlikely to be successful without knowing the salting key, however
-attributes suitable for salting might not be available in a database.
-Other, more experimental hardening techniques, include random bits
-and fake record injection~\cite{Dur14,Vat13}, as well as BLIP
-(BLoom-and-flIP) which flips bits (noise addition) in a BF according 
-to a differential privacy model~\cite{Sch16}. Many of these
-hardening techniques improve security against attacks at the cost
-of a reduction in linkage quality~\cite{Sch16}. In the experiments
-in Sect.~\ref{sec-data} we will investigate if balancing and
-XOR-folding make BFs more resistant to our proposed attack.
+%\subsection
 
 % --------------------------------------------------------------------
-
-\section{Frequency-based Bloom Filter Cryptanalysis}
-\label{sec-bf-attack}
-
-% define problem? or define problem in sect 3?
-
-% below explain what we aim to achieve
-
-% I think we need to explain that our attack does better than simply
-% sorting attribute values and BF by frequencies and aligning
-% them! we do find which q-grams are at which position
-
-We now describe our frequency-based attack on BFs in detail. As shown
-in Fig.~\ref{fig:outline}, the attack consists of two main steps.
-First, for each BF position we find its set of possible and
-\emph{not} possible q-grams
-%$p$, $1 \le p \le l$, we find its set of possible q-grams,
-%$\mathbf{c}[p]$
-(steps (1a) and (1b) in Fig.~\ref{fig:outline}).
-%
-Next, 
-%using the list $\mathbf{C} = [\mathbf{c}[1], \ldots, \mathbf{c}
-%[l]]$,
-we re-identify for each BF  
-%in $\mathbf{B}$
-the set of attribute values
-% $g_i$ from a set of values $\mathbf{G}$
-that possibly were hashed into this BF (step (2) in
-Fig.~\ref{fig:outline}).
-%
-Algorithms~1 and~2 show the details of steps (1a) and (1b), and (2),
-respectively, which we describe in the next two sub-sections. We then
-provide an analysis of our attack and discuss its limitations.
-
-%  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-\subsection{Candidate Q-gram Set Generation}
-\label{sec-cand-q-gram}
-
-For our attack we require as input a set of BFs, $\mathbf{B}$, that
-we assume to come from a sensitive database (the one from which we
-aim to re-identify its sensitive values), and a set of attribute
-values, $\mathbf{V}$, assumed to come from a public database. Each BF
-$\mathbf{b}_i \in \mathbf{B}$ and each attribute value $v_i \in
-\mathbf{V}$ has a frequency attached to it, denoted with
-$\mathbf{b}_i.f$ and $v_i.f$, respectively. Unlike previous attacks on
-BFs~\cite{Kuz11,Kuz13,Nie14}, we do not require any other information
-about how the BFs were encoded.
 
@@
 
-Algorithm~1 starts by initializing two empty sets for each BF
-position $p$: $\mathbf{c}^+[p]$ of possible q-grams at that
-position, and $\mathbf{c}^-[p]$ of \emph{not} possible q-grams at
-that position. Next, in lines 2 and 3, we find all BFs and attribute
-values that occur at least $m$ times. In line 4 we sort both the BFs
-and attribute values according to their frequencies in reverse order
-(most frequent first), and then we align a BF $\mathbf{b}_i$ and an
-attribute value $v_i$ into the sorted list $\mathbf{A}$ of pairs
-$(\mathbf{b}_i, v_i)$. We do this as long as both the $i$th BF
-$\mathbf{b}_i$ and $i$th attribute value $v_i$ have a unique
-frequency compared to the next, less frequent, BF or attribute
-value, respectively. This stopping criterion ensures that we do not
-have a BF that could correspond to two or more attribute values,
-and vice versa, as this would lead to more uncertainty in the
-mapping of q-grams into BF positions.
-
-Step (1a) of our approach starts in line 6 and loops over pairs of
-aligned BFs and attribute values, $(\mathbf{b}_i, v_i) \in
-\mathbf{A}$. First, we convert $v_i$ into its set of q-grams,
-$\mathbf{q}_i$, in line 7. Then we loop over all BF positions,
-$1 \le p \le l$, in line 8, and if the bit at position $p$ in BF is
-$1$  (i.e., $\mathbf{b}_i[p] = 1$), we add the q-gram set
-$\mathbf{q}_i$ to the set $\mathbf{c}^+[p]$ of possible q-grams at
-that position (line 10) because a $1$-bit means any q-gram from
-$\mathbf{q}_i$ could have been hashed to that position. Conversely,
-if the BF bit at position $p$ is $0$, then no q-gram from
-$\mathbf{q}_i$ could have been hashed to that position and so we add
-the q-grams in $\mathbf{q}_i$ to $\mathbf{c}^-[p]$ (line 12).
-
-In step (1b), line 13 onwards in Algo.~1, we get the final set of
-q-grams for each BF position $p$ as the set of possible q-grams
-($\mathbf{c}^+[p]$) minus the set of not possible q-grams
-($\mathbf{c}^-[p]$), and add these sets to the list %of q-gram sets
-$\mathbf{C}$ in line 15.
-
-%  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-\subsection{Attribute Value Re-identification}
-\label{sec-reidentify}
-
-\begin{figure}[t]
-  \begin{center}
-  \begin{scriptsize}
-%    \begin{footnotesize}
-  \begin{tabular}{ll} \hline
-\multicolumn{2}{l}{\textbf{Algorithm 2: \emph{Attribute value
-  re-identification}} -- Step (2) } \\ \hline
-\multicolumn{2}{l}{Input:} \\
-\multicolumn{2}{l}{- $\mathbf{B}$: Set of BFs from the sensitive
-  database (same as in Algo.~1)} \\
-\multicolumn{2}{l}{- $\mathbf{C}$: List of possible q-grams for
-  each BF position (from Algo.~1)} \\
-  \multicolumn{2}{l}{- $\mathbf{G}$: Set of attribute values we aim
-  to re-identify in the set of BFs $\mathbf{B}$} \\
-\noalign{\smallskip}
-\multicolumn{2}{l}{Output:} \\
-\multicolumn{2}{l}{- $\mathbf{R}$: List of possible attribute values
-  re-identified for each BF in $\mathbf{B}$} \\
-\noalign{\smallskip}
-  1:  & $\mathbf{R} = [\,]$ ~~~~~~~~~~~~~~~~~~~~ // Initialize empty
-        list of re-identified attribute values \\
-%  1:  & $\mathbf{Q} = [\,]$ ~~~~ // Initialize empty list of q-gram
-%        sets and attribute values \\
-%  2:  & \textbf{for} $v_i \in \mathbf{G}$ \textbf{do}: ~~~~~~
-%      // Loop
-%      over all attribute values \\
-%  3:  & ~~~ $\mathbf{q}_i = genQGramSet(v_i, q))$ ~ // Convert
-%        attribute value into its q-gram set \\
-%  4:  & ~~~ $\mathbf{Q}.append((\mathbf{q}_i, v_i))$ \\
-  2:  & \textbf{for} $\mathbf{b}_i \in \mathbf{B}$
-        \textbf{do}: ~~~~~~~~~ // Loop over all BFs \\
-  3:  & ~~~ $\mathbf{g}_i = \mathbf{G}$ ~~~~~~~~~~~~~~~ //
-        Initialize set of candidate attribute values as all
-        possible values \\        
-  4:  & ~~~  \textbf{for} $1 \le p \le l$ \textbf{do}: 
-        ~~~~~~~~~~~~~~~~~~~~~~~\, // Loop over all BF positions \\
-  5:  & ~~~ ~~~ \textbf{if} $\mathbf{b}_i[p] == 1$ \textbf{then}:
-        ~~~~~~~~~~~~~~~~ // Bit at position $p$ is 1 \\
-  6:  & ~~~ ~~~ ~~~ $\mathbf{c} = \mathbf{C}[p]$  ~~~~~~~~~~~~~~~~
-        ~~~~~~~~~\, // Set of possible q-grams at this bit position
-        \\
-  7:  & ~~~ ~~~ ~~~ \textbf{for} $v_j \in \mathbf{g}_i$ \textbf{do}: 
-        ~~~~~~~~~~~~~~~~~~~// Check all candidate attribute
-        values \\
-  8:  & ~~~ ~~~ ~~~ ~~~ \textbf{if} $(\forall q \in \textbf{c}:
-        q \notin v_j)$ \textbf{then}: ~ // Check if no q-gram from
-        $\mathbf{c}$ in attribute value \\
-  9:  & ~~~ ~~~ ~~~ ~~~ ~~~ $\mathbf{g}_i = \mathbf{g}_i \setminus
-        v_j$ ~~~~~~~~~~~~~~~\,// Value could not have been hashed
-        into this BF \\ 
-  10: & ~~~ $\mathbf{R}.append(\mathbf{g}_i)$ ~~~~~~~~~~~~~~~~~~~~~
-        ~~~~~~ // Append possible attribute values for this BF \\
-  11: & \textbf{return} $\mathbf{R}$ \\
-  \hline
-  \end{tabular}
-%    \end{footnotesize}
-  \end{scriptsize}
-  \end{center}
-  \label{algo:step2}
-\end{figure}
-
-The re-identification step, shown in Algo.~2, has as input the same
-set of BFs, $\mathbf{B}$, as used in the first step, as well as the
-list of candidate q-grams sets, $\mathbf{C}$, generated in Algo.~1.
-A set of frequent attribute values, $\mathbf{G}$, also needs to be
-provided. These are the values we aim to re-identify (guess) in
-$\mathbf{B}$ using $\mathbf{C}$ (as shown in
-Fig.~\ref{fig:outline}). The output of the algorithm is a set
-of one or more re-identified attribute value(s), $\mathbf{g}_i
-\subset \mathbf{G}$, for each BF $\mathbf{b}_i \in \mathbf{B}$,
-collected in the result list $\mathbf{R}$.
-
-The algorithm loops over all BFs $\mathbf{b}_i$ in $\mathbf{B}$ (line
-2 onwards), and for each $\mathbf{b}_i$ it initializes the set of
-possible candidate attribute values $\mathbf{g}_i$ (that potentially
-have been hashed into this BF) as all values in $\mathbf{G}$ (line 3).
-Next we loop over all BF positions $p$ (line  4), and for any position
-that has a $1$-bit ($\mathbf{b}_i[p] = 1$) we retrieve the set of
-possible q-grams $\mathbf{c} = \mathbf{C}[p]$ at that position
-(line 6).
-
-Using $\mathbf{c}$, we now check for each attribute value $v_j$ in
-$\mathbf{g}_i$ if \emph{at least one of the q-grams} in $\mathbf{c}$
-occurs in that value (lines 7 and 8). If $v_j$ does not contain at
-least one q-gram from $\mathbf{c}$, it could not have generated the
-$1$-bit at position $p$. If this is the case, in line 9 we remove
-$v_j$ from the set of candidates $\mathbf{g}_i$. At the end of this
-loop (line 10), the set $\mathbf{g}_i$ will contain all attribute
-values from $\mathbf{G}$ that possibly have generated the BF
-$\mathbf{b}_i$, and we append this set to the results list
-$\mathbf{R}$. Note that positions with $0$-bits do not help us to
-remove values $v_j \in \mathbf{g}_i$ because any q-gram in a value
-$v_j$ is potentially hashed into other positions.
-% Peter: not sure if this makes sense, we need to think further
-
-%\textbf{\emph{In Algo.~2, we do not consider in a bit position $p$
-%is $0$ (i.e. no else for the if in line 5). Are we missing something
-%here? Would $0$-bits allow us to do more filtering? Can you please
-%all have a think and let me know.. if there is nothing we can do with
-%$0$-bits we should add a sentence why we cannot use them in the
-%description of Algo.~2 below.}}
-
-%% old below, PC 19/10 *****************************************
-
-% Peter: Old and notes below,  maybe we can use somewhere:
-
-%As with earlier attacks on BF-based PPRL approaches~\cite{...}, we
-%assume that BFs are generated attribute-wise rather than record-wise,
-%allowing us to analyze the frequency distribution of BFs as correlated
-%with the frequency distribution of attribute values.
-
-%Many attribute types commonly used in record linkage, such as first
-%and last name, town or city names, postcodes or zipcodes, exhibit
-%highly skewed frequency distributions of their values~\cite{?}. Often
-%these distributions follow a Zipf-like distribution~\cite{}, where a
-%few values occur very frequently (like surname `Smith' and `Miller' in
-%English speaking countries), while the majority of values occur only
-%rarely or once.
-
-%As described in details below, we align the most frequent BFs from a
-%set of BFs with the most frequent attribute values. We assume that
-%an initial analysis of BF frequencies and attribute value frequencies
-%(as shown in Fig.~\ref{xx}) has allowed us to identify which set of
-%BFs corresponds to which attribute.
-
-%Unlike other attack methods (which assume an attack knows certain
-%parameters of the encoding approach used to generate the BFs~\cite{}),
-%our attack does not require any such knowledge, making the attack more
-%applicable and successful in situations where an attack only gains
-%access to a set of BFs.
-
-%Our attack exploits the fundamental properties of BFs: if a certain
-%bit in a BF is not set to 1, then a corresponding element (q-gram in
-%our case) of the set used to generate the BF cannot be in the %corresponding record.
-%
-%Using the alignment of frequent BFs and frequent attribute values,
-%for each bit position of the BF we generate two sets: (1) the set of
-%potential q-grams that \emph{cannot} be mapped to this position based
-%on the q-grams in the frequent attribute values, and (2) the set of
-%potential q-grams that \emph{can} be mapped to this position based on
-%the q-grams in the frequent attribute values.
-
-%- should we include how we find which BF correspond to which 
-%  attributes? \\
-% or is this an assumption? In section experiments we can show
-% frequency plots of attributes and corresponding BF to show this
-% step is easy
-%
-%- should we include the estimation of the number of hash functions k?
-%  it is not too accurate at the moment, and somewhat independent of
-%  our approach.
-
 %  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
@@
 \subsection{Analysis and Limitations}
+\label{sec-analysis}
 
-\label{sec-analysis}
-We now discuss the complexity and limitations of our cryptanalysis
-attack.
-\smallskip
-
-\textbf{Complexity:}
-We assume $N = |\mathbf{B}|$ is the number of BFs coming from the
-sensitive database, $Q$ is the average number of q-grams per value
-in the set of attribute values $\textbf{V}$ from the public
-database, $k$ is the number of hash functions used to hash q-grams
-into BFs, and $l$ is the length of a BF.
-
-In step (1) of our attack, we first find all frequent BFs and
-attribute values that occur at least $m$ times. This requires a
-linear scan through $\mathbf{B}$ and $\textbf{V}$, respectively,
-which has a complexity of $O(N)$ and $O(|\textbf{V}|)$. In line 4 of
-Algo.~1, the sorting function used on values in $\textbf{B}_F$ and
-$\textbf{V}_F$ has a complexity of $O(M \cdot log\,M)$, where $M =
-min(|\textbf{B}_F|, |\textbf{V}_F|)$.
-%
-In step (1a), each attribute value in $\textbf{A}$ is converted into
-its set of q-grams which are then added to the sets $\mathbf{c}^+[p]$
-and $\mathbf{c}^-[p]$ for all BF bit positions $1 \le b \le l$. This
-step has a computational complexity of $O(M \cdot Q \cdot l)$. In
-step (1b), the sets of possible q-grams are added to list
-$\mathbf{C}$ by performing a set difference operation between
-$\mathbf{c}^+[p]$ and $\mathbf{c}^-[p]$. Assuming $\mathbf{Q}$ is
-the set of all unique q-grams generated from all values $v_i \in
-\mathbf{V}$, then on average $t=|\mathbf{Q}|*k/l$ q-grams are hashed
-into each bit position. Therefore each set difference is of $O(t)$,
-and so step (1b) has an overall complexity of $O(t \cdot l)$ for $l$
-bit positions.
-%The memory requirements of our algorithm is of $O(\mathbf{C})$ at
-%the end of step 1.
-
-In step (2) of our attack, Algo.~2 iterates through each BF
-$\mathbf{b}_i \in \mathbf{B}$ to re-identify the possible attribute
-values $\mathbf{g}_i \in \mathbf{G}$ that map to $\mathbf{b}_i$. For
-each bit position $p$ that is set to $1$, we check if any q-gram in
-$\mathbf{C}[p]$ occurs in a value $v_j \in \mathbf{g}_i$. This leads
-to a total complexity of $O(N \cdot |\mathbf{G}| \cdot l \cdot t)$,
-assuming the average size of a $\mathbf{C}[p]$ is $t$. Finally, the
-worst case space complexity of the list $\mathbf{R}$ returned by
-Algo.~2 is $O(N \cdot |\mathbf{G}|)$ if every value in $\mathbf{G}$
-is mapped to each $\mathbf{b}_i \in \mathbf{B}$.
-
-%the set of possible attribute values $g_i$ are identified from
-%$\mathbf{G}$ if any of their q-grams are available in the set of 
-%candidate q-grams $C[p]$. Thus, step 2 of our attacks has a
-%computational complexity of $O(N\cdot|G|\cdot l)$. Finally, a list
-%$R$ of possible attribute values for $B$ is returned from Algorithm
-%$2$ which requires a worst case space complexity of $O(G\cdot B)$ if
-%all attribute values of $G$ are mapped to each $b\in B$.
-\smallskip
-
-%What are the limitations?
-\textbf{Limitations:} Our frequency-based cryptanalysis attack on
-BFs depends on several assumptions. First, we assume the attacker
-has access to a publicly available population database from where
-the set $\mathbf{V}$ of attribute values and their frequencies can
-be extracted. We also assume that $\mathbf{B}$ does contain a sub-set
-of BFs that occur several times, and that the frequency distribution
-of BFs in $\mathbf{B}$ is similar to the frequency
-%can be aligned by an attacker to the
-distribution of a single or a sub-set of attribute values in
-$\mathbf{V}$. Without such frequency information, that can be
-aligned, our attack (like previous cryptanalysis attacks on
-BFs~\cite{Kro15,Kuz11,Kuz13,Nie14}) would not be possible.
-
-%$G$ which is a superset of the set of attribute values which are
-%encoded in BFs in $B$. We also assume based on the frequency
-%distribution of BFs, the attacker can identify the list of
-%attributes that are encoded in $B$ which he used to extract the
-%sets of frequent values from $G$.
+maybe a complexity analysis? maybe an analysis of expected 
+linkage quality?
 
@@
 \label{sec-data}
+
+plot coverage as defined above for different data sets and the different blocking techniques
+
+x-axis: similarity threshold from 0 to 1, y-axis is coverage as a
+percentage (percentage of how many record pairs with a certain similarity value are generated by a blocking technique compared to all record pairs with that similarity - how to do this?
+we can't do full pair-wise comparison?
+can we get all pairs from m-tree for a certain distance, then run them through blocking?
+
+For M-tree etc coverage will always be 1.0, while for other blocking techniques the coverage will be < 1.0 especially as similarities
+become lower.
+
+Plots should include average coverages for a certain similarity threshold over various parameter settings - e.g.
+for LHS or sorted neighbourhood, as well as min/max similarities (or std-dev bands in a transparent colour), as in the below example:
+
+%http://www.randalolson.com/2014/06/28/how-to-make-beautiful-data-visualizations-in-python-with-matplotlib/
+
+%% old text below  ***************************************
 
@@
 %  {reident-bf_len-ncvoter_20140619_1}
-%  ~~
+  ~~
 %  \includegraphics[width=0.31\textwidth]
 %  {reident-hash_type-ncvoter_20140619_1}
-%  ~~
+  ~~
 %  \includegraphics[width=0.31\textwidth]
 %  {reident-q-ncvoter_20140619_1}
-  %\\ ~ \\
-  %\includegraphics[width=0.31\textwidth]
-  %{reident-bf_len-ncvoter_20140619_3}
-  %~~
-  %\includegraphics[width=0.31\textwidth]
-  %{reident-hash_type-ncvoter_20140619_3}
-  %~~
-  %\includegraphics[width=0.31\textwidth]
-  %{reident-q-ncvoter_20140619_3}
-  %\\ ~ \\
-  %\includegraphics[width=0.31\textwidth]
-  %{reident-bf_len-ncvoter_20140619_6}
-  %~~
-  %\includegraphics[width=0.31\textwidth]
-  %{reident-hash_type-ncvoter_20140619_6}
-  %~~
-  %\includegraphics[width=0.31\textwidth]
-  %{reident-q-ncvoter_20140619_6}
-%  \\ ~ \\
+  \\ ~ \\
 %\includegraphics[width=0.31\textwidth]
 %  {reident-bf_len-ncvoter_20140619_1-3}
-%  ~~
+  ~~
 %  \includegraphics[width=0.31\textwidth]
 %  {reident-hash_type-ncvoter_20140619_1-3}
-%  ~~
+  ~~
 %  \includegraphics[width=0.31\textwidth]
 %  {reident-q-ncvoter_20140619_1-3}
-%
-% results with no hardening below
-%~ \\~\\~\\
-  \includegraphics[width=0.31\textwidth]
-  {reident-bf_len-ncvoter_20140619_1}
-  ~~
-  \includegraphics[width=0.31\textwidth]
-  {reident-hash_type-ncvoter_20140619_1}
-  ~~
-  \includegraphics[width=0.31\textwidth]
-  {reident-q-ncvoter_20140619_1}
-  \\ ~ \\
-\includegraphics[width=0.31\textwidth]
-  {reident-bf_len-ncvoter_20140619_1-3}
-  ~~
-  \includegraphics[width=0.31\textwidth]
-  {reident-hash_type-ncvoter_20140619_1-3}
-  ~~
-  \includegraphics[width=0.31\textwidth]
-  {reident-q-ncvoter_20140619_1-3}
 \caption{Results for the NCVR data sets with first name (top)
@@
 \end{figure}
-
-The second are a pair of census data sets (named UKCD) collected
-from the years 1851 and 1901 (used as the sensitive and public
-databases, respectively) for the town of Rawtenstall in
-England~\cite{Fu14b}. These data sets contain around 50,000
-records with personal details of individuals. We again use the
-first name attribute, and the concatenation of the first name and
-surname attributes.
-%surname and address attributes.
 
@@
 %most frequent values from the public data set to be guessed.
-
-We evaluate the accuracy of our attack by calculating (1) the
-percentage of correct guesses with 1-to-1 matching, (2) the
-percentage of correct guesses with 1-to-m (many) matching, (3) the
-percentage of wrong guesses, and (4) the percentage of no guesses,
-where these four percentages sum to 100. These four categories are
-labeled as `1-1 corr', `1-m corr', `Wrong', and `No' in the plots,
-respectively. We evaluate the efficiency of our attack using run
-time.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[width=0.31\textwidth]
-    {reident-bf_len-1851_allcleaned_1901_allcleaned_6}
-  ~~
-  \includegraphics[width=0.31\textwidth]
-    {reident-hash_type-1851_allcleaned_1901_allcleaned_6}
-  ~~
-  \includegraphics[width=0.31\textwidth]
-    {reident-bf_harden-1851_allcleaned_1901_allcleaned_6}
-    \\ ~ \\
-  \includegraphics[width=0.31\textwidth]
-    {reident-bf_len-1851_allcleaned_1901_allcleaned_5-6}
-  ~~
-  \includegraphics[width=0.31\textwidth]
-    {reident-hash_type-1851_allcleaned_1901_allcleaned_5-6}
-  ~~
-  \includegraphics[width=0.31\textwidth]
-    {reident-bf_harden-1851_allcleaned_1901_allcleaned_5-6}
-  \caption{Results for the UKCD data sets with first name (top) and
-  combination of first and surname (bottom) for different BF lengths
-  (left), different hashing methods (middle), and different BF
-  hardening methods (right).
-  \label{fig:ukcd}}
-\end{figure}
-
-We implemented our attack using Python 2.7 and ran all experiments on
-a server with 64-bit Intel Xeon 2.4 GHz CPUs, 128 GBytes of memory
-and running Ubuntu 14.04. The programs and data sets are available
-from the authors.
-
-%Toto: generate a similar figure for UKCS? add data set table
-%(num rec, attributes use?, num matches? what do we need?
-%discuss data sets, table with details - UKC, NCVR balanced,
-%NVR 2011 to 2016 (select
-%a few, give number of records, number of values in selected attributes)
-%show plots with freq distributions of 3 selected attribute values
-%(first name, last name, postcode, city?) and their corresponding BF distributions
-%If room, show for different data sets.
-%experimental setup
-%also show if possible for pairs of attributes mapped into 1 BF
-
-\smallskip
 
@@
 values can still be re-identified even when values from two
-attributes are combined.
-%
-The middle column
-%in Fig.~\ref{fig:ncvr}
-shows that random hashing
-(which supposedly improves privacy on BFs compared to double
-hashing~\cite{Kro15,Nie14}) does not provide improved protection
-against our attack, as a similar percentage of values can be
-correctly re-identified for both hashing approaches. As for using
-different values of $q$ (right column)
-%in Fig.~\ref{fig:ncvr}),
-the accuracy of an attack somewhat improves when $q$ is increased
-because larger values of $q$ result in more unique q-grams.
-
-The accuracy results for the UKCD data sets are shown in
-Fig.~\ref{fig:ukcd} for the same attributes as for the NCVR data
-sets. Similar re-identification patterns as with the NCVR data sets
-can be seen for different BF lengths (left) and hashing methods
-(middle). We also study how different BF hardening methods affect the
-re-identification accuracy, as shown in the right column plots in
-Fig.~\ref{fig:ukcd}. For the single attribute case (top right),
-neither of the hardening techniques is capable of reducing the
-re-identification accuracy, however for the combined attribute
-case both hardening techniques improve the privacy of BF encoding.
+attributes
 
@@
 
-% -------------------
-%DV: Peter, for your question on what is the difference between 2 and
-%3 in table 1:
-%In that paper they first tried to re-identify 40 most frequent
-%values ant it took more than a week without any results (0 
-%re-identification percentage), and then they used 2 most frequent
-%names and it results in 20% (i.e. 4 values are re-identified)
-%accuracy in few seconds (they didn't mention the actual number of
-%seconds).
-%
-%That's why I included both entries in the table to show the %scalability of our proposed approach.
-
-% --------------
-
-In Table~\ref{table:exist_approaches} we compare our attack method
-with existing approaches in terms of accuracy and efficiency (with
-results taken from the corresponding papers). As can be seen, our
-method is both more efficient and effective in re-identification,
-having both higher accuracy and reduced computational requirements.
-% than existing approaches.
-
-These results show the vulnerability of basic BF encoding to our
-novel attack. They highlight the need for improved hardening
-techniques to overcome such attacks. Our attack provides data
-custodians with an efficient method to evaluate the privacy of their
-BF encoded databases before using them for PPRL.
-
-\smallskip
-\textbf{Recommendations:}~
-As a set of guidelines for the practical application of BF based PPRL
-systems, to limit the vulnerability of such systems to known attack
-methods we recommend to use record-level BF encoding (CLK or RBF),
-apply advanced BF hardening methods~\cite{Sch16}, and reduce the
-frequency of bit patterns (for example, by salting) to prevent any
-frequency analysis.
-
 % --------------------------------------------------------------------
@@
 \label{sec-concl}
-
-We have presented a novel efficient frequency-based attack on Bloom
-filters (BFs) that contain encoded sensitive attribute values
-intended for privacy-preserving record linkage (PPRL). Unlike earlier
-attacks on BFs for PPRL, our approach only requires an attacker to
-have access to a public database of attribute values, but no
-information about the BF encoding used.
-%Most importantly, it is independent of the kind of hash function
-%used for BF encoding.
-Our approach is faster than earlier attacks, making it feasible for
-database owners to efficiently validate the security of their
-encoded sensitive databases before they are being sent to other
-parties for conducting PPRL. We believe our attack is an important 
-component to making PPRL more secure for practical applications.
-
-As future work we will study how our approach can be modified for
-attacking composite and record-level BFs. We also plan to investigate
-the risk of re-identification when advanced hardening techniques, such
-as BLIP or BF salting, have been applied. Finally, our attack can be
-accelerated by further analyzing the re-identified attribute values,
-while correlations between $1$-bits and q-gram sets across BFs could
-be identified using association rule mining
-techniques~\cite{Ceg06}.
-
-% --------------------------------------------------------------------
-
-%\section*{Acknowledgments}
-%
-%Peter Christen was supported by a grant from the Simons Foundation.
-%The authors would also like to thank the Isaac Newton Institute for
-%Mathematical Sciences, Cambridge, for support and hospitality during
-%the programme \emph{Data Linkage and Anonymisation} where this work
-%was conducted (EPSRC grant EP/K032208/1). This work was partially
-%funded by the Australian Research Council under Discovery Project
-%DP130101801.
 

### 2017-09-26 src/common.py
Doing a full editdistance computation on CORA

@@ -10,8 +10,8 @@ import sys
 import time
 
 # febrl
-# sys.path.append(os.path.join(sys.path[0],'..','..','tools','febrl','febrl-0.4.2'))
-# import stringcmp
+sys.path.append(os.path.expanduser("~/repos/github/stacs-srg/linkage-py/tools/febrl/febrl-0.4.2"))
+import stringcmp
 
 
 
@@ -157,12 +157,14 @@ def loadCSV(filepath, delimiter=",", verbose=0):
     Returns
         fieldNames : [String]                   -- headings
         values     : [Dict String String]       -- values, each entry is a dictionary with fieldNames as the keys
+        valuesArray: [String]
     """
     with open(filepath, newline='\n', encoding="latin-1") as f:
         reader = csv.reader(f, delimiter=delimiter)
         fieldNames = [ x.strip() for x in next(reader) ]
         lenFieldNames = len(fieldNames)
         rawValues = list()
+        rawValuesArray = list()
         for row in reader:
             lenValues = len(row)
             if lenFieldNames != lenValues:
@@ -179,13 +181,14 @@ def loadCSV(filepath, delimiter=",", verbose=0):
             # do not insert into the flatTable if all entries are None
             if not allNone:
                 rawValues.append(row2)
+                rawValuesArray.append(row)
         if verbose >= 1:
             print("Loaded data from %s, %d fields, %s entries." % (filepath, len(fieldNames), len(rawValues)))
         if verbose >= 2:
             print("Following are the fields.")
             for fieldName in fieldNames:
                 print(" - %s" % fieldName)
-        return (fieldNames, rawValues)
+        return (fieldNames, rawValues, rawValuesArray)
 
 
 

### 2017-09-26 src/dataset-CORA.py

@@ -1,0 +1,26 @@
+#!/usr/bin/env python3
+
+import common
+import statistics
+
+(_, _, cora_lines) = common.loadCSV("data/cora.csv")
+
+for (i,l1) in enumerate(cora_lines):
+    for (j,l2) in enumerate(cora_lines):
+        if i != j:
+            similarities = [ common.mkStringComparison("editdistance", a, b)
+                             for (a,b) in zip(l1, l2)
+                           ]
+            mean = statistics.mean(similarities)
+            similarities_nonzero = list(filter(lambda x: x > 0, similarities))
+            mean_nonzero = statistics.mean(similarities_nonzero)
+            if mean_nonzero >= 0.5:
+                print(i,j)
+                print(l1)
+                print(l2)
+                print(mean, similarities)
+                print(mean_nonzero, similarities_nonzero)
+                common.tick("")
+                print()
+
+common.tick("Done!")

### 2017-10-30 paper.tex
Update on Overleaf.

@@
 
-\title{Efficient and accurate record linkage using similarity spaces}
-%  \thanks{The authors would like to thank the Isaac Newton Institute
-%  for Mathematical Sciences, Cambridge, for support and hospitality
-%  during the programme \emph{Data Linkage and Anonymisation} where
-%  this work was conducted (EPSRC grant EP/K032208/1). Peter Christen
-%  was also supported by a grant from the Simons Foundation. The work
-%  was also partially funded by the Australian Research Council
-%  (DP130101801).}}
+\title{Exploring approaches ro probablistic record linkage using similarity searching}
+%  \thanks{The authors would like to thank the ...|
 
 \author{Submitted for double-blind review}
-%\author{Peter Christen\inst{1} \and Rainer Schnell\inst{2}
-%        \and Dinusha Vatsalan\inst{1} \and Thilina
-%        Ranbaduge\inst{1}}
+%\author{Özgür Akgün\inst{1} and Alan Dearle\inst{1} and Graham Kirby\inst{1} Peter Christen\inst{2}
+
+%\institute{School of Computer Science,
+%           University of St Andrews, \\
+%           St Andrews, Scotland.~
+%           Contact: \email{ozgur.akgun@st-andrews.ac.uk}
+
 %\institute{Research School of Computer Science,
@@
 %           Contact: \email{peter.christen@anu.edu.au}
-%           \and
-%           Methodology Research Group, %Institute for Sociology,
-%           University Duisburg-Essen, Duisburg, Germany.} %~
-%%           \email{rainer.schnell@uni-due.de}}
 
@@
 
-\keywords Entity resolution; deduplication; metric spaces; etc.
+\keywords Probablistic record Linkage; similarity search; metric spaces; etc.
 
@@
 
-We tried to develop a plan with PeterC. Here is the photo of the white-board.
-
-\includegraphics[width=\textwidth]{2017-08-01_12_08_43.jpg}
-
 \vspace{5mm}
 
-But, even better, here it is in a bulleted list.
+Plan in a bulleted list:
 
@@
 
+Note: The rest of this paper should be regarded as a placeholder for now.
+
 \section{Introduction}
@@
 
+The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977.
+This data set will, when complete, include around 14 million birth records, 11 million death records and 4 million marriage records.
+As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.
+
 % --------------------------------------------------------------------
@@
 
-\begin{figure}[t]
-  \centering
-%  \includegraphics[width=1.0\textwidth]{xyz.eps}
-  \caption{Descriptive caption
-           %as outlined in Sect.~\ref{sec-overview} and detailed in
-           %Sect.~\ref{sec-bf-attack}.
-           \label{fig:xyz}}
-\end{figure}
+\section{Datasets}
+\label{sec-data}
+
+We conducted our experimental study using four real data sets from three
+domains.
+The first is the CORA data set  (\url{https://linqs-data.soe.ucsc.edu/public/lbc/cora.tgz}) \textbf{ which consists of 2708 scientific publications classified into one of seven classes.
+The citation network consists of 5429 links.}
+
+%Fields and true links here
+
+The second are a pair of sets of records drawn from the North Carolina Voter Registration (NCVR) data
+sets (\url{ftp://alt.ncsbe.gov/data/}) collected in June 2014 and October 2016
+ \textbf{( who collected? where is the dataset published?)}.
+These data sets contain over five million records of voters encoded into 11 fields including
+their first names, surnames, and addresses. Ground truth is provided via an \textit{voter\_id} field which is invariant across the data sets.
+In our experiments we use two subsets of this data comprising 224074 and 224062 records drawn from the original five million plus records. \textbf{chosen how?}
+
+We also experiment over two datasets prepared by researchers pursuing previous projects \cite{reid2002} and \cite{reid2006}, which act as pseudo subsets of the Digitising Scotland data set.
+One data set contains records of vital events registered on the Isle of Skye, a rural district, while the other contains records from Kilmarnock, an industrial town.
+Both cover the period 1861-1901.
+These are different types of communities, with different family structures and name distributions \cite{reid2002}.
+Both datasets are encoded with the same schemata and include names, sex, addresses and the names of each person's father and mother.
+Ground truth is provided using demographer encoded identifiers which provide (incomplete) linkage between the records.
+The Isle of Skye dataset comprises around 17,600 birth records, 12,300 death records, and 2,700 marriage records.
+The demographers  identified:
+\begin{itemize}
+\item 4,300 family groups of siblings, with a mean family size of 3.9 siblings and a maximum family size of 16,
+\item 2,900 links from a birth record to the child's death record,
+\item 60 links from a marriage record to the groom's death record, and
+\item 100 links from a marriage record to the bride's death record.
+\end{itemize}
+The Kilmarnock data set contains around 38,400 birth records, 23,700 death records, and 8,700 marriage records. Here, the demographers identified:
+\begin{itemize}
+\item 13,100 family groups of siblings, with a mean family size of 2.8 siblings and a  maximum family size of 16,
+\item 8,300 links from a birth record to the child's death record,
+\item 30 links from a marriage record to the groom's death record, and
+\item 50 links from a marriage record to the bride's death record.
+\end{itemize}
+
+% --------------------------------------------------------------------
 
@@
 
-\begin{figure}[t]
-  \begin{center}
-  \begin{scriptsize}
-%    \begin{footnotesize}
-  \begin{tabular}{ll} \hline
-\multicolumn{2}{l}{\textbf{Algorithm 1: \emph{Candidate q-gram set
-  generation}} -- Steps (1a) and (1b) } \\ \hline
-\multicolumn{2}{l}{Input:} \\
-\multicolumn{2}{l}{- $\mathbf{V}$: Set of attribute values and their
-  frequencies from a public database} \\
-\multicolumn{2}{l}{- $\mathbf{B}$: Set of BFs and their
-  frequencies from the sensitive database} \\
-\multicolumn{2}{l}{- $l$:~\, Length of Bloom filters} \\
-\multicolumn{2}{l}{- $q$:~ Length of sub-strings to extract from
-  attribute values} \\  
-\multicolumn{2}{l}{- $m$: Minimum frequency for BFs and attribute
-  values} \\
-\noalign{\smallskip}
-\multicolumn{2}{l}{Output:} \\
-\multicolumn{2}{l}{- $\mathbf{C}$: List of possible q-grams for each
-  BF position} \\
-\noalign{\smallskip}
-  1:  & $\mathbf{c}^+[p] = \{\}$, $\mathbf{c}^-[p] = \{\}$,
-        $1 \le p \le l$ ~~~ // Initialize list of candidate q-gram
-        sets \\
-  2:  & $\mathbf{V}_F = \{v \in \mathbf{V}: v.f \ge m \}$
-        ~~~~~~~~~~~~~~ // Attribute values with frequency of at
-        least $m$ \\
-  3:  & $\mathbf{B}_F = \{\mathbf{b} \in \mathbf{B}:
-        \mathbf{b}.f \ge m \}$
-        ~~~~~~~~~~~~~~\,// BFs with frequency of at least $m$ \\
-  4:  & $revSort(\mathbf{B}_F)$, $revSort(\mathbf{V}_F)$ ~~~~~~~~~~ 
-        // Sort according to frequencies, highest first \\
-  5:  & $\mathbf{A} = [(\mathbf{b}_i, v_i) : 
-        \mathbf{b}_i \in \mathbf{B}_F, v_i \in \mathbf{V}_F:
-        \mathbf{b}_i.f > \mathbf{b}_j.f \land v_i.f > v_j.f:
-        1 \le i < j \le min(|\mathbf{V}_F |, |\mathbf{B}_F|)]$ \\
-  ~  &   ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ // Align $\mathbf{V}_F$
-        and $\mathbf{B}_F$ as long as their frequencies are
-        unique \\  
-  6:  & \textbf{for} $(\mathbf{b}_i, v_i) \in \mathbf{A}$
-        \textbf{do}: ~~~~~~~~~~~~~~~~~~~~ //  Step (1a): Get
-        candidate sets of q-grams \\
-  7:  & ~~~ $\mathbf{q}_i = genQGramSet(v_i, q)$ ~~~~~~~~\,//
-        Convert attribute value into its q-gram set \\
-  8:  & ~~~  \textbf{for} $1 \le p \le l$ \textbf{do}: 
-        ~~~~~~~~~~~~~~~~~~~\, // Loop over all BF positions \\
-  9:  & ~~~ ~~~ \textbf{if} $\mathbf{b}_i[p] == 1$ \textbf{then}:
-        ~~~~~~~~~~~~\, // Bit at position p is 1 \\
-  10: & ~~~ ~~~ ~~~ $\mathbf{c}^+[p] = \mathbf{c}^+[p] \cup
-        \mathbf{q}_i$ ~~~~~~~~~~~\,// Add to set of possible
-        q-grams \\
-  11: & ~~~ ~~~ \textbf{else}: ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ //
-        Bit at position $p$ is $0$ \\
-  12: & ~~~ ~~~ ~~~ $\mathbf{c}^-[p] = \mathbf{c}^-[p] \cup
-        \mathbf{q}_i$  ~~~~~~~~~~~// Add to set of \emph{not}
-        possible q-grams \\
-  13: & $\mathbf{C} = [\,]$ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~\,
-        // Step (1b): Initialize empty list of q-gram sets \\
-  14: & \textbf{for} $1 \le p \le l$ \textbf{do}: 
-        ~~~~~~~~~~~~~~~~~~~~~~~~\,// Loop over all BF positions to
-        combine q-gram sets \\
-  15: & ~~~ $\mathbf{C}.append(\mathbf{c}^+[p] \setminus
-        \mathbf{c}^-[p])$ ~~~~~~~~~~~\,// Remove not possible from
-        possible q-grams \\
-%  15: & ~~~ $\mathbf{c}[p] = \mathbf{c}^+[p] \setminus
-%        \mathbf{c}^-[p]$ ~~~~~~~~~~~~~~~~~ // Remove not possible
-%        from possible q-grams \\
-%  16: & ~~~ $\mathbf{C}.append(\mathbf{c}[p])$
-%        ~~~~~~~~~~~~~~~~~~~~~~ // Add to list of candidates for all
-%        BF positions\\
-16: & \textbf{return} $\mathbf{C}$ \\
-  \hline
-  \end{tabular}
-%  \end{footnotesize}
- \end{scriptsize}
-  \end{center}
-  \label{algo:step1}
-\end{figure}
-
 %  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
@@
 
-plot coverage as defined above for different data sets and the different blocking techniques
-
-x-axis: similarity threshold from 0 to 1, y-axis is coverage as a
-percentage (percentage of how many record pairs with a certain similarity value are generated by a blocking technique compared to all record pairs with that similarity - how to do this?
-we can't do full pair-wise comparison?
-can we get all pairs from m-tree for a certain distance, then run them through blocking?
-
-For M-tree etc coverage will always be 1.0, while for other blocking techniques the coverage will be < 1.0 especially as similarities
-become lower.
-
-Plots should include average coverages for a certain similarity threshold over various parameter settings - e.g.
-for LHS or sorted neighbourhood, as well as min/max similarities (or std-dev bands in a transparent colour), as in the below example:
-
-%http://www.randalolson.com/2014/06/28/how-to-make-beautiful-data-visualizations-in-python-with-matplotlib/
-
-%% old text below  ***************************************
-
-We conducted our experimental study using real data sets from two
-domains.
-%
-The first are a pair of North Carolina Voter Registration (NCVR) data
-sets (\url{ftp://alt.ncsbe.gov/data/})
-%~\footnote{Available from \url{ftp://alt.ncsbe.gov/data/}}
-collected in June 2014 (the database to be attacked) and October 2016
-(the public database).
-%, that have increasingly large differences in their time collection
-%to see how this influences re-identification accuracy.
-These data sets contain over five million records of voters including
-their first names, surnames, and addresses. We present results of our
-attack individually on the first name
-%, last name, and address attributes, 
-attribute, as well as on the concatenation of the first name and
-surname attributes.
-
-\begin{figure}[!t]
-  \centering
-%  \includegraphics[width=0.31\textwidth]
-%  {reident-bf_len-ncvoter_20140619_1}
-  ~~
-%  \includegraphics[width=0.31\textwidth]
-%  {reident-hash_type-ncvoter_20140619_1}
-  ~~
-%  \includegraphics[width=0.31\textwidth]
-%  {reident-q-ncvoter_20140619_1}
-  \\ ~ \\
-%\includegraphics[width=0.31\textwidth]
-%  {reident-bf_len-ncvoter_20140619_1-3}
-  ~~
-%  \includegraphics[width=0.31\textwidth]
-%  {reident-hash_type-ncvoter_20140619_1-3}
-  ~~
-%  \includegraphics[width=0.31\textwidth]
-%  {reident-q-ncvoter_20140619_1-3}
-\caption{Results for the NCVR data sets with first name (top)
-    and combination of first and surname (bottom) for different BF
-    lengths (left), different hashing methods (middle), and different
-    values of $q$ (right). No BF hardening was applied.
-   \label{fig:ncvr}}
-\end{figure}
-
-The parameter settings we use in our experiments are $q = [2,3,4]$
-(length of sub-strings used in BF encoding), BF length
-$l = [250, 500, 1000,$ $2000]$, and either the double or random
-hashing method~\cite{Sch16} as described in
-Sect.~\ref{sec-overview}. We calculate the number of hash functions
-$k$ based on $l$ and $q$ such that the false positive rate is
-minimized~\cite{Vat14c}. We also apply the BF hardening techniques
-balancing and XOR-folding~\cite{Sch16}.
-% as described in Sect.~\ref{sec-overview}.
-We use different numbers for the most frequent attribute values
-(ranked by their frequencies) to be re-identified: $|\mathbf{G}| =
-[10,20,50,100]$.
-%most frequent values from the public data set to be guessed.
-
-\textbf{Discussion:}~
-In Fig.~\ref{fig:ncvr} we show the results for the NCVR data sets.
-% for one individual attribute (first name) in the top row, and
-%three individual attributes (first name, last name, and address in
-%the first three rows, respectively), and
-%combined attributes of first and surnames in the bottom row.
-% BFs are hardened using the balanced method.
-%
-As can be seen, when the BF length $l$ increases (left column) the
-percentage of correct re-identifications mostly increases, as with
-larger $l$ the number of q-grams mapped to a certain bit position
-decreases. All values can be correctly re-identified for individual
-attributes
-%nearly $90\%$ of values can be correctly re-identified
-%%as 1-to-1 or 1-to-many matches
-when the number of frequent values is 10.
-%or 20. 
-Around half of all
-values can still be re-identified even when values from two
-attributes
-
-\begin{table}[!t]
- %\addtolength{\tabcolsep}{-3pt}
-  \caption{Comparison of re-identification results with existing BF
-     attack methods.}
-     \label{table:exist_approaches}
-  \centering
-  \begin{scriptsize}
-    \begin{tabular}{llccc}
-    \hline\noalign{\smallskip}
-   Publication & Data set & Num BFs~ & ~1-1 corr~ & Run time
-     \\ \noalign{\smallskip}\hline\noalign{\smallskip}
-  Kuzu et al. (2011)~\cite{Kuz11} & NCVR first names & 3,500 & 400 &
-    1,000 sec \\
-  Kuzu et al. (2013)~\cite{Kuz13} & Patient names & 20 & 4 & 
-    few sec \\
-  ~~~~ " & ~~~~ " & 42 & 0 & $>$ week \\
-  Niedermeyer et al. (2014)~\cite{Nie14}~ & German surnames & 7,580 &
-    934 & $>$ days  \\
-  Kroll and Steinmetzer (2015)~\cite{Kro15}~ & German names and
-    locations & 100K & 44K & $>$ days \\
-  Our approach & NCVR first names & 10--100 & 7--10 & 0.73--0.75 sec
-    \\
-  ~~~~ " & NCVR first and surnames & 10--100 & 3--6 & 1.5--1.9
-  sec  \\
-%  ~~~~ " & UKCD surname and address~ & 10 & 10 & 0.05 sec \\
-  \noalign{\smallskip} \hline
-  \end{tabular}
-  \end{scriptsize}
-\end{table}
-
 % --------------------------------------------------------------------
@@
 
-%\bibliographystyle{abbrv}
-\bibliographystyle{splncs03}
-\bibliography{paper}
+\bibliographystyle{abbrv}
+\bibliography{paper.bib}
 

### 2017-10-30 paper.tex
Update on Overleaf.

@@
 
-\section{approach section?}
-\label{sec-overview}
+\section{Approach}
+\label{sec-approach}
+
+In this paper we explore the efficacy of the use of different similarity search algorithms in probabilistic record linkage.
+We explore two inexact algorithms: LSH and MIFile \cite{amato2014mi} one exact method: mTrees \cite{paolociaccia2m}.
+We also use a simple brute force exact technique as a baseline for comparison although this approach can only be applied to the smallest of our datasets.
 

### 2017-10-31 paper.tex
Update on Overleaf.

@@
 
-\title{Exploring approaches ro probablistic record linkage using similarity searching}
+% Peter 31 Oct: Current title does not really reflect the topic I
+% think:
+%\title{Exploring approaches to probablistic record linkage using
+%        similarity searching}
+% Peter 31 Oct: What about
+\title{Efficient and Effective Metric Space Indexing for Complete
+       Record Linkage
+       Al: Exploring approaches to probabilistic record linkage using similarity searching (metric space indexing?) }
 %  \thanks{The authors would like to thank the ...|
@@
 
+% MAX 200 words
 \begin{abstract}
-% usually about 200 words max- best to check
+Probabilistic record linkage is the process of identifying records that refer to
+the same real-world entities across two or more databases in situations where entity identifiers are unavailable.
+It requires  similarities between records to be approximated using common attributes (such as names, addresses, phone numbers) and for pairs of records to be classified as matches or non-matches.
+Record linkage is usually done as a two step process;  in the
+first step a blocking or indexing is employed to efficiently identify
+candidate record pairs which are then compared in more detail and
+classified in the second step.
+
+As we show in this paper, even state-of-the-art blocking and
+indexing techniques such as locality sensitive hashing, have two
+major drawbacks. First, they often remove some true matching record
+pairs that have high similarities due to their heuristic nature,
+leading to a loss of recall. Second, they include many record pairs
+with low similarity leading to high computational requirements.
+Combined, this means a reduction of both effectiveness and efficiency.
+% * <al@st-andrews.ac.uk> 2017-10-31T08:46:25.069Z:
+%
+% > efficiency
+%
+% ^.
+
+We propose an metric indexing approach to \emph{complete} record
+linkage, which ensures all record pairs that have a similarity above
+a certain similarity threshold are being \textbf{compared}, while no pairs
+with a similarity below the threshold are \textbf{compared}. Our approach
+leads to a one-step record linkage process which combines indexing
+with comparison and classification resulting in improved
+effectiveness and efficiency as we experimentally evaluate on several
+real-world databases.
 
@@
 
-\keywords Probablistic record Linkage; similarity search; metric spaces; etc.
+\keywords Entity resolution; data matching; similarity search;
+         blocking.
 
@@
 \item LSH-blocking
-\item Traditional blocking (e.g. Region, Surname)
-\item Sorted Neighbourhood blocking
+\item Traditional blocking (e.g. Region, Surname) - \textbf{why this? - this was done in the paper: A Comparison of Blocking Methods for Record Linkage}
+\item MIFile
+\item Sorted Neighbourhood blocking - \textbf{why this?}
 \end{itemize}
@@
 
-\section{Datasets}
-\label{sec-data}
-
-We conducted our experimental study using four real data sets from three
-domains.
-The first is the CORA data set  (\url{https://linqs-data.soe.ucsc.edu/public/lbc/cora.tgz}) \textbf{ which consists of 2708 scientific publications classified into one of seven classes.
-The citation network consists of 5429 links.}
-
-%Fields and true links here
-
-The second are a pair of sets of records drawn from the North Carolina Voter Registration (NCVR) data
-sets (\url{ftp://alt.ncsbe.gov/data/}) collected in June 2014 and October 2016
- \textbf{( who collected? where is the dataset published?)}.
-These data sets contain over five million records of voters encoded into 11 fields including
-their first names, surnames, and addresses. Ground truth is provided via an \textit{voter\_id} field which is invariant across the data sets.
-In our experiments we use two subsets of this data comprising 224074 and 224062 records drawn from the original five million plus records. \textbf{chosen how?}
-
-We also experiment over two datasets prepared by researchers pursuing previous projects \cite{reid2002} and \cite{reid2006}, which act as pseudo subsets of the Digitising Scotland data set.
-One data set contains records of vital events registered on the Isle of Skye, a rural district, while the other contains records from Kilmarnock, an industrial town.
-Both cover the period 1861-1901.
-These are different types of communities, with different family structures and name distributions \cite{reid2002}.
-Both datasets are encoded with the same schemata and include names, sex, addresses and the names of each person's father and mother.
-Ground truth is provided using demographer encoded identifiers which provide (incomplete) linkage between the records.
-The Isle of Skye dataset comprises around 17,600 birth records, 12,300 death records, and 2,700 marriage records.
-The demographers  identified:
-\begin{itemize}
-\item 4,300 family groups of siblings, with a mean family size of 3.9 siblings and a maximum family size of 16,
-\item 2,900 links from a birth record to the child's death record,
-\item 60 links from a marriage record to the groom's death record, and
-\item 100 links from a marriage record to the bride's death record.
-\end{itemize}
-The Kilmarnock data set contains around 38,400 birth records, 23,700 death records, and 8,700 marriage records. Here, the demographers identified:
-\begin{itemize}
-\item 13,100 family groups of siblings, with a mean family size of 2.8 siblings and a  maximum family size of 16,
-\item 8,300 links from a birth record to the child's death record,
-\item 30 links from a marriage record to the groom's death record, and
-\item 50 links from a marriage record to the bride's death record.
-\end{itemize}
-
-% --------------------------------------------------------------------
-
 \section{Approach}
@@
 
+We now describe the data sets we use in our evaluation and the
+experimental setup we employed to evaluate our proposed metric
+indexing approach.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+% Peter 31 Oct: table style follows LNCS style guide
+
+\begin{table}[t]
+\caption{Characteristics of data sets used in experiments.
+  \emph{Peter, 31 Oct: what else should we show here?}}
+ \label{table-datasets}
+  \centering
+  \begin{scriptsize}
+  %\addtolength{\tabcolsep}{-0.5pt}
+  \begin{tabular}{ccccc}
+  \hline\noalign{\smallskip}
+  Data set~ & ~Number of~ & ~Number of true~
+   & Attributes used? & ~?? \\
+  name(s)  & records   & matching pairs &
+    for indexing & ?? \\
+  \noalign{\smallskip}
+  \hline
+  \noalign{\smallskip}
+  NCVR  & ~224,073 / 224,061~ & ~148,036~ & ? & ? \\
+  CORA  & 1,295             & title, authors, venue, year ? & ?
+    & ? \\
+  Isle of Skye & ? & ?
+    & ? & ? \\
+  Kilmarnock  & ? & ? & ? & ? \\
+  \hline
+  \end{tabular}
+  \end{scriptsize}
+\end{table}
+
+% Peter, 31 Oct: To save space, don't use sub-sections but simply
+% \textbf{}, such as:
+% \smallskip
+%\textbf{Data sets}:~
+
+\subsection{Data Sets}
+\label{sec-data}
+
+We used four real data sets from three domains in our experiments. The
+first is the CORA data set~\footnote{Available from:
+\texttt{http://secondstring.sourceforge.net}}, which contains 1,295
+(?) records that refer to XX machine learning publications.
+% Peter, 31 Oct: not sure about the following - which Cora version do
+% you use? ****
+ (\url{https://linqs-data.soe.ucsc.edu/public/lbc/cora.tgz}) \textbf{ which consists of 2708 scientific publications classified into one of seven classes. The citation network consists of 5429 links.}
+
+The second are a pair of data sets based on the North Carolina Voter
+Registration (NCVR) data sets\footnote{Available from: \texttt{http://dl.ncsbe.gov/}}, which we collected in June 2014 and
+October 2016. Each of these two data sets contain over five million
+records of voters and include attributes with their first names,
+surnames, and addresses. Ground truth is provided via a \emph{NCID}
+identifier which uniquely identifies a voter. In our experiments we
+use two randomly selected sub-set of $224,073$ and $224,061$ records,
+respectively, where $148,036$ of those refer to voters that occurred
+in both original NCVR data sets but had name and/or address changes
+over time (thus leading to around $66\%$ matching records).
+
+\emph{Peter, 31 Oct: We need to be careful not to reveal our identities
+as the submission is double-blind - so no mention of DS.}
+
+% Peter, 31 Oct: can the following be shortened to 1 paragraph?
+
+We also experiment over two datasets prepared by researchers pursuing previous projects \cite{reid2002} and \cite{reid2006}, which act as pseudo subsets of the Digitising Scotland data set.
+One data set contains records of vital events registered on the Isle of Skye, a rural district, while the other contains records from Kilmarnock, an industrial town.
+Both cover the period 1861-1901.
+These are different types of communities, with different family structures and name distributions \cite{reid2002}.
+Both datasets are encoded with the same schemata and include names, sex, addresses and the names of each person's father and mother.
+Ground truth is provided using demographer encoded identifiers which provide (incomplete) linkage between the records.
+The Isle of Skye dataset comprises around 17,600 birth records, 12,300 death records, and 2,700 marriage records.
+The demographers  identified:
+\begin{itemize}
+\item 4,300 family groups of siblings, with a mean family size of 3.9 siblings and a maximum family size of 16,
+\item 2,900 links from a birth record to the child's death record,
+\item 60 links from a marriage record to the groom's death record, and
+\item 100 links from a marriage record to the bride's death record.
+\end{itemize}
+\emph{Peter 31 Oct: what are you actually using in your experiments?
+All certificates or only birth?}
+
+The Kilmarnock data set contains around 38,400 birth records, 23,700 death records, and 8,700 marriage records. Here, the demographers identified:
+\begin{itemize}
+\item 13,100 family groups of siblings, with a mean family size of 2.8 siblings and a  maximum family size of 16,
+\item 8,300 links from a birth record to the child's death record,
+\item 30 links from a marriage record to the groom's death record, and
+\item 50 links from a marriage record to the bride's death record.
+\end{itemize}
+
 % --------------------------------------------------------------------

### 2017-11-06 paper.tex
Update on Overleaf.

@@
 It requires  similarities between records to be approximated using common attributes (such as names, addresses, phone numbers) and for pairs of records to be classified as matches or non-matches.
-Record linkage is usually done as a two step process;  in the
-first step a blocking or indexing is employed to efficiently identify
+Record linkage is usually performed in a two step process;  in the
+first step a blocking or indexing is employed to efficiently group similar
 candidate record pairs which are then compared in more detail and
@@
 real-world databases.
+Parameter space: good linkage bad linkage don't know which without ground truth. M Trees always high no config -> high confidence.
 

### 2017-11-10 paper.tex

@@
 \documentclass{llncs}
-% * <peter.christen@anu.edu.au> 2016-08-19T15:07:29.567Z:
-%
-% ^.
 %\usepackage{llncsdoc}
@@
 
+% --------------------------------------------------------------------
+
 \section{Introduction}
@@
 
+\textbf{Traditional first - Peter?}
+
+Steorts et al.
+\cite{Steorts2014} examined various approaches to blocking records raging from traditional blocking to variations on Locality Sensitive Hashing (LSH).
+They assert that traditional blocking based on human selected fields has two major drawbacks.
+The first is that the blocks that are formed are so large that it remains computationally impractical to compare all the records in the block.
+Secondly, since similarity is based on a subset of the available fields, records in the same block may differ greatly in the fields that are not considered.
+\textbf{{this is true of our technique - explore somewhere}}.
+The paper examines two LSH variations - transitive LSH {ref needed} (TLSH) and K-Means LSH (KLSH).
+{\textbf{describe these where?}}
+Using simulated data based on people, their data of birth and postal address to evaluate the efficacy of these approaches, they demonstrate that KLSH performs as well or better than commonly used approaches to blocking such as those described \textbf{above}.
+However, they point out that, like all LSH methods, in order to get good results, it must be tuned to the particular dataset being linked.
+This in turn requires good ground truth about the dataset which may be unavailable, impossible or very expensive to obtain.
+\textbf{We examine and address these problems in this paper.}
+
+In \cite{Yu2016} Yu et al.
+have written an extensive survey on string similarity and join in which the problem is defined as "Given two sets of objects, similarity join aims to find all similar pairs from the two collections".
+Part of the survey is concerned the  algorithms used to compare two records which are characterised as being character-based, token-based or hybrid.
+The first, which includes approaches such as Levenshtein {ref needed} use character differences in the strings being compared, the second, which includes techniques such as Jaccard similarity {ref needed}, treats the records as sets of tokens and utilises set similarity, the third hybridises the two approaches.
+Yu's divides join algorithms (which unify records - also known as linkage) into three categories: a number of techniques based on filtering, verification based and threshold-based.
+Filtering involves generating signatures for each record and creating inverted lists containing records with common signatures and pruning records with no common signatures.
+Verification algorithms involve calculating the exact distances of a prefix of the data, and using another technique to estimate the edit distance between records.
+If the estimated distance is larger than the desired similarity threshold, the candidate may be discarded.
+The last category (threshold based) includes M-tree, Trie-Join, Pass-Join and Partenum.
+These approaches all require indexing structures to be constructed that are then used to support comparison.
+
+{\textbf{do we need to describe these?}}
+
+ \cite{Li2006} describes a method of performing linkage using R-trees {ref needed} that fall into the threshold based category of Yu et al.
+Their approach demonstrates that high precision and recall may be achieved by measuring Jaccard similarity over selected fields from the records being compared.
+In their evaluation , their approach achieved a recall of 99 percent in experiments in which conjunctions of similarities over fields were used.
+In a later paper \cite{Ciaccia97indexingmetric}  Ciaccia et al.
+demonstrate that using M-trees are almost always more efficient than R-trees.
+
 summarise related work in metric spaces as used in record linkage
@@
 In this paper we explore the efficacy of the use of different similarity search algorithms in probabilistic record linkage.
-We explore two inexact algorithms: LSH and MIFile \cite{amato2014mi} one exact method: mTrees \cite{paolociaccia2m}.
-We also use a simple brute force exact technique as a baseline for comparison although this approach can only be applied to the smallest of our datasets.
+We explore two inexact algorithms: LSH-minhash and MIFile \cite{amato2014mi} one exact method: mTrees \cite{paolociaccia2m}.
+We also use a simple brute force (nested-loop) exact technique as a baseline for comparison although this approach can only be applied to the smallest of our datasets.
+All of our experiments have a number of configuration parameters with which we can configure the search space and algorithm behaviours.
+All have a \textit{distance threshold} below which two records are deemed to represent a link.
+
+In the brute force approach two nested loops are used to compare every record with every other non identical record.
+The complexity of this approach is n2.
+If a link can be found with whatever distance metric is used, this approach will find it.
+Each of our datasets contains ground truth with which we can establish the records that are true links and those that are not.
+
+The LSH-minhash approach is a fast but inexact method of performing similarity search.
+It has a number of tuning parameters: \textit{shingle\_size}, the \textit{number\_of\_bands} and the \textit{band\_size}.
+These are explained in the following algorithm description which follows a standard LSH approach.
+Firstly all the fields of the record are turned into strings and concatenated.
+Next the strings are \textit{shingled} into a set of n-grams according to the \textit{shingle\_size} parameter.
+Next a set of randomly, deterministically generated hash functions are applied to each of the sets of n-grams and the smallest (the minhash) of each application is added to a signature.
+The number of hashes used and thus the size of the signature is determined by the product of the {number\_of\_bands} and the {band\_size}.
+Lastly the signature is divided into {number\_of\_bands} segments and the elements from the band are hashed again to create an key.
+The original record is added to a hashmap associated with this by this key.
+It can be shown that the similarities between signatures generated in the second step have very close Jaccard similarity to the original string and that if an arbitrary record is run through this process the records found in the map are close to the record being searched with good precision and recall.
+We will further explore these properties in this paper.
+
+The MIFile \cite{amato2014mi} approach is another inexact technique for performing similarity search.
+Using  MI_file,  each record is represented by the ordering of distances from a collection of reference objects.
+It is based on the idea that if two records are close to each other  will have similar neighbours and will thus be represented by similar of distances from the chosen reference objects.
+An inverted file is created mapping from...
+
+In order to compare two objects in the dataset, we compare the two corresponding orderings of the reference objects
+
+{\textbf{need to say something about the operations supported somewhere}}
 
@@
 (?) records that refer to XX machine learning publications.
-% Peter, 31 Oct: not sure about the following - which Cora version do
-% you use? ****
- (\url{https://linqs-data.soe.ucsc.edu/public/lbc/cora.tgz}) \textbf{ which consists of 2708 scientific publications classified into one of seven classes. The citation network consists of 5429 links.}
 
@@
 as the submission is double-blind - so no mention of DS.}
+\textbf{I think we need to say something about this to provide context; there is none currently.}
 
@@
 \end{itemize}
-\emph{Peter 31 Oct: what are you actually using in your experiments?
-All certificates or only birth?}
+In the experiments described in this paper we examine links between birth and death records since these provided the highest levels of ground truth.
 
@@
 \item 50 links from a marriage record to the bride's death record.
+\item 
 \end{itemize}
+Again only links between birth and death have been used in the experiments in this paper.
 

### 2017-11-11 paper.tex
Update on Overleaf.

@@
 % think:
-%\title{Exploring approaches to probablistic record linkage using
+%\title{Exploring approaches to probabilistic record linkage using
 %        similarity searching}
@@
 
-something about record linkage and its importance, application examples
+% Peter, 20171111:
 
-drawback of existing approaches: blocking and comparisons/classification
-done separately, sometimes if bad linkage quality means to go back to redo
-blocking.
+Record linkage, also known as entity resolution, data matching and
+duplicate detection~\cite{Chr12}, is the process of identifying and
+matching records that refer to the same real-world entities within and
+across databases. The entities to be linked are often people (such as 
+patients in hospital databases or customers in business databases),
+but record linkage can also be applied to link consumer products or
+bibliographic records~\cite{Chr12}. Commonly, record linkage is
+challenged by the lack of unique entity identifiers (keys) in the
+databases to be linked, which prevents the use of a database join.
+Instead, the linkage of records requires the comparisons of the
+attributes (or fields) available in common in the databases to be
+linked. For databases that contain information about individuals,
+the common attributes used often include the names, address,
+dates of birth, and so on, of individuals.
+
+To overcome data quality issues such as typographical errors and
+variations (which are common in name and address values~\cite{Chr12}),
+approximate string comparison functions (such as edit distance, the
+Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used
+to compare pairs of records, leading to a vector of similarities (one
+similarity per attribute/field compared) for each pair. These
+similarity vectors are then used to classify the compared record
+pairs into matches (where it is assumed the records in a pair
+correspond to the same real-world entity) and non-matches (where the
+records are assumed to correspond to two different entities). Various
+classification methods have been employed in record linkage, ranging
+from simple threshold based to sophisticated clustering approaches
+and supervised classifiers~\cite{}.
+
+Besides the two challenges of a lack of unique entity identifiers and
+data quality (which will affect linkage quality), record linkage is
+challenged by the increasing sizes of the databases to be
+linked~\cite{Don15} in today's Big Data era. To prevent the full
+pair-wise comparison of all possible record pairs (which is quadratic
+in the sizes of the databases to be linked), blocking, indexing and
+filtering techniques are commonly used~\cite{Chr12}. The basic idea
+of techniques is to split the databases into smaller blocks in a
+computationally efficient way such that records that likely correspond
+to the same entity are inserted into the same block. Only records that
+are in the same block are then compared with each other in more
+details. While blocking techniques facilitate the efficient linkage
+of very large databases~\cite{Don15}, they generally have the drawback
+of achieving scalability at the cost of (a possibly small) loss in
+linkage quality, as potentially true matching record pairs are
+removed in the blocking or indexing step. Blocking or indexing often
+leads to a reduction in recall of the final linkage
+result~\cite{Chr12}. A variety of blocking and indexing techniques,
+as discussed in more details in the following section, have been
+proposed, ranging from simple phonetic based blocking~\cite{Chr12} and
+sorting of the databases~\cite{Dra12} to locality sensitive hashing
+based techniques~\cite{Kim10} and even unsupervised~\cite{Kej13,Ram15}
+or supervised~\cite{Bil06,Mic06} learning of optimal blocking schemes.
+
+~ \\
+\emph{Peter 20171111: Could we have 2 diagrams: traditional RL process
+(block - compare - classify) versus our proposed 1 step approach
+(metric indexing) to illustrate our idea?}
+~ \\
+
+The overall process of a record linkage application, where blocking
+or indexing is preformed prior to the comparison and classification
+steps as illustrated in Fig.~\ref{fig-rl-process}, adds a further
+practical complexity. Often blocking, comparison and classification
+are conducted using algorithms selected (and their parameters set)
+based on technical and domain expertise of the record linkage user,
+followed by a manual assessment of the linkage outcomes. If the
+quality of the resulting links is not good enough for a certain
+application, the linkage process needs to be repeated with different
+parameter settings and possibly also alternative algorithms
+selected. This can result in a time-consuming and labor intensive
+process. A major challenge is often the heuristic nature of blocking,
+where the choice of a certain blocking technique and its parameters
+will determine the final outcome of a linkage. If a certain blocking
+approach results in a loss of recall in the final linkage results,
+the blocking needs to be redone, an often time consuming process.
+
+In this paper we explore how existing blocking and indexing technique
+can lead to loss in recall, even for record pairs that are highly
+similar and likely correspond to true matches. As a result, using
+such techniques can lead to \emph{incomplete} record linkage
+results. To overcome this problem, we advocate the use of
+\emph{metric space indexing} (MSI) approaches, where it is possible to
+guarantee that all record pairs, up-to a certain similarity threshold,
+are included in the record linkage results. Furthermore, MSI allows
+us to combine the indexing and comparison steps (as traditionally
+conducted independently) into one single step, making the overall
+record linkage process both more efficient and more effective.
+
+% peter 20171111: do we also need to mention we do not waste
+% comparisons of record pairs with low similarities?
+
+~ \\
+\emph{Peter 20171111: here we need the figure showing that traditional
+blocking techniques result in incomplete record linkage, while metric
+indexing does not.}
+~ \\
+
+While MSI has been used extensively in areas such as database indexing
+and general nearest neighbor search, its use in record linkage so far
+has been limited. 
+% Peter 20171111: a dangerous statement. if wrong we're stuffed.. 
+% but I am not really aware of MSI approaches in RL..
+Many different MSI approaches have been developed, such as various
+metric tree structures~\cite{}. 
+% Peter 20171111: needs same text from you guys on MSI..
+In this paper, we will use the MSI techniques of M-tree and M-file (?)
+and show how their use can facilitate an improvement of both
+efficiency and effectiveness of the record linkage process.
+
+\smallskip
+\textbf{Contributions:} We make the following specific contributions
+in the paper: (1) We investigate how existing indexing and blocking
+techniques can result in \emph{incomplete} record linkage results,
+as well as in an in-efficient record linkage process. (2) We explore
+MSI for record linkage and propose a MSI based approach that achieves
+\emph{complete}, efficient and effective record linkage. (3) We
+evaluate our approach on several real-world data sets from diverse
+domains and show its advantages over existing indexing and
+blocking techniques for record linkage.
+
+% old stuff below
 
@@
 
+\iffalse  % *******************************************************
+% we can't write this as it reveals us in the double-blind review
+% we could ad to final paper
 The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977.
@@
 As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.
+\fi  % *********************************************************
 
@@
 The MIFile \cite{amato2014mi} approach is another inexact technique for performing similarity search.
-Using  MI_file,  each record is represented by the ordering of distances from a collection of reference objects.
+Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
 It is based on the idea that if two records are close to each other  will have similar neighbours and will thus be represented by similar of distances from the chosen reference objects.

### 2017-11-12 paper.tex
Update on Overleaf.

--- 2017-11-10 paper.tex
@@
 
-something about record linkage and its importance, application examples
+% Peter, 20171112:
 
-drawback of existing approaches: blocking and comparisons/classification
-done separately, sometimes if bad linkage quality means to go back to redo
-blocking.
+Record linkage, also known as entity resolution, data matching and
+duplicate detection~\cite{Chr12}, is the process of identifying and
+matching records that refer to the same real-world entities within or
+across databases. The entities to be linked are often people (such as 
+patients in hospital or customers in business databases), but record
+linkage can also be applied to link consumer products or bibliographic
+records~\cite{Chr12}. Record linkage is commonly challenged by the
+lack of unique entity identifiers (keys) in the databases to be linked,
+which prevents the use of a database join. Instead, the linkage of
+records requires the comparison of the attributes (or fields) that are
+available in common in the databases to be linked. For databases that
+contain information about individuals, these attributes include the
+names, address, dates of birth, and so on, of individuals.
 
-something about similarity spaces / metric spaces, M-trees and its
+To overcome data quality issues such as typographical errors and
+variations (which are common in name and address values~\cite{Chr12}),
+approximate string comparison functions (such as edit distance, the
+Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used
+to compare pairs of records, leading to a vector of similarities (one
+similarity per attribute/field compared) for each pair. These
+similarity vectors are then used to classify the compared record
+pairs into matches (where it is assumed both records in a pair
+correspond to the same real-world entity) and non-matches (where the
+records are assumed to correspond to two different entities). Various
+classification methods have been employed in record
+linkage~\cite{Chr12,Don15}, ranging from simple threshold based to
+sophisticated clustering and supervised classification techniques, as
+well as active learning approaches~\cite{Wan15}.
+
+Besides the two challenges of a lack of unique entity identifiers and
+data quality (which will affect linkage quality), record linkage is
+also challenged by the increasing sizes of the databases to be
+linked~\cite{Don15}. To prevent the full pair-wise comparison of all
+possible record pairs (which is quadratic in the sizes of the
+databases to be linked), blocking techniques (commonly known as
+\emph{indexing}~\cite{Chr12b}) are used. The basic idea of these
+techniques is to split the databases into smaller blocks in a
+computationally efficient way such that records that likely correspond
+to the same entity are inserted into the same block. Only records that
+are in the same block are then compared with each other in more
+details.
+
+While indexing techniques facilitate the efficient linkage of very
+large data-bases~\cite{Don15}, they generally achieve scalability at
+the cost of loss in linkage quality, because potentially true matching
+record pairs are removed in the indexing step, leading to a reduction
+in recall of the final linkage result~\cite{Chr12}. A variety of
+indexing techniques, as discussed in more details in the following
+section, have been proposed, ranging from simple phonetic based
+blocking~\cite{Chr12} and sorting of the databases~\cite{Dra12} to
+locality sensitive hashing based techniques~\cite{Kim10,Steorts2014}.
+Unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06} 
+learning techniques of optimal blocking schemes have also been
+proposed.
+
+\begin{figure}[!t]
+  \centering
+  %\scalebox{1.0}[1.0]
+  \includegraphics[width=1.0\textwidth]{figures/linkage-process}
+  \caption{Overview of the steps of the traditional record linkage
+           process (left-hand size) and our proposed metric space
+           indexing approach (right-hand size), as described in
+           Sect.~\ref{sec-intro}, where records from two databases,
+           $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
+           \label{fig-rl-process}
+\end{figure}
+
+The overall process of a record linkage application, where indexing is
+preformed prior to the comparison and classification steps, as
+illustrated in Fig.~\ref{fig-rl-process}, adds a further practical
+complexity. Indexing, comparison and classification are often conducted
+using algorithms selected (and their parameters set) based on technical
+and domain expertise of the record linkage user, followed by a manual
+assessment of the linkage outcomes~\cite{Chr12}. If the quality of the
+resulting links is not good enough for a certain application, the
+linkage process needs to be repeated with different parameter settings
+and possibly also alternative algorithms selected. This can result in
+a time-consuming and labor intensive iterative process. A major
+challenge is often the heuristic nature of indexing, where the choice
+of a certain indexing technique and its parameters (including which
+attributes to use in indexing) will determine the final outcome of a
+linkage.
+% Repeat:
+% If a certain blocking approach results in a loss of recall in the
+%final linkage results, the blocking needs to be redone, an often time
+%consuming process.
+
+In this paper we explore how existing indexing techniques can lead to
+loss in recall, even for record pairs that are highly similar and
+likely correspond to true matches. As a result, using such techniques
+can lead to \emph{incomplete} record linkage results. To overcome this
+problem, we advocate the use of \emph{metric space indexing} (MSI)
+approaches, where it is possible to guarantee that all record pairs,
+up-to a certain distance threshold, are included in the record linkage
+results. MSI also prevents the expensive detailed comparison of record
+pairs that have low similarities but that need to be compared because
+they have been inserted into the same block with traditional indexing
+techniques. Furthermore, MSI allows us to combine the indexing,
+comparison and classification steps (as traditionally conducted
+independently) into one single step, as shown in
+Fig.~\ref{fig-rl-process}, making the overall record linkage process
+both more efficient and more effective.
+
+% Peter 20171112: do we also need to mention we do not waste
+% comparisons of record pairs with low similarities (i.e. those with
+% a distance > max_dist)?
+
+~ \\
+\emph{Peter 20171111: here we need the figure showing that traditional
+blocking techniques result in incomplete record linkage, while metric
+indexing does not.}
+~ \\
+
+While MSI has been used extensively in areas such as database indexing
+and general nearest neighbor search, its use in record linkage so far
+has been limited. 
+% Peter 20171112: a pretty dangerous statement. if wrong we're stuffed.. 
+% but I am not really aware of MSI approaches in RL..
+Many different MSI approaches have been developed, such as various
+metric tree structures~\cite{}. 
+% Peter 20171112: needs same text from you guys on MSI..
+In this paper, we will use the MSI techniques of M-tree and M-file (?)
+and show how their use can facilitate an improvement of both
+efficiency and effectiveness of the record linkage process.
+
+~ \\
+\emph{
+to add: something about similarity spaces / metric spaces, M-trees and its
 more recent advanced/improved versions, and how they have been used in
-similarity search.
+similarity search.}
 
-How our work is different from these other metric space works, and
-description of our contributions
+\emph{to add: How our work is different from these other metric space works, and
+description of our contributions.
+}
+~ \\
 
+\smallskip
+\textbf{Contributions:} We make three specific contributions in this
+paper: (1) We investigate how existing indexing techniques can result
+in \emph{incomplete} record linkage results, as well as in an
+in-efficient record linkage process. (2) We explore MSI for record
+linkage and propose a MSI based linkage approach that achieves
+\emph{complete}, efficient and effective record linkage. (3) We
+evaluate our approach on several real-world data sets from diverse
+domains and show its advantages over existing indexing techniques for
+record linkage.
+
+% old stuff below
+
+\iffalse  % *******************************************************
+% we can't write this as it reveals us in the double-blind review
+% we could ad to final paper
 The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977.
@@
 
-summarise related work in blocking/indexing
+We now briefly review relevant related work in the areas of indexing
+for record linkage (for a recent survey see Christen~\cite{Chr12b}),
+and metric space indexing.
+
+\smallskip
+\textbf{Indexing for Record Linkage:}
+Techniques to link records across databases have been developed over
+the past five decades~\cite{Fel69}, and scaling the linkage of large
+databases has been an ongoing challenge as databases are continue to
+grow in size and complexity. Traditional blocking techniques use a
+single or a combination of attributes (known as a \emph{blocking key})
+to insert records that have the same value(s) in such a blocking key
+into the same block. Only records in the same block are then combined
+with each other. To overcome variations and misspellings, the values
+used in blocking keys are commonly phonetically encoded using functions
+such as Soundex, NYSSI, or Double-Metaphone~\cite{Chr12}. These
+techniques convert a string (assumed to be a name) into a code
+according to how the name sounds, assigning the same code to similar
+sounding names (such as `Gail' and `Gayle' or `Dickson' and `Dixon').
+
+A different approach to indexing is the sorted neighborhood 
+method~\cite{Mon96}, where the databases to be linked are sorted
+according to a \emph{sorting key} (usually a concatenation of the
+values from several attributes). A sliding window is then moved over
+the databases and only records within the same window are being
+compared. A fixed window size, $w$, leads to a number of record
+pair comparison of $O(n \times w)$, where $n = |\mathbf{D}_A| +
+|\mathbf{D}_b|$, is the number of records in the two databases
+$\mathbf{D}_A$ and $\mathbf{D}_B$ to be linked. Techniques that
+adaptively shrink or expand the window size based on the
+characteristics of the sorting key values have shown to improve both
+linkage efficiency and quality~\cite{Dra12}. The drawback of sorting
+based
+
+\emph{peter 20171112 until here, to do: LSH Harra, Kim10, Stoerts, drawbacks of exiting techniques}
+blocking framework papadakis TKDE 2013
+
+% describe LSH: harra (Kim10)
+% Stoerts
+
+% learning blocking keys
+
+Drawbacks of existing techniques: errors in blocking/sorting keys,
+manually selected, is learned need manual training data, or are
+expensive (if unsupervised)
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+\smallskip
+\textbf{Metric Space Indexing:}
+
+% Al's writing below
 

### 2017-11-13 paper.tex
Update on Overleaf.

@@
 
-We propose an metric indexing approach to \emph{complete} record
+We propose a metric indexing approach to \emph{complete} record
 linkage, which ensures all record pairs that have a similarity above
@@
 
+% 1 1/2 pages maximum!
+
 \section{Introduction}
@@
 in recall of the final linkage result~\cite{Chr12}. A variety of
-indexing techniques, as discussed in more details in the following
+indexing techniques, as discussed in more detail in the following
 section, have been proposed, ranging from simple phonetic based
@@
 using algorithms selected (and their parameters set) based on technical
-and domain expertise of the record linkage user, followed by a manual
-assessment of the linkage outcomes~\cite{Chr12}. If the quality of the
+and domain expertise of the user of the record linkage system, followed by a manual
+assessment (auditing) of the linkage outcomes~\cite{Chr12}. If the quality of the
 resulting links is not good enough for a certain application, the
@@
 they have been inserted into the same block with traditional indexing
-techniques. Furthermore, MSI allows us to combine the indexing,
+techniques. Furthermore, MSI allows the combination of the indexing,
 comparison and classification steps (as traditionally conducted
@@
 both more efficient and more effective.
+
+% Al: but is it? be careful of last sentence in final version.
 
@@
 has been limited. 
+% Al - So take it out - I recently found and cited a paper that uses R trees later in the paper?
+
 % Peter 20171112: a pretty dangerous statement. if wrong we're stuffed.. 
@@
 
+% 1 page maximum!
+
 \section{Related Work}
@@
 We now briefly review relevant related work in the areas of indexing
-for record linkage (for a recent survey see Christen~\cite{Chr12b}),
-and metric space indexing.
+for record linkage (for a recent survey see~\cite{Chr12b,Pap16}),
+and metric space indexing~\cite{Zezula2010}.
 
@@
 \textbf{Indexing for Record Linkage:}
-Techniques to link records across databases have been developed over
-the past five decades~\cite{Fel69}, and scaling the linkage of large
-databases has been an ongoing challenge as databases are continue to
-grow in size and complexity. Traditional blocking techniques use a
-single or a combination of attributes (known as a \emph{blocking key})
-to insert records that have the same value(s) in such a blocking key
-into the same block. Only records in the same block are then combined
+Techniques to link records across databases have been developed for
+over five decades~\cite{Fel69,New59}, and the scalability of linking
+has been an ongoing challenge as databases are continue to grow in
+size and complexity. Traditional blocking~\cite{Chr12b} use a single
+or a combination of attributes (known as a \emph{blocking key}) to
+insert records that have the same value(s) in a blocking key into
+the same block. Only records within the same block are then combined
 with each other. To overcome variations and misspellings, the values
-used in blocking keys are commonly phonetically encoded using functions
-such as Soundex, NYSSI, or Double-Metaphone~\cite{Chr12}. These
+used in blocking keys can be phonetically encoded using functions
+such as Soundex, NYSIIS, or Double-Metaphone~\cite{Chr12}. These
 techniques convert a string (assumed to be a name) into a code
-according to how the name sounds, assigning the same code to similar
-sounding names (such as `Gail' and `Gayle' or `Dickson' and `Dixon').
+according to its pronunciation, assigning the same code to similar
+sounding names (such as `Gail' and `Gayle').
 
@@
 the databases and only records within the same window are being
-compared. A fixed window size, $w$, leads to a number of record
-pair comparison of $O(n \times w)$, where $n = |\mathbf{D}_A| +
-|\mathbf{D}_b|$, is the number of records in the two databases
-$\mathbf{D}_A$ and $\mathbf{D}_B$ to be linked. Techniques that
-adaptively shrink or expand the window size based on the
-characteristics of the sorting key values have shown to improve both
-linkage efficiency and quality~\cite{Dra12}. The drawback of sorting
-based
+compared. Techniques that adaptively shrink or expand the window size based on
+the characteristics of the sorting key values have shown to improve
+both linkage efficiency and quality~\cite{Dra12}.
 
-\emph{peter 20171112 until here, to do: LSH Harra, Kim10, Stoerts, drawbacks of exiting techniques}
-blocking framework papadakis TKDE 2013
+% should we mention canopy clustering? best to only briefly mention
+% techniques we do not use in our comparison
 
-% describe LSH: harra (Kim10)
-% Stoerts
+Locality sensitive hashing (LSH), as originally proposed to allow
+efficient nearest-neighbor search in high-dimensional
+spaces~\cite{Ind98}, has been employed in record linkage as a blocking
+technique where attribute values are hashed multiple times, and
+blocks are created from those records that share some hash values.
+\emph{HARRA} is a record linkage approach based on Min-hash and LSH
+which blocks, compares, and then merges linked records in an iterative
+fashion, where merged records are re-hashed to improve overall linkage
+quality. Steorts et al.~\cite{Steorts2014} recently evaluated two LSH
+variations, with their main conclusion being that in order to get good
+results LSH methods must be tuned to the particular databases being
+linked. This requires good quality ground truth data which may be
+unavailable or very expensive to obtain.
 
-% learning blocking keys
+Existing blocking techniques have the drawbacks that they are
+heuristics and commonly require domain knowledge, such as the choice
+of blocking or sorting key used. Errors and variations in the
+attributes used for blocking also result in records being inserted
+into a wrong block and thus true matches being missed. As a result,
+existing blocking techniques can lead to \emph{incomplete} linkage,
+and that many of the record pairs compared in a block turn out to be
+pairs with low similarities that correspond to non-matches, resulting
+in \emph{inefficient} linkage.
 
-Drawbacks of existing techniques: errors in blocking/sorting keys,
-manually selected, is learned need manual training data, or are
-expensive (if unsupervised)
+% Peter until here, 20171113 ************
 
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+\iffalse
 
-\smallskip
-\textbf{Metric Space Indexing:}
-
-% Al's writing below
-
-\textbf{Traditional first - Peter?}
+% commented, as overall related works should be around 1 page max
 
@@
 \textbf{We examine and address these problems in this paper.}
+
+\fi
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+\smallskip
+\textbf{Metric Space Indexing:}
+
+% Al's writing below
 
@@
 
+% end of related work should be around end of page 3!
+
 % --------------------------------------------------------------------
+
+% do we need a section on preliminaries, notation, etc?
+% ideally techniques are described formally, possibly via pseudo-code
+% algorithmic descriptions
 
@@
 The number of hashes used and thus the size of the signature is determined by the product of the {number\_of\_bands} and the {band\_size}.
-Lastly the signature is divided into {number\_of\_bands} segments and the elements from the band are hashed again to create an key.
-The original record is added to a hashmap associated with this by this key.
+Lastly the signature is divided into {number\_of\_bands} segments and the elements from the band are hashed again to create a number of keys.
+The original record is added to a hashmap associated with each of the keys.
 It can be shown that the similarities between signatures generated in the second step have very close Jaccard similarity to the original string and that if an arbitrary record is run through this process the records found in the map are close to the record being searched with good precision and recall.
 We will further explore these properties in this paper.
+To perform a {\textit{nearest-neighbour}} lookup, the query value is hashed using the procedure described above and the union of set of of values found for all hashes in the hashmap are returned.
+If desired, these may then be filtered by performing distance calculations on the results.
 
-The MIFile \cite{amato2014mi} approach is another inexact technique for performing similarity search.
+The MIFile \cite{amato2014mi} approach is inexact technique for performing similarity search based on metric spaces.
 Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
-It is based on the idea that if two records are close to each other  will have similar neighbours and will thus be represented by similar of distances from the chosen reference objects.
-An inverted file is created mapping from...
+It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances from the chosen reference objects.
+In order to create the indexes, for each value to be added to the data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3...
+based on the reference object's position from the value).
+Next for each score, a  mapping is created in an inverted index mapping from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
+In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
+Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
+Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
 
-In order to compare two objects in the dataset, we compare the two corresponding orderings of the reference objects
+M-trees support exact search operations over a set of values over a metric space (distances must support  triangle inequality - distance(i,j) + distance(j,k ) >= distance(i,k)).
+In addition to supporting exact search the operations supported by M-trees are richer than the other data structures described above and include \textit{range search} (all values within distance d and some query q), \textit{nearest-neighbour} (the closest value to q), and \textit{nearestN} (the N closest n objects to q).
+An M-tree is a balanced tree structure comprised of fixed sized nodes that can each store up to \textit{M} children.
+In addition to the children, each node contains a reference to an object being indexed, a pointer and the distance to its parent, and a radius (described below).
+The radius of an internal node called a \textit{routing-node} represents the  furthest distance to any of its children.
+Thus all the children may be visualised as being contained within a ball of radius \textit{r} from the parent.
+M-trees are grown from the roots upward and require dynamic tree rebalancing when set of children become full (of size \textit{M}).
 
-{\textbf{need to say something about the operations supported somewhere}}
-
-coverage: proportion of all rec pairs with a certain similarity
-(independent of their match status) that are being generated by
-a blocking  technique
-
-transitive closure is a problem if we have a blocking technique with overlapping blocks.
-
-define blocking: only aimed at improving computational aspects by reducing the number of record pairs to be compared
+% TODO coverage: proportion of all rec pairs with a certain similarity (independent of their match status) that are being generated by a blocking  technique - what does above refer to? al
 
@@
 \section{Experiments and Results}
-\label{sec-data}
+\label{sec-expt}
 
@@
 
-\emph{Peter, 31 Oct: We need to be careful not to reveal our identities
-as the submission is double-blind - so no mention of DS.}
-\textbf{I think we need to say something about this to provide context; there is none currently.}
+% Peter, 31 Oct: can the following be shortened to 1 paragraph - tick - al.
 
-% Peter, 31 Oct: can the following be shortened to 1 paragraph?
-
-We also experiment over two datasets prepared by researchers pursuing previous projects \cite{reid2002} and \cite{reid2006}, which act as pseudo subsets of the Digitising Scotland data set.
+We also experiment over two datasets prepared by researchers pursuing previous projects \cite{reid2002} and \cite{reid2006}, which act as pseudo subsets for the demographic linkage on which our work is primarily focused .
 One data set contains records of vital events registered on the Isle of Skye, a rural district, while the other contains records from Kilmarnock, an industrial town.
@@
 The Isle of Skye dataset comprises around 17,600 birth records, 12,300 death records, and 2,700 marriage records.
-The demographers  identified:
-\begin{itemize}
-\item 4,300 family groups of siblings, with a mean family size of 3.9 siblings and a maximum family size of 16,
-\item 2,900 links from a birth record to the child's death record,
-\item 60 links from a marriage record to the groom's death record, and
-\item 100 links from a marriage record to the bride's death record.
-\end{itemize}
-In the experiments described in this paper we examine links between birth and death records since these provided the highest levels of ground truth.
-
-The Kilmarnock data set contains around 38,400 birth records, 23,700 death records, and 8,700 marriage records. Here, the demographers identified:
-\begin{itemize}
-\item 13,100 family groups of siblings, with a mean family size of 2.8 siblings and a  maximum family size of 16,
-\item 8,300 links from a birth record to the child's death record,
-\item 30 links from a marriage record to the groom's death record, and
-\item 50 links from a marriage record to the bride's death record.
-\item 
-\end{itemize}
-Again only links between birth and death have been used in the experiments in this paper.
+Relating to the experiments described below, this dataset contains 2,900 links from a birth record to the child's death record.
+The Kilmarnock data set contains around 38,400 birth records, 23,700 death records, and 8,700 marriage records with a similar schema to the Skye data.
+In this dataset demographers identified  2,900 links from a birth record to the child's death record.
 

### 2017-11-15 paper.tex
Update on Overleaf.

@@
 
-In this paper we explore the efficacy of the use of different similarity search algorithms in probabilistic record linkage.
-We explore two inexact algorithms: LSH-minhash and MIFile \cite{amato2014mi} one exact method: mTrees \cite{paolociaccia2m}.
-We also use a simple brute force (nested-loop) exact technique as a baseline for comparison although this approach can only be applied to the smallest of our datasets.
-All of our experiments have a number of configuration parameters with which we can configure the search space and algorithm behaviours.
-All have a \textit{distance threshold} below which two records are deemed to represent a link.
+% peter: We need one paragraph describing the overall 'approach' /
+% methodology, as well as the aim, then a short description of
+% each method/technique used
 
-In the brute force approach two nested loops are used to compare every record with every other non identical record.
-The complexity of this approach is n2.
-If a link can be found with whatever distance metric is used, this approach will find it.
-Each of our datasets contains ground truth with which we can establish the records that are true links and those that are not.
+In this paper we explore the efficacy of the use of different
+similarity search algorithms in probabilistic record linkage.
+% Peter: remove probabilistic - hat we do is general to RL, the
+% probabilistic aspect is not really relevant
+We explore two inexact algorithms: LSH-minhash and MIFile \cite{amato2014mi} and one exact method: mTrees \cite{paolociaccia2m}. 
+We also use a simple brute force (nested-loop) exact technique as a 
+baseline for comparison although this approach can only be applied to 
+the smallest of our datasets. All of our experiments have a number of 
+configuration parameters with which we can configure the search space and algorithm behaviours, as we describe in Sect.~\ref{sec-exp}. All
+have a \emph{distance threshold}, $d$, below which two records are
+deemed to represent a link.
+% Peter: are classified to be link (i.e.\ refer to the same entity)
 
-The LSH-minhash approach is a fast but inexact method of performing similarity search.
-It has a number of tuning parameters: \textit{shingle\_size}, the \textit{number\_of\_bands} and the \textit{band\_size}.
-These are explained in the following algorithm description which follows a standard LSH approach.
-Firstly all the fields of the record are turned into strings and concatenated.
-Next the strings are \textit{shingled} into a set of n-grams according to the \textit{shingle\_size} parameter.
-Next a set of randomly, deterministically generated hash functions are applied to each of the sets of n-grams and the smallest (the minhash) of each application is added to a signature.
-The number of hashes used and thus the size of the signature is determined by the product of the {number\_of\_bands} and the {band\_size}.
-Lastly the signature is divided into {number\_of\_bands} segments and the elements from the band are hashed again to create a number of keys.
+In the brute force approach two nested loops are used to compare every 
+record with every other non identical record. The complexity of this 
+approach is n2. If a link can be found with whatever distance metric is 
+% Peter: $O(n^2)$, where $n$ is the number of records in the dataset to
+% be linked.
+used, this approach will find it.
+% Peter: find -> identify  -- better actually clarify: This approach
+% will identify all links of records that have a distance up-to the
+% given distance threshold, $d$.
+Each of our datasets contains ground truth with which we can establish
+the records that are true links and those that are not.
+
+% peter: below fast -> I would remove fast, as it might be slow for
+% certain parameter settings
+The LSH-minhash approach is a fast but inexact method of performing similarity search. It has three tuning parameters:
+\emph{shingle size}, $l_{ss}$, the \emph{band size}, $l_{bs}$, and the
+\emph{number of bands}, $l_{nb}$. These are explained in the following algorithm description which follows a standard LSH approach~\cite{??}.
+Firstly all the fields of the record are turned into strings and 
+concatenated. Next the strings are \emph{shingled} into a set of
+n-grams (sub-strings of length $n$) according to the $l_{ss}$
+parameter (i.e.\ $n = l_{ss}$). Next a set of randomly, 
+deterministically generated hash functions are applied to each of the
+sets of n-grams and the smallest (the minhash) of each application is
+added to a signature. The number of hashes used and thus the size of
+the signature is set to $l_{nb} \times l_{bs}$.
+Lastly the signature is divided into $l_{nb}$ segments and the elements from the band are hashed again to create a number of keys.
 The original record is added to a hashmap associated with each of the keys.
-It can be shown that the similarities between signatures generated in the second step have very close Jaccard similarity to the original string and that if an arbitrary record is run through this process the records found in the map are close to the record being searched with good precision and recall.
-We will further explore these properties in this paper.
+It can be shown that the similarities between signatures generated in the second step have very close Jaccard similarity to the original string~\cite{lsh ref} and that if an arbitrary record is run through
+% Peter: run through.. -> reword: is converted into such a LSH
+% signature then..  records with the same key(s) are...
+% Peter: close to.. -> is a bit vague
+this process the records found in the map are close to the record being searched with good precision and recall. We will further explore these properties in this paper.
+% peter: we will evaluate the performance of LSH based blocking with
+% regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
+% Sect.~\ref{sec-exp}.
+% peter: below - we want to do record linkage (slightly different
+% from nn search) - better to describe RL below not NN search! 
 To perform a {\textit{nearest-neighbour}} lookup, the query value is hashed using the procedure described above and the union of set of of values found for all hashes in the hashmap are returned.
@@
 
-The MIFile \cite{amato2014mi} approach is inexact technique for performing similarity search based on metric spaces.
+The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces.
 Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
-It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances from the chosen reference objects.
-In order to create the indexes, for each value to be added to the data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3...
-based on the reference object's position from the value).
-Next for each score, a  mapping is created in an inverted index mapping from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
+% Peter: are these reference object also records? 
+It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances
+% Peter: unclear what you mean with: similar of distances
+from the chosen reference objects. 
+In order to create the indexes, for each value to be added to the
+% peter: value -> record ?
+data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3... based on the reference object's position from the value).
+% peter: this is unclear .. I think I understand k-NN: for each
+% record you find the k nearest reference objects and you enumerate
+% them
+Next for each score, a  mapping is created in an inverted index mapping
+% peter: unclear..? a mapping in an inverted index mapping?
+from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
+% peter: unclear: what is "the data" associated with the reference
+% object.
 In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
+% peter: this sentence reads incomplete..?
 Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
+% Peter: sorry, this is unclear as well. "the objects closest" ->
+% records closest? And why highest scores..? Is the closest record
+% to a reference object given the highest score (not the lowest
+% score - i.e. not 1-closest, 2-closest etc)
 Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
+% peter: so I guess you use the triangular inequality and the reference
+% objects to prune the search space.
 
-M-trees support exact search operations over a set of values over a metric space (distances must support  triangle inequality - distance(i,j) + distance(j,k ) >= distance(i,k)).
-In addition to supporting exact search the operations supported by M-trees are richer than the other data structures described above and include \textit{range search} (all values within distance d and some query q), \textit{nearest-neighbour} (the closest value to q), and \textit{nearestN} (the N closest n objects to q).
+M-trees support exact search operations over a set of values over a metric space (distances must support triangle inequality -
+distance(i,j) + distance(j,k ) >= distance(i,k)).
+% peter: (distances must support triangle inequality, i.e.\ for three
+% objects $i$, $j$ and $k$, $dist(i,j) + dist(j,k) \ge dist(i,k)$).
+%
+In addition to supporting exact search the operations supported by M-trees are richer than the other data structures described above and include \textit{range search} (all values within distance d and some query q),
+% peter: where all objects $i$ within a distance $d$ for a query object
+% $q$ are identified, i.e.\ $dist(i,q) \le d$)
+\textit{nearest-neighbour} (the closest value to q), and \textit{nearestN} (the N closest n objects to q).
+% peter: we can add proper mathematical ways of defining those if
+% needed
 An M-tree is a balanced tree structure comprised of fixed sized nodes that can each store up to \textit{M} children.
+% not sure if we should use M here - the M in M-tree is not because
+% each node can have m children! In Sect 5 I used $t_b$ as name for
+% the branching factor.
 In addition to the children, each node contains a reference to an object being indexed, a pointer and the distance to its parent, and a radius (described below).
@@
 M-trees are grown from the roots upward and require dynamic tree rebalancing when set of children become full (of size \textit{M}).
+% Peter: and how is an M-tree used in record linkage?
+% TODO coverage: proportion of all rec pairs with a certain similarity (independent of their match status) that are being generated by a blocking  technique - what does above refer to? al
 
-% TODO coverage: proportion of all rec pairs with a certain similarity (independent of their match status) that are being generated by a blocking  technique - what does above refer to? al
+% peter: could we have some pseudo-code descriptions of how these
+% techniques are used for RL blocking?
 
@@
 
-\subsection{Analysis and Limitations}
-\label{sec-analysis}
+%\subsection{Analysis and Limitations}
+%\label{sec-analysis}
 
-maybe a complexity analysis? maybe an analysis of expected 
-linkage quality?
+%maybe a complexity analysis? maybe an analysis of expected 
+%linkage quality?
 
@@
 \section{Experiments and Results}
-\label{sec-expt}
+\label{sec-exp}
 
-We now describe the data sets we use in our evaluation and the
-experimental setup we employed to evaluate our proposed metric
-indexing approach.
+We now describe the data sets we use in our evaluation, the setup we
+employed to evaluate our proposed metric indexing approach and
+compare it with traditional blocking techniques, and we then present
+and discuss our results.
 
@@
 \begin{table}[t]
-\caption{Characteristics of data sets used in experiments.
-  \emph{Peter, 31 Oct: what else should we show here?}}
+\caption{Characteristics of data sets used in the experiments.}
  \label{table-datasets}
@@
   %\addtolength{\tabcolsep}{-0.5pt}
-  \begin{tabular}{ccccc}
+  \begin{tabular}{cccc}
   \hline\noalign{\smallskip}
-  Data set~ & ~Number of~ & ~Number of true~
-   & Attributes used? & ~?? \\
-  name(s)  & records   & matching pairs &
-    for indexing & ?? \\
-  \noalign{\smallskip}
-  \hline
-  \noalign{\smallskip}
-  NCVR  & ~224,073 / 224,061~ & ~148,036~ & ? & ? \\
-  CORA  & 1,295             & title, authors, venue, year ? & ?
-    & ? \\
-  Isle of Skye & ? & ?
-    & ? & ? \\
-  Kilmarnock  & ? & ? & ? & ? \\
-  \hline
+  Data set~ & ~Number of~ & ~Number of true~ & Attributes used \\
+  name(s)  & records  & matching pairs & for blocking \\
+  \noalign{\smallskip} \hline \noalign{\smallskip}
+  CORA  & 1,295 & 17,184 & Title, authors, venue, year ? \\
+  NCVR  & ~224,073 / 224,061~ & ~148,036~ & ? \\
+  Isle of Skye & 32,567 & 2,900 & ? \\
+  Kilmarnock  & 70852 & 2,900 ? why same number & ? \\
+  \noalign{\smallskip} \hline
   \end{tabular}
@@
 
-\subsection{Data Sets}
-\label{sec-data}
+\smallskip
+\textbf{Data Sets:}
+%\subsection{Data Sets and Experimental Setup}
+%\label{sec-data}
+%
+We used four real data sets from three domains in our experiments as
+summarized in Table~\ref{table-datasets}. The first is
+\emph{CORA}~\footnote{Available from:
+\texttt{http://secondstring.sourceforge.net}}, which contains 1,295
+records that refer to 112 machine learning publications.
 
-We used four real data sets from three domains in our experiments. The
-first is the CORA data set~\footnote{Available from:
-\texttt{http://secondstring.sourceforge.net}}, which contains 1,295
-(?) records that refer to XX machine learning publications.
+The second are two snapshots of the North Carolina Voter Registration
+database (\emph{NCVR})~\footnote{Available from: \texttt{http://dl.ncsbe.gov/}}, which we collected in April and June
+2014. Each of these contains records of voters and includes their
+first names, surnames, and addresses. Ground truth is provided via a
+\emph{NCID} identifier which uniquely identifies a voter. In our
+experiments we use two randomly selected sub-set of $224,073$ and
+$224,061$ records, respectively, where $148,036$ of those refer to
+voters that occurred in both original NCVR snapshots but had name
+and/or address changes over time (thus leading to around $66\%$
+matching records).
 
-The second are a pair of data sets based on the North Carolina Voter
-Registration (NCVR) data sets\footnote{Available from: \texttt{http://dl.ncsbe.gov/}}, which we collected in June 2014 and
-October 2016. Each of these two data sets contain over five million
-records of voters and include attributes with their first names,
-surnames, and addresses. Ground truth is provided via a \emph{NCID}
-identifier which uniquely identifies a voter. In our experiments we
-use two randomly selected sub-set of $224,073$ and $224,061$ records,
-respectively, where $148,036$ of those refer to voters that occurred
-in both original NCVR data sets but had name and/or address changes
-over time (thus leading to around $66\%$ matching records).
+The last two data sets are historical Scottish records of vital
+events (birth, marriages and deaths) registered on the
+\emph{Isle of Skye}, a rural district, while the other contains
+records from \emph{Kilmarnock}, an industrial town. These data sets
+were provided to us by the demographer and historians who extensively
+curated and linked both data sets~\cite{reid2002,reid2006}. Both data
+sets include the  names, gender, addresses of individuals and their
+parents. Ground truth was generated by the demographers based on their
+extensive domain knowledge.
 
@@
 
+\iffalse
 We also experiment over two datasets prepared by researchers pursuing previous projects \cite{reid2002} and \cite{reid2006}, which act as pseudo subsets for the demographic linkage on which our work is primarily focused .
@@
 
+% Peter: Why same number of links for Skye and Killi? *****************
+\fi
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+\begin{table}[t]
+\caption{Parameter settings used for the different data sets used in
+   the experiments,with $d$ the distance threshold used, $t_b$ the
+   M-tree branching factor, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
+   shingle size, and the size and number of bands, respectively, and
+   $f_{ki}$ and $f_{ks}$ the MIfile... \emph{please add...}.}
+ \label{table-parameters}
+  \centering
+  \begin{scriptsize}
+  %\addtolength{\tabcolsep}{-0.5pt}
+  \begin{tabular}{cccccccc}
+  \hline\noalign{\smallskip}
+  Data set~ & $d$ & $t_b$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ & 
+    $f_{ki}$ & $f_{ks}$ \\
+  name(s) & & & & & & & \\
+  \noalign{\smallskip} \hline \noalign{\smallskip}
+  CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$[5,10,25,50]$~ & ~$2$~ &
+    ~$[2,5,10]$~ & ~$[2,5,10]$~ & ~$[30,35,40]$~ & ~$[30,35,40]$ \\
+  NCVR  &  \\
+  Isle of Skye &  \\
+  Kilmarnock  &  \\
+  \noalign{\smallskip} \hline
+  \end{tabular}
+  \end{scriptsize}
+\end{table}
+
+% the narrow values for ki and ks need some justification!
+
+\smallskip
+\textbf{Experimental Setup:}
+%
+We implemented all techniques described in Sect.~\ref{sec-approach}
+using Java (version x.y) and ran an extensive et of experiments on
+a compute server with XX GBytes of main memory and XX TBytes of disk
+space, and running (operating system and version).
+
+The parameter settings used in our experiments are summarized in
+Table~\ref{table-parameters}. ... \emph{describe and justify selected
+values..}
+
+discuss measures used. PC, PQ, num comparisons, time? etc.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+\smallskip
+\textbf{Results and Discussion}
+
+%\subsection{Results and Discussion}
+%\label{sec-results}
+
+table or plots?
+
 % --------------------------------------------------------------------
@@
 
-\bibliographystyle{abbrv}
+\bibliographystyle{splncs03}
+%\bibliographystyle{abbrv}
 \bibliography{paper.bib}

### 2017-11-17 paper.tex
2017-11-17

@@
 \emph{shingle size}, $l_{ss}$, the \emph{band size}, $l_{bs}$, and the
-\emph{number of bands}, $l_{nb}$. These are explained in the following algorithm description which follows a standard LSH approach~\cite{??}.
+\emph{number of bands}, $l_{nb}$. These are explained in the following algorithm description which follows a standard LSH approach [CITATION].
 Firstly all the fields of the record are turned into strings and 
@@
 The original record is added to a hashmap associated with each of the keys.
-It can be shown that the similarities between signatures generated in the second step have very close Jaccard similarity to the original string~\cite{lsh ref} and that if an arbitrary record is run through
+It can be shown that the similarities between signatures generated in the second step have very close Jaccard similarity to the original string [LSH CITATION] and that if an arbitrary record is run through
 % Peter: run through.. -> reword: is converted into such a LSH

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 
-% Peter 31 Oct: Current title does not really reflect the topic I
-% think:
-%\title{Exploring approaches to probabilistic record linkage using
-%        similarity searching}
+% ====================================================================
+% Previous title versions
+\iffalse
+
+% Peter 31 Oct: Current title does not really reflect the topic I think:
+%\title{Exploring approaches to probabilistic record linkage using similarity searching}
 % Peter 31 Oct: What about
-\title{Efficient and Effective Metric Space Indexing for Complete
-       Record Linkage
-       Al: Exploring approaches to probabilistic record linkage using similarity searching (metric space indexing?) }
-%  \thanks{The authors would like to thank the ...|
+\title{Efficient and Effective Metric Space Indexing for Complete Record Linkage}
+
+\fi
+% ====================================================================
+
+\title{Using Metric Space Indexing for Efficient Complete Record Linkage}
 
 \author{Submitted for double-blind review}
-%\author{Özgür Akgün\inst{1} and Alan Dearle\inst{1} and Graham Kirby\inst{1} Peter Christen\inst{2}
 
-%\institute{School of Computer Science,
-%           University of St Andrews, \\
-%           St Andrews, Scotland.~
-%           Contact: \email{ozgur.akgun@st-andrews.ac.uk}
+% ====================================================================
+%  Hidden For Blind Review
+\iffalse
 
-%\institute{Research School of Computer Science,
-%           The Australian National University, \\
-%           Canberra, Australia.~
-%           Contact: \email{peter.christen@anu.edu.au}
+\author{{\"O}zg{\"u}r Akg{\"u}n\inst{1} \and Alan Dearle\inst{1} \and Graham Kirby\inst{1}
+\and Peter Christen\inst{2}}
+
+\institute{School of Computer Science, University of St Andrews, St Andrews, Scotland.\\
+Contact: \email{ozgur.akgun@st-andrews.ac.uk}
+\and
+Research School of Computer Science, The Australian National University, Canberra, Australia.\\
+Contact: \email{peter.christen@anu.edu.au}}
+
+\fi
+% ====================================================================
 
@@
 
-% MAX 200 words
 \begin{abstract}
-Probabilistic record linkage is the process of identifying records that refer to
-the same real-world entities across two or more databases in situations where entity identifiers are unavailable.
-It requires  similarities between records to be approximated using common attributes (such as names, addresses, phone numbers) and for pairs of records to be classified as matches or non-matches.
-Record linkage is usually performed in a two step process;  in the
-first step a blocking or indexing is employed to efficiently group similar
-candidate record pairs which are then compared in more detail and
-classified in the second step.
-
-As we show in this paper, even state-of-the-art blocking and
-indexing techniques such as locality sensitive hashing, have two
-major drawbacks. First, they often remove some true matching record
-pairs that have high similarities due to their heuristic nature,
-leading to a loss of recall. Second, they include many record pairs
-with low similarity leading to high computational requirements.
-Combined, this means a reduction of both effectiveness and efficiency.
-% * <al@st-andrews.ac.uk> 2017-10-31T08:46:25.069Z:
+%Probabilistic
+Record linkage is the process of identifying records that refer to the
+same real-world entities, in situations where entity identifiers are
+unavailable. Records are linked on the basis of similarity between
+common attributes, with every pair of records being classified as a
+match or non-match depending on the degree of similarity between them.
+Record linkage is usually performed in a three-step process, where first
+groups of similar candidate records are identified using indexing. Only
+pairs within the same group are then compared in more detail, and
+finally classified.
 %
-% > efficiency
+Even state-of-the-art indexing techniques, such as Locality Sensitive
+Hashing, have potential drawbacks. They may fail to group together some
+true matching records with high similarity. Conversely, they may group
+records with low similarity, leading to high computational overhead.
 %
-% ^.
-
-We propose a metric indexing approach to \emph{complete} record
-linkage, which ensures all record pairs that have a similarity above
-a certain similarity threshold are being \textbf{compared}, while no pairs
-with a similarity below the threshold are \textbf{compared}. Our approach
-leads to a one-step record linkage process which combines indexing
-with comparison and classification resulting in improved
-effectiveness and efficiency as we experimentally evaluate on several
-real-world databases.
-Parameter space: good linkage bad linkage don't know which without ground truth. M Trees always high no config -> high confidence.
+We propose using metric space indexing to perform \emph{complete} record
+linkage, which results in a parameterless record linkage process
+combining indexing, comparison and classification into a single step
+delivering high effectiveness and efficiency.
 
@@
 % ====================================================================
-
-\section{The Plan}
-
-\vspace{5mm}
-
-Plan in a bulleted list:
-
-\begin{itemize}
-\item Design: Traditionally it has two stages.
-A blocking method needs to be developed separately from a similarity method.
-With similarity search based record linkage, there is only one method, and that is the similarity method.
-\item Efficiency: I am not convinced of this, but Peter thinks there may be efficiency benefits.
-The reasoning behind this is that with blocking-based methods we would have to calculate some sort of a similarity value first.
-This value will only be used for blocking, and later a similarity value between candidate pairs will need to be calculated again.
-\item Completeness: Similarity search based methods will be complete (with respect to the given similarity method) whereas blocking based methods (including LSH) are likely to be incomplete (there will be high-similarity pairs which are not places in the same block)
-\item Data sets
-\begin{itemize}
-\item Kilmarnock Birth-Death
-\item Skye Birth-Death
-\item CORA
-\item North Carolina Voter Registration DB
-\end{itemize}
-\item The following is a list of all methods we thought we would have to compare.
-\begin{itemize}
-\item Similarity search (M-Tree, others?)
-\item LSH-blocking
-\item Traditional blocking (e.g. Region, Surname) - \textbf{why this? - this was done in the paper: A Comparison of Blocking Methods for Record Linkage}
-\item MIFile
-\item Sorted Neighbourhood blocking - \textbf{why this?}
-\end{itemize}
-\item Pairs-completeness and pairs-quality (w.r.t high similarity instead of true-link status). I will expand on this a bit more later.
-\item Comparison (for blocking based methods)
-		Precision/Recall/F1-Measure
-\item Which similarity methods are metric?
-\item Run time and memory usage comparisons
-\item Theorem. Precision might get better or worse with similarity search based methods, but recall should never get worse.
-\end{itemize}
-
-Note: The rest of this paper should be regarded as a placeholder for now.
-
-% --------------------------------------------------------------------
 
@@
 linkage can also be applied to link consumer products or bibliographic
-records~\cite{Chr12}. Record linkage is commonly challenged by the
-lack of unique entity identifiers (keys) in the databases to be linked,
-which prevents the use of a database join. Instead, the linkage of
-records requires the comparison of the attributes (or fields) that are
-available in common in the databases to be linked. For databases that
-contain information about individuals, these attributes include the
-names, address, dates of birth, and so on, of individuals.
+records~\cite{Chr12}. Record linkage is commonly challenged by the lack
+of unique entity identifiers (keys) in the databases to be linked, which
+prevents the use of a database join. Instead, the linkage of records
+requires the comparison of the common attributes (or fields) that are available
+in the datasets to be linked. For datasets that contain
+information about individuals, these attributes include the names,
+address, dates of birth, and so on, of individuals.
 
@@
 approximate string comparison functions (such as edit distance, the
-Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used
-to compare pairs of records, leading to a vector of similarities (one
-similarity per attribute/field compared) for each pair. These
-similarity vectors are then used to classify the compared record
-pairs into matches (where it is assumed both records in a pair
-correspond to the same real-world entity) and non-matches (where the
-records are assumed to correspond to two different entities). Various
-classification methods have been employed in record
-linkage~\cite{Chr12,Don15}, ranging from simple threshold based to
-sophisticated clustering and supervised classification techniques, as
-well as active learning approaches~\cite{Wan15}.
+Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
+compare pairs of records, leading to a vector of similarities (one
+similarity per attribute/field compared) for each pair. These similarity
+vectors are then used to classify the compared record pairs into matches
+(where it is assumed both records in a pair correspond to the same
+real-world entity) and non-matches (where the records are assumed to
+correspond to two different entities). Various classification methods
+have been employed in record linkage~\cite{Chr12,Don15}, ranging from
+simple threshold based to sophisticated clustering, supervised
+classification techniques, and active learning approaches~\cite{Wan15}.
 
-Besides the two challenges of a lack of unique entity identifiers and
-data quality (which will affect linkage quality), record linkage is
-also challenged by the increasing sizes of the databases to be
-linked~\cite{Don15}. To prevent the full pair-wise comparison of all
-possible record pairs (which is quadratic in the sizes of the
-databases to be linked), blocking techniques (commonly known as
-\emph{indexing}~\cite{Chr12b}) are used. The basic idea of these
-techniques is to split the databases into smaller blocks in a
-computationally efficient way such that records that likely correspond
-to the same entity are inserted into the same block. Only records that
-are in the same block are then compared with each other in more
-details.
+Besides the issues of a lack of unique entity identifiers, and data
+quality (which will affect linkage quality), record linkage is also
+challenged by the increasing sizes of the datasets to be
+linked~\cite{Don15}. To avoid full pair-wise comparison of all possible
+record pairs (quadratic in the sizes of the datasets to be linked),
+blocking techniques, commonly known as \emph{indexing}~\cite{Chr12b},
+are used. These split the datasets into smaller blocks in a
+computationally efficient way, such that records that are likely to
+correspond to the same entity are grouped into the same block. Only
+records within the same block are then compared in more detail.
 
-While indexing techniques facilitate the efficient linkage of very
-large data-bases~\cite{Don15}, they generally achieve scalability at
-the cost of loss in linkage quality, because potentially true matching
-record pairs are removed in the indexing step, leading to a reduction
-in recall of the final linkage result~\cite{Chr12}. A variety of
-indexing techniques, as discussed in more detail in the following
-section, have been proposed, ranging from simple phonetic based
-blocking~\cite{Chr12} and sorting of the databases~\cite{Dra12} to
-locality sensitive hashing based techniques~\cite{Kim10,Steorts2014}.
-Unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06} 
-learning techniques of optimal blocking schemes have also been
-proposed.
+While indexing techniques facilitate efficient linkage of very large
+datasets~\cite{Don15}, they generally achieve scalability at the cost
+of reduced linkage quality, because potentially true matching record
+pairs are removed in the indexing step, leading to a reduction in recall
+of the final linkage result~\cite{Chr12}. A variety of indexing
+techniques, discussed in more detail in the following section, have
+been proposed, ranging from simple phonetic based blocking~\cite{Chr12}
+and sorting of the datasets~\cite{Dra12} to locality sensitive hashing
+based techniques~\cite{Kim10,Steorts2014}. Techniques for
+unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06}
+learning of optimal blocking schemes have also been proposed.
 
@@
   \caption{Overview of the steps of the traditional record linkage
-           process (left-hand size) and our proposed metric space
-           indexing approach (right-hand size), as described in
+           process (left-hand side) and our proposed metric space
+           indexing based approach (right-hand side), as described in
            Sect.~\ref{sec-intro}, where records from two databases,
@@
 
-The overall process of a record linkage application, where indexing is
-preformed prior to the comparison and classification steps, as
-illustrated in Fig.~\ref{fig-rl-process}, adds a further practical
-complexity. Indexing, comparison and classification are often conducted
-using algorithms selected (and their parameters set) based on technical
-and domain expertise of the user of the record linkage system, followed by a manual
-assessment (auditing) of the linkage outcomes~\cite{Chr12}. If the quality of the
-resulting links is not good enough for a certain application, the
-linkage process needs to be repeated with different parameter settings
-and possibly also alternative algorithms selected. This can result in
-a time-consuming and labor intensive iterative process. A major
-challenge is often the heuristic nature of indexing, where the choice
-of a certain indexing technique and its parameters (including which
-attributes to use in indexing) will determine the final outcome of a
-linkage.
+Systems that perform indexing prior to comparison and classification, as
+illustrated in Fig.~\ref{fig-rl-process}, add a further practical
+complexity to the process. Indexing, comparison and classification are
+often conducted using algorithms and parameters selected based on
+technical and domain expertise of the user of the record linkage system,
+followed by a manual assessment (auditing) of the linkage
+outcomes~\cite{Chr12}. If the quality of the resulting links is not good
+enough for a certain application, the linkage process needs to be
+repeated with different parameter settings and possibly also alternative
+algorithms. This can result in a time-consuming and labor-intensive
+iterative process. A major challenge is often the heuristic nature of
+indexing, where the choice of a certain indexing technique and its
+parameters (including which attributes to use in indexing) will
+determine the final outcome of a linkage.
+
 % Repeat:
 % If a certain blocking approach results in a loss of recall in the
-%final linkage results, the blocking needs to be redone, an often time
-%consuming process.
+% final linkage results, as determined by post auditing, the blocking
+% needs to be redone, an often time consuming process.
 
-In this paper we explore how existing indexing techniques can lead to
-loss in recall, even for record pairs that are highly similar and
-likely correspond to true matches. As a result, using such techniques
-can lead to \emph{incomplete} record linkage results. To overcome this
-problem, we advocate the use of \emph{metric space indexing} (MSI)
-approaches, where it is possible to guarantee that all record pairs,
-up-to a certain distance threshold, are included in the record linkage
-results. MSI also prevents the expensive detailed comparison of record
-pairs that have low similarities but that need to be compared because
-they have been inserted into the same block with traditional indexing
-techniques. Furthermore, MSI allows the combination of the indexing,
-comparison and classification steps (as traditionally conducted
-independently) into one single step, as shown in
-Fig.~\ref{fig-rl-process}, making the overall record linkage process
-both more efficient and more effective.
+Here we explore how existing indexing techniques can lead to
+reduced recall, even for record pairs that are highly similar and likely
+correspond to true matches. As a result, such techniques can lead
+to \emph{incomplete} record linkage results. To address this problem,
+\emph{metric space indexing} (MSI) can be used to guarantee that all
+record pairs, up to a certain distance threshold, are considered in the
+record linkage results. MSI also avoids the need for costly detailed
+comparison of record pairs with low similarities, which may be inserted
+into the same block with traditional indexing techniques. It allows
+indexing, comparison and classification to be combined into a
+single step, as shown in Fig.~\ref{fig-rl-process}, making the overall
+record linkage process simpler, more efficient and more effective than
+the traditional approach.
 
-% Al: but is it? be careful of last sentence in final version.
+% Sanity check this claim in final version.
 
@@
 
-~ \\
-\emph{Peter 20171111: here we need the figure showing that traditional
-blocking techniques result in incomplete record linkage, while metric
-indexing does not.}
-~ \\
-
-While MSI has been used extensively in areas such as database indexing
-and general nearest neighbor search, its use in record linkage so far
-has been limited. 
-% Al - So take it out - I recently found and cited a paper that uses R trees later in the paper?
-
-% Peter 20171112: a pretty dangerous statement. if wrong we're stuffed.. 
-% but I am not really aware of MSI approaches in RL..
-Many different MSI approaches have been developed, such as various
-metric tree structures~\cite{}. 
-% Peter 20171112: needs same text from you guys on MSI..
-In this paper, we will use the MSI techniques of M-tree and M-file (?)
-and show how their use can facilitate an improvement of both
-efficiency and effectiveness of the record linkage process.
-
-~ \\
-\emph{
-to add: something about similarity spaces / metric spaces, M-trees and its
-more recent advanced/improved versions, and how they have been used in
-similarity search.}
-
-\emph{to add: How our work is different from these other metric space works, and
-description of our contributions.
-}
-~ \\
+%to add: something about similarity spaces / metric spaces, M-trees and its more recent advanced/improved versions, and how they have been used in similarity search.}
+%to add: How our work is different from these other metric space works, and description of our contributions.
 
 \smallskip
-\textbf{Contributions:} We make three specific contributions in this
-paper: (1) We investigate how existing indexing techniques can result
-in \emph{incomplete} record linkage results, as well as in an
-in-efficient record linkage process. (2) We explore MSI for record
-linkage and propose a MSI based linkage approach that achieves
-\emph{complete}, efficient and effective record linkage. (3) We
-evaluate our approach on several real-world data sets from diverse
-domains and show its advantages over existing indexing techniques for
-record linkage.
+%Al - I have beefed this up and added a fourth contribution.
+\textbf{Contributions:} We make four specific contributions: (1) We
+investigate how existing indexing techniques can result in
+\emph{incomplete} record linkage results. (2) We show that these
+techniques are highly sensitive to parameter settings and can produce
+very low quality linkage. (3) We explore MSI for record linkage and
+propose an approach that achieves consistently \emph{complete},
+efficient and effective record linkage. (4) We evaluate our approach on
+several data sets from diverse domains and show its advantages over
+existing indexing techniques for record linkage.
 
-% old stuff below
+% TODO: check that the above claims have been addressed.
 
-\iffalse  % *******************************************************
+% *******************************************************
 % we can't write this as it reveals us in the double-blind review
-% we could ad to final paper
-The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977.
+% we can add to final paper:
+% The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977.
 This data set will, when complete, include around 14 million birth records, 11 million death records and 4 million marriage records.
 As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.
-\fi  % *********************************************************
+% *********************************************************
 
 % --------------------------------------------------------------------
-
-% 1 page maximum!
 
@@
 
-We now briefly review relevant related work in the areas of indexing
-for record linkage (for a recent survey see~\cite{Chr12b,Pap16}),
+We now briefly review relevant work in the areas of indexing
+for record linkage (for recent surveys see~\cite{Chr12b,Pap16}),
 and metric space indexing~\cite{Zezula2010}.
@@
 \textbf{Indexing for Record Linkage:}
-Techniques to link records across databases have been developed for
-over five decades~\cite{Fel69,New59}, and the scalability of linking
-has been an ongoing challenge as databases are continue to grow in
-size and complexity. Traditional blocking~\cite{Chr12b} use a single
-or a combination of attributes (known as a \emph{blocking key}) to
-insert records that have the same value(s) in a blocking key into
-the same block. Only records within the same block are then combined
-with each other. To overcome variations and misspellings, the values
-used in blocking keys can be phonetically encoded using functions
-such as Soundex, NYSIIS, or Double-Metaphone~\cite{Chr12}. These
-techniques convert a string (assumed to be a name) into a code
-according to its pronunciation, assigning the same code to similar
+Techniques to link records across datasets have been investigated for
+over five decades~\cite{Fel69,New59} with the scalability of linking
+being an ongoing challenge as datasets continue to grow in size and
+complexity. Traditional blocking~\cite{Chr12b} uses a set of attributes
+(known as a \emph{blocking key}) to insert records that share the same
+value(s) in their blocking key into the same block. Only records within
+the same block are  compared to each other. To overcome variations and
+misspellings, the values used in blocking keys can be phonetically
+encoded using functions such as Soundex, NYSIIS, or
+Double-Metaphone~\cite{Chr12}. These techniques convert a string  into a
+code according to its pronunciation, assigning the same code to similar
 sounding names (such as `Gail' and `Gayle').
@@
 A different approach to indexing is the sorted neighborhood
-method~\cite{Mon96}, where the databases to be linked are sorted
-according to a \emph{sorting key} (usually a concatenation of the
-values from several attributes). A sliding window is then moved over
-the databases and only records within the same window are being
-compared. Techniques that adaptively shrink or expand the window size based on
-the characteristics of the sorting key values have shown to improve
-both linkage efficiency and quality~\cite{Dra12}.
+method~\cite{Mon96}, where the datasets to be linked are sorted
+according to a \emph{sorting key} (usually a concatenation of the values
+from several attributes). A sliding window is then moved over the
+datasets and only records within the window are compared. Techniques
+that adaptively shrink or expand the window size based on the
+characteristics of the sorting key values have shown to improve both
+linkage efficiency and quality~\cite{Dra12} over other sliding window
+approaches.
 
-% should we mention canopy clustering? best to only briefly mention
-% techniques we do not use in our comparison
+These blocking techniques are heuristics, commonly requiring domain
+knowledge, such as the choice of appropriate blocking or sorting keys.
+Poor choices of blocking attributes result in records being inserted
+into a wrong block and thus true matches being missed. As a result,
+existing blocking techniques can lead to \emph{incomplete} linkage.
+Conversely, many of the record pairs compared in a block turn out to be
+pairs with low similarities, corresponding to non-matches, resulting in
+\emph{inefficient} linkage.
 
-Locality sensitive hashing (LSH), as originally proposed to allow
+Locality sensitive hashing (LSH), originally proposed to allow
 efficient nearest-neighbor search in high-dimensional
@@
 variations, with their main conclusion being that in order to get good
-results LSH methods must be tuned to the particular databases being
+results LSH methods must be tuned to the particular datasets being
 linked. This requires good quality ground truth data which may be
@@
 
-Existing blocking techniques have the drawbacks that they are
-heuristics and commonly require domain knowledge, such as the choice
-of blocking or sorting key used. Errors and variations in the
-attributes used for blocking also result in records being inserted
-into a wrong block and thus true matches being missed. As a result,
-existing blocking techniques can lead to \emph{incomplete} linkage,
-and that many of the record pairs compared in a block turn out to be
-pairs with low similarities that correspond to non-matches, resulting
-in \emph{inefficient} linkage.
+Metric space indexing (MSI) techniques require indexing structures to be constructed over a set of records that are then used to support comparison with another set.
+MSI data structures often support a rich set of similarity search operations over a set of values over a metric space (distances must support the triangle inequality).
+The operations supported include \textit{range search} where all objects $i$ within a distance $d$ for a query object $q$ are identified, i.e.\ $dist(i,q) \le d$) \textit{nearest-neighbour} (the closest value to q), and \textit{nearestN} (the N closest n objects to q).
+An overview of the many Metric space search techniques that are available may be found in \cite{Zezula2010}.
+In this paper we choose one metric space data structure the M-Tree \cite{paolociaccia2m} and investigate its efficacy for record linkage.
 
-% Peter until here, 20171113 ************
+An M-tree is a balanced tree structure comprised of fixed sized nodes each have a fixed number of children.
+In addition to the children, each node contains a reference to an object being indexed, a pointer and the distance to its parent, and a radius (described below).
+The radius of an internal node called a \textit{routing-node} represents the  furthest distance to any of its children.
+Thus all the children may be visualised as being contained within a ball of radius \textit{r} from the parent.
+M-trees are grown from the roots upward and require dynamic tree rebalancing when set of children become full (of size \textit{M}).
 
-\iffalse
-
-% commented, as overall related works should be around 1 page max
-
-Steorts et al.
-\cite{Steorts2014} examined various approaches to blocking records raging from traditional blocking to variations on Locality Sensitive Hashing (LSH).
-They assert that traditional blocking based on human selected fields has two major drawbacks.
-The first is that the blocks that are formed are so large that it remains computationally impractical to compare all the records in the block.
-Secondly, since similarity is based on a subset of the available fields, records in the same block may differ greatly in the fields that are not considered.
-\textbf{{this is true of our technique - explore somewhere}}.
-The paper examines two LSH variations - transitive LSH {ref needed} (TLSH) and K-Means LSH (KLSH).
-{\textbf{describe these where?}}
-Using simulated data based on people, their data of birth and postal address to evaluate the efficacy of these approaches, they demonstrate that KLSH performs as well or better than commonly used approaches to blocking such as those described \textbf{above}.
-However, they point out that, like all LSH methods, in order to get good results, it must be tuned to the particular dataset being linked.
-This in turn requires good ground truth about the dataset which may be unavailable, impossible or very expensive to obtain.
-\textbf{We examine and address these problems in this paper.}
-
-\fi
-
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-\smallskip
-\textbf{Metric Space Indexing:}
-
-% Al's writing below
-
-In \cite{Yu2016} Yu et al.
-have written an extensive survey on string similarity and join in which the problem is defined as "Given two sets of objects, similarity join aims to find all similar pairs from the two collections".
+Yu et al.
+\cite{Yu2016} have written an extensive survey on the use of string similarity and join for record linkage and define the problem as "Given two sets of objects, similarity join aims to find all similar pairs from the two collections".
 Part of the survey is concerned the algorithms used to compare two records which are characterised as being character-based, token-based or hybrid.
-The first, which includes approaches such as Levenshtein {ref needed} use character differences in the strings being compared, the second, which includes techniques such as Jaccard similarity {ref needed}, treats the records as sets of tokens and utilises set similarity, the third hybridises the two approaches.
+The first, which includes approaches such as Levenshtein~\cite{Levenshtein66}, use character differences in the strings being compared, the second, which includes techniques such as Jaccard similarity, which treat the records as sets of tokens and utilises set similarity, the third hybridises the two approaches.
 Yu's divides join algorithms (which unify records - also known as linkage) into three categories: a number of techniques based on filtering, verification based and threshold-based.
@@
 If the estimated distance is larger than the desired similarity threshold, the candidate may be discarded.
-The last category (threshold based) includes M-tree, Trie-Join, Pass-Join and Partenum.
-These approaches all require indexing structures to be constructed that are then used to support comparison.
+The last category (threshold based) includes M-trees, Trie-Join, Pass-Join and Partenum.
 
-{\textbf{do we need to describe these?}}
-
- \cite{Li2006} describes a method of performing linkage using R-trees {ref needed} that fall into the threshold based category of Yu et al.
-Their approach demonstrates that high precision and recall may be achieved by measuring Jaccard similarity over selected fields from the records being compared.
-In their evaluation , their approach achieved a recall of 99 percent in experiments in which conjunctions of similarities over fields were used.
+\cite{Li2006} describes a method of performing linkage using R-trees~\cite{Hjaltason1998} that fall into the threshold based category of Yu et al.
+Their approach demonstrated that high precision and recall may be achieved by measuring Jaccard similarity over selected fields from set of records being compared.
+In their evaluation, their approach achieved a recall of 99 percent in experiments in which conjunctions of similarities over fields were used.
 In a later paper \cite{Ciaccia97indexingmetric}  Ciaccia et al.
-demonstrate that using M-trees are almost always more efficient than R-trees.
-
-summarise related work in metric spaces as used in record linkage
-
-other related works as relevant
+demonstrate that using M-trees are almost always more efficient than R-trees and hence their use in the experiments described in this paper.
 
@@
 
-In this paper we explore the efficacy of the use of different
-similarity search algorithms in probabilistic record linkage.
-% Peter: remove probabilistic - hat we do is general to RL, the
-% probabilistic aspect is not really relevant
-We explore two inexact algorithms: LSH-minhash and MIFile \cite{amato2014mi} and one exact method: mTrees \cite{paolociaccia2m}. 
-We also use a simple brute force (nested-loop) exact technique as a 
-baseline for comparison although this approach can only be applied to 
-the smallest of our datasets. All of our experiments have a number of 
-configuration parameters with which we can configure the search space and algorithm behaviours, as we describe in Sect.~\ref{sec-exp}. All
-have a \emph{distance threshold}, $d$, below which two records are
-deemed to represent a link.
-% Peter: are classified to be link (i.e.\ refer to the same entity)
+% would be good to have a figure if possible illustrating these
+% approaches
 
+% Graham - should it be complete/incomplete rather than exact/inexact below?
+We explore the efficacy of several similarity search algorithms in
+record linkage: one popular inexact algorithm, LSH-minhash, and one exact
+method, M-tree~\cite{paolociaccia2m}. We also use a simple brute force
+(nested-loop) exact technique as a baseline for comparison, although
+this approach can only feasibly be applied to the smallest of our
+datasets. All experiments have a number of parameters with which we can
+configure the search space and algorithm behaviour, as described in
+Sect.~\ref{sec-exp}. A \emph{distance threshold}, $d$, specifies a
+maximum distance, equivalent to a minimum similarity, for two records to
+be classified as a link (i.e.\ referring to the same entity).
+
+\smallskip
+\textbf{Nested loop:}
 In the brute force approach two nested loops are used to compare every 
-record with every other non identical record. The complexity of this 
-approach is n2. If a link can be found with whatever distance metric is 
-% Peter: $O(n^2)$, where $n$ is the number of records in the dataset to
-% be linked.
-used, this approach will find it.
-% Peter: find -> identify  -- better actually clarify: This approach
-% will identify all links of records that have a distance up-to the
-% given distance threshold, $d$.
-Each of our datasets contains ground truth with which we can establish
-the records that are true links and those that are not.
+record with every other record. The complexity of this 
+approach is quadratic in the number of records to
+be linked. This approach is guaranteed to identify all links, i.e. record pairs within the given distance threshold, $d$.
 
-% peter: below fast -> I would remove fast, as it might be slow for
-% certain parameter settings
-The LSH-minhash approach is a fast but inexact method of performing similarity search. It has three tuning parameters:
+% Needed? Apart from anything else, haven't introduced the datasets yet.
+% Each of our datasets contains ground truth with which we can establish the records that are true links and those that are not.
+Clearly this approach can only be applied to relatively small datasets and we restrict its usage to the Cora dataset as a baseline for the other experiments.
+
+\smallskip
+\textbf{LSH-MinHash:}
+The LSH-minhash algorithm used has three tuning parameters:
 \emph{shingle size}, $l_{ss}$, the \emph{band size}, $l_{bs}$, and the
-\emph{number of bands}, $l_{nb}$. These are explained in the following algorithm description which follows a standard LSH approach [CITATION].
-Firstly all the fields of the record are turned into strings and 
+\emph{number of bands}, $l_{nb}$.
+These are explained in the following algorithm description which follows a standard LSH approach [CITATION].
+Using this approach, each record drawn from a set of records is placed in a \emph{LSH-MinHash} data structure.
+To do this, for each record, firstly all the fields of the record are turned into strings and 
 concatenated. Next the strings are \emph{shingled} into a set of
@@
 The original record is added to a hashmap associated with each of the keys. 
-It can be shown that the similarities between signatures generated in the second step have very close Jaccard similarity to the original string [LSH CITATION] and that if an arbitrary record is run through
-% Peter: run through.. -> reword: is converted into such a LSH
-% signature then..  records with the same key(s) are...
-% Peter: close to.. -> is a bit vague
-this process the records found in the map are close to the record being searched with good precision and recall. We will further explore these properties in this paper.
-% peter: we will evaluate the performance of LSH based blocking with
-% regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
-% Sect.~\ref{sec-exp}.
-% peter: below - we want to do record linkage (slightly different
-% from nn search) - better to describe RL below not NN search! 
-To perform a {\textit{nearest-neighbour}} lookup, the query value is hashed using the procedure described above and the union of set of of values found for all hashes in the hashmap are returned.
-If desired, these may then be filtered by performing distance calculations on the results.
+We we will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
+Sect.~\ref{sec-exp}.
+To perform linkage the following procedure is followed: for each query value is hashed using the procedure described above and the union of set of of values found for all hashes in the hashmap are returned.
+These are then be filtered by performing distance calculations on the results to eliminate false positives.
 
-The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces.
-Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
-% Peter: are these reference object also records? 
-It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances
-% Peter: unclear what you mean with: similar of distances
-from the chosen reference objects. 
-In order to create the indexes, for each value to be added to the
-% peter: value -> record ?
-data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3... based on the reference object's position from the value).
-% peter: this is unclear .. I think I understand k-NN: for each
-% record you find the k nearest reference objects and you enumerate
-% them
-Next for each score, a  mapping is created in an inverted index mapping
-% peter: unclear..? a mapping in an inverted index mapping?
-from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
-% peter: unclear: what is "the data" associated with the reference
-% object.
-In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
-% peter: this sentence reads incomplete..?
-Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
-% Peter: sorry, this is unclear as well. "the objects closest" ->
-% records closest? And why highest scores..? Is the closest record
-% to a reference object given the highest score (not the lowest
-% score - i.e. not 1-closest, 2-closest etc)
-Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
-% peter: so I guess you use the triangular inequality and the reference
-% objects to prune the search space.
-
-M-trees support exact search operations over a set of values over a metric space (distances must support triangle inequality -
-distance(i,j) + distance(j,k ) >= distance(i,k)).
-% peter: (distances must support triangle inequality, i.e.\ for three
-% objects $i$, $j$ and $k$, $dist(i,j) + dist(j,k) \ge dist(i,k)$).
-%
-In addition to supporting exact search the operations supported by M-trees are richer than the other data structures described above and include \textit{range search} (all values within distance d and some query q),
-% peter: where all objects $i$ within a distance $d$ for a query object
-% $q$ are identified, i.e.\ $dist(i,q) \le d$)
-\textit{nearest-neighbour} (the closest value to q), and \textit{nearestN} (the N closest n objects to q).
-% peter: we can add proper mathematical ways of defining those if
-% needed
-An M-tree is a balanced tree structure comprised of fixed sized nodes that can each store up to \textit{M} children.
-% not sure if we should use M here - the M in M-tree is not because
-% each node can have m children! In Sect 5 I used $t_b$ as name for
-% the branching factor.
-In addition to the children, each node contains a reference to an object being indexed, a pointer and the distance to its parent, and a radius (described below).
-The radius of an internal node called a \textit{routing-node} represents the  furthest distance to any of its children.
-Thus all the children may be visualised as being contained within a ball of radius \textit{r} from the parent.
-M-trees are grown from the roots upward and require dynamic tree rebalancing when set of children become full (of size \textit{M}).
-% Peter: and how is an M-tree used in record linkage?
-% TODO coverage: proportion of all rec pairs with a certain similarity (independent of their match status) that are being generated by a blocking  technique - what does above refer to? al
-
-% peter: could we have some pseudo-code descriptions of how these
-% techniques are used for RL blocking?
+\smallskip 
+\textbf{M-Tree:} The M-tree experiments have no parameters.
+In a similar manner to the \emph{LSH-MinHash} approach, M-trees are constructed by supplying the implementation with a set of records and a \emph{distance function}.
+The distance functions used are the same as those used for \emph{Nested-loop} and \emph{LSH-MinHash} experiments.
+To perform linkage the data-structure is searched using a \emph{range-search} query which yield all records within distance $d$ of the query object and all the returned records are directly recorded as links.
 
@@
   Data set~ & ~Number of~ & ~Number of true~ & Attributes used \\
-  name(s)  & records  & matching pairs & for blocking \\
+  name(s)  & records  & matching pairs & for traditional blocking \\
   \noalign{\smallskip} \hline \noalign{\smallskip}
@@
   Isle of Skye & 32,567 & 2,900 & ? \\
-  Kilmarnock  & 70852 & 2,900 ? why same number & ? \\
+  Kilmarnock  & 70852 & 8,300 & ? \\
   \noalign{\smallskip} \hline
@@
 
-% Peter, 31 Oct: To save space, don't use sub-sections but simply
-% \textbf{}, such as:
-% \smallskip
-%\textbf{Data sets}:~
-
 \smallskip
 \textbf{Data Sets:}
-%\subsection{Data Sets and Experimental Setup}
 %\label{sec-data}
-%
 We used four real data sets from three domains in our experiments as
@@
 \texttt{http://secondstring.sourceforge.net}}, which contains 1,295
-records that refer to 112 machine learning publications.
+records that refer to 112 machine learning publications. Ground truth is provided via a unique
+\emph{paper\_id} identifier of the form "blum1993".
 
 The second are two snapshots of the North Carolina Voter Registration
-database (\emph{NCVR})~\footnote{Available from: \texttt{http://dl.ncsbe.gov/}}, which we collected in April and June
+database (\emph{NCVR})~\footnote{Available from: \texttt{http://dl.ncsbe.gov/}}, which were collected in April and June
 2014. Each of these contains records of voters and includes their
@@
 \emph{NCID} identifier which uniquely identifies a voter. In our
-experiments we use two randomly selected sub-set of $224,073$ and
+experiments we use two randomly selected sub-sets of $224,073$ and
 $224,061$ records, respectively, where $148,036$ of those refer to
@@
 
-% Peter, 31 Oct: can the following be shortened to 1 paragraph - tick - al.
-
-\iffalse
-We also experiment over two datasets prepared by researchers pursuing previous projects \cite{reid2002} and \cite{reid2006}, which act as pseudo subsets for the demographic linkage on which our work is primarily focused .
-One data set contains records of vital events registered on the Isle of Skye, a rural district, while the other contains records from Kilmarnock, an industrial town.
-Both cover the period 1861-1901.
-These are different types of communities, with different family structures and name distributions \cite{reid2002}.
-Both datasets are encoded with the same schemata and include names, sex, addresses and the names of each person's father and mother.
-Ground truth is provided using demographer encoded identifiers which provide (incomplete) linkage between the records.
-The Isle of Skye dataset comprises around 17,600 birth records, 12,300 death records, and 2,700 marriage records.
-Relating to the experiments described below, this dataset contains 2,900 links from a birth record to the child's death record.
-The Kilmarnock data set contains around 38,400 birth records, 23,700 death records, and 8,700 marriage records with a similar schema to the Skye data.
-In this dataset demographers identified  2,900 links from a birth record to the child's death record.
-
-% Peter: Why same number of links for Skye and Killi? *****************
-\fi
-
 % - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
@@
    M-tree branching factor, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
-   shingle size, and the size and number of bands, respectively, and
-   $f_{ki}$ and $f_{ks}$ the MIfile... \emph{please add...}.}
+   shingle size, and the size and number of bands, respectively.
  \label{table-parameters}
@@
 
-% the narrow values for ki and ks need some justification!
-
 \smallskip
@@
 We implemented all techniques described in Sect.~\ref{sec-approach}
-using Java (version x.y) and ran an extensive et of experiments on
-a compute server with XX GBytes of main memory and XX TBytes of disk
-space, and running (operating system and version).
+using Java (version 1.8.0\_144) and ran an extensive et of experiments on
+a compute server with 24 Intel(R) Xeon(R) CPU E5-2630 processors running at 2.30GHz, 
+128 GBytes of main memory and running Red Hat 4.8.5-11.
 
@@
 discuss measures used. PC, PQ, num comparisons, time? etc.
+
+\subsection{Cora}
+
+Do BruteForce, LSH and M-Tree
+
+1.1 BruteForce
 
@@
 
+% ====================================================================
+
 \end{document}
@@
 % ====================================================================
+
+OLD STUFF
+
+\section{The Plan}
+
+\vspace{5mm}
+
+Plan in a bulleted list:
+
+\begin{itemize}
+\item Design: Traditionally it has two stages.
+A blocking method needs to be developed separately from a similarity method.
+With similarity search based record linkage, there is only one method, and that is the similarity method.
+\item Efficiency: I am not convinced of this, but Peter thinks there may be efficiency benefits.
+The reasoning behind this is that with blocking-based methods we would have to calculate some sort of a similarity value first.
+This value will only be used for blocking, and later a similarity value between candidate pairs will need to be calculated again.
+\item Completeness: Similarity search based methods will be complete (with respect to the given similarity method) whereas blocking based methods (including LSH) are likely to be incomplete (there will be high-similarity pairs which are not places in the same block)
+\item Data sets
+\begin{itemize}
+\item Kilmarnock Birth-Death
+\item Skye Birth-Death
+\item CORA
+\item North Carolina Voter Registration DB
+\end{itemize}
+\item The following is a list of all methods we thought we would have to compare.
+\begin{itemize}
+\item Similarity search (M-Tree, others?)
+\item LSH-blocking
+\item Traditional blocking (e.g. Region, Surname) - \textbf{why this? - this was done in the paper: A Comparison of Blocking Methods for Record Linkage}
+\item Sorted Neighbourhood blocking - \textbf{why this?}
+\end{itemize}
+\item Pairs-completeness and pairs-quality (w.r.t high similarity instead of true-link status). I will expand on this a bit more later.
+\item Comparison (for blocking based methods)
+		Precision/Recall/F1-Measure
+\item Which similarity methods are metric?
+\item Run time and memory usage comparisons
+\item Theorem. Precision might get better or worse with similarity search based methods, but recall should never get worse.
+\end{itemize}
+
+Note: The rest of this paper should be regarded as a placeholder for now.
+
+% ====================================================================
+
+\smallskip
+\textbf{MIfile:}
+The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces.
+Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
+% Peter: are these reference object also records? 
+It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances
+% Peter: unclear what you mean with: similar of distances
+from the chosen reference objects. 
+In order to create the indexes, for each value to be added to the
+% peter: value -> record ?
+data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3... based on the reference object's position from the value).
+% peter: this is unclear .. I think I understand k-NN: for each
+% record you find the k nearest reference objects and you enumerate
+% them
+Next for each score, a  mapping is created in an inverted index mapping
+% peter: unclear..? a mapping in an inverted index mapping?
+from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
+% peter: unclear: what is "the data" associated with the reference
+% object.
+In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
+% peter: this sentence reads incomplete..?
+Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
+% Peter: sorry, this is unclear as well. "the objects closest" ->
+% records closest? And why highest scores..? Is the closest record
+% to a reference object given the highest score (not the lowest
+% score - i.e. not 1-closest, 2-closest etc)
+Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
+% peter: so I guess you use the triangular inequality and the reference
+% objects to prune the search space.
+
+\if

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 matching records that refer to the same real-world entities within or
-across databases. The entities to be linked are often people (such as
-patients in hospital or customers in business databases), but record
+across datasets. The entities to be linked are often people (such as
+patients in hospital or customers in business datasets), but record
 linkage can also be applied to link consumer products or bibliographic
 records~\cite{Chr12}. Record linkage is commonly challenged by the lack
-of unique entity identifiers (keys) in the databases to be linked, which
+of unique entity identifiers (keys) in the datasets to be linked, which
 prevents the use of a database join. Instead, the linkage of records
@@
            indexing based approach (right-hand side), as described in
-           Sect.~\ref{sec-intro}, where records from two databases,
+           Sect.~\ref{sec-intro}, where records from two datasets,
            $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
@@
 
-% do we need a section on preliminaries, notation, etc?
-% ideally techniques are described formally, possibly via pseudo-code
-% algorithmic descriptions
-
 \section{Approach}
@@
 
-% peter: We need one paragraph describing the overall 'approach' /
-% methodology, as well as the aim, then a short description of
-% each method/technique used
-
-% would be good to have a figure if possible illustrating these
-% approaches
-
-% Graham - should it be complete/incomplete rather than exact/inexact below?
-We explore the efficacy of several similarity search algorithms in
-record linkage: one popular inexact algorithm, LSH-minhash, and one exact
+We explore the efficacy of several similarity search algorithms for
+record linkage: traditional blocking, one popular incomplete algorithm, LSH-minhash, and one complete
 method, M-tree~\cite{paolociaccia2m}. We also use a simple brute force
@@
 
-% Needed? Apart from anything else, haven't introduced the datasets yet.
-% Each of our datasets contains ground truth with which we can establish the records that are true links and those that are not.
-Clearly this approach can only be applied to relatively small datasets and we restrict its usage to the Cora dataset as a baseline for the other experiments.
+\smallskip
+\textbf{Traditional  blocking:}
+In the traditional blocking technique records are blocked using hand selected keys as described in \cite{Chr12b}.
+The keys are chosen with the intention of giving the best possible results for the dataset.
+Multiple blocking keys are commonly used to address the problem of missing and erroneous values preventing matches from being found.
 
@@
 Sect.~\ref{sec-exp}.
-To perform linkage the following procedure is followed: for each query value is hashed using the procedure described above and the union of set of of values found for all hashes in the hashmap are returned.
+
+To perform linkage the following procedure is followed: for each query, the value is hashed using the procedure described above and the union of the set of values found for all hashes in the hashmap are returned.
 These are then be filtered by performing distance calculations on the results to eliminate false positives.
@@
 
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-% subsection
-
-%  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-%\subsection
-
-% --------------------------------------------------------------------
-
-%  - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-%\subsection{Analysis and Limitations}
-%\label{sec-analysis}
-
-%maybe a complexity analysis? maybe an analysis of expected 
-%linkage quality?
-
 % --------------------------------------------------------------------
@@
 
-We now describe the data sets we use in our evaluation, the setup we
+We now describe the data sets we use in our evaluation, the setup
 employed to evaluate our proposed metric indexing approach and
@@
 
+%The experiments for each technique have been run within a generic Java framework which permits the actual linkage method used to be supplied.
+This ensures that the algorithms are consistently executed and that true comparisons of the their behaviour can be reported.
+It supplies each algorithm with the source records to be matched using an \emph{addSourceRecord} method which takes a record as a parameter.
+For each of the records to be matched a  \emph{getCandidateMatches} method is called which is expected to return a list of the linked records from the sources.
+Each of our datasets contains ground truth with which we can establish the records that are true links and those that are not (discussed below).
+From this ground truth the framework can report the number of links make, linkage quality etc.
+
 % - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
@@
    M-tree branching factor, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
-   shingle size, and the size and number of bands, respectively.
+   shingle size, and the size and number of bands, respectively.}
  \label{table-parameters}

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 Multiple blocking keys are commonly used to address the problem of missing and erroneous values preventing matches from being found.
+Often the choice of blocking keys requires intimate knowledge of the domain and of the records being linked.
+fucky fuck fuck.
 
@@
   \noalign{\smallskip} \hline \noalign{\smallskip}
-  CORA  & 1,295 & 17,184 & Title, authors, venue, year ? \\
-  NCVR  & ~224,073 / 224,061~ & ~148,036~ & ? \\
-  Isle of Skye & 32,567 & 2,900 & ? \\
-  Kilmarnock  & 70852 & 8,300 & ? \\
+  CORA  & 1,295 & 17,184 & many-many & publication-publication\\
+  NCVR  & ~448,134~ & ~148,036~ & 1:1  & voter-voter\\
+  Isle of Skye & 32,567 & 2,900 & 1:1 * b\\
+  Kilmarnock  & 70,852 & 8,300 & 1:1\\
   \noalign{\smallskip} \hline

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 Often the choice of blocking keys requires intimate knowledge of the domain and of the records being linked.
-fucky fuck fuck.
 
@@
 \emph{number of bands}, $l_{nb}$.
-These are explained in the following algorithm description which follows a standard LSH approach [CITATION].
+These are explained in the following algorithm description which follows a standard LSH approach \cite{Ind98}.
 Using this approach, each record drawn from a set of records is placed in a \emph{LSH-MinHash} data structure.
@@
 the signature is set to $l_{nb} \times l_{bs}$.
-Lastly the signature is divided into $l_{nb}$ segments and the elements from the band are hashed again to create a number of keys.
-The original record is added to a hashmap associated with each of the keys. 
-We we will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
-Sect.~\ref{sec-exp}.
+Lastly the signature is divided into $l_{nb}$ bands and the values  from each band are hashed again to create a number of keys.
+The original record is added to a map associated with each of the keys.
 
-To perform linkage the following procedure is followed: for each query, the value is hashed using the procedure described above and the union of the set of values found for all hashes in the hashmap are returned.
-These are then be filtered by performing distance calculations on the results to eliminate false positives.
+To perform linkage the following procedure is followed: for each query, the value is hashed using the procedure described above and the union of the set of values found for all hashes in the map are returned.
+These are then filtered by performing distance calculations on the results to eliminate false positives.
 
@@
   %\addtolength{\tabcolsep}{-0.5pt}
-  \begin{tabular}{cccc}
+  \begin{tabular}{ccccc}
   \hline\noalign{\smallskip}
-  Data set~ & ~Number of~ & ~Number of true~ & Attributes used \\
-  name(s)  & records  & matching pairs & for traditional blocking \\
+  Data set~ & ~Number of~ & ~Number of true~ & ~Linkage~ & ~Entities\\
+  name(s)  & records  & matching pairs & type & linked \\
   \noalign{\smallskip} \hline \noalign{\smallskip}
@@
   NCVR  & ~448,134~ & ~148,036~ & 1:1  & voter-voter\\
-  Isle of Skye & 32,567 & 2,900 & 1:1 * b\\
-  Kilmarnock  & 70,852 & 8,300 & 1:1\\
+  Isle of Skye & 32,567 & 2,900 & 1:1 & birth-death\\
+  Kilmarnock  & 70,852 & 8,300 & 1:1 & birth-death\\
   \noalign{\smallskip} \hline
@@
 From this ground truth the framework can report the number of links make, linkage quality etc.
+
+FFFXXXTTT
+We we will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
+Sect.~\ref{sec-exp}.
 

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 
+The general problem we address is that for two datasets A and B, we wish
+to find all the records from dataset B that match each record from dataset A.
 We explore the efficacy of several similarity search algorithms for
@@
   \hline\noalign{\smallskip}
-  Data set~ & ~Number of~ & ~Number of true~ & ~Linkage~ & ~Entities\\
-  name(s)  & records  & matching pairs & type & linked \\
+  Data set~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities\\
+  name(s)  & dataset1  & dataset2  & matching pairs & linked \\
   \noalign{\smallskip} \hline \noalign{\smallskip}
-  CORA  & 1,295 & 17,184 & many-many & publication-publication\\
-  NCVR  & ~448,134~ & ~148,036~ & 1:1  & voter-voter\\
-  Isle of Skye & 32,567 & 2,900 & 1:1 & birth-death\\
-  Kilmarnock  & 70,852 & 8,300 & 1:1 & birth-death\\
+  CORA  & 1,295 & 1295 & 17,184 & publication-publication\\
+  NCVR  & ~224,073~ & 224,061 & ~148,036~ & voter-voter\\
+  Isle of Skye & 17,612 & 12,284& 2,900 & birth-death\\
+  Kilmarnock  & 38,430 & 23,714 & 8,300 & birth-death\\
   \noalign{\smallskip} \hline
@@
 records that refer to 112 machine learning publications. Ground truth is provided via a unique
-\emph{paper\_id} identifier of the form "blum1993".
+\emph{paper\_id} identifier of the form "blum1993". In this experiment linkage is performed over the same set of records.
 
-The second are two snapshots of the North Carolina Voter Registration
+The second are two randomly selected sub-sets extracted from the North Carolina Voter Registration
 database (\emph{NCVR})~\footnote{Available from: \texttt{http://dl.ncsbe.gov/}}, which were collected in April and June
@@
 first names, surnames, and addresses. Ground truth is provided via a
-\emph{NCID} identifier which uniquely identifies a voter. In our
-experiments we use two randomly selected sub-sets of $224,073$ and
-$224,061$ records, respectively, where $148,036$ of those refer to
-voters that occurred in both original NCVR snapshots but had name
-and/or address changes over time (thus leading to around $66\%$
-matching records).
+\emph{NCID} identifier which uniquely identifies a voter.
 
@@
 
-%The experiments for each technique have been run within a generic Java framework which permits the actual linkage method used to be supplied.
-This ensures that the algorithms are consistently executed and that true comparisons of the their behaviour can be reported.
-It supplies each algorithm with the source records to be matched using an \emph{addSourceRecord} method which takes a record as a parameter.
-For each of the records to be matched a  \emph{getCandidateMatches} method is called which is expected to return a list of the linked records from the sources.
-Each of our datasets contains ground truth with which we can establish the records that are true links and those that are not (discussed below).
-From this ground truth the framework can report the number of links make, linkage quality etc.
-
-FFFXXXTTT
-We we will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
-Sect.~\ref{sec-exp}.
-
 % - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
@@
 \caption{Parameter settings used for the different data sets used in
-   the experiments,with $d$ the distance threshold used, $t_b$ the
-   M-tree branching factor, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
+   the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
    shingle size, and the size and number of bands, respectively.}
@@
   %\addtolength{\tabcolsep}{-0.5pt}
-  \begin{tabular}{cccccccc}
+  \begin{tabular}{ccccc}
   \hline\noalign{\smallskip}
-  Data set~ & $d$ & $t_b$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ & 
-    $f_{ki}$ & $f_{ks}$ \\
-  name(s) & & & & & & & \\
+  Data set~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
+  name(s) & & & &  \\
   \noalign{\smallskip} \hline \noalign{\smallskip}
-  CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$[5,10,25,50]$~ & ~$2$~ &
-    ~$[2,5,10]$~ & ~$[2,5,10]$~ & ~$[30,35,40]$~ & ~$[30,35,40]$ \\
+  CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
+    ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
   NCVR  &  \\
@@
 We implemented all techniques described in Sect.~\ref{sec-approach}
-using Java (version 1.8.0\_144) and ran an extensive et of experiments on
-a compute server with 24 Intel(R) Xeon(R) CPU E5-2630 processors running at 2.30GHz, 
+using Java (version 1.8.0\_144) and ran an extensive set of experiments on
+a compute server with 24 Intel(R) Xeon(R) processors running at 2.30GHz, 
 128 GBytes of main memory and running Red Hat 4.8.5-11.
@@
 1.1 BruteForce
+
+%the following is a sentence we culled earlier but may be needed.
+We we will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
+Sect.~\ref{sec-exp}.
 

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 common attributes, with every pair of records being classified as a
-match or non-match depending on the degree of similarity between them.
+link or non-link depending on the degree of similarity between them.
 Record linkage is usually performed in a three-step process, where first
@@
 similarity per attribute/field compared) for each pair. These similarity
-vectors are then used to classify the compared record pairs into matches
+vectors are then used to classify the compared record pairs into links
 (where it is assumed both records in a pair correspond to the same
-real-world entity) and non-matches (where the records are assumed to
+real-world entity) and non-links (where the records are assumed to
 correspond to two different entities). Various classification methods
 have been employed in record linkage~\cite{Chr12,Don15}, ranging from
-simple threshold based to sophisticated clustering, supervised
+simple threshold-based to sophisticated clustering, supervised
 classification techniques, and active learning approaches~\cite{Wan15}.
@@
 
-\smallskip
-\textbf{Indexing for Record Linkage:}
 Techniques to link records across datasets have been investigated for
-over five decades~\cite{Fel69,New59} with the scalability of linking
-being an ongoing challenge as datasets continue to grow in size and
+over five decades~\cite{Fel69,New59}, with the scalability of linking
+being an ongoing challenge as datasets grow in size and
 complexity. Traditional blocking~\cite{Chr12b} uses a set of attributes
@@
 encoded using functions such as Soundex, NYSIIS, or
-Double-Metaphone~\cite{Chr12}. These techniques convert a string  into a
+Double-Metaphone~\cite{Chr12}. These convert a string into a
 code according to its pronunciation, assigning the same code to similar
-sounding names (such as `Gail' and `Gayle').
+sounding names (such as `Gail' and `Gayle'). Multiple blocking keys may
+be used to deal with the problem of missing attribute values.
 

### 2017-11-20 scripts/plots.py
Add plotting script

@@ -1,0 +1,252 @@
+#!/usr/bin/env python3
+
+import matplotlib
+matplotlib.use("cairo")
+import matplotlib.pyplot as plt
+
+import csv, sys
+
+def loadCSV(filepath, delimiter=",", verbose=0):
+    """
+    Load a table from a given CSV/TSV file.
+
+    Input
+        filepath   : String
+        delimiter  : String
+        verbose    : Int{0..2}
+
+    Returns
+        fieldNames : [String]                   -- headings
+        values     : [Dict String String]       -- values, each entry is a dictionary with fieldNames as the keys
+    """
+    with open(filepath, newline="\n", encoding="latin-1") as f:
+        reader = csv.reader(f, delimiter=delimiter)
+        try:
+            fieldNames = [ x.strip() for x in next(reader) ]
+        except:
+            exit("Cannot read line from: %s" % filepath)
+        lenFieldNames = len(fieldNames)
+        rawValues = list()
+        for row in reader:
+            lenValues = len(row)
+            if lenFieldNames != lenValues:
+                exit("Mismatch in the number of fields, %d vs %d\nfilepath: %s" % (lenFieldNames, lenValues, filepath))
+            row2 = {}
+            allNone = True
+            for i in range(0, lenFieldNames):
+                col = row[i].strip()
+                if col == "n/e" or col == "":
+                    row2[fieldNames[i]] = None
+                else:
+                    row2[fieldNames[i]] = col
+                    allNone = False
+            # do not insert into the flatTable if all entries are None
+            if not allNone:
+                rawValues.append(row2)
+        if verbose >= 1:
+            print("Loaded data from %s, %d fields, %s entries." % (filepath, len(fieldNames), len(rawValues)))
+        if verbose >= 2:
+            print("Following are the fields.")
+            for fieldName in fieldNames:
+                print(" - %s" % fieldName)
+        return (fieldNames, rawValues)
+
+
+(fields, stats) = loadCSV("stats.tsv", delimiter="\t")
+
+for f in fields:
+    print("Field: " + f)
+
+
+################################################################################
+# Plotting Linkage Quality
+################################################################################
+
+def plotLQ(title, predicate):
+    plt.clf()
+
+    w,h = plt.figaspect(0.5)
[…]

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 Poor choices of blocking attributes result in records being inserted
-into a wrong block and thus true matches being missed. As a result,
-existing blocking techniques can lead to \emph{incomplete} linkage.
-Conversely, many of the record pairs compared in a block turn out to be
-pairs with low similarities, corresponding to non-matches, resulting in
+into inappropriate blocks, and thus true matches being missed. As a
+result, such techniques can lead to \emph{incomplete} linkage.
+Conversely, many of the record pairs compared in a block may turn
+out to have low similarity, corresponding to non-matches, resulting in
 \emph{inefficient} linkage.
@@
 efficient nearest-neighbor search in high-dimensional
-spaces~\cite{Ind98}, has been employed in record linkage as a blocking
+spaces~\cite{Ind98}, has been employed in record linkage as an indexing
 technique where attribute values are hashed multiple times, and
 blocks are created from those records that share some hash values.
-\emph{HARRA} is a record linkage approach based on Min-hash and LSH
+\emph{HARRA} [CITATION?] is a record linkage approach based on Min-hash
+[CITATION?] and LSH
 which blocks, compares, and then merges linked records in an iterative

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 
-Metric space indexing (MSI) techniques require indexing structures to be constructed over a set of records that are then used to support comparison with another set.
-MSI data structures often support a rich set of similarity search operations over a set of values over a metric space (distances must support the triangle inequality).
-The operations supported include \textit{range search} where all objects $i$ within a distance $d$ for a query object $q$ are identified, i.e.\ $dist(i,q) \le d$) \textit{nearest-neighbour} (the closest value to q), and \textit{nearestN} (the N closest n objects to q).
-An overview of the many Metric space search techniques that are available may be found in \cite{Zezula2010}.
-In this paper we choose one metric space data structure the M-Tree \cite{paolociaccia2m} and investigate its efficacy for record linkage.
+Metric space indexing (MSI) techniques provide indexing structures to
+support the comparison of records in one set with those in another,
+typically also offering similarity search operations. To create an MSI
+data structure, it is necessary to define a distance measure between
+records, with certain properties including the \emph{triangle
+inequality}. Similarity search operations include
+\textit{range-search(q,d)}, where all records within a distance $d$ of a
+query record $q$ are identified; \textit{nearest-neighbour(q)},
+returning the record with smallest distance to $q$; and
+\textit{nearest-n(q,n)}, returning the $n$ closest records to $q$.
+\cite{Zezula2010} gives an overview of MSI techniques.
+Here we choose one metric space indexing structure, the M-tree
+\cite{paolociaccia2m}, and investigate its efficacy for record linkage.
 
-An M-tree is a balanced tree structure comprised of fixed sized nodes each have a fixed number of children.
+The M-tree is a balanced tree structure comprised of fixed sized nodes each have a fixed number of children.
 In addition to the children, each node contains a reference to an object being indexed, a pointer and the distance to its parent, and a radius (described below).

### 2017-11-20 scripts/plots.py
more plotting

@@ -83,7 +83,6 @@ def plotLQ(title, predicate):
     plt.grid()
 
     data = [ line for line in stats if predicate(line) ]
-    print(len(data))
 
     plt.plot( [ line["Distance Threshold"]  for line in data ]
             , [ line["Precision"]           for line in data ]
@@ -109,7 +108,8 @@ def plotLQ(title, predicate):
     plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
 
     plt.savefig("plotLQ -- %s.png" % title, bbox_inches="tight")
-
+    plt.close()
+    print("plotLC -- %s.png" % title)
 
 
 dataset = "Cora"
@@ -119,7 +119,6 @@ linker = "Brute Force"
 title = " - ".join([dataset, linker])
 predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker
 plotLQ(title, predicate)
-print(title)
 
 
 linker = "TradBlocking"
@@ -129,14 +128,12 @@ for blockingMethod in blockingMethods:
     title = " - ".join([dataset, linker, blockingMethod])
     predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker and line["Blocking Method"] == blockingMethod
     plotLQ(title, predicate)
-    print(title)
 
 
 linker = "MTree"
 title = " - ".join([dataset, linker])
 predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker
 plotLQ(title, predicate)
-print(title)
 
 
 linker = "LSH"
@@ -150,7 +147,6 @@ for shingleSize in shingleSizes:
             title = " - ".join([dataset, linker, shingleSize, nbBands, bandSize])
             predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker and line["Shingle size"] == shingleSize and line["Number of bands"] == nbBands and line["Band size"] == bandSize
             plotLQ(title, predicate)
-            print(title)
 
 
 
@@ -161,7 +157,7 @@ for shingleSize in shingleSizes:
 # Plotting Blocking Completeness
 ################################################################################
 
-def plotBQC(title, predicate):
+def plotPQC(title, predicate):
     plt.clf()
 
     w,h = plt.figaspect(0.5)
@@ -175,20 +171,13 @@ def plotBQC(title, predicate):
     plt.suptitle(title, y = 1.02, fontsize=28)
 
[…]

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 
-The M-tree is a balanced tree structure comprised of fixed sized nodes each have a fixed number of children.
-In addition to the children, each node contains a reference to an object being indexed, a pointer and the distance to its parent, and a radius (described below).
-The radius of an internal node called a \textit{routing-node} represents the  furthest distance to any of its children.
-Thus all the children may be visualised as being contained within a ball of radius \textit{r} from the parent.
-M-trees are grown from the roots upward and require dynamic tree rebalancing when set of children become full (of size \textit{M}).
+% Do we need to spell out that the inter-node distances are the
+% distances between the corresponding indexed records?
 
-Yu et al.
-\cite{Yu2016} have written an extensive survey on the use of string similarity and join for record linkage and define the problem as "Given two sets of objects, similarity join aims to find all similar pairs from the two collections".
-Part of the survey is concerned the algorithms used to compare two records which are characterised as being character-based, token-based or hybrid.
-The first, which includes approaches such as Levenshtein~\cite{Levenshtein66}, use character differences in the strings being compared, the second, which includes techniques such as Jaccard similarity, which treat the records as sets of tokens and utilises set similarity, the third hybridises the two approaches.
-Yu's divides join algorithms (which unify records - also known as linkage) into three categories: a number of techniques based on filtering, verification based and threshold-based.
-Filtering involves generating signatures for each record and creating inverted lists containing records with common signatures and pruning records with no common signatures.
-Verification algorithms involve calculating the exact distances of a prefix of the data, and using another technique to estimate the edit distance between records.
-If the estimated distance is larger than the desired similarity threshold, the candidate may be discarded.
-The last category (threshold based) includes M-trees, Trie-Join, Pass-Join and Partenum.
+The M-tree is a dynamically balanced tree structure. Every node contains
+a reference to a record being indexed, a pointer to its parent, the
+distance to its parent, and the node's radius. The radius of a node is
+the distance from it to its furthest child. For a parent node with
+radius \textit{r}, all its children may be visualised as being contained
+within a ball of radius \textit{r} from it.
+
+\cite{Yu2016} surveys the use of string similarity in record linkage.
+Algorithms used to compare two records are characterised as being
+character-based or token-based. The first category, including approaches
+such as Levenshtein~\cite{Levenshtein66}, uses character differences in
+the strings being compared. The second, including techniques such as
+Jaccard similarity, treats records as sets of tokens, and utilises set
+similarity. There are also hybrids of the two approaches.
+% Do we need the rest of this paragraph?
+\cite{Yu2016} categorises linkage algorithms as being based on filtering,
+verification, or thresholds.
+% This is missing some context as to what is going on.
+Filtering algorithms generate signatures for each record, construct
+inverted lists containing records with common signatures, and prune
+records with no common signatures.
+% Distance of prefix to what? What is the data in this context?
+Verification algorithms calculate the exact distances of a prefix of the
+data, and use another technique to estimate the edit distance between
+records. If the estimated distance is larger than the desired similarity
+threshold, the candidate may be discarded.
+% This seems a bit throw-away.
+Threshold-based algorithms include M-trees, Trie-Join, Pass-Join and
+Partenum.
 

### 2017-11-20 paper.tex
Update on Overleaf.

@@
 blocks are created from those records that share some hash values.
-\emph{HARRA} [CITATION?] is a record linkage approach based on Min-hash
+\emph{HARRA}~\cite{Kim10} is a record linkage approach based on Min-hash
 [CITATION?] and LSH
@@
 % This seems a bit throw-away.
+% Also, these aren't themselves linkage algorithms are they?
 Threshold-based algorithms include M-trees, Trie-Join, Pass-Join and
@@
 
-\cite{Li2006} describes a method of performing linkage using R-trees~\cite{Hjaltason1998} that fall into the threshold based category of Yu et al.
-Their approach demonstrated that high precision and recall may be achieved by measuring Jaccard similarity over selected fields from set of records being compared.
-In their evaluation, their approach achieved a recall of 99 percent in experiments in which conjunctions of similarities over fields were used.
-In a later paper \cite{Ciaccia97indexingmetric}  Ciaccia et al.
-demonstrate that using M-trees are almost always more efficient than R-trees and hence their use in the experiments described in this paper.
+\cite{Li2006} describes a threshold-based method of performing linkage
+using R-trees~\cite{Hjaltason1998}. It demonstrates that high quality
+may be achieved using Jaccard similarity over selected record fields.
+% Is next sentence necessary?
+The approach achieved a recall of 99 percent in experiments in which
+conjunctions of similarities over fields were used.
+% Is this execution-time efficiency?
+A later paper \cite{Ciaccia97indexingmetric} shows that M-trees are
+almost always more efficient than R-trees, hence their use in the
+experiments described in this paper.
 

### 2017-11-21 scripts/plots.py
more plotting

@@ -70,7 +70,7 @@ def plotLQ(title, predicate):
 
     font = {"family":"sans-serif", "weight":"normal", "size":24}
     plt.rc("font", **font)
-    plt.rcParams["text.usetex"] = True
+    # plt.rcParams["text.usetex"] = True
 
     plt.title("Linkage Quality w.r.t. Threshold", fontsize=22)
     plt.suptitle(title, y = 1.02, fontsize=28)
@@ -78,11 +78,13 @@ def plotLQ(title, predicate):
     plt.xlabel("Threshold")
     plt.ylabel("Linkage Quality")
 
-    plt.xlim(0, 100)
+    data = [ line for line in stats if predicate(line) ]
+
+    plt.xlim( min([ float(line["Distance Threshold"]) for line in data ])
+            , max([ float(line["Distance Threshold"]) for line in data ])
+            )
     plt.ylim(-0.05, 1.05)
     plt.grid()
-
-    data = [ line for line in stats if predicate(line) ]
 
     plt.plot( [ line["Distance Threshold"]  for line in data ]
             , [ line["Precision"]           for line in data ]
@@ -109,11 +111,10 @@ def plotLQ(title, predicate):
 
     plt.savefig("plotLQ -- %s.png" % title, bbox_inches="tight")
     plt.close()
-    print("plotLC -- %s.png" % title)
+    print("plotLQ -- %s.png" % title)
 
 
 dataset = "Cora"
-
 
 linker = "Brute Force"
 title = " - ".join([dataset, linker])
@@ -157,6 +158,45 @@ for shingleSize in shingleSizes:
 # Plotting Blocking Completeness
 ################################################################################
 
+def plotBQ(title, predicate):
+    plt.clf()
+
+    w,h = plt.figaspect(0.5)
+    plt.figure(figsize=(w,h))
+
+    font = {"family":"sans-serif", "weight":"normal", "size":24}
+    plt.rc("font", **font)
+    # plt.rcParams["text.usetex"] = True
+
+    plt.title("Block Quality", fontsize=22)
+    plt.suptitle(title, y = 1.02, fontsize=28)
+
+    plt.xlabel("Threshold")
+    plt.ylabel("Proportion")
+
+    data = [ line for line in stats if predicate(line) ]
+
+    plt.xlim( min([ float(line["Distance Threshold"]) for line in data ])
+            , max([ float(line["Distance Threshold"]) for line in data ])
+            )
[…]

### 2017-11-21 scripts/plots.py
more plotting

@@ -118,27 +118,27 @@ dataset = "Cora"
 
 linker = "Brute Force"
 title = " - ".join([dataset, linker])
-predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker
+predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and line["Linker"] == linker
 plotLQ(title, predicate)
 
 
 linker = "TradBlocking"
-predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker
+predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and line["Linker"] == linker
 blockingMethods = set([ line["Blocking Method"] for line in stats if predicate(line) ])
 for blockingMethod in blockingMethods:
     title = " - ".join([dataset, linker, blockingMethod])
-    predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker and line["Blocking Method"] == blockingMethod
+    predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and line["Linker"] == linker and line["Blocking Method"] == blockingMethod
     plotLQ(title, predicate)
 
 
 linker = "MTree"
 title = " - ".join([dataset, linker])
-predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker
+predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and line["Linker"] == linker
 plotLQ(title, predicate)
 
 
 linker = "LSH"
-predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker
+predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and line["Linker"] == linker
 shingleSizes = set([ line["Shingle size"] for line in stats if predicate(line) ])
 nbBandss = set([ line["Number of bands"] for line in stats if predicate(line) ])
 bandSizes = set([ line["Band size"] for line in stats if predicate(line) ])
@@ -146,7 +146,49 @@ for shingleSize in shingleSizes:
     for nbBands in nbBandss:
         for bandSize in bandSizes:
             title = " - ".join([dataset, linker, shingleSize, nbBands, bandSize])
-            predicate = lambda line: line["Data Set (Source)"] == dataset and line["Linker"] == linker and line["Shingle size"] == shingleSize and line["Number of bands"] == nbBands and line["Band size"] == bandSize
[…]

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 \documentclass{llncs}
-%\usepackage{llncsdoc}
 \usepackage{epsfig}
 \usepackage{graphicx}
-
-%\usepackage{caption}
-%\usepackage{subcaption}
 
@@
 \usepackage{hyperref}
+\usepackage{cleveref}
 
-\pagestyle{empty}
+%\pagestyle{empty}
+\pagestyle{headings}
+
+% ====================================================================
+
+% TODO:
+% - clearly point out novelty of this work
+% - contribution compared to earlier work
+
+\begin{document}
+
+% ====================================================================
+
+\title{Using Metric Space Indexing for Complete and Efficient Record Linkage}
+
+% ====================================================================
+
+\author{Submitted for double-blind review}
+
+%  Hidden For Blind Review
+\iffalse
+
+\author{{\"O}zg{\"u}r Akg{\"u}n\inst{1} \and Alan Dearle\inst{1} \and\\
+Graham Kirby\inst{1} \and Peter Christen\inst{2}}
+
+\institute{School of Computer Science, University of St Andrews,\\
+St Andrews, Scotland. Contact: \email{ozgur.akgun@st-andrews.ac.uk}
+\and
+Research School of Computer Science, The Australian National University,\\
+Canberra, Australia. Contact: \email{peter.christen@anu.edu.au}}
+
+\fi
+% ====================================================================
+
+\maketitle
+
+\begin{abstract}
+
+Record linkage is the process of identifying records that refer to the
+same real-world entities, in situations where entity identifiers are
+unavailable. Records are linked on the basis of similarity between
+common attributes, with every pair of records being classified as a
+link or non-link depending on the degree of similarity between them.
+Record linkage is usually performed in a three-step process: first
+groups of similar candidate records are identified using indexing; 
+pairs within the same group are then compared in more detail, and
+finally classified.
+%
+Even state-of-the-art indexing techniques, such as Locality Sensitive
+Hashing, have potential drawbacks. They may fail to group together some
+true matching records with high similarity. Conversely, they may group
+records with low similarity, leading to high computational overhead.
+%
+We propose using metric space indexing to perform \emph{complete} record
+linkage, which results in a parameterless record linkage process
+combining indexing, comparison and classification into a single step
+delivering complete and efficient record linkage.
+
+\end{abstract}
+
+\keywords Entity resolution; data matching; similarity search;
+         blocking.
+
+% ====================================================================
+
+\section{Introduction}
+\label{sec-intro}
+
+Record linkage, also known as entity resolution, data matching and
+duplicate detection~\cite{Chr12}, is the process of identifying and
+matching records that refer to the same real-world entities within or
+across datasets. The entities to be linked are often people (such as
+patients in hospital or customers in business datasets), but record
+linkage can also be applied to link consumer products or bibliographic
+records~\cite{Chr12}. Record linkage is commonly challenged by the lack
+of unique entity identifiers (keys) in the datasets to be linked, which
+prevents the use of a database join. Instead, the linkage of records
+requires the comparison of the common attributes (or fields) that are
+available within the datasets. For datasets that contain information
+about individuals, these attributes include the names, address, dates of
+birth, and so on, of individuals.
+
+To overcome data quality issues such as typographical errors and
+variations (which are common in name and address values~\cite{Chr12}),
+approximate string comparison functions (such as edit distance, the
+Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
+compare pairs of records, leading to a vector of similarities (one
+similarity per attribute/field compared) for each pair. These similarity
+vectors are then used to classify the compared record pairs into links
+(where it is assumed both records in a pair correspond to the same
+real-world entity) and non-links (where the records are assumed to
+correspond to two different entities). Various classification methods
+have been employed in record linkage~\cite{Chr12,Don15}, ranging from
+simple threshold-based to sophisticated clustering, supervised
+classification techniques, and active learning approaches~\cite{Wan15}.
+
+Besides the issues of a lack of unique entity identifiers, and data
+quality (which will affect linkage quality), record linkage is also
+challenged by the increasing sizes of the datasets to be
+linked~\cite{Don15}. To avoid full pair-wise comparison of all possible
+record pairs (quadratic in the sizes of the datasets to be linked),
+blocking techniques, commonly known as \emph{indexing}~\cite{Chr12b},
+are used. These split the datasets into smaller blocks in a
+computationally efficient way, such that records that are likely to
+correspond to the same entity are grouped into the same block. Only
+records within the same block are then compared in more detail.
+
+While indexing techniques facilitate efficient linkage of very large
+datasets~\cite{Don15}, they generally achieve scalability at the cost
+of reduced linkage quality, because potentially true matching record
+pairs are removed in the indexing step, leading to a reduction in recall
+of the final linkage result~\cite{Chr12}. A variety of indexing
+techniques, discussed in more detail in the following section, have
+been proposed, ranging from simple phonetic based blocking~\cite{Chr12}
+and sorting of the datasets~\cite{Dra12} to locality sensitive hashing
+based techniques~\cite{Kim10,Steorts2014}. Techniques for
+unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06}
+learning of optimal blocking schemes have also been proposed.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[width=1.0\textwidth]{figures/linkage-process}
+  \caption{Overview of the steps of the traditional record linkage
+           process (left-hand side) and our proposed metric space
+           indexing based approach (right-hand side), as described in
+           Sect.~\ref{sec-intro}, where records from two datasets,
+           $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
+           \label{fig-rl-process}
+\end{figure}
+
+Systems that perform indexing prior to comparison and classification, as
+illustrated in Fig.~\ref{fig-rl-process}, add a further practical
+complexity to the process. Indexing, comparison and classification are
+often conducted using algorithms and parameters selected based on
+technical and domain expertise of the user of the record linkage system,
+followed by a manual assessment (auditing) of the linkage
+outcomes~\cite{Chr12}. If the quality of the resulting links is not good
+enough for a certain application, the linkage process needs to be
+repeated with different parameter settings and possibly also alternative
+algorithms. This can result in a time-consuming and labor-intensive
+iterative process. A major challenge is often the heuristic nature of
+indexing, where the choice of a certain indexing technique and its
+parameters (including which attributes to use in indexing) will
+determine the final outcome of a linkage.
+
+\textbf{TODO: GK rewrite this paragraph. Make the connection between completeness and high recall. Define completenes.}
+Here we explore how existing indexing techniques can lead to
+low recall, even for record pairs that are highly similar and likely
+correspond to true matches. As a result, such techniques can lead
+to \emph{incomplete} record linkage results. To address this problem,
+\emph{metric space indexing} (MSI) can be used to guarantee that all
+record pairs, up to a certain distance threshold, are considered in the
+record linkage results. MSI also avoids the need for costly detailed
+comparison of record pairs with low similarities, which may be inserted
+into the same block with traditional indexing techniques. It allows
+indexing, comparison and classification to be combined into a
+single step, as shown in Fig.~\ref{fig-rl-process}, making the overall
+record linkage process simpler, more efficient and more effective than
+the traditional approach.
+
+\textbf{Contribution:} The primary contribution of this paper is the novel application of MSI to achieve complete and efficient record linkage, without the need for complex parameter tuning.
+We evaluate our approach on several data sets from diverse domains and demonstrate its advantages over
+existing indexing techniques for record linkage.
+
+% For final paper:
+% The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977.
+This data set will, when complete, include around 14 million birth records, 11 million death records and 4 million marriage records.
+As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.
+
+% --------------------------------------------------------------------
+
+\section{Related Work}
+\label{sec-related}
+
+We now briefly review relevant work in the areas of indexing
+for record linkage (for recent surveys see~\cite{Chr12b,Pap16}),
+and metric space indexing~\cite{Zezula2010}.
+
+Techniques to link records across datasets have been investigated for
+over five decades~\cite{Fel69,New59}, with the scalability of linking
+being an ongoing challenge as datasets grow in size and complexity.
+Traditional blocking~\cite{Chr12b} uses a set of attributes (known as a
+\emph{blocking key}) to insert records that share the same value(s) in
+their blocking key into the same block. Only records within the same
+block are  compared to each other. To overcome variations and
+misspellings, the values used in blocking keys can be phonetically
+encoded using functions such as Soundex, NYSIIS, or
+Double-Metaphone~\cite{Chr12}. These convert a string into a code
+according to its pronunciation, assigning the same code to similar
+sounding names (such as `Gail' and `Gayle'). Multiple blocking keys may
+be used to deal with the problem of missing attribute values.
+
+A different approach to indexing is the sorted neighborhood
+method~\cite{Mon96}, where the datasets to be linked are sorted
+according to a \emph{sorting key} (usually a concatenation of the values
+from several attributes). A sliding window is then moved over the
+datasets and only records within the window are compared. Techniques
+that adaptively shrink or expand the window size based on the
+characteristics of the sorting key values have shown to improve both
+linkage efficiency and quality~\cite{Dra12} over other sliding window
+approaches.
+
+These blocking techniques are heuristics, commonly requiring domain
+knowledge, such as the choice of appropriate blocking or sorting keys.
+Poor choices of blocking attributes result in records being inserted
+into inappropriate blocks, and thus true matches being missed. As a
+result, such techniques can lead to \emph{incomplete} linkage.
+Conversely, many of the record pairs compared in a block may turn
+out to have low similarity, corresponding to non-matches, resulting in
+\emph{inefficient} linkage.
+
+Locality sensitive hashing (LSH), originally proposed to allow efficient
+nearest-neighbor search in high-dimensional spaces~\cite{Ind98}, has
+been employed in record linkage as an indexing technique where attribute
+values are hashed multiple times, and blocks are created from those
+records that share some hash values. \emph{HARRA}~\cite{Kim10} is a
+record linkage approach based on MinHash~\cite{Broder1997} and LSH which
+blocks, compares, and then merges linked records in an iterative
+fashion, where merged records are re-hashed to improve overall linkage
+quality. ~\cite{Steorts2014} evaluates two LSH variations, concluding
+that in order to get good results, LSH methods must be tuned to the
+particular datasets being linked. This requires good quality ground
+truth data which may be unavailable or expensive to obtain.
+
+Metric space indexing (MSI) techniques provide indexing structures to
+support the comparison of records in one set with those in another,
+typically also offering similarity search operations. To create an MSI
+data structure, it is necessary to define a distance measure between
+records, with certain properties including the \emph{triangle
+inequality}. Similarity search operations include
+\textit{range-search(q,d)}, where all records within a distance $d$ of a
+query record $q$ are identified; \textit{nearest-neighbour(q)},
+returning the record with smallest distance to $q$; and
+\textit{nearest-n(q,n)}, returning the $n$ closest records to $q$.
+\cite{Zezula2010} gives an overview of MSI techniques.
+Here we choose one metric space indexing structure, the M-tree
+\cite{paolociaccia2m}, and investigate its efficacy for record linkage.
+
+The M-tree is a dynamically balanced tree structure. Every node contains
+a reference to a record being indexed, a pointer to its parent, the
+distance to its parent, and the node's radius. The radius of a node is
+the distance from it to its furthest child. For a parent node with
+radius \textit{r}, all its children may be visualised as being contained
+within a ball of radius \textit{r} from it.
+
+\cite{Li2006} describes a threshold-based linkage method
+using R-trees~\cite{Hjaltason1998}. It demonstrates that high quality
+may be achieved using Jaccard similarity over selected record fields.
+\cite{Ciaccia97indexingmetric} shows that M-trees are
+almost always more efficient than R-trees, hence their use in the
+experiments described in this paper.
+
+% --------------------------------------------------------------------
+
+\section{Approach}
+\label{sec-approach}
+
+We address the following general record linkage problem: for two
+datasets $A$ and $B$, we wish to find, for each record in $A$, all the
+records in $B$ that match it. We compare several algorithms for record
+linkage: traditional blocking, one incomplete similarity search method,
+LSH-MinHash, and one complete method, M-tree~\cite{paolociaccia2m}. We
+also use a simple brute-force complete technique as a baseline for
+comparison, although this can only feasibly be applied to the smallest
+of our datasets. All experiments have a number of parameters to
+configure the search space and algorithm behaviour, including a
+\emph{distance threshold} specifying a maximum distance, equivalent to a
+minimum similarity, for two records to be classified as a link (i.e.\
+referring to the same entity).
+
+\textbf{Brute-force:}
+Two nested loops are used to compare every record in $A$ with every
+record in $B$. Each pair is classified as a link if the distance between
+the records is less than the threshold. This approach is guaranteed to
+identify all links, with complexity $ O(|A| \times |B|) $.
+
+\textbf{Traditional  blocking:}
+The additional parameters are the set of blocking keys and the
+normalization encodings applied to each field. These are selected as
+described in \cite{Chr12b}, exploiting knowledge of the domain and of
+the records being linked, and chosen with the intention of giving the
+best possible results for the datasets.
+%
+Each record in $A$ is placed into the appropriate block based on its
+blocking key value. The algorithm then iterates over the records in $B$,
+and for each one, compares it with each of the records from $A$ in the
+block with the same blocking key value.
+
+\textbf{LSH-MinHash:}
+The additional parameters are \emph{shingle size} ($l_{ss}$), \emph{band
+size} ($l_{bs}$) and \emph{number of bands} ($l_{nb}$). Each record in
+$A$ is placed in an \emph{LSH-MinHash} data structure. First, all the
+fields of the record are concatenated, and the result \emph{shingled}
+into a set of n-grams with $n = l_{ss}$. Next, a set of 
+deterministically generated hash functions are applied to
+each n-gram in the set
+and the smallest result (the MinHash) of each hash application is added
+to a signature for the record. The number of hashes used, and thus the
+size of the signature, is set to $l_{nb} \times l_{bs}$. Finally, the
+signature is split into $l_{nb}$ bands and the values from each band
+are hashed again to create a number of keys. The original record is
+added to a map associated with each of the keys.
+
+To perform linkage, the algorithm iterates over the records in $B$. Each
+record is hashed using the procedure described above, to obtain a set of
+keys. For each of these keys, the key is looked up in the LSH-MinHash
+data structure, and the records from $A$ associated with it added to the
+result set. Finally, the record from $B$ is compared in turn with each
+record in the result set, with the pair being classified as a link or
+non-link based on their distance.
+
+\textbf{M-Tree:} The M-tree linkage algorithm has no additional parameters.
+In a similar manner to the \emph{LSH-MinHash} approach, M-trees are constructed by supplying the implementation with a set of records and a \emph{distance function}.
+The distance functions used are the same as those used for \emph{Nested-loop} and \emph{LSH-MinHash} experiments.
+To perform linkage, the data-structure is searched using a \emph{range-search} query which yields all records within distance $d$ of the query object and all the returned records are directly recorded as links.
+
+% --------------------------------------------------------------------
+
+\section{Experiments and Results}
+\label{sec-exp}
+
+We now describe the data sets we use in our evaluation, the setup
+employed to evaluate our proposed metric indexing approach and
+compare it with traditional blocking techniques, and we then present
+and discuss our results~\footnote{The raw data, additional figures and the source code required to run these experiments can be downloaded from here. (Link removed for double blind review)}.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+% Peter 31 Oct: table style follows LNCS style guide
+
+\begin{table}[t]
+\caption{Characteristics of data sets used in the experiments.}
+ \label{table-datasets}
+  \centering
+  \begin{scriptsize}
+  %\addtolength{\tabcolsep}{-0.5pt}
+  \begin{tabular}{ccccc}
+  \hline\noalign{\smallskip}
+  Data set~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities\\
+  name(s)  & dataset1  & dataset2  & matching pairs & linked \\
+  \noalign{\smallskip} \hline \noalign{\smallskip}
+  CORA  & 1,295 & 1295 & 17,184 & publication-publication\\
+  NCVR  & ~224,073~ & 224,061 & ~148,036~ & voter-voter\\
+  Isle of Skye & 17,612 & 12,284& 2,900 & birth-death\\
+  Kilmarnock  & 38,430 & 23,714 & 8,300 & birth-death\\
+  \noalign{\smallskip} \hline
+  \end{tabular}
+  \end{scriptsize}
+\end{table}
+
+\smallskip
+\textbf{Data Sets:}
+%\label{sec-data}
+We used four data sets from three domains in our experiments, as
+summarized in Table~\ref{table-datasets}. The first is
+\emph{CORA}~\footnote{Available from:
+\texttt{http://secondstring.sourceforge.net}}, which contains 1,295
+records that refer to 112 machine learning publications. 
+It is used as a benchmark data set in the literature for assessing linkage algorithms. Ground truth is provided via a unique
+\emph{paper\_id} identifier of the form "blum1993". In this experiment linkage is performed over the same set of records.
+
+The second are two randomly selected sub-sets extracted from the North Carolina Voter Registration
+database (\emph{NCVR})~\footnote{Available from: \texttt{http://dl.ncsbe.gov/}}, which were collected in April and June
+2014. Each of these contains records of voters including their
+first names, surnames, and addresses. Ground truth is provided via a
+\emph{NCID} identifier which uniquely identifies a voter.
+
+The last two data sets are historical Scottish records of vital
+events (birth, marriages and deaths) one registered on the
+\emph{Isle of Skye}, a rural district, and the other
+records from \emph{Kilmarnock}, an industrial town. These data sets
+were provided to us by historical demographers who extensively
+curated and linked both data sets~\cite{reid2002,reid2006}. Both data
+sets include the  names, gender, addresses of individuals and their
+parents. Ground truth was generated by the demographers based on their
+extensive domain knowledge.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+\begin{table}[t]
+\caption{Parameter settings used for the different data sets used in
+   the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
+   shingle size, and the size and number of bands, respectively.}
+ \label{table-parameters}
+  \centering
+  \begin{scriptsize}
+  %\addtolength{\tabcolsep}{-0.5pt}
+  \begin{tabular}{ccccc}
+  \hline\noalign{\smallskip}
+  Data set~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
+  name(s) & & & &  \\
+  \noalign{\smallskip} \hline \noalign{\smallskip}
+  CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
+    ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
+  NCVR  &  \\
+  Isle of Skye &  \\
+  Kilmarnock  &  \\
+  \noalign{\smallskip} \hline
+  \end{tabular}
+  \end{scriptsize}
+\end{table}
+
+% \smallskip
+% \textbf{Experimental Setup:}
+% %
+% We implemented all techniques described in Sect.~\ref{sec-approach}
+% using Java (version 1.8.0\_144) and ran an extensive set of experiments on
+% a compute server with 24 Intel(R) Xeon(R) processors running at 2.30GHz, 
+% 128 GBytes of main memory and running Red Hat 4.8.5-11.
+
+The parameter settings used in our experiments are summarized in
+Table~\ref{table-parameters}.
+
+In all of our experiments we use a single distance metric which is the sum of the Levenshtein distances between the set of fields being compared.
+
+discuss measures used. PC, PQ, num comparisons, time? etc.
+
+\subsection{Cora}
+
+In this section, we perform linkage on the Cora data set using all four approaches presented in this paper: brute force, traditional blocking, LSH and M-Tree.
+The brute-force approach is used to establish the threshold value that yields the best possible linkage for the Cora dataset.
+The threshold value is varied between 0 and 250, \Cref{fig-cora-quality} presents the precision, recall, and F-measure values.
+The linkage results have high precision and low recall for low threshold values, and high recall and low precision for high threshold values.
+For this data set, the maximum F-measure value is achieved using a threshold value of 70.
+This value is data dependent; for different datasets the maximum F-measure will correspond to different thresholds.
+In the rest of this section we fix the threshold value to be 70 which maximises the F-measure for this dataset.
+
+\begin{figure}
+\begin{verbatim}
+Figure 3 will be a table of Comparisons, F1, precision recall and pairs stuff.
+12 ROWS ISH
+\end{verbatim}
+\caption{Linkage Quality using Brute Force linking on the Cora data set\label{fig-cora-quality}}
+\end{figure}
+
+\subsubsection{Traditional Blocking}
+
+Traditionally, blocking is a method for choosing candidates for linkage using manually designed criteria.
+Thanks to blocking, a distance is only computed between promising pairs of records.
+A major disadvantage of traditional blocking is the need for extra engineering effort.
+Bespoke blocking strategies need to be designed for a given data set and application domain.
+In this section we experiment with using each key field in the Cora data set as a blocking key one by one, followed by a very permissive blocking key which takes the union of all the blocks attained by using individual key fields.
+
+\emph{AL: The above needs more explanation and not the first bit?}
+
+TODO: Explain pairs-quality and pairs-completeness~\Cref{cora-traditional}.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/plotFixedThresholdTraditional2_--_Cora_-_TradBlocking_-_70}
+\caption{Blocking Quality using traditional blocking on the Cora data set\label{cora-traditional}}
+\end{figure}
+
+\Cref{cora-traditional-prf} plots the linkage quality values obtained by running traditional blocking based linkage at a threshold level of 70.
+Here, we see that high precision values are possible using blocking based linkage, however the recall (and as a result the F-measure) is poor in comparison to brute-force.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/plotFixedThresholdTraditional1_--_Cora_-_TradBlocking_-_70}
+\caption{Linkage Quality using traditional blocking on the Cora data set\label{cora-traditional-prf}}
+\end{figure}
+
+\subsection{LSH}
+
+LSH is a method for performing blocking without the engineering overhead of traditional blocking.
+It is parameterised over two integer values: number of bands, and band sizes~\footnote{LSH can also be parameterised over the shingle size, but in this paper we set the shingle size to 2.}.
+
+\subsection{M-Tree}
+
+M-Tree is a complete method that avoids brute-force search.
+In addition, it does not require any additional engineering work, it works with any distance metric.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_Brute_Force}
+\caption{Block Quality - Brute Force\label{cora-block-quality-brute}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
+\caption{Block Quality - M-Tree\label{cora-block-quality-mtree}}
+\end{figure}
+
+\subsection{Historical Demography results}
+\begin{itemize}
+\item Fig X -1 Repeat of figure 2 with precision recall comparisons etc.
+\item Fig X - Kilmarnock M tree plot Precision,recall and F
+\item Fig X + 1 - Skye M tree plot Precision,recall and F
+\item Fig X + 2 - Mtree F measure and a bunch LSH F measures.
+\end{itemize}
+
+\subsection{NCVR}
+
+% --------------------------------------------------------------------
+
+\section{Conclusions and Future Work}
+
+\textbf{Contribution:} The primary contribution of this paper is the novel application of MSI to achieve complete and efficient record linkage, without the need for complex parameter tuning.
+We evaluate our approach on several data sets from diverse domains and demonstrate its advantages over
+existing indexing techniques for record linkage.
+
+% TODO: check that the above claims have been addressed.
+
+\label{sec-concl}
+
+% --------------------------------------------------------------------
+
+\bibliographystyle{splncs03}
+\bibliography{paper.bib}
+
+% ====================================================================
+
+\end{document}
+
+% ====================================================================
+
+OLD STUFF
+
+\section{The Plan}
+
+\vspace{5mm}
+
+Plan in a bulleted list:
+
+\begin{itemize}
+\item Design: Traditionally it has two stages.
+A blocking method needs to be developed separately from a similarity method.
+With similarity search based record linkage, there is only one method, and that is the similarity method.
+\item Efficiency: I am not convinced of this, but Peter thinks there may be efficiency benefits.
+The reasoning behind this is that with blocking-based methods we would have to calculate some sort of a similarity value first.
+This value will only be used for blocking, and later a similarity value between candidate pairs will need to be calculated again.
+\item Completeness: Similarity search based methods will be complete (with respect to the given similarity method) whereas blocking based methods (including LSH) are likely to be incomplete (there will be high-similarity pairs which are not places in the same block)
+\item Data sets
+\begin{itemize}
+\item Kilmarnock Birth-Death
+\item Skye Birth-Death
+\item CORA
+\item North Carolina Voter Registration DB
+\end{itemize}
+\item The following is a list of all methods we thought we would have to compare.
+\begin{itemize}
+\item Similarity search (M-Tree, others?)
+\item LSH-blocking
+\item Traditional blocking (e.g. Region, Surname) - \textbf{why this? - this was done in the paper: A Comparison of Blocking Methods for Record Linkage}
+\item Sorted Neighbourhood blocking - \textbf{why this?}
+\end{itemize}
+\item Pairs-completeness and pairs-quality (w.r.t high similarity instead of true-link status). I will expand on this a bit more later.
+\item Comparison (for blocking based methods)
+		Precision/Recall/F1-Measure
+\item Which similarity methods are metric?
+\item Run time and memory usage comparisons
+\item Theorem. Precision might get better or worse with similarity search based methods, but recall should never get worse.
+\end{itemize}
+
+Note: The rest of this paper should be regarded as a placeholder for now.
+
+% ====================================================================
+
+\smallskip
+\textbf{MIfile:}
+The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces.
+Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
+% Peter: are these reference object also records? 
+It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances
+% Peter: unclear what you mean with: similar of distances
+from the chosen reference objects. 
+In order to create the indexes, for each value to be added to the
+% peter: value -> record ?
+data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3... based on the reference object's position from the value).
+% peter: this is unclear .. I think I understand k-NN: for each
+% record you find the k nearest reference objects and you enumerate
+% them
+Next for each score, a  mapping is created in an inverted index mapping
+% peter: unclear..? a mapping in an inverted index mapping?
+from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
+% peter: unclear: what is "the data" associated with the reference
+% object.
+In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
+% peter: this sentence reads incomplete..?
+Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
+% Peter: sorry, this is unclear as well. "the objects closest" ->
+% records closest? And why highest scores..? Is the closest record
+% to a reference object given the highest score (not the lowest
+% score - i.e. not 1-closest, 2-closest etc)
+Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
+% peter: so I guess you use the triangular inequality and the reference
+% objects to prune the search space.
+
+\cite{Yu2016} surveys the use of string similarity in record linkage.
+Algorithms used to compare two records are characterised as being
+character-based or token-based. The first category, including approaches
+such as Levenshtein~\cite{Levenshtein66}, uses character differences in
+the strings being compared. The second, including techniques such as
+Jaccard similarity, treats records as sets of tokens, and utilises set
+similarity. There are also hybrids of the two approaches.
+% Do we need the rest of this paragraph?
+\cite{Yu2016} categorises linkage algorithms as being based on filtering,
+verification, or thresholds.
+% This is missing some context as to what is going on.
+Filtering algorithms generate signatures for each record, construct
+inverted lists containing records with common signatures, and prune
+records with no common signatures.
+% Distance of prefix to what? What is the data in this context?
+Verification algorithms calculate the exact distances of a prefix of the
+data, and use another technique to estimate the edit distance between
+records. If the estimated distance is larger than the desired similarity
+threshold, the candidate may be discarded.
+% This seems a bit throw-away.
+% Also, these aren't themselves linkage algorithms are they?
+Threshold-based algorithms include M-trees, Trie-Join, Pass-Join and
+Partenum.
+
+%the following is a sentence we culled earlier but may be needed.
+We will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
+Sect.~\ref{sec-exp}.
+
+\begin{verbatim}
+for each record a in A
+    for each record b in B
+        classify (a,b) as link if distance(a,b) < threshold
+\end{verbatim}
 
@@
 % 15.General carelessness
-
-% ====================================================================
-
-% TODO:
-% - clearly point out novelty of this work
-% - contribution compared to earlier work
-
-\begin{document}
-
-% ====================================================================
-% Previous title versions
-\iffalse
-
-% Peter 31 Oct: Current title does not really reflect the topic I think:
-%\title{Exploring approaches to probabilistic record linkage using similarity searching}
-% Peter 31 Oct: What about
-\title{Efficient and Effective Metric Space Indexing for Complete Record Linkage}
-
-\fi
-% ====================================================================
-
-\title{Using Metric Space Indexing for Efficient Complete Record Linkage}
-
-\author{Submitted for double-blind review}
-
-% ====================================================================
-%  Hidden For Blind Review
-\iffalse
-
-\author{{\"O}zg{\"u}r Akg{\"u}n\inst{1} \and Alan Dearle\inst{1} \and Graham Kirby\inst{1}
-\and Peter Christen\inst{2}}
-
-\institute{School of Computer Science, University of St Andrews, St Andrews, Scotland.\\
-Contact: \email{ozgur.akgun@st-andrews.ac.uk}
-\and
-Research School of Computer Science, The Australian National University, Canberra, Australia.\\
-Contact: \email{peter.christen@anu.edu.au}}
-
-\fi
-% ====================================================================
-
-\maketitle
-
-\begin{abstract}
-%Probabilistic
-Record linkage is the process of identifying records that refer to the
-same real-world entities, in situations where entity identifiers are
-unavailable. Records are linked on the basis of similarity between
-common attributes, with every pair of records being classified as a
-link or non-link depending on the degree of similarity between them.
-Record linkage is usually performed in a three-step process, where first
-groups of similar candidate records are identified using indexing. Only
-pairs within the same group are then compared in more detail, and
-finally classified.
-%
-Even state-of-the-art indexing techniques, such as Locality Sensitive
-Hashing, have potential drawbacks. They may fail to group together some
-true matching records with high similarity. Conversely, they may group
-records with low similarity, leading to high computational overhead.
-%
-We propose using metric space indexing to perform \emph{complete} record
-linkage, which results in a parameterless record linkage process
-combining indexing, comparison and classification into a single step
-delivering high effectiveness and efficiency.
-
-\end{abstract}
-
-\keywords Entity resolution; data matching; similarity search;
-         blocking.
-
-% ====================================================================
-
-% 1 1/2 pages maximum!
-
-\section{Introduction}
-\label{sec-intro}
-
-% Peter, 20171112:
-
-Record linkage, also known as entity resolution, data matching and
-duplicate detection~\cite{Chr12}, is the process of identifying and
-matching records that refer to the same real-world entities within or
-across datasets. The entities to be linked are often people (such as
-patients in hospital or customers in business datasets), but record
-linkage can also be applied to link consumer products or bibliographic
-records~\cite{Chr12}. Record linkage is commonly challenged by the lack
-of unique entity identifiers (keys) in the datasets to be linked, which
-prevents the use of a database join. Instead, the linkage of records
-requires the comparison of the common attributes (or fields) that are available
-in the datasets to be linked. For datasets that contain
-information about individuals, these attributes include the names,
-address, dates of birth, and so on, of individuals.
-
-To overcome data quality issues such as typographical errors and
-variations (which are common in name and address values~\cite{Chr12}),
-approximate string comparison functions (such as edit distance, the
-Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
-compare pairs of records, leading to a vector of similarities (one
-similarity per attribute/field compared) for each pair. These similarity
-vectors are then used to classify the compared record pairs into links
-(where it is assumed both records in a pair correspond to the same
-real-world entity) and non-links (where the records are assumed to
-correspond to two different entities). Various classification methods
-have been employed in record linkage~\cite{Chr12,Don15}, ranging from
-simple threshold-based to sophisticated clustering, supervised
-classification techniques, and active learning approaches~\cite{Wan15}.
-
-Besides the issues of a lack of unique entity identifiers, and data
-quality (which will affect linkage quality), record linkage is also
-challenged by the increasing sizes of the datasets to be
-linked~\cite{Don15}. To avoid full pair-wise comparison of all possible
-record pairs (quadratic in the sizes of the datasets to be linked),
-blocking techniques, commonly known as \emph{indexing}~\cite{Chr12b},
-are used. These split the datasets into smaller blocks in a
-computationally efficient way, such that records that are likely to
-correspond to the same entity are grouped into the same block. Only
-records within the same block are then compared in more detail.
-
-While indexing techniques facilitate efficient linkage of very large
-datasets~\cite{Don15}, they generally achieve scalability at the cost
-of reduced linkage quality, because potentially true matching record
-pairs are removed in the indexing step, leading to a reduction in recall
-of the final linkage result~\cite{Chr12}. A variety of indexing
-techniques, discussed in more detail in the following section, have
-been proposed, ranging from simple phonetic based blocking~\cite{Chr12}
-and sorting of the datasets~\cite{Dra12} to locality sensitive hashing
-based techniques~\cite{Kim10,Steorts2014}. Techniques for
-unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06}
-learning of optimal blocking schemes have also been proposed.
-
-\begin{figure}[!t]
-  \centering
-  %\scalebox{1.0}[1.0]
-  \includegraphics[width=1.0\textwidth]{figures/linkage-process}
-  \caption{Overview of the steps of the traditional record linkage
-           process (left-hand side) and our proposed metric space
-           indexing based approach (right-hand side), as described in
-           Sect.~\ref{sec-intro}, where records from two datasets,
-           $\mathbf{D}_A$ and $\mathbf{D}_B$, are being linked.}
-           \label{fig-rl-process}
-\end{figure}
-
-Systems that perform indexing prior to comparison and classification, as
-illustrated in Fig.~\ref{fig-rl-process}, add a further practical
-complexity to the process. Indexing, comparison and classification are
-often conducted using algorithms and parameters selected based on
-technical and domain expertise of the user of the record linkage system,
-followed by a manual assessment (auditing) of the linkage
-outcomes~\cite{Chr12}. If the quality of the resulting links is not good
-enough for a certain application, the linkage process needs to be
-repeated with different parameter settings and possibly also alternative
-algorithms. This can result in a time-consuming and labor-intensive
-iterative process. A major challenge is often the heuristic nature of
-indexing, where the choice of a certain indexing technique and its
-parameters (including which attributes to use in indexing) will
-determine the final outcome of a linkage.
-
-% Repeat:
-% If a certain blocking approach results in a loss of recall in the
-% final linkage results, as determined by post auditing, the blocking
-% needs to be redone, an often time consuming process.
-
-Here we explore how existing indexing techniques can lead to
-reduced recall, even for record pairs that are highly similar and likely
-correspond to true matches. As a result, such techniques can lead
-to \emph{incomplete} record linkage results. To address this problem,
-\emph{metric space indexing} (MSI) can be used to guarantee that all
-record pairs, up to a certain distance threshold, are considered in the
-record linkage results. MSI also avoids the need for costly detailed
-comparison of record pairs with low similarities, which may be inserted
-into the same block with traditional indexing techniques. It allows
-indexing, comparison and classification to be combined into a
-single step, as shown in Fig.~\ref{fig-rl-process}, making the overall
-record linkage process simpler, more efficient and more effective than
-the traditional approach.
-
-% Sanity check this claim in final version.
-
-% Peter 20171112: do we also need to mention we do not waste
-% comparisons of record pairs with low similarities (i.e. those with
-% a distance > max_dist)?
-
-%to add: something about similarity spaces / metric spaces, M-trees and its more recent advanced/improved versions, and how they have been used in similarity search.}
-%to add: How our work is different from these other metric space works, and description of our contributions.
-
-\smallskip
-%Al - I have beefed this up and added a fourth contribution.
-\textbf{Contributions:} We make four specific contributions: (1) We
-investigate how existing indexing techniques can result in
-\emph{incomplete} record linkage results. (2) We show that these
-techniques are highly sensitive to parameter settings and can produce
-very low quality linkage. (3) We explore MSI for record linkage and
-propose an approach that achieves consistently \emph{complete},
-efficient and effective record linkage. (4) We evaluate our approach on
-several data sets from diverse domains and show its advantages over
-existing indexing techniques for record linkage.
-
-% TODO: check that the above claims have been addressed.
-
-% *******************************************************
-% we can't write this as it reveals us in the double-blind review
-% we can add to final paper:
-% The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977.
-This data set will, when complete, include around 14 million birth records, 11 million death records and 4 million marriage records.
-As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.
-% *********************************************************
-
-% --------------------------------------------------------------------
-
-\section{Related Work}
-\label{sec-related}
-
-We now briefly review relevant work in the areas of indexing
-for record linkage (for recent surveys see~\cite{Chr12b,Pap16}),
-and metric space indexing~\cite{Zezula2010}.
-
-Techniques to link records across datasets have been investigated for
-over five decades~\cite{Fel69,New59}, with the scalability of linking
-being an ongoing challenge as datasets grow in size and
-complexity. Traditional blocking~\cite{Chr12b} uses a set of attributes
-(known as a \emph{blocking key}) to insert records that share the same
-value(s) in their blocking key into the same block. Only records within
-the same block are  compared to each other. To overcome variations and
-misspellings, the values used in blocking keys can be phonetically
-encoded using functions such as Soundex, NYSIIS, or
-Double-Metaphone~\cite{Chr12}. These convert a string into a
-code according to its pronunciation, assigning the same code to similar
-sounding names (such as `Gail' and `Gayle'). Multiple blocking keys may
-be used to deal with the problem of missing attribute values.
-
-A different approach to indexing is the sorted neighborhood
-method~\cite{Mon96}, where the datasets to be linked are sorted
-according to a \emph{sorting key} (usually a concatenation of the values
-from several attributes). A sliding window is then moved over the
-datasets and only records within the window are compared. Techniques
-that adaptively shrink or expand the window size based on the
-characteristics of the sorting key values have shown to improve both
-linkage efficiency and quality~\cite{Dra12} over other sliding window
-approaches.
-
-These blocking techniques are heuristics, commonly requiring domain
-knowledge, such as the choice of appropriate blocking or sorting keys.
-Poor choices of blocking attributes result in records being inserted
-into inappropriate blocks, and thus true matches being missed. As a
-result, such techniques can lead to \emph{incomplete} linkage.
-Conversely, many of the record pairs compared in a block may turn
-out to have low similarity, corresponding to non-matches, resulting in
-\emph{inefficient} linkage.
-
-Locality sensitive hashing (LSH), originally proposed to allow
-efficient nearest-neighbor search in high-dimensional
-spaces~\cite{Ind98}, has been employed in record linkage as an indexing
-technique where attribute values are hashed multiple times, and
-blocks are created from those records that share some hash values.
-\emph{HARRA}~\cite{Kim10} is a record linkage approach based on Min-hash
-[CITATION?] and LSH
-which blocks, compares, and then merges linked records in an iterative
-fashion, where merged records are re-hashed to improve overall linkage
-quality. Steorts et al.~\cite{Steorts2014} recently evaluated two LSH
-variations, with their main conclusion being that in order to get good
-results LSH methods must be tuned to the particular datasets being
-linked. This requires good quality ground truth data which may be
-unavailable or very expensive to obtain.
-
-Metric space indexing (MSI) techniques provide indexing structures to
-support the comparison of records in one set with those in another,
-typically also offering similarity search operations. To create an MSI
-data structure, it is necessary to define a distance measure between
-records, with certain properties including the \emph{triangle
-inequality}. Similarity search operations include
-\textit{range-search(q,d)}, where all records within a distance $d$ of a
-query record $q$ are identified; \textit{nearest-neighbour(q)},
-returning the record with smallest distance to $q$; and
-\textit{nearest-n(q,n)}, returning the $n$ closest records to $q$.
-\cite{Zezula2010} gives an overview of MSI techniques.
-Here we choose one metric space indexing structure, the M-tree
-\cite{paolociaccia2m}, and investigate its efficacy for record linkage.
-
-% Do we need to spell out that the inter-node distances are the
-% distances between the corresponding indexed records?
-
-The M-tree is a dynamically balanced tree structure. Every node contains
-a reference to a record being indexed, a pointer to its parent, the
-distance to its parent, and the node's radius. The radius of a node is
-the distance from it to its furthest child. For a parent node with
-radius \textit{r}, all its children may be visualised as being contained
-within a ball of radius \textit{r} from it.
-
-\cite{Yu2016} surveys the use of string similarity in record linkage.
-Algorithms used to compare two records are characterised as being
-character-based or token-based. The first category, including approaches
-such as Levenshtein~\cite{Levenshtein66}, uses character differences in
-the strings being compared. The second, including techniques such as
-Jaccard similarity, treats records as sets of tokens, and utilises set
-similarity. There are also hybrids of the two approaches.
-% Do we need the rest of this paragraph?
-\cite{Yu2016} categorises linkage algorithms as being based on filtering,
-verification, or thresholds.
-% This is missing some context as to what is going on.
-Filtering algorithms generate signatures for each record, construct
-inverted lists containing records with common signatures, and prune
-records with no common signatures.
-% Distance of prefix to what? What is the data in this context?
-Verification algorithms calculate the exact distances of a prefix of the
-data, and use another technique to estimate the edit distance between
-records. If the estimated distance is larger than the desired similarity
-threshold, the candidate may be discarded.
-% This seems a bit throw-away.
-% Also, these aren't themselves linkage algorithms are they?
-Threshold-based algorithms include M-trees, Trie-Join, Pass-Join and
-Partenum.
-
-\cite{Li2006} describes a threshold-based method of performing linkage
-using R-trees~\cite{Hjaltason1998}. It demonstrates that high quality
-may be achieved using Jaccard similarity over selected record fields.
-% Is next sentence necessary?
-The approach achieved a recall of 99 percent in experiments in which
-conjunctions of similarities over fields were used.
-% Is this execution-time efficiency?
-A later paper \cite{Ciaccia97indexingmetric} shows that M-trees are
-almost always more efficient than R-trees, hence their use in the
-experiments described in this paper.
-
-% end of related work should be around end of page 3!
-
-% --------------------------------------------------------------------
-
-\section{Approach}
-\label{sec-approach}
-
-The general problem we address is that for two datasets A and B, we wish
-to find all the records from dataset B that match each record from dataset A.
-We explore the efficacy of several similarity search algorithms for
-record linkage: traditional blocking, one popular incomplete algorithm, LSH-minhash, and one complete
-method, M-tree~\cite{paolociaccia2m}. We also use a simple brute force
-(nested-loop) exact technique as a baseline for comparison, although
-this approach can only feasibly be applied to the smallest of our
-datasets. All experiments have a number of parameters with which we can
-configure the search space and algorithm behaviour, as described in
-Sect.~\ref{sec-exp}. A \emph{distance threshold}, $d$, specifies a
-maximum distance, equivalent to a minimum similarity, for two records to
-be classified as a link (i.e.\ referring to the same entity).
-
-\smallskip
-\textbf{Nested loop:}
-In the brute force approach two nested loops are used to compare every 
-record with every other record. The complexity of this 
-approach is quadratic in the number of records to
-be linked. This approach is guaranteed to identify all links, i.e. record pairs within the given distance threshold, $d$.
-
-\smallskip
-\textbf{Traditional  blocking:}
-In the traditional blocking technique records are blocked using hand selected keys as described in \cite{Chr12b}.
-The keys are chosen with the intention of giving the best possible results for the dataset.
-Multiple blocking keys are commonly used to address the problem of missing and erroneous values preventing matches from being found.
-Often the choice of blocking keys requires intimate knowledge of the domain and of the records being linked.
-
-\smallskip
-\textbf{LSH-MinHash:}
-The LSH-minhash algorithm used has three tuning parameters:
-\emph{shingle size}, $l_{ss}$, the \emph{band size}, $l_{bs}$, and the
-\emph{number of bands}, $l_{nb}$.
-These are explained in the following algorithm description which follows a standard LSH approach \cite{Ind98}.
-Using this approach, each record drawn from a set of records is placed in a \emph{LSH-MinHash} data structure.
-To do this, for each record, firstly all the fields of the record are turned into strings and 
-concatenated. Next the strings are \emph{shingled} into a set of
-n-grams (sub-strings of length $n$) according to the $l_{ss}$
-parameter (i.e.\ $n = l_{ss}$). Next a set of randomly, 
-deterministically generated hash functions are applied to each of the
-sets of n-grams and the smallest (the minhash) of each application is
-added to a signature. The number of hashes used and thus the size of
-the signature is set to $l_{nb} \times l_{bs}$.
-Lastly the signature is divided into $l_{nb}$ bands and the values  from each band are hashed again to create a number of keys.
-The original record is added to a map associated with each of the keys.
-
-To perform linkage the following procedure is followed: for each query, the value is hashed using the procedure described above and the union of the set of values found for all hashes in the map are returned.
-These are then filtered by performing distance calculations on the results to eliminate false positives.
-
-\smallskip 
-\textbf{M-Tree:} The M-tree experiments have no parameters.
-In a similar manner to the \emph{LSH-MinHash} approach, M-trees are constructed by supplying the implementation with a set of records and a \emph{distance function}.
-The distance functions used are the same as those used for \emph{Nested-loop} and \emph{LSH-MinHash} experiments.
-To perform linkage the data-structure is searched using a \emph{range-search} query which yield all records within distance $d$ of the query object and all the returned records are directly recorded as links.
-
-% --------------------------------------------------------------------
-
-\section{Experiments and Results}
-\label{sec-exp}
-
-We now describe the data sets we use in our evaluation, the setup
-employed to evaluate our proposed metric indexing approach and
-compare it with traditional blocking techniques, and we then present
-and discuss our results.
-
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-% Peter 31 Oct: table style follows LNCS style guide
-
-\begin{table}[t]
-\caption{Characteristics of data sets used in the experiments.}
- \label{table-datasets}
-  \centering
-  \begin{scriptsize}
-  %\addtolength{\tabcolsep}{-0.5pt}
-  \begin{tabular}{ccccc}
-  \hline\noalign{\smallskip}
-  Data set~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities\\
-  name(s)  & dataset1  & dataset2  & matching pairs & linked \\
-  \noalign{\smallskip} \hline \noalign{\smallskip}
-  CORA  & 1,295 & 1295 & 17,184 & publication-publication\\
-  NCVR  & ~224,073~ & 224,061 & ~148,036~ & voter-voter\\
-  Isle of Skye & 17,612 & 12,284& 2,900 & birth-death\\
-  Kilmarnock  & 38,430 & 23,714 & 8,300 & birth-death\\
-  \noalign{\smallskip} \hline
-  \end{tabular}
-  \end{scriptsize}
-\end{table}
-
-\smallskip
-\textbf{Data Sets:}
-%\label{sec-data}
-We used four real data sets from three domains in our experiments as
-summarized in Table~\ref{table-datasets}. The first is
-\emph{CORA}~\footnote{Available from:
-\texttt{http://secondstring.sourceforge.net}}, which contains 1,295
-records that refer to 112 machine learning publications. Ground truth is provided via a unique
-\emph{paper\_id} identifier of the form "blum1993". In this experiment linkage is performed over the same set of records.
-
-The second are two randomly selected sub-sets extracted from the North Carolina Voter Registration
-database (\emph{NCVR})~\footnote{Available from: \texttt{http://dl.ncsbe.gov/}}, which were collected in April and June
-2014. Each of these contains records of voters and includes their
-first names, surnames, and addresses. Ground truth is provided via a
-\emph{NCID} identifier which uniquely identifies a voter.
-
-The last two data sets are historical Scottish records of vital
-events (birth, marriages and deaths) registered on the
-\emph{Isle of Skye}, a rural district, while the other contains
-records from \emph{Kilmarnock}, an industrial town. These data sets
-were provided to us by the demographer and historians who extensively
-curated and linked both data sets~\cite{reid2002,reid2006}. Both data
-sets include the  names, gender, addresses of individuals and their
-parents. Ground truth was generated by the demographers based on their
-extensive domain knowledge.
-
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-\begin{table}[t]
-\caption{Parameter settings used for the different data sets used in
-   the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
-   shingle size, and the size and number of bands, respectively.}
- \label{table-parameters}
-  \centering
-  \begin{scriptsize}
-  %\addtolength{\tabcolsep}{-0.5pt}
-  \begin{tabular}{ccccc}
-  \hline\noalign{\smallskip}
-  Data set~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
-  name(s) & & & &  \\
-  \noalign{\smallskip} \hline \noalign{\smallskip}
-  CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
-    ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
-  NCVR  &  \\
-  Isle of Skye &  \\
-  Kilmarnock  &  \\
-  \noalign{\smallskip} \hline
-  \end{tabular}
-  \end{scriptsize}
-\end{table}
-
-\smallskip
-\textbf{Experimental Setup:}
-%
-We implemented all techniques described in Sect.~\ref{sec-approach}
-using Java (version 1.8.0\_144) and ran an extensive set of experiments on
-a compute server with 24 Intel(R) Xeon(R) processors running at 2.30GHz, 
-128 GBytes of main memory and running Red Hat 4.8.5-11.
-
-The parameter settings used in our experiments are summarized in
-Table~\ref{table-parameters}. ... \emph{describe and justify selected
-values..}
-
-discuss measures used. PC, PQ, num comparisons, time? etc.
-
-\subsection{Cora}
-
-Do BruteForce, LSH and M-Tree
-
-1.1 BruteForce
-
-%the following is a sentence we culled earlier but may be needed.
-We we will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
-Sect.~\ref{sec-exp}.
-
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-\smallskip
-\textbf{Results and Discussion}
-
-%\subsection{Results and Discussion}
-%\label{sec-results}
-
-table or plots?
-
-% --------------------------------------------------------------------
-
-\section{Conclusions and Future Work}
-\label{sec-concl}
-
-% --------------------------------------------------------------------
-
-\bibliographystyle{splncs03}
-%\bibliographystyle{abbrv}
-\bibliography{paper.bib}
-
-% ====================================================================
-
-\end{document}
-
-% ====================================================================
-
-OLD STUFF
-
-\section{The Plan}
-
-\vspace{5mm}
-
-Plan in a bulleted list:
-
-\begin{itemize}
-\item Design: Traditionally it has two stages.
-A blocking method needs to be developed separately from a similarity method.
-With similarity search based record linkage, there is only one method, and that is the similarity method.
-\item Efficiency: I am not convinced of this, but Peter thinks there may be efficiency benefits.
-The reasoning behind this is that with blocking-based methods we would have to calculate some sort of a similarity value first.
-This value will only be used for blocking, and later a similarity value between candidate pairs will need to be calculated again.
-\item Completeness: Similarity search based methods will be complete (with respect to the given similarity method) whereas blocking based methods (including LSH) are likely to be incomplete (there will be high-similarity pairs which are not places in the same block)
-\item Data sets
-\begin{itemize}
-\item Kilmarnock Birth-Death
-\item Skye Birth-Death
-\item CORA
-\item North Carolina Voter Registration DB
-\end{itemize}
-\item The following is a list of all methods we thought we would have to compare.
-\begin{itemize}
-\item Similarity search (M-Tree, others?)
-\item LSH-blocking
-\item Traditional blocking (e.g. Region, Surname) - \textbf{why this? - this was done in the paper: A Comparison of Blocking Methods for Record Linkage}
-\item Sorted Neighbourhood blocking - \textbf{why this?}
-\end{itemize}
-\item Pairs-completeness and pairs-quality (w.r.t high similarity instead of true-link status). I will expand on this a bit more later.
-\item Comparison (for blocking based methods)
-		Precision/Recall/F1-Measure
-\item Which similarity methods are metric?
-\item Run time and memory usage comparisons
-\item Theorem. Precision might get better or worse with similarity search based methods, but recall should never get worse.
-\end{itemize}
-
-Note: The rest of this paper should be regarded as a placeholder for now.
-
-% ====================================================================
-
-\smallskip
-\textbf{MIfile:}
-The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces.
-Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
-% Peter: are these reference object also records? 
-It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances
-% Peter: unclear what you mean with: similar of distances
-from the chosen reference objects. 
-In order to create the indexes, for each value to be added to the
-% peter: value -> record ?
-data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3... based on the reference object's position from the value).
-% peter: this is unclear .. I think I understand k-NN: for each
-% record you find the k nearest reference objects and you enumerate
-% them
-Next for each score, a  mapping is created in an inverted index mapping
-% peter: unclear..? a mapping in an inverted index mapping?
-from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
-% peter: unclear: what is "the data" associated with the reference
-% object.
-In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
-% peter: this sentence reads incomplete..?
-Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
-% Peter: sorry, this is unclear as well. "the objects closest" ->
-% records closest? And why highest scores..? Is the closest record
-% to a reference object given the highest score (not the lowest
-% score - i.e. not 1-closest, 2-closest etc)
-Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
-% peter: so I guess you use the triangular inequality and the reference
-% objects to prune the search space.
-
-\if

### 2017-11-21 scripts/plots.py
Rename Threshold to Distance Threshold

@@ -75,7 +75,7 @@ def plotLQ(title, predicate):
     plt.title("Linkage Quality w.r.t. Threshold", fontsize=22)
     plt.suptitle(title, y = 1.02, fontsize=28)
 
-    plt.xlabel("Threshold")
+    plt.xlabel("Distance Threshold")
     plt.ylabel("Linkage Quality")
 
     data = [ line for line in stats if predicate(line) ]
@@ -213,7 +213,7 @@ def plotBQ(title, predicate):
     plt.title("Block Quality", fontsize=22)
     plt.suptitle(title, y = 1.02, fontsize=28)
 
-    plt.xlabel("Threshold")
+    plt.xlabel("Distance Threshold")
     plt.ylabel("Proportion")
 
     data = [ line for line in stats if predicate(line) ]
@@ -252,7 +252,7 @@ def plotPQC(title, predicate):
     plt.title("Pairs Quality and Completeness", fontsize=22)
     plt.suptitle(title, y = 1.02, fontsize=28)
 
-    plt.xlabel("Threshold")
+    plt.xlabel("Distance Threshold")
     plt.ylabel("Proportion")
 
     data = [ line for line in stats if predicate(line) ]

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 best possible results for the datasets.
-%
 Each record in $A$ is placed into the appropriate block based on its
@@
 block with the same blocking key value.
+
+\textbf{THIS TEXT WAS LIFTED FROM BELOW USE HERE? Traditionally, blocking is a method for choosing candidates for linkage using manually designed criteria.
+Thanks to blocking, a distance is only computed between promising pairs of records.
+A major disadvantage of traditional blocking is the need for extra engineering effort.
+Bespoke blocking strategies need to be designed for a given data set and application domain.}
 
@@
 
-% Peter 31 Oct: table style follows LNCS style guide
-
 \begin{table}[t]
@@
   CORA  & 1,295 & 1295 & 17,184 & publication-publication\\
-  NCVR  & ~224,073~ & 224,061 & ~148,036~ & voter-voter\\
   Isle of Skye & 17,612 & 12,284& 2,900 & birth-death\\
@@
 \emph{paper\_id} identifier of the form "blum1993". In this experiment linkage is performed over the same set of records.
-
-The second are two randomly selected sub-sets extracted from the North Carolina Voter Registration
-database (\emph{NCVR})~\footnote{Available from: \texttt{http://dl.ncsbe.gov/}}, which were collected in April and June
-2014. Each of these contains records of voters including their
-first names, surnames, and addresses. Ground truth is provided via a
-\emph{NCID} identifier which uniquely identifies a voter.
 
@@
     ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
-  NCVR  &  \\
   Isle of Skye &  \\
@@
 
-% \smallskip
 % \textbf{Experimental Setup:}
@@
 
-discuss measures used. PC, PQ, num comparisons, time? etc.
-
 \subsection{Cora}
@@
 The brute-force approach is used to establish the threshold value that yields the best possible linkage for the Cora dataset.
-The threshold value is varied between 0 and 250, \Cref{fig-cora-quality} presents the precision, recall, and F-measure values.
+The threshold value is varied between 0 and 250, \Cref{fig-cora-brute} presents the precision, recall, and F-measure values.
 The linkage results have high precision and low recall for low threshold values, and high recall and low precision for high threshold values.
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{figures/plotLQ_--_Cora_-_Brute_Force.png}
+\caption{Linkage Quality using Brute Force linking on the Cora data set\label{fig-cora-brute}}
+\end{figure}
+
+In the traditional blocking experiment the following fields from the Cora data set are used as a blocking key: \textit{author}, \textit{title}, \textit{venue}, \textit{location}, \textit{publisher} and \textit{year}.
+We also use a very permissive blocking key which takes the union of all the values for these fields.
+
+\Cref{comparison-of-results} shows the results of the Cora experiments.
+The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the pairs quality and pair completeness achieved by each algorithm.
+
+TODO: Explain pairs-quality and pairs-completeness.
+\emph{\textbf{TODO: discuss measures used. PC, PQ, num comparisons, time? etc.}}
+
+\begin{figure}
 \begin{verbatim}
@@
 \end{verbatim}
-\caption{Linkage Quality using Brute Force linking on the Cora data set\label{fig-cora-quality}}
-\end{figure}
-
-\subsubsection{Traditional Blocking}
-
-Traditionally, blocking is a method for choosing candidates for linkage using manually designed criteria.
-Thanks to blocking, a distance is only computed between promising pairs of records.
-A major disadvantage of traditional blocking is the need for extra engineering effort.
-Bespoke blocking strategies need to be designed for a given data set and application domain.
-In this section we experiment with using each key field in the Cora data set as a blocking key one by one, followed by a very permissive blocking key which takes the union of all the blocks attained by using individual key fields.
-
-\emph{AL: The above needs more explanation and not the first bit?}
-
-TODO: Explain pairs-quality and pairs-completeness~\Cref{cora-traditional}.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/plotFixedThresholdTraditional2_--_Cora_-_TradBlocking_-_70}
-\caption{Blocking Quality using traditional blocking on the Cora data set\label{cora-traditional}}
+\caption{Linkage Quality using Brute Force linking on the Cora data set\label{comparison-of-results}}
 \end{figure}
@@
 \end{figure}
-
-\subsection{LSH}
-
-LSH is a method for performing blocking without the engineering overhead of traditional blocking.
-It is parameterised over two integer values: number of bands, and band sizes~\footnote{LSH can also be parameterised over the shingle size, but in this paper we set the shingle size to 2.}.
 
@@
 \end{itemize}
-
-\subsection{NCVR}
 

### 2017-11-21 scripts/plots.py
toFloat, to account for "NA" entries

@@ -52,6 +52,13 @@ def loadCSV(filepath, delimiter=",", verbose=0):
         return (fieldNames, rawValues)
 
 
+def toFloat(s):
+    if s == "NA":
+        return 0
+    else:
+        return float(s)
+
+
 (fields, stats) = loadCSV("stats.tsv", delimiter="\t")
 
 for f in fields:
@@ -80,8 +87,8 @@ def plotLQ(title, predicate):
 
     data = [ line for line in stats if predicate(line) ]
 
-    plt.xlim( min([ float(line["Distance Threshold"]) for line in data ])
-            , max([ float(line["Distance Threshold"]) for line in data ])
+    plt.xlim( min([ toFloat(line["Distance Threshold"]) for line in data ])
+            , max([ toFloat(line["Distance Threshold"]) for line in data ])
             )
     plt.ylim(-0.05, 1.05)
     plt.grid()
@@ -218,14 +225,14 @@ def plotBQ(title, predicate):
 
     data = [ line for line in stats if predicate(line) ]
 
-    plt.xlim( min([ float(line["Distance Threshold"]) for line in data ])
-            , max([ float(line["Distance Threshold"]) for line in data ])
+    plt.xlim( min([ toFloat(line["Distance Threshold"]) for line in data ])
+            , max([ toFloat(line["Distance Threshold"]) for line in data ])
             )
     plt.ylim(-0.05, 1.05)
     plt.grid()
 
-    plt.plot( [ line["Distance Threshold"]          for line in data ]
-            , [ line["Average links quality"]       for line in data ]
+    plt.plot( [         line["Distance Threshold"]          for line in data ]
+            , [ toFloat(line["Average links quality"])       for line in data ]
             , color="green"
             , marker="o"
             , label="Block Quality"
@@ -257,21 +264,21 @@ def plotPQC(title, predicate):
 
     data = [ line for line in stats if predicate(line) ]
 
-    plt.xlim( min([ float(line["Distance Threshold"]) for line in data ])
-            , max([ float(line["Distance Threshold"]) for line in data ])
+    plt.xlim( min([ toFloat(line["Distance Threshold"]) for line in data ])
+            , max([ toFloat(line["Distance Threshold"]) for line in data ])
             )
     plt.ylim(-0.05, 1.05)
     plt.grid()
 
-    plt.plot( [ line["Distance Threshold"]          for line in data ]
-            , [ line["Average pairs quality"]       for line in data ]
[…]

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 \begin{verbatim}
-Figure 3 will be a table of Comparisons, F1, precision recall and pairs stuff.
+This will be a table of Comparisons, F1, precision recall and pairs stuff.
 12 ROWS ISH
 \end{verbatim}
-\caption{Linkage Quality using Brute Force linking on the Cora data set\label{comparison-of-results}}
+\caption{Linkage Quality on the Cora data set\label{comparison-of-results}}
 \end{figure}
 
-\Cref{cora-traditional-prf} plots the linkage quality values obtained by running traditional blocking based linkage at a threshold level of 70.
-Here, we see that high precision values are possible using blocking based linkage, however the recall (and as a result the F-measure) is poor in comparison to brute-force.
+\subsection{Historical Demography results}
 
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/plotFixedThresholdTraditional1_--_Cora_-_TradBlocking_-_70}
-\caption{Linkage Quality using traditional blocking on the Cora data set\label{cora-traditional-prf}}
+\begin{verbatim}
+This will be a table of Comparisons, F1, precision recall and pairs stuff for Skye/Kilmarnock
+12 ROWS ISH
+\end{verbatim}
+\caption{Linkage Quality on the Cora data set\label{comparison-of-results}}
 \end{figure}
-
-\subsection{M-Tree}
-
-M-Tree is a complete method that avoids brute-force search.
-In addition, it does not require any additional engineering work, it works with any distance metric.
 
@@
 \includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_Brute_Force}
-\caption{Block Quality - Brute Force\label{cora-block-quality-brute}}
+\caption{Will be Linkage quality with M-Tree Skye\label{cora-block-quality-brute}}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
-\caption{Block Quality - M-Tree\label{cora-block-quality-mtree}}
+\caption{Will be Linkage quality with M-Tree Kilmarnock\label{cora-block-quality-mtree}}
 \end{figure}
 
-\subsection{Historical Demography results}
-\begin{itemize}
-\item Fig X -1 Repeat of figure 2 with precision recall comparisons etc.
-\item Fig X - Kilmarnock M tree plot Precision,recall and F
-\item Fig X + 1 - Skye M tree plot Precision,recall and F
-\item Fig X + 2 - Mtree F measure and a bunch LSH F measures.
-\end{itemize}
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
+\caption{Will be Mtree F measure and a bunch LSH F measures\label{cora-block-quality-mtree}}
+\end{figure}
 

### 2017-11-21 paper.tex

@@
 \begin{figure}
-\begin{verbatim}
-This will be a table of Comparisons, F1, precision recall and pairs stuff.
-12 ROWS ISH
-\end{verbatim}
+\begin{table}
+\centering
+\begin{tabular}{lllllllll}
+Linker      & Linker             & Total comparisons & Number of links & Average pairs quality & Average pairs completeness & Precision & Recall & F1 Measure \\
+Brute Force & Brute Force        & 1677025           & 12474           & 0.02                  & 0.33                       & 0.84      & 0.57   & 0.68       \\
+MTree       & MTree              & 902693            & 12474           & 0.89                  & 0.22                       & 0.84      & 0.57   & 0.68       \\
+LSH         & LSH-2-2            & 175435            & 10443           & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH         & LSH-5-2            & 175435            & 10443           & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH         & LSH-10-2           & 175435            & 10443           & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH         & LSH-2-5            & 175435            & 10443           & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH         & LSH-5-5            & 175435            & 10443           & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH         & LSH-10-5           & 175435            & 10443           & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH         & LSH-2-10           & 175435            & 10443           & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH         & LSH-5-10           & 175435            & 10443           & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH         & LSH-10-10          & 175435            & 10443           & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+Blocking    & Blocking-year      & 115893            & 6504            & 0.22                  & 0.19                       & 0.99      & 0.35   & 0.51       \\
+Blocking    & Blocking-authors   & 11039             & 3219            & 0.74                  & 0.1                        & 0.94      & 0.16   & 0.28       \\
+Blocking    & Blocking-title     & 27407             & 8225            & 0.84                  & 0.23                       & 0.95      & 0.42   & 0.58       \\
+Blocking    & Blocking-venue     & 36647             & 6383            & 0.7                   & 0.12                       & 0.85      & 0.29   & 0.44       \\
+Blocking    & Blocking-location  & 1009957           & 9551            & 0.17                  & 0.24                       & 0.83      & 0.43   & 0.57       \\
+Blocking    & Blocking-publisher & 833079            & 9634            & 0.12                  & 0.23                       & 0.85      & 0.44   & 0.58       \\
+Blocking    & Blocking-combined  & 1214269           & 12324           & 0.05                  & 0.32                       & 0.84      & 0.56   & 0.67      
+\end{tabular}
+\end{table}
 \caption{Linkage Quality on the Cora data set\label{comparison-of-results}}
@@
 
-\textbf{Contribution:} The primary contribution of this paper is the novel application of MSI to achieve complete and efficient record linkage, without the need for complex parameter tuning.
-We evaluate our approach on several data sets from diverse domains and demonstrate its advantages over
-existing indexing techniques for record linkage.
+In this paper we have demonstrated the efficacy of the application of MSI to achieve complete and efficient record linkage, without the need for complex parameter tuning.
+In conclusion, this assertion deserves some careful unpacking.
+It is always possible to achieve high quality linkage using a brute-force approach.
+However the n-squared complexity of this approach prevents its practical application for datasets of even moderate size.
+The use of MSI techniques such as M-tree delivers the same results as the brute-force approach at the cost of
 

### 2017-11-21 paper.tex

--- 2017-11-21 paper.tex
@@
 \usepackage{cleveref}
+\usepackage{hhline}
 
@@
 repeated with different parameter settings and possibly also alternative
-algorithms. This can result in a time-consuming and labor-intensive
+algorithms. This can result in a time-consuming and labour-intensive
 iterative process. A major challenge is often the heuristic nature of
@@
 
-\textbf{TODO: GK rewrite this paragraph. Make the connection between completeness and high recall. Define completenes.}
-Here we explore how existing indexing techniques can lead to
-low recall, even for record pairs that are highly similar and likely
-correspond to true matches. As a result, such techniques can lead
-to \emph{incomplete} record linkage results. To address this problem,
-\emph{metric space indexing} (MSI) can be used to guarantee that all
-record pairs, up to a certain distance threshold, are considered in the
-record linkage results. MSI also avoids the need for costly detailed
-comparison of record pairs with low similarities, which may be inserted
-into the same block with traditional indexing techniques. It allows
-indexing, comparison and classification to be combined into a
-single step, as shown in Fig.~\ref{fig-rl-process}, making the overall
-record linkage process simpler, more efficient and more effective than
-the traditional approach.
+This paper focuses on approaches that use a similarity threshold to
+classify links. Such techniques are fundamentally limited by the extent
+to which true matching records are similar, and true non-matches are
+dissimilar---this is dataset dependent. Within this domain, we define
+a technique to be \emph{complete} if it is guaranteed to identify all
+pairs within the specified similarity threshold. Commonly used indexing
+techniques are incomplete, since they reduce computational cost at the
+expense of potentially overlooking some true matches. By definition,
+incomplete techniques yield lower recall than complete techniques.
+Conversely, and perhaps counter-intuitively, complete techniques can
+yield lower precision with some datasets. This is discussed further in
+Sect.~\ref{sec-approach}.
 
-\textbf{Contribution:} The primary contribution of this paper is the novel application of MSI to achieve complete and efficient record linkage, without the need for complex parameter tuning.
-We evaluate our approach on several data sets from diverse domains and demonstrate its advantages over
-existing indexing techniques for record linkage.
+\emph{Metric space indexing} (MSI) is a complete technique with much
+lower computational overheads than a brute-force approach. It allows
+indexing, comparison and classification to be combined into a single
+step, as shown in Fig.~\ref{fig-rl-process}, making the overall record
+linkage process simpler, more efficient and more effective than
+incomplete indexing approaches.
+
+\textbf{Contribution:} the primary contribution of this paper is the
+novel application of MSI to achieve complete and efficient record
+linkage, without the need for complex parameter tuning. We evaluate our
+approach on several data sets from diverse domains and demonstrate its
+advantages over existing indexing techniques for record linkage.
 
 % For final paper:
-% The motivation for this work is the Digitising Scotland project (Dibben 2012), which is in the process of transcribing and linking all the vital events recorded in Scotland between 1856 and 1977.
-This data set will, when complete, include around 14 million birth records, 11 million death records and 4 million marriage records.
-As part of the work, certain data fields (locations, occupations and causes of death) are also being classified to the relevant standard coding schemes.
+% The motivation for this work is the Digitising Scotland project
+% (Dibben 2012), which is in the process of transcribing and linking all
+% the vital events recorded in Scotland between 1856 and 1977. This data
+% set will, when complete, include around 14 million birth records, 11
+% million death records and 4 million marriage records. As part of the
+% work, certain data fields (locations, occupations and causes of death)
+% are also being classified to the relevant standard coding schemes.
 
@@
 
+TODO: add stuff about how incomplete methods may yield higher precision than complete
+
 \textbf{THIS TEXT WAS LIFTED FROM BELOW USE HERE? Traditionally, blocking is a method for choosing candidates for linkage using manually designed criteria.
@@
 into a set of n-grams with $n = l_{ss}$. Next, a set of
-deterministically generated hash functions are applied to
-each n-gram in the set
-and the smallest result (the MinHash) of each hash application is added
-to a signature for the record. The number of hashes used, and thus the
-size of the signature, is set to $l_{nb} \times l_{bs}$. Finally, the
-signature is split into $l_{nb}$ bands and the values from each band
-are hashed again to create a number of keys. The original record is
-added to a map associated with each of the keys.
+deterministically generated hash functions are applied to each n-gram in
+the set and the smallest result (the MinHash) of each hash application
+is added to a signature for the record. The number of hashes used, and
+thus the size of the signature, is set to $l_{nb} \times l_{bs}$.
+Finally, the signature is split into $l_{nb}$ bands and the values from
+each band are hashed again to create a number of keys. The original
+record is added to a map associated with each of the keys.
 
@@
 summarized in Table~\ref{table-datasets}. The first is
-\emph{CORA}~\footnote{Available from:
-\texttt{http://secondstring.sourceforge.net}}, which contains 1,295
+\emph{Cora}~\cite{Cora2017}, which contains 1,295
 records that refer to 112 machine learning publications. 
@@
 
-\Cref{comparison-of-results} shows the results of the Cora experiments.
+\Cref{comparison-of-results-cora} shows the results of the Cora experiments.
 The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the pairs quality and pair completeness achieved by each algorithm.
+The linker labels may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the size of the bands followed by the number of bands.
+For traditional blocking the parameters are the fields used for blocking.
 
@@
 
-\begin{figure}
-\begin{verbatim}
-This will be a table of Comparisons, F1, precision recall and pairs stuff.
-12 ROWS ISH
-\end{verbatim}
-\caption{Linkage Quality on the Cora data set\label{comparison-of-results}}
-\end{figure}
+\begin{table}[ht]
+\centering
+\begin{tabular}{l|r|r|r|r|r|r}
+Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline{=#=#=}
+Brute Force        & 1,677,025                    & 0.02                  & 0.33                       & 0.84      & 0.57   & 0.68       \\ \hline
+MTree              &  902,693                     & 0.89                  & 0.22                       & 0.84      & 0.57   & 0.68       \\ \hline
+LSH-2-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH-5-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH-10-2           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH-2-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH-5-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH-10-5           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH-2-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH-5-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
+LSH-10-10          &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\  \hline
+Blocking-year      &  115,893                     & 0.22                  & 0.19                       & 0.99      & 0.35   & 0.51       \\
+Blocking-authors   &   11,039                      & 0.74                  & 0.1                        & 0.94      & 0.16   & 0.28       \\
+Blocking-title     &   27,407                      & 0.84                  & 0.23                       & 0.95      & 0.42   & 0.58       \\
+Blocking-venue     &   36,647                      & 0.7                   & 0.12                       & 0.85      & 0.29   & 0.44       \\
+Blocking-location  & 1,009,957                    & 0.17                  & 0.24                       & 0.83      & 0.43   & 0.57       \\
+Blocking-publisher &  833,079                     & 0.12                  & 0.23                       & 0.85      & 0.44   & 0.58       \\
+Blocking-combined  & 1,214,269                    & 0.05                  & 0.32                       & 0.84      & 0.56   & 0.67       \\ \hline
+\end{tabular}
+\caption{Linkage Quality on the Cora data set distance=70\label{comparison-of-results-cora}
+\textbf{TODO: The table needs some rows removed. Note that this table is with a threshold value of 70.}
+}
+\end{table}
 
@@
 
-\begin{figure}
-\begin{verbatim}
-This will be a table of Comparisons, F1, precision recall and pairs stuff for Skye/Kilmarnock
-12 ROWS ISH
-\end{verbatim}
-\caption{Linkage Quality on the Cora data set\label{comparison-of-results}}
-\end{figure}
+\begin{table}[ht]
+\centering
+\begin{tabular}{l|r|r|r|r|r|r}
+Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hhline{=#=#=}
+MTree    & 102,318,525         & 0.84                  & 0.86                       & 0.65      & 0.46   & 0.54      \\ \hline
+LSH-2-2  & 3,109,250           & 0                     & 0.05                       & 0.63      & 0.03   & 0.06       \\
+LSH-5-2  & 10,412,496          & 0                     & 0.16                       & 0.64      & 0.11   & 0.19       \\
+LSH-10-2 & 53,874,127          & 0                     & 0.46                       & 0.68      & 0.36   & 0.47       \\
+LSH-5-5  & 36,566             & 0                     & 0.01                       & 0.76      & 0.01   & 0.01       \\
+LSH-10-5 & 129,873            & 0                     & 0.01                       & 0.72      & 0.01   & 0.02       \\ \hline
+\end{tabular}
+\caption{Linkage Quality on demography datasets \textbf{This is for Skye. Threshold value is 2.}
+\label{comparison-of-results-demography-skye}
+}
+\end{table}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{l|r|r|r|r|r|r}
+Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline
+MTree     & 514,871,153         & 0.89                  & 0.82                       & 0.76      & 0.45   & 0.57      \\ \hline
+LSH-2-2   & 99,145,887          & 0                     & 0.23                       & 0.81      & 0.16   & 0.27       \\
+LSH-5-2   & 130,721,338         & 0                     & 0.32                       & 0.79      & 0.23   & 0.36       \\
+LSH-10-2  & 177,168,848         & 0                     & 0.48                       & 0.79      & 0.36   & 0.49       \\
+LSH-5-5   & 239,368            & 0.01                  & 0.02                       & 0.84      & 0.01   & 0.02       \\
+LSH-10-5  & 855,431            & 0.01                  & 0.03                       & 0.87      & 0.02   & 0.03 \\ \hline
+\end{tabular}
+\caption{Linkage Quality on demography datasets\textbf{This is for Kilmarnock. Threshold value is 2.
+\label{comparison-of-results-demography-kili}
+}
+\end{table}
 
@@
 \includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_Brute_Force}
-\caption{Will be Linkage quality with M-Tree Skye\label{cora-block-quality-brute}}
+\caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
-\caption{Will be Linkage quality with M-Tree Kilmarnock\label{cora-block-quality-mtree}}
+\caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
-\caption{Will be Mtree F measure and a bunch LSH F measures\label{cora-block-quality-mtree}}
+\caption{Comparison of quality attainable on Skye dataset using M-tree and various LSH configurations\label{mtree-lsh-quality}}
 \end{figure}
@@
 
-\section{Conclusions and Future Work}
+\section{Conclusions and Future Work\label{sec-concl}}
 
-\textbf{Contribution:} The primary contribution of this paper is the novel application of MSI to achieve complete and efficient record linkage, without the need for complex parameter tuning.
-We evaluate our approach on several data sets from diverse domains and demonstrate its advantages over
-existing indexing techniques for record linkage.
+In this paper we have demonstrated the efficacy of MSI in achieving consistently complete and efficient record linkage, without the need for complex parameter tuning.
+In conclusion, this assertion deserves some careful unpacking.
+It is always possible to achieve high quality linkage using a brute-force approach.
+However the n-squared complexity of this approach prevents its practical application for datasets of even moderate size.
+The use of MSI techniques such as M-tree delivers the same results as the brute-force approach with fewer comparisons.
+We have shown that the M-tree approach delivers high precision, high recall results that are the same as those delivered by the brute-force approach.
+Furthermore this is achieved consistently without the need for complex parameter tuning.
+
+We contrast this to traditional blocking and LSH based approaches.
+The major drawback of their use is that whilst they can produce extremely good results, they can also produce extremely poor results.
+It was our observations of low recall given by these approaches that originally led us to experiment with M-trees.
+Without extensive ground truth or large amounts of clerking effort by human domain experts it is impossible to determine the quality of the links produced.
+
+However, this paper also highlights some unexpected results.
+Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase.
+A second unexpected result is that shown in \Cref{fig-cora-brute}, namely that LSH can yield higher precision than is achieved by a complete method such as M-Tree.
+This is due to complex reasons ....
+\emph{TODO EXPAND HERE}.
+
+\emph{Some sentence or two about run-time?}
+
+The observation that LSH can yield results quickly at the cost of some recall suggests a potentially profitable future direction for exploration, namely the use of hybrid algorithms.
+We hypothesise that it may be possible to use an incomplete technique like LSH to form indices into a data structure such as an M-tree.
+In this manner, M-tree nodes could be used as the starting point for range-search queries and, in so doing, prune the search space considerably to yield highly quality complete results quickly.
 
 % TODO: check that the above claims have been addressed.
-
-\label{sec-concl}
 
@@
 
-OLD STUFF
+%text cut from earlier versions:
 
-\section{The Plan}
-
-\vspace{5mm}
-
-Plan in a bulleted list:
-
-\begin{itemize}
-\item Design: Traditionally it has two stages.
-A blocking method needs to be developed separately from a similarity method.
-With similarity search based record linkage, there is only one method, and that is the similarity method.
-\item Efficiency: I am not convinced of this, but Peter thinks there may be efficiency benefits.
-The reasoning behind this is that with blocking-based methods we would have to calculate some sort of a similarity value first.
-This value will only be used for blocking, and later a similarity value between candidate pairs will need to be calculated again.
-\item Completeness: Similarity search based methods will be complete (with respect to the given similarity method) whereas blocking based methods (including LSH) are likely to be incomplete (there will be high-similarity pairs which are not places in the same block)
-\item Data sets
-\begin{itemize}
-\item Kilmarnock Birth-Death
-\item Skye Birth-Death
-\item CORA
-\item North Carolina Voter Registration DB
-\end{itemize}
-\item The following is a list of all methods we thought we would have to compare.
-\begin{itemize}
-\item Similarity search (M-Tree, others?)
-\item LSH-blocking
-\item Traditional blocking (e.g. Region, Surname) - \textbf{why this? - this was done in the paper: A Comparison of Blocking Methods for Record Linkage}
-\item Sorted Neighbourhood blocking - \textbf{why this?}
-\end{itemize}
-\item Pairs-completeness and pairs-quality (w.r.t high similarity instead of true-link status). I will expand on this a bit more later.
-\item Comparison (for blocking based methods)
-		Precision/Recall/F1-Measure
-\item Which similarity methods are metric?
-\item Run time and memory usage comparisons
-\item Theorem. Precision might get better or worse with similarity search based methods, but recall should never get worse.
-\end{itemize}
-
-Note: The rest of this paper should be regarded as a placeholder for now.
-
-% ====================================================================
-
-\smallskip
-\textbf{MIfile:}
-The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces.
+%The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces.
 Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
@@
 
-%the following is a sentence we culled earlier but may be needed.
-We will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
+%We will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
 Sect.~\ref{sec-exp}.
-
-\begin{verbatim}
-for each record a in A
-    for each record b in B
-        classify (a,b) as link if distance(a,b) < threshold
-\end{verbatim}
-
-% PAKDD 2018: maximum 12 pages
-% Abstract max 200 words
-% title, abstract: 1/2 page
-% introduction: 1 1/2 page (2 pages so far)
-% related work: 1 to 1 1/2 pages (3 pages so far)
-% method: 4 pages (7 pages so far)
-% experiments: 3 (10 pages so far)
-% discussion /conclusion: 1 page
-% citations 1
-
-% ====================================================================
-
-% Eamonn ICDM'10 tutorial slides
-% - clear problem statement in abstract
-% - To convince a reviewer, you must think like a reviewer
-
-% Writing the paper:
-% - Make a working title
-% - Introduce the topic and define (informally at this stage)
-%   terminology
-% - Motivation: Emphasize why is the topic important
-% - Relate to current knowledge: what’s been done
-% - Indicate the gap: what need’s to be done?
-% - Formally pose research questions
-% - Explain any necessary background material.
-% - Introduce formal definitions.
-% - Introduce your novel algorithm/representation/data structure etc.
-% - Describe experimental set-up, explain what the experiments will
-%   show
-% - Describe the datasets
-% - Summarize results with figures/tables
-% - Discuss results
-% - Explain conflicting results, unexpected findings and discrepancies
-%   with other research
-% - State limitations of the study
-% - State importance of findings
-% - Announce directions for further research
-% - Acknowledgements
-% - References
-%
-% - Don’t make the reviewer of your paper think!
-% - Reviewers make an initial impression on the first page and don’t
-%   change 80% of the time
-% - A good introduction with a good motivation is half your success
-% - By the end of the introduction the reviewer mustknow.
-%   - What is the problem?
-%   - Why is it interesting and important?
-%   - Why is it hard?why do naive approaches fail?
-%   - Why hasn't it been solved before?(Or, what's wrong with previous
-%     proposed solutions?)
-%   - What are the key components of my approach and results?Also
-%     include any specific limitations.
-%   - A final paragraph or subsection: “Summary of Contributions”.
-%     It should list the major contributions in bullet form,
-%     mentioning in which sections they can be found. This material
-%     doubles as an outline of the rest of the paper, saving space and
-%     eliminating redundancy
-% - Unjustified Choices (are bad)
-% - Optimal: Does not mean `very good'
-% - Proved: Does not mean `demonstrated'
-% - Significant: There is a danger of confusing the informal statement
-%   and the statistical claim
-% - Use all the Space Available
-% - Avoid Weak Language: aim, attempt, might, etc.
-% - Use the Active Voice
-% - ALWAYS put some variance estimate on performance measures (do
-%   everything 10 times and give me the variance of whatever you are
-%   reporting)
-
-% Figures:
-% - Don't cover the data with the labels!
-% - Color helps -Direct labeling helps -Meaningful captions help
-%
-% Common problem with figures:
-% 1.Too many patterns on bars
-% 2.Use of both different symbols and different lines
-% 3.Too many shades of gray on bars
-% 4.Lines too thin (or thick)
-% 5.Use of three-dimensional bars for only two variables
-% 6.Lettering too small and font difficult to read
-% 7.Symbols too small or difficult to distinguish
-% 8.Redundant title printed on graph
-% 9.Use of gray symbols or lines
-% 10.Key outside the graph
-% 11.Unnecessary numbers in the axis
-% 12.Multiple colors map to the same shade of gray
-% 13.Unnecessary shading in background
-% 14.Using bitmap graphics (instead of vector graphics)
-% 15.General carelessness

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 to which true matching records are similar, and true non-matches are
-dissimilar---this is dataset dependent. Within this domain, we define
+dissimilar---this is dataset-dependent. Within this domain, we define
 a technique to be \emph{complete} if it is guaranteed to identify all
@@
 
-We now briefly review relevant work in the areas of indexing
-for record linkage (for recent surveys see~\cite{Chr12b,Pap16}),
-and metric space indexing~\cite{Zezula2010}.
+We now briefly review relevant work in the areas of indexing for record
+linkage (for recent surveys see~\cite{Chr12b,Pap16}), and metric space
+indexing~\cite{Zezula2010}.
 
@@
 
-A different approach to indexing is the sorted neighborhood
+A different approach to indexing is the sorted neighbourhood
 method~\cite{Mon96}, where the datasets to be linked are sorted
@@
 Locality sensitive hashing (LSH), originally proposed to allow efficient
-nearest-neighbor search in high-dimensional spaces~\cite{Ind98}, has
+nearest-neighbour search in high-dimensional spaces~\cite{Ind98}, has
 been employed in record linkage as an indexing technique where attribute
@@
 LSH-MinHash, and one complete method, M-tree~\cite{paolociaccia2m}. We
-also use a simple brute-force complete technique as a baseline for
+also use a simple complete brute-force technique as a baseline for
 comparison, although this can only feasibly be applied to the smallest
@@
 block with the same blocking key value.
+In some circumstances, incomplete approaches such as blocking can yield
+higher precision than complete techniques. This can occur when the
+datasets contain a significant number
+of non-matches that nonetheless have high similarity. In this situation,
+the fact that an incomplete technique omits consideration of some
+potential links can serve to improve precision, since the similarity
+distance
 
 TODO: add stuff about how incomplete methods may yield higher precision than complete
-
-\textbf{THIS TEXT WAS LIFTED FROM BELOW USE HERE? Traditionally, blocking is a method for choosing candidates for linkage using manually designed criteria.
-Thanks to blocking, a distance is only computed between promising pairs of records.
-A major disadvantage of traditional blocking is the need for extra engineering effort.
-Bespoke blocking strategies need to be designed for a given data set and application domain.}
 
@@
 
-\begin{table}[t]
-\caption{Parameter settings used for the different data sets used in
-   the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
-   shingle size, and the size and number of bands, respectively.}
- \label{table-parameters}
-  \centering
-  \begin{scriptsize}
-  %\addtolength{\tabcolsep}{-0.5pt}
-  \begin{tabular}{ccccc}
-  \hline\noalign{\smallskip}
-  Data set~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
-  name(s) & & & &  \\
-  \noalign{\smallskip} \hline \noalign{\smallskip}
-  CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
-    ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
-  Isle of Skye &  \\
-  Kilmarnock  &  \\
-  \noalign{\smallskip} \hline
-  \end{tabular}
-  \end{scriptsize}
-\end{table}
+% \begin{table}[t]
+% \caption{Parameter settings used for the different data sets used in
+%    the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
+%    shingle size, and the size and number of bands, respectively.}
+%  \label{table-parameters}
+%   \centering
+%   \begin{scriptsize}
+%   %\addtolength{\tabcolsep}{-0.5pt}
+%   \begin{tabular}{ccccc}
+%   \hline\noalign{\smallskip}
+%   Data set~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
+%   name(s) & & & &  \\
+%   \noalign{\smallskip} \hline \noalign{\smallskip}
+%   CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
+%     ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
+%   Isle of Skye &  \\
+%   Kilmarnock  &  \\
+%   \noalign{\smallskip} \hline
+%   \end{tabular}
+%   \end{scriptsize}
+% \end{table}
 
@@
 The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the pairs quality and pair completeness achieved by each algorithm.
-The linker labels may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the size of the bands followed by the number of bands.
+The linker labels may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the number of the bands followed by the size of bands.
 For traditional blocking the parameters are the fields used for blocking.
@@
 TODO: Explain pairs-quality and pairs-completeness.
-\emph{\textbf{TODO: discuss measures used. PC, PQ, num comparisons, time? etc.}}
+\emph{\textbf{TODO: discuss measures used. num comparisons, time? etc.}}
 
@@
 \begin{tabular}{l|r|r|r|r|r|r}
-Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline{=#=#=}
+Linker & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\
+\hline
 Brute Force        & 1,677,025                    & 0.02                  & 0.33                       & 0.84      & 0.57   & 0.68       \\ \hline
@@
 \begin{tabular}{l|r|r|r|r|r|r}
-Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hhline{=#=#=}
+Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline
 MTree    & 102,318,525         & 0.84                  & 0.86                       & 0.65      & 0.46   & 0.54      \\ \hline
@@
 
-\begin{table}[h]
+\begin{table}[ht]
 \centering
@@
 \end{tabular}
-\caption{Linkage Quality on demography datasets\textbf{This is for Kilmarnock. Threshold value is 2.
+\caption{Linkage Quality on demography datasets\textbf{This is for Kilmarnock. Threshold value is 2.}
 \label{comparison-of-results-demography-kili}

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 linkage, without the need for complex parameter tuning. We evaluate our
-approach on several data sets from diverse domains and demonstrate its
+approach on several datasets from diverse domains and demonstrate its
 advantages over existing indexing techniques for record linkage.
@@
 the records being linked, and chosen with the intention of giving the
-best possible results for the datasets.
-Each record in $A$ is placed into the appropriate block based on its
-blocking key value. The algorithm then iterates over the records in $B$,
-and for each one, compares it with each of the records from $A$ in the
-block with the same blocking key value.
+best possible results for the datasets. Each record in $A$ is placed
+into the appropriate block based on its blocking key value. The
+algorithm then iterates over the records in $B$, and for each one,
+compares it with each of the records from $A$ in the block with the same
+blocking key value.
+
 In some circumstances, incomplete approaches such as blocking can yield
 higher precision than complete techniques. This can occur when the
-datasets contain a significant number
-of non-matches that nonetheless have high similarity. In this situation,
-the fact that an incomplete technique omits consideration of some
-potential links can serve to improve precision, since the similarity
-distance
-
-TODO: add stuff about how incomplete methods may yield higher precision than complete
+datasets contain a significant number of non-matches that nonetheless
+have high similarity. In this situation, the fact that an incomplete
+technique omits consideration of some potential links can serve to
+improve precision, since a classification decision based on similarity
+threshold is incorrect for high-similarity non-matches. By definition,
+recall is never higher for incomplete techniques.
 
@@
 
-\textbf{M-Tree:} The M-tree linkage algorithm has no additional parameters.
-In a similar manner to the \emph{LSH-MinHash} approach, M-trees are constructed by supplying the implementation with a set of records and a \emph{distance function}.
-The distance functions used are the same as those used for \emph{Nested-loop} and \emph{LSH-MinHash} experiments.
-To perform linkage, the data-structure is searched using a \emph{range-search} query which yields all records within distance $d$ of the query object and all the returned records are directly recorded as links.
+\textbf{M-tree:} The M-tree linkage algorithm has no additional
+parameters. In a similar manner to the \emph{LSH-MinHash} approach, each
+record in A is inserted into an M-tree. To perform linkage, the
+algorithm iterates over each record $b$ in $B$. A $range-search(b,d)$
+operation is performed, passing the distance threshold as the second
+parameter. All the returned records are directly recorded as links.
 
@@
 
-We now describe the data sets we use in our evaluation, the setup
-employed to evaluate our proposed metric indexing approach and
-compare it with traditional blocking techniques, and we then present
-and discuss our results~\footnote{The raw data, additional figures and the source code required to run these experiments can be downloaded from here. (Link removed for double blind review)}.
+We now describe the datasets we use in our evaluation, the set-up
+employed to evaluate our proposed metric indexing approach and compare
+it with traditional blocking techniques, and we then present and discuss
+our results~\footnote{The raw data, additional figures and the source
+code required to run these experiments can be downloaded from here.
+(Link removed for double blind review)}.
 
@@
 \begin{table}[t]
-\caption{Characteristics of data sets used in the experiments.}
+\caption{Characteristics of datasets used in the experiments.}
  \label{table-datasets}
@@
   \hline\noalign{\smallskip}
-  Data set~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities\\
+  Dataset~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities\\
   name(s)  & dataset1  & dataset2  & matching pairs & linked \\
@@
 \smallskip
-\textbf{Data Sets:}
-%\label{sec-data}
-We used four data sets from three domains in our experiments, as
-summarized in Table~\ref{table-datasets}. The first is
-\emph{Cora}~\cite{Cora2017}, which contains 1,295
-records that refer to 112 machine learning publications. 
-It is used as a benchmark data set in the literature for assessing linkage algorithms. Ground truth is provided via a unique
-\emph{paper\_id} identifier of the form "blum1993". In this experiment linkage is performed over the same set of records.
+\textbf{Datasets:}
+\label{sec-data}
+We used three datasets from two domains in our experiments, as
+summarised in Table~\ref{table-datasets}. The first is
+\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer to
+112 machine learning publications. It is used as a benchmark dataset in
+the literature for assessing linkage algorithms. Ground truth is
+provided via a unique \emph{paper\_id} identifier of the form
+``blum1993". In this experiment linkage is performed over the same set
+of records.
 
-The last two data sets are historical Scottish records of vital
+The other datasets are historical Scottish records of vital
 events (birth, marriages and deaths) one registered on the
 \emph{Isle of Skye}, a rural district, and the other
-records from \emph{Kilmarnock}, an industrial town. These data sets
-were provided to us by historical demographers who extensively
-curated and linked both data sets~\cite{reid2002,reid2006}. Both data
+records from \emph{Kilmarnock}, an industrial town. These datasets
+were create by historical demographers who extensively
+curated and linked both datasets~\cite{reid2002,reid2006}. Both data
 sets include the  names, gender, addresses of individuals and their
@@
 % \begin{table}[t]
-% \caption{Parameter settings used for the different data sets used in
+% \caption{Parameter settings used for the different datasets used in
 %    the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
@@
 %   \hline\noalign{\smallskip}
-%   Data set~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
+%   Dataset~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
 %   name(s) & & & &  \\
@@
 
-In this section, we perform linkage on the Cora data set using all four approaches presented in this paper: brute force, traditional blocking, LSH and M-Tree.
+In this section, we perform linkage on the Cora dataset using all four approaches presented in this paper: brute force, traditional blocking, LSH and M-Tree.
 The brute-force approach is used to establish the threshold value that yields the best possible linkage for the Cora dataset.
@@
 The linkage results have high precision and low recall for low threshold values, and high recall and low precision for high threshold values.
-For this data set, the maximum F-measure value is achieved using a threshold value of 70.
+For this dataset, the maximum F-measure value is achieved using a threshold value of 70.
 This value is data dependent; for different datasets the maximum F-measure will correspond to different thresholds.
@@
 \includegraphics[width=\textwidth]{figures/plotLQ_--_Cora_-_Brute_Force.png}
-\caption{Linkage Quality using Brute Force linking on the Cora data set\label{fig-cora-brute}}
+\caption{Linkage Quality using Brute Force linking on the Cora dataset\label{fig-cora-brute}}
 \end{figure}
 
-In the traditional blocking experiment the following fields from the Cora data set are used as a blocking key: \textit{author}, \textit{title}, \textit{venue}, \textit{location}, \textit{publisher} and \textit{year}.
+In the traditional blocking experiment the following fields from the Cora dataset are used as a blocking key: \textit{author}, \textit{title}, \textit{venue}, \textit{location}, \textit{publisher} and \textit{year}.
 We also use a very permissive blocking key which takes the union of all the values for these fields.
@@
 \Cref{comparison-of-results-cora} shows the results of the Cora experiments.
-The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the pairs quality and pair completeness achieved by each algorithm.
-The linker labels may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the number of the bands followed by the size of bands.
+The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the \textit{pairs quality} and\textit{ pairs completeness} achieved by each algorithm.
+The linker column may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the number of the bands followed by the size of bands.
 For traditional blocking the parameters are the fields used for blocking.
+\textit{Pairs quality} and \textit{pairs completeness}  are given to illustrate the efficacy of the algorithms and are defined as follows: XXXXXXXXXXX.
+
+\textit{Pairs quality}  relates to recall and gives an indication of the number of (correct) matches amongst the links. 
+Pairs completeness BLAH BLAH.
+
+As can be seen in \Cref{comparison-of-results-cora}, M-tree reports the same results as brute-force....
 
@@
 \begin{tabular}{l|r|r|r|r|r|r}
-Linker & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\
-\hline
+\parbox{0.18\linewidth}{\centering Linker} &
+\parbox{0.16\linewidth}{\centering Comparisons} &
+\parbox{0.11\linewidth}{\centering Pairs Quality} &
+\parbox{0.17\linewidth}{\centering Pairs \\ Completeness} &
+\parbox{0.11\linewidth}{\centering Precision} &
+\parbox{0.08\linewidth}{\centering Recall} &
+\parbox{0.11\linewidth}{\centering F1 Measure} \\ \hline \hline
 Brute Force        & 1,677,025                    & 0.02                  & 0.33                       & 0.84      & 0.57   & 0.68       \\ \hline
@@
 LSH-10-10          &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\  \hline
-Blocking-year      &  115,893                     & 0.22                  & 0.19                       & 0.99      & 0.35   & 0.51       \\
-Blocking-authors   &   11,039                      & 0.74                  & 0.1                        & 0.94      & 0.16   & 0.28       \\
-Blocking-title     &   27,407                      & 0.84                  & 0.23                       & 0.95      & 0.42   & 0.58       \\
-Blocking-venue     &   36,647                      & 0.7                   & 0.12                       & 0.85      & 0.29   & 0.44       \\
-Blocking-location  & 1,009,957                    & 0.17                  & 0.24                       & 0.83      & 0.43   & 0.57       \\
-Blocking-publisher &  833,079                     & 0.12                  & 0.23                       & 0.85      & 0.44   & 0.58       \\
-Blocking-combined  & 1,214,269                    & 0.05                  & 0.32                       & 0.84      & 0.56   & 0.67       \\ \hline
+Block-year      &  115,893                     & 0.22                  & 0.19                       & 0.99      & 0.35   & 0.51       \\
+Block-authors   &   11,039                      & 0.74                  & 0.1                        & 0.94      & 0.16   & 0.28       \\
+Block-title     &   27,407                      & 0.84                  & 0.23                       & 0.95      & 0.42   & 0.58       \\
+Block-venue     &   36,647                      & 0.7                   & 0.12                       & 0.85      & 0.29   & 0.44       \\
+Block-location  & 1,009,957                    & 0.17                  & 0.24                       & 0.83      & 0.43   & 0.57       \\
+Block-publisher &  833,079                     & 0.12                  & 0.23                       & 0.85      & 0.44   & 0.58       \\
+Block-combined  & 1,214,269                    & 0.05                  & 0.32                       & 0.84      & 0.56   & 0.67       \\ \hline
 \end{tabular}
-\caption{Linkage Quality on the Cora data set distance=70\label{comparison-of-results-cora}
+\caption{Linkage Quality on the Cora dataset distance=70\label{comparison-of-results-cora}
 \textbf{TODO: The table needs some rows removed. Note that this table is with a threshold value of 70.}
@@
 \begin{tabular}{l|r|r|r|r|r|r}
-Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline
+\parbox{0.18\linewidth}{\centering Linker} &
+\parbox{0.16\linewidth}{\centering Comparisons} &
+\parbox{0.11\linewidth}{\centering Pairs Quality} &
+\parbox{0.17\linewidth}{\centering Pairs \\ Completeness} &
+\parbox{0.11\linewidth}{\centering Precision} &
+\parbox{0.08\linewidth}{\centering Recall} &
+\parbox{0.11\linewidth}{\centering F1 Measure} \\ \hline \hline
 MTree    & 102,318,525         & 0.84                  & 0.86                       & 0.65      & 0.46   & 0.54      \\ \hline
@@
 \begin{tabular}{l|r|r|r|r|r|r}
-Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline
+\parbox{0.18\linewidth}{\centering Linker} &
+\parbox{0.16\linewidth}{\centering Comparisons} &
+\parbox{0.11\linewidth}{\centering Pairs Quality} &
+\parbox{0.17\linewidth}{\centering Pairs \\ Completeness} &
+\parbox{0.11\linewidth}{\centering Precision} &
+\parbox{0.08\linewidth}{\centering Recall} &
+\parbox{0.11\linewidth}{\centering F1 Measure} \\ \hline \hline
 MTree     & 514,871,153         & 0.89                  & 0.82                       & 0.76      & 0.45   & 0.57      \\ \hline

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 \usepackage{hhline}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 
-In some circumstances, incomplete approaches such as blocking can yield
-higher precision than complete techniques. This can occur when the
-datasets contain a significant number of non-matches that nonetheless
-have high similarity. In this situation, the fact that an incomplete
-technique omits consideration of some potential links can serve to
-improve precision, since a classification decision based on similarity
-threshold is incorrect for high-similarity non-matches. By definition,
-recall is never higher for incomplete techniques.
-
 \textbf{LSH-MinHash:}
@@
 non-link based on their distance.
+
+In some circumstances, incomplete approaches such as traditional blocking and LSH-MinHash can yield
+higher precision than complete techniques. This can occur when the
+datasets contain a significant number of non-matches that nonetheless
+have high similarity. In this situation, the fact that an incomplete
+technique omits consideration of some potential links can serve to
+improve precision, since a classification decision based on similarity
+threshold is incorrect for high-similarity non-matches. By definition,
+recall is never higher for incomplete techniques.
 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_Brute_Force}
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/skye-mtree}
 \caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/kili-mtree}
 \caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
+\end{subfigure}
+\caption{M-Tree based linkage on the demographic data sets}
+\label{demography-quality-mtree}
 \end{figure}
@@
 
-In this paper we have demonstrated the efficacy of MSI in achieving consistently complete and efficient record linkage, without the need for complex parameter tuning.
+In this paper we have demonstrated the efficacy of MSI in achieving  complete and efficient record linkage, without the need for complex parameter tuning.
 In conclusion, this assertion deserves some careful unpacking.
@@
 Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase.
-A second unexpected result is that shown in \Cref{fig-cora-brute}, namely that LSH can yield higher precision than is achieved by a complete method such as M-Tree.
-This is due to complex reasons ....
-\emph{TODO EXPAND HERE}.
+A second unexpected result is that shown in \Cref{fig-cora-brute}, namely that LSH can yield higher precision than is achieved by a complete method such as M-Tree as was discussed in Sect.~\ref{sec-approach}.
+
+\emph{IS THIS RIGHT?}.
 

### 2017-11-21 scripts/plots.py
fiddle with legends

@@ -79,8 +79,8 @@ def plotLQ(title, predicate):
     plt.rc("font", **font)
     # plt.rcParams["text.usetex"] = True
 
-    plt.title("Linkage Quality w.r.t. Threshold", fontsize=22)
-    plt.suptitle(title, y = 1.02, fontsize=28)
+    # plt.title("Linkage Quality w.r.t. Threshold", fontsize=22)
+    # plt.suptitle(title, y = 1.02, fontsize=28)
 
     plt.xlabel("Distance Threshold")
     plt.ylabel("Linkage Quality")
@@ -114,7 +114,8 @@ def plotLQ(title, predicate):
             , label="F1 Measure"
             )
 
-    plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
+    # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
+    plt.legend()
 
     plt.savefig("plotLQ -- %s.png" % title, bbox_inches="tight")
     plt.close()
@@ -238,7 +239,8 @@ def plotBQ(title, predicate):
             , label="Block Quality"
             )
 
-    plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
+    # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
+    plt.legend()
 
     plt.savefig("plotBQ -- %s.png" % title, bbox_inches="tight")
     plt.close()
@@ -284,7 +286,8 @@ def plotPQC(title, predicate):
             , label="Pairs Completeness"
             )
 
-    plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
+    # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
+    plt.legend()
 
     plt.savefig("plotPQC -- %s.png" % title, bbox_inches="tight")
     plt.close()
@@ -438,7 +441,8 @@ def plotFixedThresholdTraditional1(title, blockingMethods, predicate):
             , label="F1 Measure"
             )
 
-    plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
+    # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
+    plt.legend()
 
[…]

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 \pagestyle{headings}
-
-% ====================================================================
-
-% TODO:
-% - clearly point out novelty of this work
-% - contribution compared to earlier work
 
@@
 
-\cite{Li2006} describes a threshold-based linkage method
-using R-trees~\cite{Hjaltason1998}. It demonstrates that high quality
-may be achieved using Jaccard similarity over selected record fields.
-\cite{Ciaccia97indexingmetric} shows that M-trees are
-almost always more efficient than R-trees, hence their use in the
-experiments described in this paper.
+\cite{Li2006} describes a threshold-based linkage method using
+R-trees~\cite{Hjaltason1998}. It demonstrates that high quality may be
+achieved using Jaccard similarity over selected record fields.
+\cite{Ciaccia97indexingmetric} shows that M-trees are almost always more
+efficient than R-trees, hence their use in the experiments described in
+this paper.
 
@@
 
-In some circumstances, incomplete approaches such as traditional blocking and LSH-MinHash can yield
-higher precision than complete techniques. This can occur when the
-datasets contain a significant number of non-matches that nonetheless
-have high similarity. In this situation, the fact that an incomplete
-technique omits consideration of some potential links can serve to
-improve precision, since a classification decision based on similarity
-threshold is incorrect for high-similarity non-matches. By definition,
-recall is never higher for incomplete techniques.
+In some circumstances, incomplete approaches such as traditional
+blocking and LSH-MinHash can yield higher precision than complete
+techniques. This can occur when the datasets contain a significant
+number of non-matches that nonetheless have high similarity. In this
+situation, the fact that an incomplete technique omits consideration of
+some potential links can serve to improve precision, since a
+classification decision based on similarity threshold is incorrect for
+high-similarity non-matches. By definition, recall is never higher for
+incomplete techniques.
 
@@
 record in A is inserted into an M-tree. To perform linkage, the
-algorithm iterates over each record $b$ in $B$. A $range-search(b,d)$
+algorithm iterates over each record $b$ in $B$. A \emph{range-search(b,d)}
 operation is performed, passing the distance threshold as the second
@@
   Dataset~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities\\
-  name(s)  & dataset1  & dataset2  & matching pairs & linked \\
+  name(s)  & dataset 1  & dataset 2  & matching pairs & linked \\
   \noalign{\smallskip} \hline \noalign{\smallskip}
-  CORA  & 1,295 & 1295 & 17,184 & publication-publication\\
+  Cora & 1,295 & 1,295 & 17,184 & publication-publication\\
   Isle of Skye & 17,612 & 12,284& 2,900 & birth-death\\
@@
 
-The other datasets are historical Scottish records of vital
-events (birth, marriages and deaths) one registered on the
-\emph{Isle of Skye}, a rural district, and the other
-records from \emph{Kilmarnock}, an industrial town. These datasets
-were create by historical demographers who extensively
-curated and linked both datasets~\cite{reid2002,reid2006}. Both data
-sets include the  names, gender, addresses of individuals and their
-parents. Ground truth was generated by the demographers based on their
-extensive domain knowledge.
+The other datasets are historical Scottish records of vital events
+(birth, marriages and deaths) one registered on the \emph{Isle of Skye},
+a rural district, and the other records from \emph{Kilmarnock}, an
+industrial town. These datasets were create by historical demographers
+who extensively curated and linked both
+datasets~\cite{reid2002,reid2006}. Both data sets include the names,
+gender, addresses of individuals and their parents. Ground truth was
+generated by the demographers based on their extensive domain knowledge.
 
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-% \begin{table}[t]
-% \caption{Parameter settings used for the different datasets used in
-%    the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
-%    shingle size, and the size and number of bands, respectively.}
-%  \label{table-parameters}
-%   \centering
-%   \begin{scriptsize}
-%   %\addtolength{\tabcolsep}{-0.5pt}
-%   \begin{tabular}{ccccc}
-%   \hline\noalign{\smallskip}
-%   Dataset~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
-%   name(s) & & & &  \\
-%   \noalign{\smallskip} \hline \noalign{\smallskip}
-%   CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
-%     ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
-%   Isle of Skye &  \\
-%   Kilmarnock  &  \\
-%   \noalign{\smallskip} \hline
-%   \end{tabular}
-%   \end{scriptsize}
-% \end{table}
-
-% \textbf{Experimental Setup:}
-% %
-% We implemented all techniques described in Sect.~\ref{sec-approach}
-% using Java (version 1.8.0\_144) and ran an extensive set of experiments on
-% a compute server with 24 Intel(R) Xeon(R) processors running at 2.30GHz, 
-% 128 GBytes of main memory and running Red Hat 4.8.5-11.
-
-The parameter settings used in our experiments are summarized in
-Table~\ref{table-parameters}.
-
-In all of our experiments we use a single distance metric which is the sum of the Levenshtein distances between the set of fields being compared.
+In all of our experiments we use a single distance metric which is the
+sum of the Levenshtein~\cite{Levenshtein66} distances between the set of
+fields being compared.
 
@@
 
-In this section, we perform linkage on the Cora dataset using all four approaches presented in this paper: brute force, traditional blocking, LSH and M-Tree.
-The brute-force approach is used to establish the threshold value that yields the best possible linkage for the Cora dataset.
-The threshold value is varied between 0 and 250, \Cref{fig-cora-brute} presents the precision, recall, and F-measure values.
-The linkage results have high precision and low recall for low threshold values, and high recall and low precision for high threshold values.
-For this dataset, the maximum F-measure value is achieved using a threshold value of 70.
-This value is data dependent; for different datasets the maximum F-measure will correspond to different thresholds.
-In the rest of this section we fix the threshold value to be 70 which maximises the F-measure for this dataset.
+In this section, we perform linkage on the Cora dataset using all four
+approaches presented in this paper: brute force, traditional blocking,
+LSH and M-Tree. The brute-force approach is used to establish the
+threshold value that yields the best possible linkage for the Cora
+dataset. The threshold value is varied between 0 and 250,
+\Cref{fig-cora-brute} presents the precision, recall, and F-measure
+values. The linkage results have high precision and low recall for low
+threshold values, and high recall and low precision for high threshold
+values. For this dataset, the maximum F-measure value is achieved using
+a threshold value of 70. This value is data dependent; for different
+datasets the maximum F-measure will correspond to different thresholds.
+In the rest of this section we fix the threshold value to be 70 which
+maximises the F-measure for this dataset.
 
@@
 
-In the traditional blocking experiment the following fields from the Cora dataset are used as a blocking key: \textit{author}, \textit{title}, \textit{venue}, \textit{location}, \textit{publisher} and \textit{year}.
-We also use a very permissive blocking key which takes the union of all the values for these fields.
+In the traditional blocking experiment the following fields from the Cora dataset are used as individually as blocking keys: \textit{author}, \textit{title}, \textit{venue}, \textit{location}, \textit{publisher} and \textit{year}.
+We also use a combined blocking key comprising these fields.
 
 \Cref{comparison-of-results-cora} shows the results of the Cora experiments.
-The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the \textit{pairs quality} and\textit{ pairs completeness} achieved by each algorithm.
+The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and the F1 measure achieved by each algorithm.
 The linker column may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the number of the bands followed by the size of bands.
 For traditional blocking the parameters are the fields used for blocking.
-\textit{Pairs quality} and \textit{pairs completeness}  are given to illustrate the efficacy of the algorithms and are defined as follows: XXXXXXXXXXX.
-
-\textit{Pairs quality}  relates to recall and gives an indication of the number of (correct) matches amongst the links. 
-Pairs completeness BLAH BLAH.
 
@@
 
-TODO: Explain pairs-quality and pairs-completeness.
 \emph{\textbf{TODO: discuss measures used. num comparisons, time? etc.}}
@@
 \centering
-\begin{tabular}{l|r|r|r|r|r|r}
-\parbox{0.18\linewidth}{\centering Linker} &
-\parbox{0.16\linewidth}{\centering Comparisons} &
-\parbox{0.11\linewidth}{\centering Pairs Quality} &
-\parbox{0.17\linewidth}{\centering Pairs \\ Completeness} &
-\parbox{0.11\linewidth}{\centering Precision} &
-\parbox{0.08\linewidth}{\centering Recall} &
-\parbox{0.11\linewidth}{\centering F1 Measure} \\ \hline \hline
-Brute Force        & 1,677,025                    & 0.02                  & 0.33                       & 0.84      & 0.57   & 0.68       \\ \hline
-MTree              &  902,693                     & 0.89                  & 0.22                       & 0.84      & 0.57   & 0.68       \\ \hline
-LSH-2-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
-LSH-5-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
-LSH-10-2           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
-LSH-2-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
-LSH-5-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
-LSH-10-5           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
-LSH-2-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
-LSH-5-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
-LSH-10-10          &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\  \hline
-Block-year      &  115,893                     & 0.22                  & 0.19                       & 0.99      & 0.35   & 0.51       \\
-Block-authors   &   11,039                      & 0.74                  & 0.1                        & 0.94      & 0.16   & 0.28       \\
-Block-title     &   27,407                      & 0.84                  & 0.23                       & 0.95      & 0.42   & 0.58       \\
-Block-venue     &   36,647                      & 0.7                   & 0.12                       & 0.85      & 0.29   & 0.44       \\
-Block-location  & 1,009,957                    & 0.17                  & 0.24                       & 0.83      & 0.43   & 0.57       \\
-Block-publisher &  833,079                     & 0.12                  & 0.23                       & 0.85      & 0.44   & 0.58       \\
-Block-combined  & 1,214,269                    & 0.05                  & 0.32                       & 0.84      & 0.56   & 0.67       \\ \hline
+\begin{tabular}{l|r|r|r|r}
+\parbox{0.2\textwidth}{\centering Linker} &
+\parbox{0.2\textwidth}{\centering Comparisons} &
+\parbox{0.17\textwidth}{\centering Precision} &
+\parbox{0.17\textwidth}{\centering Recall} &
+\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
+Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68       \\ \hline
+MTree              &  902,693                  & 0.84      & 0.57   & 0.68       \\ \hline
+LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
+LSH-5-2            &  342,849	& 0.91	& 0.55	& 0.69       \\
+LSH-10-2           &  513,947	& 0.88	& 0.57	& 0.69       \\
+LSH-2-5            &  14,329	& 0.99	& 0.28	& 0.43      \\
+LSH-5-5            &  22,057	& 0.99	& 0.36	& 0.53       \\
+LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
+LSH-2-10           &  4,711	& 1	& 0.15	& 0.27    \\
+LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
+LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\  \hline
+Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
+Block-authors      &   11,039                   & 0.94      & 0.16   & 0.28       \\
+Block-title        &   27,407                   & 0.95      & 0.42   & 0.58       \\
+Block-venue        &   36,647                   & 0.85      & 0.29   & 0.44       \\
+Block-location     & 1,009,957                 & 0.83      & 0.43   & 0.57       \\
+Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
+Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\ \hline
 \end{tabular}
@@
 \centering
-\begin{tabular}{l|r|r|r|r|r|r}
-\parbox{0.18\linewidth}{\centering Linker} &
-\parbox{0.16\linewidth}{\centering Comparisons} &
-\parbox{0.11\linewidth}{\centering Pairs Quality} &
-\parbox{0.17\linewidth}{\centering Pairs \\ Completeness} &
-\parbox{0.11\linewidth}{\centering Precision} &
-\parbox{0.08\linewidth}{\centering Recall} &
-\parbox{0.11\linewidth}{\centering F1 Measure} \\ \hline \hline
-MTree    & 102,318,525         & 0.84                  & 0.86                       & 0.65      & 0.46   & 0.54      \\ \hline
-LSH-2-2  & 3,109,250           & 0                     & 0.05                       & 0.63      & 0.03   & 0.06       \\
-LSH-5-2  & 10,412,496          & 0                     & 0.16                       & 0.64      & 0.11   & 0.19       \\
-LSH-10-2 & 53,874,127          & 0                     & 0.46                       & 0.68      & 0.36   & 0.47       \\
-LSH-5-5  & 36,566             & 0                     & 0.01                       & 0.76      & 0.01   & 0.01       \\
-LSH-10-5 & 129,873            & 0                     & 0.01                       & 0.72      & 0.01   & 0.02       \\ \hline
+\begin{tabular}{l|r|r|r|r}
+\parbox{0.2\textwidth}{\centering Linker} &
+\parbox{0.2\textwidth}{\centering Comparisons} &
+\parbox{0.17\textwidth}{\centering Precision} &
+\parbox{0.17\textwidth}{\centering Recall} &
+\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
+MTree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline
+
+LSH-2-2  & 3,109,250                   & 0.63      & 0.03   & 0.06       \\
+
+LSH-5-2  & 10,412,496                  & 0.64      & 0.11   & 0.19       \\
+
+LSH-10-2 & 53,874,127                  & 0.68      & 0.36   & 0.47       \\
+
+LSH-5-5  & 36,566                     & 0.76      & 0.01   & 0.01       \\
+
+LSH-10-5 & 129,873                    & 0.72      & 0.01   & 0.02       \\ \hline
 \end{tabular}
@@
 
+\Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} shows the results for LSH-Minhash and M-tree applied to The Isle of Skye and Kilmarnock datasets.
+The shingle-size was set to 2 for all the LSH experiments shown as this was found to give good results and LSH was not especially sensitive to this parameter.
+Results for other shingle sizes were omitted from this paper for brevity.
+In these experiments the linkage performed is between birth and death records.
+In both cases the F1 measure achieved is better for M-tree than any of the experimental settings chosen for LSH.
+This is largely due to the recall for M-tree being much higher than any of the LSH experiments.
+For all LSH parameters LSH out performs M-tree in terms of precision.
+
+Perhaps more important than the fact that M-tree gives better F1 measures than all of the LSH experiments is the fact that the LSH performance is extremely variable.
+For a relatively plausible settings for the number of bands and band-size the F1 measurements vary from 0.01 (extremely poor) to 0.47 (relatively good) for Skye and between 0.03 and 0.49 for the Kilmarnock dataset.
+In both cases LSH-10-2 performs best, but since this is data dependent there is no guarantee these parameters would work well with another dataset.
+Furthermore, without good ground truth, which is usually unavailable, there is no way of knowing if good linkage results are being obtained or not.
+
+Another observation that is worthy of comment is the number of (distance) comparisons being performed by the different algorithms.
+For M-trees the number of comparisons are closely related to the complexity of the search being performed since the algorithms are performing (range) searches over distance measures to determine the query results.
+This is not true of LSH which is performing Jaccard similarity using Hashes over the datasets.
+Comparisons are only used at the final step to determine if a candidate is a match or not.
+The LSH results which provide the best results by F1 measure are performing between a quarter and half the number of matches performed by M-trees.
+From this we can conclude that the hash algorithms are returning many candidates that are outwith the search distance.
+Thus, in order to get good results, LSH is tending towards a brute force search over the candidate results.
+Despite this LSH is much faster due to the efficiency of the hashing mechanisms.
+
 \begin{table}[ht]
@@
 \begin{tabular}{l|r|r|r|r|r|r}
-\parbox{0.18\linewidth}{\centering Linker} &
-\parbox{0.16\linewidth}{\centering Comparisons} &
-\parbox{0.11\linewidth}{\centering Pairs Quality} &
-\parbox{0.17\linewidth}{\centering Pairs \\ Completeness} &
-\parbox{0.11\linewidth}{\centering Precision} &
-\parbox{0.08\linewidth}{\centering Recall} &
-\parbox{0.11\linewidth}{\centering F1 Measure} \\ \hline \hline
-MTree     & 514,871,153         & 0.89                  & 0.82                       & 0.76      & 0.45   & 0.57      \\ \hline
-LSH-2-2   & 99,145,887          & 0                     & 0.23                       & 0.81      & 0.16   & 0.27       \\
-LSH-5-2   & 130,721,338         & 0                     & 0.32                       & 0.79      & 0.23   & 0.36       \\
-LSH-10-2  & 177,168,848         & 0                     & 0.48                       & 0.79      & 0.36   & 0.49       \\
-LSH-5-5   & 239,368            & 0.01                  & 0.02                       & 0.84      & 0.01   & 0.02       \\
-LSH-10-5  & 855,431            & 0.01                  & 0.03                       & 0.87      & 0.02   & 0.03 \\ \hline
+\parbox{0.2\textwidth}{\centering Linker} &
+\parbox{0.2\textwidth}{\centering Comparisons} &
+\parbox{0.17\textwidth}{\centering Precision} &
+\parbox{0.17\textwidth}{\centering Recall} &
+\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
+MTree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\ \hline
+LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
+LSH-5-2   & 130,721,338                   & 0.79      & 0.23   & 0.36       \\
+LSH-10-2  & 177,168,848                   & 0.79      & 0.36   & 0.49       \\
+LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
+LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ \hline
 \end{tabular}
@@
 Sect.~\ref{sec-exp}.
+
+% \begin{table}[t]
+% \caption{Parameter settings used for the different datasets used in
+%    the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
+%    shingle size, and the size and number of bands, respectively.}
+%  \label{table-parameters}
+%   \centering
+%   \begin{scriptsize}
+%   %\addtolength{\tabcolsep}{-0.5pt}
+%   \begin{tabular}{ccccc}
+%   \hline\noalign{\smallskip}
+%   Dataset~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
+%   name(s) & & & &  \\
+%   \noalign{\smallskip} \hline \noalign{\smallskip}
+%   CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
+%     ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
+%   Isle of Skye &  \\
+%   Kilmarnock  &  \\
+%   \noalign{\smallskip} \hline
+%   \end{tabular}
+%   \end{scriptsize}
+% \end{table}
+
+% \textbf{Experimental Setup:}
+% %
+% We implemented all techniques described in Sect.~\ref{sec-approach}
+% using Java (version 1.8.0\_144) and ran an extensive set of experiments on
+% a compute server with 24 Intel(R) Xeon(R) processors running at 2.30GHz, 
+% 128 GBytes of main memory and running Red Hat 4.8.5-11.

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 \begin{figure}
-\includegraphics[width=\textwidth]{figures/plotLQ_--_Cora_-_Brute_Force.png}
-\caption{Linkage Quality using Brute Force linking on the Cora dataset\label{fig-cora-brute}}
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
+\caption{Cora - Brute Force}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
+\caption{Cora - M-Tree}
+\end{subfigure}
+
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
+\caption{Cora - Traditional Blocking on the `authors' field}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
+\caption{Cora - Traditional Blocking on the `title' field}
+\end{subfigure}
+
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
+\caption{Cora - Traditional Blocking on the `publication year' field}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
+\caption{Cora - Traditional Blocking using the union of blocking on all fields}
+\end{subfigure}
+
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
+\caption{Cora - LSH Blocking using 2 bands of size 2}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
+\caption{Cora - LSH Blocking using 2 bands of size 10}
+\end{subfigure}
+
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
+\caption{Cora - LSH Blocking using 10 bands of size 2}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
+\caption{Cora - LSH Blocking using 10 bands of size 10}
+\end{subfigure}
+
+\caption{Linkage results on the Cora data set}
+\label{demography-quality-mtree}
 \end{figure}
 
-In the traditional blocking experiment the following fields from the Cora dataset are used as individually as blocking keys: \textit{author}, \textit{title}, \textit{venue}, \textit{location}, \textit{publisher} and \textit{year}.
-We also use a combined blocking key comprising these fields.
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
+\caption{Linkage quality using brute-force linking on the Cora dataset\label{fig-cora-brute}}
+\end{figure}
 
-\Cref{comparison-of-results-cora} shows the results of the Cora experiments.
-The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and the F1 measure achieved by each algorithm.
-The linker column may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the number of the bands followed by the size of bands.
-For traditional blocking the parameters are the fields used for blocking.
+In the traditional blocking experiment the following fields from the
+Cora dataset are used individually as blocking keys: \textit{author},
+\textit{title}, \textit{venue}, \textit{location}, \textit{publisher}
+and \textit{year}. We also use a combined blocking key comprising all
+fields.
 
-As can be seen in \Cref{comparison-of-results-cora}, M-tree reports the same results as brute-force....
+\Cref{comparison-of-results-cora} shows the results of the Cora
+experiments. The table shows the parameters for the experiment, the
+number of distance comparisons made, the precision, recall and the F1
+measure achieved by each algorithm.  In the \emph{Linker} column the algorithm name is followed by its
+parameters: for LSH the number of the bands followed by the band size, and for traditional blocking the fields used
+for blocking.
 
-\emph{\textbf{TODO: discuss measures used. num comparisons, time? etc.}}
+M-tree yields the same linkage quality as brute-force, although using a significantly lower
+number of comparisons.
 
@@
 \end{tabular}
-\caption{Linkage Quality on the Cora dataset distance=70\label{comparison-of-results-cora}
-\textbf{TODO: The table needs some rows removed. Note that this table is with a threshold value of 70.}
+\caption{Linkage quality on Cora dataset with distance threshold = 70
+\label{comparison-of-results-cora}
 }
@@
 \subsection{Historical Demography results}
+
+\Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} show the results for LSH-Minhash and M-tree on the Skye and Kilmarnock datasets.
+Linkage was performed between birth and death records.
+The shingle size was set to 2 for all the LSH experiments reported, as this was found to give good results and LSH was not especially sensitive to this parameter.
+Results for other shingle sizes were omitted from this paper for brevity.
+In both cases the F1 measure achieved is better for M-tree than any of the experimental settings chosen for LSH.
+This is largely due to the recall for M-tree being much higher than any of the LSH experiments.
+For all LSH parameters LSH out performs M-tree in terms of precision.
+
+Perhaps more important than the fact that M-tree gives better F1 measures than all of the LSH experiments is the fact that the LSH performance is extremely variable.
+For a relatively plausible settings for the number of bands and band-size the F1 measurements vary from 0.01 (extremely poor) to 0.47 (relatively good) for Skye and between 0.03 and 0.49 for the Kilmarnock dataset.
+In both cases LSH-10-2 performs best, but since this is data dependent there is no guarantee these parameters would work well with another dataset.
+Furthermore, without good ground truth, which is usually unavailable, there is no way of knowing if good linkage results are being obtained or not.
+
+Another observation that is worthy of comment is the number of (distance) comparisons being performed by the different algorithms.
+For M-trees the number of comparisons are closely related to the complexity of the search being performed since the algorithms are performing (range) searches over distance measures to determine the query results.
+This is not true of LSH which is performing Jaccard similarity using Hashes over the datasets.
+Comparisons are only used at the final step to determine if a candidate is a match or not.
+The LSH results which provide the best results by F1 measure are performing between a quarter and half the number of matches performed by M-trees.
+From this we can conclude that the hash algorithms are returning many candidates that are outwith the search distance.
+Thus, in order to get good results, LSH is tending towards a brute force search over the candidate results.
+Despite this LSH is much faster due to the efficiency of the hashing mechanisms.
 
@@
 \end{tabular}
-\caption{Linkage Quality on demography datasets \textbf{This is for Skye. Threshold value is 2.}
+\caption{Linkage quality on Skye dataset with distance threshold = 2
 \label{comparison-of-results-demography-skye}
@@
 
-\Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} shows the results for LSH-Minhash and M-tree applied to The Isle of Skye and Kilmarnock datasets.
-The shingle-size was set to 2 for all the LSH experiments shown as this was found to give good results and LSH was not especially sensitive to this parameter.
-Results for other shingle sizes were omitted from this paper for brevity.
-In these experiments the linkage performed is between birth and death records.
-In both cases the F1 measure achieved is better for M-tree than any of the experimental settings chosen for LSH.
-This is largely due to the recall for M-tree being much higher than any of the LSH experiments.
-For all LSH parameters LSH out performs M-tree in terms of precision.
-
-Perhaps more important than the fact that M-tree gives better F1 measures than all of the LSH experiments is the fact that the LSH performance is extremely variable.
-For a relatively plausible settings for the number of bands and band-size the F1 measurements vary from 0.01 (extremely poor) to 0.47 (relatively good) for Skye and between 0.03 and 0.49 for the Kilmarnock dataset.
-In both cases LSH-10-2 performs best, but since this is data dependent there is no guarantee these parameters would work well with another dataset.
-Furthermore, without good ground truth, which is usually unavailable, there is no way of knowing if good linkage results are being obtained or not.
-
-Another observation that is worthy of comment is the number of (distance) comparisons being performed by the different algorithms.
-For M-trees the number of comparisons are closely related to the complexity of the search being performed since the algorithms are performing (range) searches over distance measures to determine the query results.
-This is not true of LSH which is performing Jaccard similarity using Hashes over the datasets.
-Comparisons are only used at the final step to determine if a candidate is a match or not.
-The LSH results which provide the best results by F1 measure are performing between a quarter and half the number of matches performed by M-trees.
-From this we can conclude that the hash algorithms are returning many candidates that are outwith the search distance.
-Thus, in order to get good results, LSH is tending towards a brute force search over the candidate results.
-Despite this LSH is much faster due to the efficiency of the hashing mechanisms.
+%-------------------------------------------------------------------------------------
 
@@
 \end{tabular}
-\caption{Linkage Quality on demography datasets\textbf{This is for Kilmarnock. Threshold value is 2.}
+\caption{Linkage quality on Kilmarnock dataset with distance threshold = 2
 \label{comparison-of-results-demography-kili}
@@
 \end{table}
+
+%-------------------------------------------------------------------------------------
 
@@
 A second unexpected result is that shown in \Cref{fig-cora-brute}, namely that LSH can yield higher precision than is achieved by a complete method such as M-Tree as was discussed in Sect.~\ref{sec-approach}.
-
-\emph{IS THIS RIGHT?}.
 

### 2017-11-21 scripts/plots.py
Generate eps

@@ -117,9 +117,9 @@ def plotLQ(title, predicate):
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
     plt.legend()
 
-    plt.savefig("plotLQ -- %s.png" % title, bbox_inches="tight")
+    plt.savefig("plotLQ -- %s.eps" % title, bbox_inches="tight")
     plt.close()
-    print("plotLQ -- %s.png" % title)
+    print("plotLQ -- %s.eps" % title)
 
 
 dataset = "Cora"
@@ -242,9 +242,9 @@ def plotBQ(title, predicate):
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
     plt.legend()
 
-    plt.savefig("plotBQ -- %s.png" % title, bbox_inches="tight")
+    plt.savefig("plotBQ -- %s.eps" % title, bbox_inches="tight")
     plt.close()
-    print("plotBQ -- %s.png" % title)
+    print("plotBQ -- %s.eps" % title)
 
 
 
@@ -289,9 +289,9 @@ def plotPQC(title, predicate):
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
     plt.legend()
 
-    plt.savefig("plotPQC -- %s.png" % title, bbox_inches="tight")
+    plt.savefig("plotPQC -- %s.eps" % title, bbox_inches="tight")
     plt.close()
-    print("plotPQC -- %s.png" % title)
+    print("plotPQC -- %s.eps" % title)
 
 
 
@@ -444,9 +444,9 @@ def plotFixedThresholdTraditional1(title, blockingMethods, predicate):
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
     plt.legend()
 
-    plt.savefig("plotFixedThresholdTraditional1 -- %s.png" % title, bbox_inches="tight")
+    plt.savefig("plotFixedThresholdTraditional1 -- %s.eps" % title, bbox_inches="tight")
     plt.close()
-    print("plotFixedThresholdTraditional1 -- %s.png" % title)
+    print("plotFixedThresholdTraditional1 -- %s.eps" % title)
 
 
 def plotFixedThresholdTraditional2(title, blockingMethods, predicate):
@@ -494,9 +494,9 @@ def plotFixedThresholdTraditional2(title, blockingMethods, predicate):
[…]

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
-\caption{Linkage quality using brute-force linking on the Cora dataset\label{fig-cora-brute}}
-\end{figure}
+% \begin{figure}
+% \includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
+% \caption{Linkage quality using brute-force linking on the Cora dataset\label{fig-cora-brute}}
+% \end{figure}
 
@@
 \end{subfigure}
-\caption{M-Tree based linkage on the demographic data sets}
+
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
+\caption{F-Measure values of all LSH variations vs M-Tree for the Skye data set}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
+\caption{F-Measure values of all LSH variations vs M-Tree for the Kilmarnock data set}
+\end{subfigure}
+
+\caption{Linkage results on the demographic data set}
 \label{demography-quality-mtree}
@@
 
-\begin{figure}
-\includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
-\caption{Comparison of quality attainable on Skye dataset using M-tree and various LSH configurations\label{mtree-lsh-quality}}
-\end{figure}
+% \begin{figure}
+% \includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
+% \caption{Comparison of quality attainable on Skye dataset using M-tree and various LSH configurations\label{mtree-lsh-quality}}
+% \end{figure}
 

### 2017-11-21 scripts/plots.py
multiple f measures plot

@@ -408,7 +408,7 @@ def plotFixedThresholdTraditional1(title, blockingMethods, predicate):
     plt.ylabel("Proportion")
 
     data = [ line for line in stats if predicate(line) ]
-    
+
     # plt.xlim([ line["Blocking Method"]       for line in data ]            )
     # plt.xlim([ line["Blocking Method"] for line in data ])
     plt.ylim(-0.05, 1.05)
@@ -466,7 +466,7 @@ def plotFixedThresholdTraditional2(title, blockingMethods, predicate):
     plt.ylabel("Proportion")
 
     data = [ line for line in stats if predicate(line) ]
-    
+
     # plt.xlim([ line["Blocking Method"]       for line in data ]            )
     # plt.xlim([ line["Blocking Method"] for line in data ])
     plt.ylim(-0.05, 1.05)
@@ -516,7 +516,7 @@ def plotFixedThresholdLSH(title, predicate):
     plt.ylabel("Proportion")
 
     data = [ line for line in stats if predicate(line) ]
-    
+
     # plt.xlim([ line["Blocking Method"]       for line in data ]            )
     # plt.xlim([ line["Blocking Method"] for line in data ])
     plt.ylim(-0.05, 1.05)
@@ -589,4 +589,99 @@ plotFixedThresholdLSH(title, predicate)
 
 
 
+
+
+
+
+
+
+################################################################################
+# Plotting F-Measures against each other
+################################################################################
+
+def plotFs(title, measure, predicate):
+    title = " - ".join([title, measure])
+    plt.clf()
+
+    w,h = plt.figaspect(0.5)
+    plt.figure(figsize=(w,h))
+
+    font = {"family":"sans-serif", "weight":"normal", "size":24}
+    plt.rc("font", **font)
+    # plt.rcParams["text.usetex"] = True
+
+    # plt.title("Linkage Quality w.r.t. Threshold", fontsize=22)
+    # plt.suptitle(title, y = 1.02, fontsize=28)
+
+    plt.xlabel("Distance Threshold")
+    plt.ylabel(measure)
+
+    data = [ line for line in stats if predicate(line) ]
+
+    plt.xlim( min([ toFloat(line["Distance Threshold"]) for line in data ])
+            , max([ toFloat(line["Distance Threshold"]) for line in data ])
+            )
+    plt.ylim(-0.05, 1.05)
+    plt.grid()
+
+    shingleSizes = set([ line["Shingle size"] for line in data if predicate(line) ])
[…]

### 2017-11-21 scripts/plots.py
Correct spelling of M-tree

@@ -646,7 +646,7 @@ def plotFs(title, measure, predicate):
             , [ line[measure]               for line in dataMTree ]
             , color="red"
             , marker="o"
-            , label="M-Tree"
+            , label="M-tree"
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 We propose using metric space indexing to perform \emph{complete} record
-linkage, which results in a parameterless record linkage process
+linkage, which results in a parameter-free record linkage process
 combining indexing, comparison and classification into a single step
@@
 approaches presented in this paper: brute force, traditional blocking,
-LSH and M-Tree. The brute-force approach is used to establish the
+LSH and M-tree. The brute-force approach is used to establish the
 threshold value that yields the best possible linkage for the Cora
 dataset. The threshold value is varied between 0 and 250,
-\Cref{fig-cora-brute} presents the precision, recall, and F-measure
+\Cref{cora-quality} presents the precision, recall, and F-measure
 values. The linkage results have high precision and low recall for low
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
-\caption{Cora - M-Tree}
+\caption{Cora - M-tree}
 \end{subfigure}
@@
 \caption{Linkage results on the Cora data set}
-\label{demography-quality-mtree}
+\label{cora-quality}
 \end{figure}
@@
 
-\Cref{comparison-of-results-cora} shows the results of the Cora
-experiments. The table shows the parameters for the experiment, the
-number of distance comparisons made, the precision, recall and the F1
-measure achieved by each algorithm.  In the \emph{Linker} column the algorithm name is followed by its
-parameters: for LSH the number of the bands followed by the band size, and for traditional blocking the fields used
-for blocking.
+\Cref{comparison-of-results-cora} shows greater detail for a subset of
+the Cora experiments, with the distance threshold fixed at 70. The table
+shows the parameters for the experiment, the number of distance
+comparisons made, the precision, recall and the F1 measure achieved by
+each algorithm.  In the \emph{Linker} column the algorithm name is
+followed by its parameters: for LSH the number of the bands followed by
+the band size, and for traditional blocking the fields used for
+blocking. The number of distance comparisons is reported as a
+machine-independent proxy for execution cost, since code profiling shows
+that distance calculations are dominant.
 
 M-tree yields the same linkage quality as brute-force, although using a significantly lower
-number of comparisons.
+number of comparisons. This is as expected, since both techniques are complete.
+
+Three of the LSH configurations give similar quality
 
@@
 \includegraphics[width=\textwidth]{figures/plotFs-skye-f}
-\caption{F-Measure values of all LSH variations vs M-Tree for the Skye data set}
+\caption{Linkage quality on Skye dataset using M-tree and all LSH configurations}
 \end{subfigure}%
@@
 \includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
-\caption{F-Measure values of all LSH variations vs M-Tree for the Kilmarnock data set}
+\caption{Linkage quality on Kilmarnock dataset using M-tree and all LSH configurations}
 \end{subfigure}
 
-\caption{Linkage results on the demographic data set}
+\caption{Linkage results on the demographic datasets}
 \label{demography-quality-mtree}
 \end{figure}
-
-% \begin{figure}
-% \includegraphics[width=\textwidth]{figures/plotBQ_--_Cora_-_MTree}
-% \caption{Comparison of quality attainable on Skye dataset using M-tree and various LSH configurations\label{mtree-lsh-quality}}
-% \end{figure}
 
@@
 Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase.
-A second unexpected result is that shown in \Cref{fig-cora-brute}, namely that LSH can yield higher precision than is achieved by a complete method such as M-Tree as was discussed in Sect.~\ref{sec-approach}.
+A second unexpected result is that shown in \Cref{quality-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.
 

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 
-% ====================================================================
+%----------------------------------------------------------------------
 
@@
 \fi
-% ====================================================================
+%----------------------------------------------------------------------
 
@@
 
+%----------------------------------------------------------------------
+
 \begin{figure}[!t]
@@
 \end{figure}
+
+%----------------------------------------------------------------------
 
@@
 
+%----------------------------------------------------------------------
+
 % For final paper:
@@
 
-% --------------------------------------------------------------------
+%----------------------------------------------------------------------
 
@@
 our results~\footnote{The raw data, additional figures and the source
-code required to run these experiments can be downloaded from here.
-(Link removed for double blind review)}.
+code required to run these experiments can be downloaded from:
+(link removed for double blind review).}.
 
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+%----------------------------------------------------------------------
 
@@
 
+%----------------------------------------------------------------------
+
 \smallskip
@@
 who extensively curated and linked both
-datasets~\cite{reid2002,reid2006}. Both data sets include the names,
+datasets~\cite{reid2002,reid2006}. Both datasets include the names,
 gender, addresses of individuals and their parents. Ground truth was
@@
 
-\subsection{Cora}
+\subsection{Cora Results}
 
@@
 approaches presented in this paper: brute force, traditional blocking,
-LSH and M-tree. The brute-force approach is used to establish the
-threshold value that yields the best possible linkage for the Cora
-dataset. The threshold value is varied between 0 and 250,
-\Cref{cora-quality} presents the precision, recall, and F-measure
-values. The linkage results have high precision and low recall for low
-threshold values, and high recall and low precision for high threshold
-values. For this dataset, the maximum F-measure value is achieved using
-a threshold value of 70. This value is data dependent; for different
-datasets the maximum F-measure will correspond to different thresholds.
-In the rest of this section we fix the threshold value to be 70 which
-maximises the F-measure for this dataset.
+LSH and M-tree, using several selected configurations for the blocking
+and LSH linkers. The distance threshold is varied between 0 and
+250~\footnote{Relatively high Levenshtein distances are included since
+Cora includes a number of low-similarity true matches.}. For traditional
+blocking, the following Cora fields are used individually as blocking
+keys: \textit{author}, \textit{title}, \textit{venue},
+\textit{location}, \textit{publisher} and \textit{year}. We also use a
+combined blocking key comprising all fields.
+
+\Cref{cora-quality} shows the precision, recall, and F-measure
+for various thresholds. As expected, low thresholds give high precision and low recall, and the reverse for high threshold
+values. As expected, brute force and
+
+For a more detailed investigation of selected linkers, the brute-force
+approach is used to establish the optimum threshold value for the Cora
+dataset. The maximum F-measure value is observed at a threshold value of
+70. This value is dataset-dependent; for different datasets the maximum
+F-measure will occur at different thresholds. In the rest of this
+section we fix the threshold value at 70.
+
+\Cref{comparison-of-results-cora} shows greater detail for selected
+linkers, with the distance threshold fixed at 70. The table shows the
+parameters for the experiment, the number of distance comparisons made,
+the precision, recall and the F1 measure achieved by each algorithm.  In
+the \emph{Linker} column the algorithm name is followed by its
+parameters: for LSH the number of the bands followed by the band size,
+and for traditional blocking the fields used for blocking. The number of
+distance comparisons is reported as a machine-independent proxy for
+execution cost, since code profiling shows that distance calculations
+are dominant.
+
+M-tree yields the same linkage quality as brute-force, although using a
+significantly lower number of comparisons. This is as expected, since
+both techniques are complete. Several of the incomplete linkers give
+similar quality, for example \emph{LSH-2-2}, \emph{LSH-5-2}, \emph{LSH-10-2} and
+\emph{Block-combined}. These, and a number of other incomplete linkers give
+better precision than the complete techniques. This is due to
+high-similarity non-matches, as discussed in Sect.~\ref{sec-approach}.
+Although several of the incomplete linkers give as good quality as
+M-tree, and in some cases at lower cost, this is offset by the need to
+select appropriate configuration parameters. Some other linkers give
+very poor results.
+
+%----------------------------------------------------------------------
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
-\caption{Cora - Traditional Blocking on the `authors' field}
+\caption{Cora - Traditional Blocking on `authors' field}
 \end{subfigure}%
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
-\caption{Cora - Traditional Blocking on the `title' field}
+\caption{Cora - Traditional Blocking on `title' field}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
-\caption{Cora - Traditional Blocking on the `publication year' field}
+\caption{Cora - Traditional Blocking on `publication year' field}
 \end{subfigure}%
@@
 
-\caption{Linkage results on the Cora data set}
+\caption{Linkage results on Cora dataset}
 \label{cora-quality}
@@
 
-% \begin{figure}
-% \includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
-% \caption{Linkage quality using brute-force linking on the Cora dataset\label{fig-cora-brute}}
-% \end{figure}
-
-In the traditional blocking experiment the following fields from the
-Cora dataset are used individually as blocking keys: \textit{author},
-\textit{title}, \textit{venue}, \textit{location}, \textit{publisher}
-and \textit{year}. We also use a combined blocking key comprising all
-fields.
-
-\Cref{comparison-of-results-cora} shows greater detail for a subset of
-the Cora experiments, with the distance threshold fixed at 70. The table
-shows the parameters for the experiment, the number of distance
-comparisons made, the precision, recall and the F1 measure achieved by
-each algorithm.  In the \emph{Linker} column the algorithm name is
-followed by its parameters: for LSH the number of the bands followed by
-the band size, and for traditional blocking the fields used for
-blocking. The number of distance comparisons is reported as a
-machine-independent proxy for execution cost, since code profiling shows
-that distance calculations are dominant.
-
-M-tree yields the same linkage quality as brute-force, although using a significantly lower
-number of comparisons. This is as expected, since both techniques are complete.
-
-Three of the LSH configurations give similar quality
+%----------------------------------------------------------------------
 
@@
 Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68       \\ \hline
-MTree              &  902,693                  & 0.84      & 0.57   & 0.68       \\ \hline
+M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\ \hline
 LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
@@
 
-\subsection{Historical Demography results}
+%----------------------------------------------------------------------
+
+\subsection{Demographic Dataset Results}
 
@@
 
+%----------------------------------------------------------------------
+
 \begin{table}[ht]
@@
 \parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
-MTree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline
+M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline
 
@@
 
-%-------------------------------------------------------------------------------------
+%----------------------------------------------------------------------
 
@@
 \parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
-MTree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\ \hline
+M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\ \hline
 LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
@@
 Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase.
-A second unexpected result is that shown in \Cref{quality-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.
+A second unexpected result is that shown in \Cref{comparison-of-results-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.
 

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 \emph{Metric space indexing} (MSI) is a complete technique with much
-lower computational overheads than a brute-force approach. It allows
+lower computational overheads than a brute force approach. It allows
 indexing, comparison and classification to be combined into a single
@@
 LSH-MinHash, and one complete method, M-tree~\cite{paolociaccia2m}. We
-also use a simple complete brute-force technique as a baseline for
+also use a simple complete brute force technique as a baseline for
 comparison, although this can only feasibly be applied to the smallest
@@
 
-\textbf{Brute-force:}
+\textbf{Brute force:}
 Two nested loops are used to compare every record in $A$ with every
@@
 for various thresholds. As expected, low thresholds give high precision and low recall, and the reverse for high threshold
-values. As expected, brute force and
+values. As expected, brute force and M-tree give identical results.
 
-For a more detailed investigation of selected linkers, the brute-force
+For a more detailed investigation of selected linkers, the brute force
 approach is used to establish the optimum threshold value for the Cora
@@
 
-M-tree yields the same linkage quality as brute-force, although using a
+M-tree yields the same linkage quality as brute force, although using a
 significantly lower number of comparisons. This is as expected, since
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
-\caption{Cora - Brute Force}
+\caption{Brute Force}
 \end{subfigure}%
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
-\caption{Cora - M-tree}
+\caption{M-tree}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
-\caption{Cora - Traditional Blocking on `authors' field}
+\caption{Blocking on `authors' field}
 \end{subfigure}%
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
-\caption{Cora - Traditional Blocking on `title' field}
+\caption{Blocking on `title' field}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
-\caption{Cora - Traditional Blocking on `publication year' field}
+\caption{Blocking on `publication year' field}
 \end{subfigure}%
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
-\caption{Cora - Traditional Blocking using the union of blocking on all fields}
+\caption{Blocking using the union of all fields}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
-\caption{Cora - LSH Blocking using 2 bands of size 2}
+\caption{LSH using 2 bands of size 2}
 \end{subfigure}%
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
-\caption{Cora - LSH Blocking using 2 bands of size 10}
+\caption{LSH using 2 bands of size 10}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
-\caption{Cora - LSH Blocking using 10 bands of size 2}
+\caption{LSH using 10 bands of size 2}
 \end{subfigure}%
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
-\caption{Cora - LSH Blocking using 10 bands of size 10}
+\caption{LSH using 10 bands of size 10}
 \end{subfigure}
 
-\caption{Linkage results on Cora dataset}
+\caption{Linkage results on the Cora dataset}
 \label{cora-quality}
@@
 
-\Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} show the results for LSH-Minhash and M-tree on the Skye and Kilmarnock datasets.
+\Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} show the results for LSH-MinHash and M-tree on the Skye and Kilmarnock datasets.
 Linkage was performed between birth and death records.
@@
 This is largely due to the recall for M-tree being much higher than any of the LSH experiments.
-For all LSH parameters LSH out performs M-tree in terms of precision.
+For most configurations, LSH out-performs M-tree in terms of precision.
 
@@
   \centering
-\includegraphics[width=\textwidth]{figures/skye-mtree}
+\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
 \caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
@@
   \centering
-\includegraphics[width=\textwidth]{figures/kili-mtree}
+\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
 \caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
@@
 In conclusion, this assertion deserves some careful unpacking.
-It is always possible to achieve high quality linkage using a brute-force approach.
+It is always possible to achieve high quality linkage using a brute force approach.
 However the n-squared complexity of this approach prevents its practical application for datasets of even moderate size.
-The use of MSI techniques such as M-tree delivers the same results as the brute-force approach with fewer comparisons.
-We have shown that the M-tree approach delivers high precision, high recall results that are the same as those delivered by the brute-force approach.
+The use of MSI techniques such as M-tree delivers the same results as brute force with fewer comparisons.
+We have shown that the M-tree approach delivers high precision, high recall results that are the same as those delivered by brute force.
 Furthermore this is achieved consistently without the need for complex parameter tuning.
@@
 However, this paper also highlights some unexpected results.
-Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute-force as the number of records in the block increase.
+Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute force as the number of records in the block increase.
 A second unexpected result is that shown in \Cref{comparison-of-results-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.

### 2017-11-21 paper.tex
Update on Overleaf.

@@
 combining indexing, comparison and classification into a single step
-delivering complete and efficient record linkage.
+delivering complete and efficient record linkage. Our experimental
+evaluation on real-world datasets from several domains shows that
+linkage using metric space indexing can yield better quality than
+current indexing techniques, with similar execution cost, without the
+need for domain knowledge or trial and error to configure the process.
 
@@
 similarity per attribute/field compared) for each pair. These similarity
-vectors are then used to classify the compared record pairs into links
-(where it is assumed both records in a pair correspond to the same
-real-world entity) and non-links (where the records are assumed to
-correspond to two different entities). Various classification methods
-have been employed in record linkage~\cite{Chr12,Don15}, ranging from
-simple threshold-based to sophisticated clustering, supervised
-classification techniques, and active learning approaches~\cite{Wan15}.
+vectors are then used to classify the compared record pairs into
+\emph{links} (where it is assumed both records in a pair correspond to
+the same real-world entity) and \emph{non-links} (where the records are
+assumed to correspond to two different entities). Various
+classification methods have been employed in record
+linkage~\cite{Chr12,Don15}, ranging from simple threshold-based to
+sophisticated clustering, supervised classification techniques, and
+active learning approaches~\cite{Wan15}.
 
@@
 of reduced linkage quality, because potentially true matching record
-pairs are removed in the indexing step, leading to a reduction in recall
-of the final linkage result~\cite{Chr12}. A variety of indexing
+pairs are removed in the indexing step, leading to a reduction in
+recall of the final linkage result~\cite{Chr12}. A variety of indexing
 techniques, discussed in more detail in the following section, have
@@
 
-Systems that perform indexing prior to comparison and classification, as
-illustrated in Fig.~\ref{fig-rl-process}, add a further practical
-complexity to the process. Indexing, comparison and classification are
-often conducted using algorithms and parameters selected based on
-technical and domain expertise of the user of the record linkage system,
-followed by a manual assessment (auditing) of the linkage
-outcomes~\cite{Chr12}. If the quality of the resulting links is not good
-enough for a certain application, the linkage process needs to be
-repeated with different parameter settings and possibly also alternative
-algorithms. This can result in a time-consuming and labour-intensive
-iterative process. A major challenge is often the heuristic nature of
-indexing, where the choice of a certain indexing technique and its
-parameters (including which attributes to use in indexing) will
-determine the final outcome of a linkage.
+Traditional record linkage systems that perform indexing prior to
+comparison and classification, as illustrated in left-hand side in
+Fig.~\ref{fig-rl-process}, add a further practical complexity to the
+process. Indexing, comparison and classification are often conducted
+using algorithms and parameters selected based on technical and
+domain expertise of the user of the record linkage system, followed
+by a manual assessment (auditing) of the linkage
+outcomes~\cite{Chr12}. If the quality of the resulting links is not
+good enough for a certain application, the linkage process needs to
+be repeated with different parameter settings and possibly also
+alternative algorithms. This can result in a time-consuming and
+labour-intensive iterative process. A major challenge is often the
+heuristic nature of indexing, where the choice of a certain indexing
+technique and its parameters (including which attributes to use in indexing) will determine the final outcome of a linkage.
 
@@
 a technique to be \emph{complete} if it is guaranteed to identify all
-pairs within the specified similarity threshold. Commonly used indexing
-techniques are incomplete, since they reduce computational cost at the
-expense of potentially overlooking some true matches. By definition,
-incomplete techniques yield lower recall than complete techniques.
-Conversely, and perhaps counter-intuitively, complete techniques can
-yield lower precision with some datasets. This is discussed further in
-Sect.~\ref{sec-approach}.
+record pairs within the specified similarity threshold. Commonly used
+indexing techniques are incomplete, since they reduce computational
+cost at the expense of potentially overlooking some true matches. By
+definition, incomplete techniques yield lower recall than complete
+techniques. Conversely, and perhaps counter-intuitively, complete
+techniques can yield lower precision with some datasets. This is
+discussed further in Sect.~\ref{sec-approach}.
 
@@
 indexing, comparison and classification to be combined into a single
-step, as shown in Fig.~\ref{fig-rl-process}, making the overall record
-linkage process simpler, more efficient and more effective than
-incomplete indexing approaches.
+step, as illustrated in the right-hand side in
+Fig.~\ref{fig-rl-process}, making the overall record linkage process
+simpler, more efficient and more effective than incomplete indexing
+approaches.
 
-\textbf{Contribution:} the primary contribution of this paper is the
+\smallskip
+\textbf{Contribution:} The primary contribution of this paper is the
 novel application of MSI to achieve complete and efficient record
 linkage, without the need for complex parameter tuning. We evaluate our
-approach on several datasets from diverse domains and demonstrate its
-advantages over existing indexing techniques for record linkage.
+approach on several real-world datasets from diverse domains and
+demonstrate its advantages over existing indexing techniques for
+record linkage.
 
@@
 according to its pronunciation, assigning the same code to similar
-sounding names (such as `Gail' and `Gayle'). Multiple blocking keys may
-be used to deal with the problem of missing attribute values.
+sounding names (such as `Gail' and `Gayle'). Multiple blocking keys
+may be used to deal with the problem of missing attribute values.
 
@@
 characteristics of the sorting key values have shown to improve both
-linkage efficiency and quality~\cite{Dra12} over other sliding window
-approaches.
+linkage efficiency and quality~\cite{Dra12} over
+%other sliding window
+approaches that use fixed sizes windows.
 
@@
 records that share some hash values. \emph{HARRA}~\cite{Kim10} is a
-record linkage approach based on MinHash~\cite{Broder1997} and LSH which
-blocks, compares, and then merges linked records in an iterative
+record linkage approach based on MinHash~\cite{Broder1997} and LSH
+which blocks, compares, and then merges linked records in an iterative
 fashion, where merged records are re-hashed to improve overall linkage
-quality. ~\cite{Steorts2014} evaluates two LSH variations, concluding
+quality. \cite{Steorts2014} evaluates two LSH variations, concluding
 that in order to get good results, LSH methods must be tuned to the
@@
 inequality}. Similarity search operations include
-\textit{range-search(q,d)}, where all records within a distance $d$ of a
-query record $q$ are identified; \textit{nearest-neighbour(q)},
-returning the record with smallest distance to $q$; and
-\textit{nearest-n(q,n)}, returning the $n$ closest records to $q$.
-\cite{Zezula2010} gives an overview of MSI techniques.
-Here we choose one metric space indexing structure, the M-tree
-\cite{paolociaccia2m}, and investigate its efficacy for record linkage.
+\emph{range-search(\textbf{q},d)}, where all records $\mathbf{r}$
+in a database $\mathbf{D}$ within a distance $d$ of a query record
+$\mathbf{q}$ are identified; \emph{nearest-neighbour(\textbf{q})},
+returning the record $\mathbf{r} \in \mathbf{D}$ with smallest
+distance to $\q$; and \emph{nearest-n(\textbf{q},n)}, returning
+the $n$ closest records in $\mathbf{D}$ to $\mathbf{q}$.
+\cite{Zezula2010} gives an overview of MSI techniques. Here we choose
+one metric space indexing structure, the M-tree~\cite{paolociaccia2m},
+and investigate its efficacy for record linkage.
 
@@
 
-\cite{Li2006} describes a threshold-based linkage method using
-R-trees~\cite{Hjaltason1998}. It demonstrates that high quality may be
-achieved using Jaccard similarity over selected record fields.
-\cite{Ciaccia97indexingmetric} shows that M-trees are almost always more
-efficient than R-trees, hence their use in the experiments described in
-this paper.
+A threshold-based linkage method using R-trees~\cite{Hjaltason1998}
+was described in~\cite{Li2006}. It demonstrates that high linkage
+quality can be achieved using Jaccard similarity over selected record
+attributes. \cite{Ciaccia97indexingmetric} shows that M-trees are
+almost always more efficient than R-trees, hence their use in the
+experiments described in this paper.
 
@@
 
-\Cref{cora-quality} shows the precision, recall, and F-measure
-for various thresholds. As expected, low thresholds give high precision and low recall, and the reverse for high threshold
-values. As expected, brute force and M-tree give identical results.
+\Cref{cora-quality} shows the precision, recall, and F-measure for
+various thresholds. As expected, low thresholds give high precision and
+low recall, and the reverse for high threshold values. Brute force and
+M-tree give identical results. The best linkage quality, with an
+F-measure of around 0.7, is achieved by a number of linkers, including
+brute force, M-tree, blocking on `authors' and blocking and on all
+fields, and two of the LSH configurations. All of these give similar
+overall results, apart from blocking on `authors', which gives much
+better quality at very high distance threshold values. This is due to
+the incomplete nature of the linker, avoiding consideration of
+significant numbers of high-similarity non-matches, thus avoiding these
+becoming false positives and keeping precision high.
 
 For a more detailed investigation of selected linkers, the brute force
-approach is used to establish the optimum threshold value for the Cora
-dataset. The maximum F-measure value is observed at a threshold value of
+approach is used to establish a good threshold value for the Cora
+dataset. The maximum F-measure is observed at a threshold value of
 70. This value is dataset-dependent; for different datasets the maximum
@@
 \Cref{comparison-of-results-cora} shows greater detail for selected
-linkers, with the distance threshold fixed at 70. The table shows the
-parameters for the experiment, the number of distance comparisons made,
-the precision, recall and the F1 measure achieved by each algorithm.  In
-the \emph{Linker} column the algorithm name is followed by its
-parameters: for LSH the number of the bands followed by the band size,
-and for traditional blocking the fields used for blocking. The number of
-distance comparisons is reported as a machine-independent proxy for
-execution cost, since code profiling shows that distance calculations
-are dominant.
+linkers, showing the parameters for the experiment, the number of
+distance comparisons made, the precision, recall and the F-measure
+achieved by each algorithm.  In the \emph{Linker} column the algorithm
+name is followed by its parameters: for LSH the number of the bands
+followed by the band size, and for traditional blocking the fields used
+for blocking. The number of distance comparisons is reported as a
+machine-independent proxy for execution cost, since code profiling shows
+that distance calculations are dominant.
 
@@
 both techniques are complete. Several of the incomplete linkers give
-similar quality, for example \emph{LSH-2-2}, \emph{LSH-5-2}, \emph{LSH-10-2} and
-\emph{Block-combined}. These, and a number of other incomplete linkers give
-better precision than the complete techniques. This is due to
-high-similarity non-matches, as discussed in Sect.~\ref{sec-approach}.
-Although several of the incomplete linkers give as good quality as
-M-tree, and in some cases at lower cost, this is offset by the need to
-select appropriate configuration parameters. Some other linkers give
-very poor results.
+similar quality, for example \emph{LSH-2-2}, \emph{LSH-5-2},
+\emph{LSH-10-2} and \emph{Block-combined}. These, and a number of other
+incomplete linkers give better precision than the complete techniques.
+This is due to high-similarity non-matches, as discussed in
+Sect.~\ref{sec-approach}. Although several of the incomplete linkers
+give as good quality as M-tree, and in some cases at lower cost, this is
+offset by the need to select appropriate configuration parameters. Some
+other linkers give very poor results.
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
+\vspace{-6mm}
 \caption{Brute Force}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
+\vspace{-6mm}
 \caption{M-tree}
-\end{subfigure}
+\end{subfigure} \vspace{5mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
+\vspace{-6mm}
 \caption{Blocking on `authors' field}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
+\vspace{-6mm}
 \caption{Blocking on `title' field}
-\end{subfigure}
+\end{subfigure} \vspace{5mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
+\vspace{-6mm}
 \caption{Blocking on `publication year' field}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
+\vspace{-6mm}
 \caption{Blocking using the union of all fields}
-\end{subfigure}
+\end{subfigure} \vspace{5mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
+\vspace{-6mm}
 \caption{LSH using 2 bands of size 2}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
+\vspace{-6mm}
 \caption{LSH using 2 bands of size 10}
-\end{subfigure}
+\end{subfigure} \vspace{5mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
+\vspace{-6mm}
 \caption{LSH using 10 bands of size 2}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
+\vspace{-6mm}
 \caption{LSH using 10 bands of size 10}
@@
 
+GRAHAM STILL WRITING HERE
+
+Birth records were linked to death records, separately for the Skye and Kilmarnock datasets, using M-tree and LSH.
+
+It was not computationally feasible to run the brute force linker. Of the incomplete linkers, LSH was selected as it gave slightly better results for Cora.
+
+The shingle size was set to 2 for all the LSH experiments reported, as this was found to give good results and LSH was not especially sensitive to this parameter.
+Results for other shingle sizes are omitted from this paper for brevity.
+
+on the Cora dataset using all four
+approaches presented in this paper: brute force, traditional blocking,
+LSH and M-tree, using several selected configurations for the blocking
+and LSH linkers. The distance threshold is varied between 0 and
+250~\footnote{Relatively high Levenshtein distances are included since
+Cora includes a number of low-similarity true matches.}. For traditional
+blocking, the following Cora fields are used individually as blocking
+keys: \textit{author}, \textit{title}, \textit{venue},
+\textit{location}, \textit{publisher} and \textit{year}. We also use a
+combined blocking key comprising all fields.
+
+\Cref{cora-quality} shows the precision, recall, and F-measure for
+various thresholds. As expected, low thresholds give high precision and
+low recall, and the reverse for high threshold values. Brute force and
+M-tree give identical results. The best linkage quality, with an
+F-measure of around 0.7, is achieved by a number of linkers, including
+brute force, M-tree, blocking on `authors' and blocking and on all
+fields, and two of the LSH configurations. All of these give similar
+overall results, apart from blocking on `authors', which gives much
+better quality at very high distance threshold values. This is due to
+the incomplete nature of the linker, avoiding consideration of
+significant numbers of high-similarity non-matches, thus avoiding these
+becoming false positives and keeping precision high.
+
+\Cref{demography-quality}
+
 \Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} show the results for LSH-MinHash and M-tree on the Skye and Kilmarnock datasets.
-Linkage was performed between birth and death records.
-The shingle size was set to 2 for all the LSH experiments reported, as this was found to give good results and LSH was not especially sensitive to this parameter.
-Results for other shingle sizes were omitted from this paper for brevity.
+
 In both cases the F1 measure achieved is better for M-tree than any of the experimental settings chosen for LSH.
@@
 Despite this LSH is much faster due to the efficiency of the hashing mechanisms.
+
+%----------------------------------------------------------------------
+
+\begin{figure}
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
+\caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
+\caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
+\end{subfigure}
+
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
+\caption{Linkage quality on Skye dataset using M-tree and all LSH configurations}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
+\caption{Linkage quality on Kilmarnock dataset using M-tree and all LSH configurations}
+\end{subfigure}
+
+\caption{Linkage results on the demographic datasets}
+\label{demography-quality}
+\end{figure}
 
@@
 \end{table}
-
-%----------------------------------------------------------------------
-
-\begin{figure}
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
-\caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
-\caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
-\end{subfigure}
-
-\begin{subfigure}{.5\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
-\caption{Linkage quality on Skye dataset using M-tree and all LSH configurations}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
-\caption{Linkage quality on Kilmarnock dataset using M-tree and all LSH configurations}
-\end{subfigure}
-
-\caption{Linkage results on the demographic datasets}
-\label{demography-quality-mtree}
-\end{figure}
 

### 2017-11-22 paper.tex
Update on Overleaf.

@@
   \caption{Overview of the steps of the traditional record linkage
-           process (left-hand side) and our proposed metric space
-           indexing based approach (right-hand side), as described in
+           process (left side) and our proposed metric space
+           indexing based approach (right side), as described in
            Sect.~\ref{sec-intro}, where records from two datasets,
@@
 returning the record $\mathbf{r} \in \mathbf{D}$ with smallest
-distance to $\q$; and \emph{nearest-n(\textbf{q},n)}, returning
+distance to $\mathbf{q}$; and \emph{nearest-n(\textbf{q},n)}, returning
 the $n$ closest records in $\mathbf{D}$ to $\mathbf{q}$.
@@
 We address the following general record linkage problem: for two
-datasets $A$ and $B$, we wish to find, for each record in $A$, all the
-records in $B$ that match it. We compare several algorithms for record
-linkage: traditional blocking, one incomplete similarity search method,
-LSH-MinHash, and one complete method, M-tree~\cite{paolociaccia2m}. We
-also use a simple complete brute force technique as a baseline for
-comparison, although this can only feasibly be applied to the smallest
-of our datasets. All experiments have a number of parameters to
-configure the search space and algorithm behaviour, including a
-\emph{distance threshold} specifying a maximum distance, equivalent to a
-minimum similarity, for two records to be classified as a link (i.e.\
-referring to the same entity).
+datasets $\mathbf{D}_A$ and $\mathbf{D}_B$, we wish to find, for each
+record in $\mathbf{D}_A$, all the records in $\mathbf{D}_B$ that match
+it. We compare several algorithms for record linkage: traditional
+blocking, one incomplete similarity search method, LSH-MinHash, and
+one complete method, M-tree~\cite{paolociaccia2m}. We also use a
+simple complete brute force technique as a baseline for comparison,
+although this can only feasibly be applied to the smallest of our
+datasets. All experiments have a number of parameters to configure
+the search space and algorithm behaviour, including a
+\emph{distance threshold}, $d$, specifying a maximum distance,
+equivalent to a minimum similarity, for two records to be classified
+as a link (i.e.\ referring to the same entity).
 
+\smallskip
 \textbf{Brute force:}
-Two nested loops are used to compare every record in $A$ with every
-record in $B$. Each pair is classified as a link if the distance between
-the records is less than the threshold. This approach is guaranteed to
-identify all links, with complexity $ O(|A| \times |B|) $.
+Two nested loops are used to compare every record in
+$\mathbf{D}_A$ with every record in $\mathbf{D}_B$. Each pair is
+classified as a link if the distance $d$ between the records is less
+than the threshold. This approach is guaranteed to identify all links,
+with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|) $.
 
+\smallskip
 \textbf{Traditional blocking:}
-The additional parameters are the set of blocking keys and the
-normalization encodings applied to each field. These are selected as
-described in \cite{Chr12b}, exploiting knowledge of the domain and of
-the records being linked, and chosen with the intention of giving the
-best possible results for the datasets. Each record in $A$ is placed
-into the appropriate block based on its blocking key value. The
-algorithm then iterates over the records in $B$, and for each one,
-compares it with each of the records from $A$ in the block with the same
-blocking key value.
+The parameters of traditional blocking are the set of blocking keys
+and the (optionally) phonetic
+% Peter: normalisation?
+encodings applied to each attribute. These are selected as described
+in~\cite{Chr12b}, exploiting knowledge of the domain and of the
+records being linked, and chosen with the intention of giving the
+best possible results for the datasets. Each record in $\mathbf{D}_A$
+is placed into the appropriate block based on its blocking key value.
+The algorithm then iterates over the records in $\mathbf{D}_B$, and
+for each one, compares it with each of the records from $\mathbf{D}_A$
+in the block with the same blocking key value.
 
+\smallskip
 \textbf{LSH-MinHash:}
-The additional parameters are \emph{shingle size} ($l_{ss}$), \emph{band
-size} ($l_{bs}$) and \emph{number of bands} ($l_{nb}$). Each record in
-$A$ is placed in an \emph{LSH-MinHash} data structure. First, all the
-fields of the record are concatenated, and the result \emph{shingled}
-into a set of n-grams with $n = l_{ss}$. Next, a set of
-deterministically generated hash functions are applied to each n-gram in
-the set and the smallest result (the MinHash) of each hash application
-is added to a signature for the record. The number of hashes used, and
-thus the size of the signature, is set to $l_{nb} \times l_{bs}$.
-Finally, the signature is split into $l_{nb}$ bands and the values from
-each band are hashed again to create a number of keys. The original
-record is added to a map associated with each of the keys.
+The parameters for LSH-Minhash are~\cite{Broder1997} \emph{shingle
+size} ($l_{ss}$), \emph{band size} ($l_{bs}$) and \emph{number of
+bands} ($l_{nb}$). Each record in $\mathbf{D}_A$ is placed in an
+\emph{LSH-MinHash} data structure. First, all the attributes of the
+record are concatenated, and the result \emph{shingled} into a set of
+n-grams with $n = l_{ss}$. Next, a set of deterministically generated
+hash functions are applied to each n-gram in the set and the smallest
+result (the MinHash) of each hash application is added to a signature
+for the record. The number of hashes used, and thus the size of the
+signature, is set to $l_{nb} \times l_{bs}$. Finally, the signature is
+split into $l_{nb}$ bands and the values from each band are hashed
+again to create a number of keys. The original record is added to a
+map associated with each of the keys.
 
-To perform linkage, the algorithm iterates over the records in $B$. Each
-record is hashed using the procedure described above, to obtain a set of
-keys. For each of these keys, the key is looked up in the LSH-MinHash
-data structure, and the records from $A$ associated with it added to the
-result set. Finally, the record from $B$ is compared in turn with each
-record in the result set, with the pair being classified as a link or
-non-link based on their distance.
+To perform linkage, the algorithm iterates over the records in
+$\mathbf{D}_B$. Each record is hashed using the procedure described
+above, to obtain a set of keys. For each of these keys, the key is
+looked up in the LSH-MinHash data structure, and the records from
+$\mathbf{D}_A$ associated with it added to the result set. Finally,
+the record from $\mathbf{D}_B$ is compared in turn with each record
+in the result set using the same distance function and threshold, $d$,
+as with the other approaches, with the pair being classified as a link
+or non-link based on their distance.
 
@@
 number of non-matches that nonetheless have high similarity. In this
-situation, the fact that an incomplete technique omits consideration of
-some potential links can serve to improve precision, since a
-classification decision based on similarity threshold is incorrect for
-high-similarity non-matches. By definition, recall is never higher for
-incomplete techniques.
+situation, the fact that an incomplete technique omits consideration
+of some potential links can serve to improve precision, since a
+classification decision based on a certain similarity threshold is
+incorrect for high-similarity non-matches. By definition, recall can
+never be higher for incomplete techniques.
 
-\textbf{M-tree:} The M-tree linkage algorithm has no additional
-parameters. In a similar manner to the \emph{LSH-MinHash} approach, each
-record in A is inserted into an M-tree. To perform linkage, the
-algorithm iterates over each record $b$ in $B$. A \emph{range-search(b,d)}
-operation is performed, passing the distance threshold as the second
-parameter. All the returned records are directly recorded as links.
+%\smallskip
+\textbf{M-tree:} Besides the distance threshold, $d$, the M-tree
+linkage algorithm has no additional parameters. In a similar manner
+to the \emph{LSH-MinHash} approach, each record in $\mathbf{D}_A$ is inserted into an M-tree. To perform linkage, the algorithm iterates
+over each record in $\mathbf{b} \in \mathbf{D}_B$. A
+\emph{range-search(\textbf{b},d)} operation is performed, passing the
+distance threshold as the second parameter. All the returned records
+are directly classified as links.
 
@@
   \hline\noalign{\smallskip}
-  Dataset~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities\\
-  name(s)  & dataset 1  & dataset 2  & matching pairs & linked \\
+  Dataset~ & ~Records in~& ~Records in~ & ~Number of true~& ~Entities \\
+    name(s)  & ~dataset $\mathbf{D}_A$~ & ~dataset $\mathbf{D}_B$~ &
+    matching pairs & linked \\
   \noalign{\smallskip} \hline \noalign{\smallskip}
-  Cora & 1,295 & 1,295 & 17,184 & publication-publication\\
-  Isle of Skye & 17,612 & 12,284& 2,900 & birth-death\\
-  Kilmarnock & 38,430 & 23,714 & 8,300 & birth-death\\
+  Cora & 1,295 & 1,295 & 17,184 & Publication--Publication\\
+  ~Isle of Skye~ & 17,612 & 12,284& 2,900 & Birth--Death \\
+  Kilmarnock & ~38,430~ & ~23,714~ & ~8,300~ & ~Birth--Death~ \\
   \noalign{\smallskip} \hline
@@
 summarised in Table~\ref{table-datasets}. The first is
-\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer to
-112 machine learning publications. It is used as a benchmark dataset in
-the literature for assessing linkage algorithms. Ground truth is
-provided via a unique \emph{paper\_id} identifier of the form
-``blum1993". In this experiment linkage is performed over the same set
-of records.
+\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer
+to 112 machine learning publications. Cora is commonly used as a
+benchmark dataset in the literature for assessing linkage algorithms.
+Ground truth is provided via a unique \emph{paper\_id} identifier of
+the form ``blum1993". In this experiment linkage is performed over
+the same set of records.
 
-The other datasets are historical Scottish records of vital events
-(birth, marriages and deaths) one registered on the \emph{Isle of Skye},
-a rural district, and the other records from \emph{Kilmarnock}, an
-industrial town. These datasets were create by historical demographers
-who extensively curated and linked both
+The other two datasets are historical Scottish records of vital events
+(birth, marriages and deaths) one registered on the \emph{Isle of
+Skye}, a rural district, and the other records from
+\emph{Kilmarnock}, an industrial town. These datasets were created by
+historical demographers who extensively curated and linked both
 datasets~\cite{reid2002,reid2006}. Both datasets include the names,
 gender, addresses of individuals and their parents. Ground truth was
-generated by the demographers based on their extensive domain knowledge.
+generated by the demographers based on their extensive domain
+knowledge.
 
 In all of our experiments we use a single distance metric which is the
-sum of the Levenshtein~\cite{Levenshtein66} distances between the set of
-fields being compared.
+sum of the Levenshtein~\cite{Levenshtein66} distances between the set
+of attributes being compared.
 
@@
 LSH and M-tree, using several selected configurations for the blocking
-and LSH linkers. The distance threshold is varied between 0 and
-250~\footnote{Relatively high Levenshtein distances are included since
-Cora includes a number of low-similarity true matches.}. For traditional
-blocking, the following Cora fields are used individually as blocking
-keys: \textit{author}, \textit{title}, \textit{venue},
-\textit{location}, \textit{publisher} and \textit{year}. We also use a
-combined blocking key comprising all fields.
+and LSH approaches. The distance threshold is varied between $0 \le d \le 250$~\footnote{Relatively high Levenshtein distances are included
+since Cora includes a number of low-similarity true matches.}. For
+traditional blocking, the following Cora attributes are used
+individually as blocking keys: \emph{author}, \emph{title},
+\emph{venue}, \emph{location}, \emph{publisher} and \emph{year}. We
+also use a combined blocking key comprising all attributes.
 
-\Cref{cora-quality} shows the precision, recall, and F-measure for
-various thresholds. As expected, low thresholds give high precision and
-low recall, and the reverse for high threshold values. Brute force and
-M-tree give identical results. The best linkage quality, with an
-F-measure of around 0.7, is achieved by a number of linkers, including
-brute force, M-tree, blocking on `authors' and blocking and on all
-fields, and two of the LSH configurations. All of these give similar
-overall results, apart from blocking on `authors', which gives much
-better quality at very high distance threshold values. This is due to
-the incomplete nature of the linker, avoiding consideration of
-significant numbers of high-similarity non-matches, thus avoiding these
-becoming false positives and keeping precision high.
+\Cref{cora-quality} shows the precision, recall, and F-measure~\cite{Chr12} for various thresholds. As expected, low 
+thresholds give high precision and low recall, and the reverse for high
+threshold values. Brute force and M-tree give identical results, as
+expected. The best linkage quality, with an F-measure of around 0.7,
+is achieved by a number of linkers, including brute force, M-tree, blocking on \emph{authors} and blocking and on all attributes, and two
+of the LSH configurations. All of these give similar overall results,
+apart from blocking on \emph{authors}, which gives much better quality
+at very high distance threshold values. This is due to the incomplete
+nature of the approach, avoiding comparisons of significant numbers
+of high-similarity non-matches and thus avoiding these becoming false
+positives and keeping precision high.
 
@@
 dataset. The maximum F-measure is observed at a threshold value of
-70. This value is dataset-dependent; for different datasets the maximum
-F-measure will occur at different thresholds. In the rest of this
-section we fix the threshold value at 70.
+$d=70$. This value is dataset-dependent; for different datasets the
+maximum F-measure will occur at different thresholds. In the rest of
+this section we fix the threshold value at $d=70$.
 
@@
 name is followed by its parameters: for LSH the number of the bands
-followed by the band size, and for traditional blocking the fields used
-for blocking. The number of distance comparisons is reported as a
-machine-independent proxy for execution cost, since code profiling shows
-that distance calculations are dominant.
+followed by the band size, and for traditional blocking the attributes
+used for blocking. The number of distance comparisons is reported as a
+machine-independent proxy for execution cost, since code profiling
+shows that distance calculations are dominant.
 
@@
 \begin{table}[ht]
+\caption{Linkage quality on Cora dataset with distance threshold
+  $d= 70$.}
+  \label{comparison-of-results-cora}
   \centering
-\begin{tabular}{l|r|r|r|r}
+  \begin{footnotesize}
+  \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
 \parbox{0.2\textwidth}{\centering Linker} &
@@
 \parbox{0.17\textwidth}{\centering Recall} &
-\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
-Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68       \\ \hline
-M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\ \hline
+\parbox{0.17\textwidth}{\centering F1 Measure} \\
+  \noalign{\smallskip} \hline \noalign{\smallskip}
+Brute Force        & 1,677,025                 & 0.84      & 0.57   & 0.68  \\
+\noalign{\smallskip} \hline \noalign{\smallskip}
+M-tree              &  902,693                  & 0.84      & 0.57   & 0.68       \\
+\noalign{\smallskip} \hline \noalign{\smallskip}
 LSH-2-2            &  192,199	& 0.95	& 0.47	& 0.63       \\
@@
 LSH-5-10           &  6,501	& 1	& 0.19	& 0.32      \\
-LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\  \hline
+LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\
+\noalign{\smallskip} \hline \noalign{\smallskip}
 Block-year         &  115,893                  & 0.99      & 0.35   & 0.51       \\
@@
 Block-publisher    &  833,079                  & 0.85      & 0.44   & 0.58       \\
-Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\ \hline
+Block-combined     & 1,214,269                 & 0.84      & 0.56   & 0.67       \\
+\noalign{\smallskip} \hline
 \end{tabular}
-\caption{Linkage quality on Cora dataset with distance threshold = 70
-\label{comparison-of-results-cora}
-}
+\end{footnotesize}
 \end{table}
@@
 
-GRAHAM STILL WRITING HERE
+Birth records were linked to death records, separately for the Skye and
+Kilmarnock datasets, using M-tree and a range of LSH configurations. It
+was not computationally feasible to run the brute force linker. Of the
+incomplete linkers, LSH was selected as it gave slightly better results
+for Cora. The shingle size was set to 2 for all the LSH experiments
+reported, as this was found to give good results and LSH was not
+especially sensitive to this parameter. Results for other shingle sizes
+are omitted from this paper for brevity. A lower range of distance
+thresholds was explored, based on domain knowledge of the datasets.
 
-Birth records were linked to death records, separately for the Skye and Kilmarnock datasets, using M-tree and LSH.
+\Cref{demography-quality} plots (a) and (b) show the M-tree precision,
+recall, and F-measure for various thresholds. In both datasets, the best
+F-measure values are obtained with a low distance threshold of 2. Plots
+(c) and (d) compare the F-measure curves for M-tree with those obtained
+from a range of LSH configurations. The best F-measure value for M-tree
+is higher than that of any of the LSH configurations, for both datasets.
+This demonstrates both the competitiveness of M-tree with respect to
+linkage quality, and its important characteristic of being
+parameter-free---the linkage quality is obtained without the need to
+tuning for the dataset.
 
-It was not computationally feasible to run the brute force linker. Of the incomplete linkers, LSH was selected as it gave slightly better results for Cora.
+\Cref{comparison-of-results-demography-skye} and
+\Cref{comparison-of-results-demography-kili} show greater detail for
+selected linkers. In both cases the F-measure achieved is better for
+M-tree than any of the LSH linkers. The better linkage quality achieved
+by M-tree is largely due to recall for M-tree being much higher than for
+any of the LSH configurations. In most cases, LSH out-performs M-tree in
+terms of precision. More significantly, LSH linkage quality is heavily
+dependent on the configuration parameters. For plausible settings for
+the number of bands and band size, F-measure varies from 0.01 (extremely
+poor) to 0.47 (relatively good) for Skye and from 0.03 to 0.49 for
+Kilmarnock. In both cases \emph{LSH-10-2} performs best, but since this
+is data-dependent there is no guarantee that these parameters would work
+well with another dataset.
 
-The shingle size was set to 2 for all the LSH experiments reported, as this was found to give good results and LSH was not especially sensitive to this parameter.
-Results for other shingle sizes are omitted from this paper for brevity.
-
-on the Cora dataset using all four
-approaches presented in this paper: brute force, traditional blocking,
-LSH and M-tree, using several selected configurations for the blocking
-and LSH linkers. The distance threshold is varied between 0 and
-250~\footnote{Relatively high Levenshtein distances are included since
-Cora includes a number of low-similarity true matches.}. For traditional
-blocking, the following Cora fields are used individually as blocking
-keys: \textit{author}, \textit{title}, \textit{venue},
-\textit{location}, \textit{publisher} and \textit{year}. We also use a
-combined blocking key comprising all fields.
-
-\Cref{cora-quality} shows the precision, recall, and F-measure for
-various thresholds. As expected, low thresholds give high precision and
-low recall, and the reverse for high threshold values. Brute force and
-M-tree give identical results. The best linkage quality, with an
-F-measure of around 0.7, is achieved by a number of linkers, including
-brute force, M-tree, blocking on `authors' and blocking and on all
-fields, and two of the LSH configurations. All of these give similar
-overall results, apart from blocking on `authors', which gives much
-better quality at very high distance threshold values. This is due to
-the incomplete nature of the linker, avoiding consideration of
-significant numbers of high-similarity non-matches, thus avoiding these
-becoming false positives and keeping precision high.
-
-\Cref{demography-quality}
-
-\Cref{comparison-of-results-demography-skye} and \Cref{comparison-of-results-demography-kili} show the results for LSH-MinHash and M-tree on the Skye and Kilmarnock datasets.
-
-In both cases the F1 measure achieved is better for M-tree than any of the experimental settings chosen for LSH.
-This is largely due to the recall for M-tree being much higher than any of the LSH experiments.
-For most configurations, LSH out-performs M-tree in terms of precision.
-
-Perhaps more important than the fact that M-tree gives better F1 measures than all of the LSH experiments is the fact that the LSH performance is extremely variable.
-For a relatively plausible settings for the number of bands and band-size the F1 measurements vary from 0.01 (extremely poor) to 0.47 (relatively good) for Skye and between 0.03 and 0.49 for the Kilmarnock dataset.
-In both cases LSH-10-2 performs best, but since this is data dependent there is no guarantee these parameters would work well with another dataset.
-Furthermore, without good ground truth, which is usually unavailable, there is no way of knowing if good linkage results are being obtained or not.
-
-Another observation that is worthy of comment is the number of (distance) comparisons being performed by the different algorithms.
-For M-trees the number of comparisons are closely related to the complexity of the search being performed since the algorithms are performing (range) searches over distance measures to determine the query results.
-This is not true of LSH which is performing Jaccard similarity using Hashes over the datasets.
-Comparisons are only used at the final step to determine if a candidate is a match or not.
-The LSH results which provide the best results by F1 measure are performing between a quarter and half the number of matches performed by M-trees.
-From this we can conclude that the hash algorithms are returning many candidates that are outwith the search distance.
-Thus, in order to get good results, LSH is tending towards a brute force search over the candidate results.
-Despite this LSH is much faster due to the efficiency of the hashing mechanisms.
+The number of distance comparisons varies dramatically among the various
+linkers. M-tree always performs the most comparisons, since they are
+intrinsic to the \emph{range-search} algorithm. The core part of the LSH
+linker performs Jaccard similarity comparisons and hashing; distance
+comparisons are only performed in the final step to determine whether a
+candidate pair is a link. The LSH configurations yielding the best
+results perform distance comparisons of the same order of magnitude as
+M-tree. This indicates that the good LSH linkers return many candidates
+beyond the distance threshold. Thus, in order to get good results, LSH
+tends towards a brute force search over the candidate results. Despite
+this, LSH is much faster due to the efficiency of the hashing process.
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
-\caption{Linkage quality on Skye dataset using M-tree\label{skye-quality-mtree}}
+\vspace{-6mm}
+\caption{M-tree linkage on Skye dataset\label{skye-quality-mtree}}
 \end{subfigure}%
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
-\caption{Linkage quality on Kilmarnock dataset using M-tree\label{kilmarnock-quality-mtree}}
-\end{subfigure}
+\vspace{-6mm}
+\caption{M-tree linkage on Kilmarnock dataset\label{kilmarnock-quality-mtree}}
+\end{subfigure} \vspace{5mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotFs-skye-f}
-\caption{Linkage quality on Skye dataset using M-tree and all LSH configurations}
+\vspace{-6mm}
+\caption{F1 Measure on Skye dataset \\ with M-tree and all LSH configurations}
 \end{subfigure}%
@@
 \includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
-\caption{Linkage quality on Kilmarnock dataset using M-tree and all LSH configurations}
+\vspace{-6mm}
+\caption{F1 Measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
 \end{subfigure}
@@
 \begin{table}[ht]
+\caption{Linkage quality on Skye dataset with distance threshold
+$d= 2$.}
+\label{comparison-of-results-demography-skye}
 \centering
-\begin{tabular}{l|r|r|r|r}
+\begin{footnotesize}
+\begin{tabular}{lrrrr} \hline\noalign{\smallskip}
 \parbox{0.2\textwidth}{\centering Linker} &
@@
 \parbox{0.17\textwidth}{\centering Recall} &
-\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
+\parbox{0.17\textwidth}{\centering F1 Measure} \\
+\noalign{\smallskip} \hline \noalign{\smallskip}
 M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline
-
+\noalign{\smallskip} \hline \noalign{\smallskip}
 LSH-2-2  & 3,109,250                   & 0.63      & 0.03   & 0.06       \\
@@
 
-LSH-10-5 & 129,873                    & 0.72      & 0.01   & 0.02       \\ \hline
+LSH-10-5 & 129,873                    & 0.72      & 0.01   & 0.02       \\
+\noalign{\smallskip} \hline
 \end{tabular}
-\caption{Linkage quality on Skye dataset with distance threshold = 2
-\label{comparison-of-results-demography-skye}
-}
+\end{footnotesize}
 \end{table}
@@
 \begin{table}[ht]
+\caption{Linkage quality on Kilmarnock dataset with distance
+threshold $d= 2$.}
+\label{comparison-of-results-demography-kili}
 \centering
-\begin{tabular}{l|r|r|r|r|r|r}
+\begin{footnotesize}
+\begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
 \parbox{0.2\textwidth}{\centering Linker} &
@@
 \parbox{0.17\textwidth}{\centering Recall} &
-\parbox{0.17\textwidth}{\centering F1 Measure} \\ \hline \hline
-M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\ \hline
+\parbox{0.17\textwidth}{\centering F1 Measure} \\
+\noalign{\smallskip} \hline \noalign{\smallskip}
+M-tree     & 514,871,153                   & 0.76      & 0.45   & 0.57      \\
+\noalign{\smallskip} \hline \noalign{\smallskip}
 LSH-2-2   & 99,145,887                    & 0.81      & 0.16   & 0.27       \\
@@
 LSH-5-5   & 239,368                      & 0.84      & 0.01   & 0.02       \\
-LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ \hline
+LSH-10-5  & 855,431                      & 0.87      & 0.02   & 0.03 \\ 
+\noalign{\smallskip} \hline
 \end{tabular}
-\caption{Linkage quality on Kilmarnock dataset with distance threshold = 2
-\label{comparison-of-results-demography-kili}
-}
+\end{footnotesize}
 \end{table}
@@
 
-In this paper we have demonstrated the efficacy of MSI in achieving  complete and efficient record linkage, without the need for complex parameter tuning.
-In conclusion, this assertion deserves some careful unpacking.
-It is always possible to achieve high quality linkage using a brute force approach.
-However the n-squared complexity of this approach prevents its practical application for datasets of even moderate size.
-The use of MSI techniques such as M-tree delivers the same results as brute force with fewer comparisons.
-We have shown that the M-tree approach delivers high precision, high recall results that are the same as those delivered by brute force.
-Furthermore this is achieved consistently without the need for complex parameter tuning.
+In this paper we have demonstrated the efficacy of MSI in achieving
+complete and efficient record linkage, without the need for complex
+parameter tuning. In conclusion, this claim deserves some careful
+unpacking. It is always possible to achieve high quality linkage using a
+brute force approach. However the quadratic complexity of this approach
+prevents its practical application for datasets of even moderate size.
+We have shown that MSI techniques such as M-tree can deliver high
+precision, high recall results that are the same as those delivered by
+brute force. Furthermore this is achieved with fewer distance
+comparisons, and consistently without the need for complex parameter
+tuning.
 
-We contrast this to traditional blocking and LSH based approaches.
-The major drawback of their use is that whilst they can produce extremely good results, they can also produce extremely poor results.
-It was our observations of low recall given by these approaches that originally led us to experiment with M-trees.
-Without extensive ground truth or large amounts of clerking effort by human domain experts it is impossible to determine the quality of the links produced.
+We contrast this to traditional blocking and LSH-based approaches. Their
+major drawback is that whilst they can produce extremely good results,
+they can also produce extremely poor results. It was our observations of
+low recall given by these approaches that originally led us to
+experiment with M-trees.
 
-However, this paper also highlights some unexpected results.
-Firstly, the good quality results yielded by both traditional blocking and LSH are partly due to the fact that (in the limit) they tend towards brute force as the number of records in the block increase.
-A second unexpected result is that shown in \Cref{comparison-of-results-cora}, namely that LSH can yield higher precision than is achieved by a complete method such as M-tree, as was discussed in Sect.~\ref{sec-approach}.
+We note that the good results yielded by both traditional blocking and
+LSH are partly due to the fact that (in the limit) they tend towards
+brute force as the number of records in the block increase. A second,
+unexpected, result is that illustrated in
+\Cref{comparison-of-results-cora}, namely that incomplete approaches
+such as LSH can in some cases yield higher precision than that achieved
+by a complete method such as M-tree. This is due to the incomplete
+linker's ability to mask the inability of a classifier based solely on
+record similarity to correctly classify high-similarity non-matches or
+low-similarity matches.
 
-\emph{Some sentence or two about run-time?}
-
-The observation that LSH can yield results quickly at the cost of some recall suggests a potentially profitable future direction for exploration, namely the use of hybrid algorithms.
-We hypothesise that it may be possible to use an incomplete technique like LSH to form indices into a data structure such as an M-tree.
-In this manner, M-tree nodes could be used as the starting point for range-search queries and, in so doing, prune the search space considerably to yield highly quality complete results quickly.
-
-% TODO: check that the above claims have been addressed.
+The observation that LSH can yield results quickly at the cost of some
+recall suggests that hybrid algorithms may be a potentially profitable
+future direction for exploration. We hypothesise that it may be possible
+to use an incomplete technique like LSH to form indices into a data
+structure such as an M-tree. In this manner, M-tree nodes could be used
+as the starting point for range-search queries and, in so doing, prune
+the search space considerably to yield high quality results quickly.
 

### 2017-11-22 paper.tex
Update on Overleaf.

@@
 unavailable. Records are linked on the basis of similarity between
-common attributes, with every pair of records being classified as a
-link or non-link depending on the degree of similarity between them.
-Record linkage is usually performed in a three-step process: first
-groups of similar candidate records are identified using indexing; 
-pairs within the same group are then compared in more detail, and
-finally classified.
+common attributes, with every pair being classified as a link or
+non-link depending on their degree of similarity. Record linkage is
+usually performed in a three-step process: first groups of similar
+candidate records are identified using indexing; pairs within the same
+group are then compared in more detail, and finally classified.
 %
@@
 pairs are removed in the indexing step, leading to a reduction in
-recall of the final linkage result~\cite{Chr12}. A variety of indexing
+recall~\cite{Chr12}. A variety of indexing
 techniques, discussed in more detail in the following section, have
@@
 Traditional record linkage systems that perform indexing prior to
-comparison and classification, as illustrated in left-hand side in
-Fig.~\ref{fig-rl-process}, add a further practical complexity to the
+comparison and classification, as illustrated on the left in
+\Cref{fig-rl-process}, add a further practical complexity to the
 process. Indexing, comparison and classification are often conducted
@@
 that adaptively shrink or expand the window size based on the
-characteristics of the sorting key values have shown to improve both
+characteristics of the sorting key values have been shown to improve both
 linkage efficiency and quality~\cite{Dra12} over
 %other sliding window
-approaches that use fixed sizes windows.
+approaches that use fixed size windows.
 
@@
 \cite{Zezula2010} gives an overview of MSI techniques. Here we choose
-one metric space indexing structure, the M-tree~\cite{paolociaccia2m},
+one metric space indexing structure, the M-tree~\cite{Ciaccia97indexingmetric},
 and investigate its efficacy for record linkage.
@@
 blocking, one incomplete similarity search method, LSH-MinHash, and
-one complete method, M-tree~\cite{paolociaccia2m}. We also use a
+one complete method, M-tree. We also use a
 simple complete brute force technique as a baseline for comparison,
@@
 \vspace{-6mm}
-\caption{Blocking on `authors' field}
+\caption{Blocking on `authors' attr.}
 \end{subfigure}%
@@
 \vspace{-6mm}
-\caption{Blocking on `title' field}
+\caption{Blocking on `title' attr.}
 \end{subfigure} \vspace{5mm}
@@
 \vspace{-6mm}
-\caption{Blocking on `publication year' field}
+\caption{Blocking on `publication year' attr.}
 \end{subfigure}%
@@
 \vspace{-6mm}
-\caption{Blocking using the union of all fields}
+\caption{Blocking using the union of all attrs.}
 \end{subfigure} \vspace{5mm}
@@
 incomplete linkers, LSH was selected as it gave slightly better results
-for Cora. The shingle size was set to 2 for all the LSH experiments
-reported, as this was found to give good results and LSH was not
-especially sensitive to this parameter. Results for other shingle sizes
-are omitted from this paper for brevity. A lower range of distance
-thresholds was explored, based on domain knowledge of the datasets.
+for Cora. The shingle size was set to $l_{ss}=2$ for all the LSH
+experiments reported, as this was found to give good results and LSH
+was not especially sensitive to this parameter. Results for other
+shingle sizes are omitted from this paper for brevity. A lower range
+of distance thresholds was explored, based on domain knowledge of the
+datasets.
 
@@
 recall, and F-measure for various thresholds. In both datasets, the best
-F-measure values are obtained with a low distance threshold of 2. Plots
-(c) and (d) compare the F-measure curves for M-tree with those obtained
-from a range of LSH configurations. The best F-measure value for M-tree
-is higher than that of any of the LSH configurations, for both datasets.
-This demonstrates both the competitiveness of M-tree with respect to
-linkage quality, and its important characteristic of being
-parameter-free---the linkage quality is obtained without the need to
-tuning for the dataset.
+F-measure values are obtained with a low distance threshold of $d=2$.
+Plots (c) and (d) compare the F-measure curves for M-tree with those
+obtained from a range of LSH configurations. The best F-measure value
+for M-tree is higher than that of any of the LSH configurations, for
+both datasets. This demonstrates both the competitiveness of M-tree
+with respect to linkage quality, and its important characteristic of
+being parameter-free---the linkage quality is obtained without the
+need to tuning for the dataset.
 
@@
 dependent on the configuration parameters. For plausible settings for
-the number of bands and band size, F-measure varies from 0.01 (extremely
-poor) to 0.47 (relatively good) for Skye and from 0.03 to 0.49 for
-Kilmarnock. In both cases \emph{LSH-10-2} performs best, but since this
-is data-dependent there is no guarantee that these parameters would work
-well with another dataset.
+the number of bands and band size, F-measure varies from $0.01$
+(extremely poor) to $0.47$ (relatively good) for Skye and from $0.03$
+to $0.49$ for Kilmarnock. In both cases \emph{LSH-10-2} performs best,
+but since this is data-dependent there is no guarantee that these
+parameters would work well with another dataset.
 
@@
 \end{subfigure}
-
 \caption{Linkage results on the demographic datasets}

### 2017-11-22 paper.tex
Update on Overleaf.

@@
 record in $\mathbf{D}_A$, all the records in $\mathbf{D}_B$ that match
-it. We compare several algorithms for record linkage: traditional
-blocking, one incomplete similarity search method, LSH-MinHash, and
-one complete method, M-tree. We also use a
+it with regard to a certain distance threshold, $d$ (i.e.\ have a
+distance of $d$ or less). We compare several algorithms for record
+linkage: traditional blocking, one incomplete similarity search method,
+LSH-MinHash, and one complete method, M-tree. We also use a
 simple complete brute force technique as a baseline for comparison,
@@
 
-%\smallskip
+\smallskip
 \textbf{M-tree:} Besides the distance threshold, $d$, the M-tree
@@
 
-\smallskip
-\textbf{Datasets:}
+%\smallskip
+%\textbf{Datasets:}
 \label{sec-data}
@@
 expected. The best linkage quality, with an F-measure of around 0.7,
-is achieved by a number of linkers, including brute force, M-tree, blocking on \emph{authors} and blocking and on all attributes, and two
+is achieved by a number of linkers, including brute force, M-tree, blocking on \emph{authors} and blocking on all attributes, and two
 of the LSH configurations. All of these give similar overall results,
@@
 \emph{LSH-10-2} and \emph{Block-combined}. These, and a number of other
-incomplete linkers give better precision than the complete techniques.
+incomplete linkers, give better precision than the complete techniques.
 This is due to high-similarity non-matches, as discussed in
@@
 being parameter-free---the linkage quality is obtained without the
-need to tuning for the dataset.
+need to tune for the dataset.
 

### 2017-11-22 paper.tex
Update on Overleaf.

@@
 usually performed in a three-step process: first groups of similar
-candidate records are identified using indexing; pairs within the same
+candidate records are identified using indexing, pairs within the same
 group are then compared in more detail, and finally classified.
@@
 available within the datasets. For datasets that contain information
-about individuals, these attributes include the names, address, dates of
-birth, and so on, of individuals.
+about individuals, these attributes include the names, addresses,
+dates of birth, and so on, of individuals.
 
@@
 compare pairs of records, leading to a vector of similarities (one
-similarity per attribute/field compared) for each pair. These similarity
+similarity per attribute compared) for each pair. These similarity
 vectors are then used to classify the compared record pairs into
@@
 
-Besides the issues of a lack of unique entity identifiers, and data
+Besides the issues of a lack of unique entity identifiers and data
 quality (which will affect linkage quality), record linkage is also
@@
 sounding names (such as `Gail' and `Gayle'). Multiple blocking keys
-may be used to deal with the problem of missing attribute values.
+may also be used to deal with the problem of missing attribute values.
 
@@
 that adaptively shrink or expand the window size based on the
-characteristics of the sorting key values have been shown to improve both
-linkage efficiency and quality~\cite{Dra12} over
+characteristics of the sorting key values have been shown to improve
+both linkage efficiency and quality~\cite{Dra12} over
 %other sliding window
@@
 result, such techniques can lead to \emph{incomplete} linkage.
-Conversely, many of the record pairs compared in a block may turn
-out to have low similarity, corresponding to non-matches, resulting in
+Conversely, many of the record pairs compared in a block may turn out
+to have low similarity, corresponding to non-matches, resulting in
 \emph{inefficient} linkage.
@@
 
-Metric space indexing (MSI) techniques provide indexing structures to
-support the comparison of records in one set with those in another,
-typically also offering similarity search operations. To create an MSI
-data structure, it is necessary to define a distance measure between
-records, with certain properties including the \emph{triangle
-inequality}. Similarity search operations include
-\emph{range-search(\textbf{q},d)}, where all records $\mathbf{r}$
-in a database $\mathbf{D}$ within a distance $d$ of a query record
-$\mathbf{q}$ are identified; \emph{nearest-neighbour(\textbf{q})},
-returning the record $\mathbf{r} \in \mathbf{D}$ with smallest
-distance to $\mathbf{q}$; and \emph{nearest-n(\textbf{q},n)}, returning
-the $n$ closest records in $\mathbf{D}$ to $\mathbf{q}$.
-\cite{Zezula2010} gives an overview of MSI techniques. Here we choose
-one metric space indexing structure, the M-tree~\cite{Ciaccia97indexingmetric},
-and investigate its efficacy for record linkage.
+Metric space indexing (MSI) techniques~\cite{Zezula2010} provide
+indexing structures to support the comparison of records in one set
+with those in another, typically also offering similarity search
+operations. To create an MSI data structure, it is necessary to define
+a distance measure between records, with certain properties including
+the \emph{triangle inequality}~\cite{Zezula2010}. Similarity search
+operations include \emph{range-search(\textbf{q},d)}, where all
+records $\mathbf{r}$ in a database $\mathbf{D}$ within a distance $d$
+of a query record $\mathbf{q}$ are identified;
+\emph{nearest-neighbour(\textbf{q})}, returning the record $\mathbf{r}
+\in \mathbf{D}$ with smallest distance to $\mathbf{q}$; and
+\emph{nearest-n(\textbf{q},n)}, returning the $n$ closest records in
+$\mathbf{D}$ to $\mathbf{q}$.
+%\cite{Zezula2010} gives an overview of MSI techniques.
+Here we choose one metric space indexing structure, the
+M-tree~\cite{Ciaccia97indexingmetric}, and investigate its efficacy
+for record linkage.
 
@@
 $\mathbf{D}_A$ with every record in $\mathbf{D}_B$. Each pair is
-classified as a link if the distance $d$ between the records is less
-than the threshold. This approach is guaranteed to identify all links,
-with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|) $.
+classified as a link if the distance between the records is equal to
+or less than the threshold $d$. This approach is guaranteed to identify
+all links, with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|)$.
 
@@
 to the \emph{LSH-MinHash} approach, each record in $\mathbf{D}_A$ is inserted into an M-tree. To perform linkage, the algorithm iterates
-over each record in $\mathbf{b} \in \mathbf{D}_B$. A
-\emph{range-search(\textbf{b},d)} operation is performed, passing the
-distance threshold as the second parameter. All the returned records
-are directly classified as links.
+over each record $\mathbf{b} \in \mathbf{D}_B$. A
+\emph{range-search(\textbf{b},d)} operation is performed on the
+M-tree, passing the distance threshold $d$ as the second parameter.
+All the returned records are directly classified as links.
 
@@
 the form ``blum1993". In this experiment linkage is performed over
-the same set of records.
+the same set of records (i.e.\ a de-duplication~\cite{Chr12}).
 
@@
 historical demographers who extensively curated and linked both
-datasets~\cite{reid2002,reid2006}. Both datasets include the names,
+datasets~\cite{reid2006,reid2002}. Both datasets include the names,
 gender, addresses of individuals and their parents. Ground truth was
@@
 In all of our experiments we use a single distance metric which is the
-sum of the Levenshtein~\cite{Levenshtein66} distances between the set
-of attributes being compared.
+sum of the Levenshtein~\cite{Levenshtein66} edit distances between the
+set of attributes being compared.
 
@@
 LSH and M-tree, using several selected configurations for the blocking
-and LSH approaches. The distance threshold is varied between $0 \le d \le 250$~\footnote{Relatively high Levenshtein distances are included
-since Cora includes a number of low-similarity true matches.}. For
-traditional blocking, the following Cora attributes are used
-individually as blocking keys: \emph{author}, \emph{title},
+and LSH approaches. The distance threshold is varied between $0 \le d
+\le 250$~\footnote{Relatively high Levenshtein edit distances are
+included since Cora includes a number of low-similarity true
+matches.}. For traditional blocking, the following Cora attributes are
+used individually as blocking keys: \emph{author}, \emph{title},
 \emph{venue}, \emph{location}, \emph{publisher} and \emph{year}. We
@@
 apart from blocking on \emph{authors}, which gives much better quality
-at very high distance threshold values. This is due to the incomplete
+at very high distance thresholds. This is due to the incomplete
 nature of the approach, avoiding comparisons of significant numbers
@@
 linkers, showing the parameters for the experiment, the number of
-distance comparisons made, the precision, recall and the F-measure
+distance comparisons made, and the precision, recall and F-measure
 achieved by each algorithm. In the \emph{Linker} column the algorithm
@@
 \vspace{-6mm}
-\caption{Blocking on `authors' attr.}
+\caption{Blocking on `authors' attribute}
 \end{subfigure}%
@@
 \vspace{-6mm}
-\caption{Blocking on `title' attr.}
+\caption{Blocking on `title' attribute}
 \end{subfigure} \vspace{5mm}
@@
 
-\caption{Linkage results on the Cora dataset}
+\caption{Linkage results on the Cora dataset.}
 \label{cora-quality}
@@
   \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
-\parbox{0.2\textwidth}{\centering Linker} &
-\parbox{0.2\textwidth}{\centering Comparisons} &
-\parbox{0.17\textwidth}{\centering Precision} &
-\parbox{0.17\textwidth}{\centering Recall} &
-\parbox{0.17\textwidth}{\centering F1 Measure} \\
+  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
+  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
+%\parbox{0.2\textwidth}{\centering Linker} &
+%\parbox{0.2\textwidth}{\centering ~~~Comparisons} &
+%\parbox{0.17\textwidth}{\centering ~~~Precision} &
+%\parbox{0.17\textwidth}{\centering ~~~Recall} &
+%\parbox{0.17\textwidth}{\centering ~~~F1 Measure} \\
   \noalign{\smallskip} \hline \noalign{\smallskip}
@@
 
-\Cref{comparison-of-results-demography-skye} and
-\Cref{comparison-of-results-demography-kili} show greater detail for
+Tables~\ref{comparison-of-results-demography-skye}
+and~\ref{comparison-of-results-demography-kili} show greater detail for
 selected linkers. In both cases the F-measure achieved is better for
@@
 tends towards a brute force search over the candidate results. Despite
-this, LSH is much faster due to the efficiency of the hashing process.
+this, LSH is faster due to the efficiency of the hashing process.
 
@@
 
-\begin{figure}
+\begin{figure}[t]
 \centering
@@
 \end{subfigure}
-\caption{Linkage results on the demographic datasets}
+\caption{Linkage results on the demographic datasets.}
 \label{demography-quality}
@@
 
-\begin{table}[ht]
+\begin{table}[t]
 \caption{Linkage quality on Skye dataset with distance threshold
@@
 \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
-\parbox{0.2\textwidth}{\centering Linker} &
-\parbox{0.2\textwidth}{\centering Comparisons} &
-\parbox{0.17\textwidth}{\centering Precision} &
-\parbox{0.17\textwidth}{\centering Recall} &
-\parbox{0.17\textwidth}{\centering F1 Measure} \\
+  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
+  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
+%\parbox{0.2\textwidth}{\centering Linker} &
+%\parbox{0.2\textwidth}{\centering Comparisons} &
+%\parbox{0.17\textwidth}{\centering Precision} &
+%\parbox{0.17\textwidth}{\centering Recall} &
+%\parbox{0.17\textwidth}{\centering F1 Measure} \\
 \noalign{\smallskip} \hline \noalign{\smallskip}
-M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\ \hline
+M-tree    & 102,318,525                 & 0.65      & 0.46   & 0.54      \\
 \noalign{\smallskip} \hline \noalign{\smallskip}
@@
 
-\begin{table}[ht]
+\begin{table}[t]
 \caption{Linkage quality on Kilmarnock dataset with distance
@@
 \begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
-\parbox{0.2\textwidth}{\centering Linker} &
-\parbox{0.2\textwidth}{\centering Comparisons} &
-\parbox{0.17\textwidth}{\centering Precision} &
-\parbox{0.17\textwidth}{\centering Recall} &
-\parbox{0.17\textwidth}{\centering F1 Measure} \\
+  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
+  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
+%\parbox{0.2\textwidth}{\centering Linker} &
+%\parbox{0.2\textwidth}{\centering Comparisons} &
+%\parbox{0.17\textwidth}{\centering Precision} &
+%\parbox{0.17\textwidth}{\centering Recall} &
+%\parbox{0.17\textwidth}{\centering F1 Measure} \\
 \noalign{\smallskip} \hline \noalign{\smallskip}
@@
 by a complete method such as M-tree. This is due to the incomplete
-linker's ability to mask the inability of a classifier based solely on
+linker masking the inability of a classifier based solely on
 record similarity to correctly classify high-similarity non-matches or
@@
 the search space considerably to yield high quality results quickly.
+We plan to develop such hybrid indexing approaches with the aim to
+achieve fully automated, complete and efficient record linkage.
 

### 2018-02-14 paper.tex
Before Final Revisions
After Final Revisions

@@
 \documentclass{llncs}
+
 \usepackage{epsfig}
 \usepackage{graphicx}
-
-\newtheorem{mydef}{Def.}
-\newtheorem{myhyp}{Hypothesis}
-
+\usepackage{hyperref}
 \usepackage{url}
-\usepackage{hyperref}
 \usepackage{cleveref}
 \usepackage{hhline}
-\usepackage{caption}
 \usepackage{subcaption}
-
-%\pagestyle{empty}
-\pagestyle{headings}
 
@@
 
-\author{Submitted for double-blind review}
+\author{{\"O}zg{\"u}r Akg{\"u}n\inst{1} \and Alan Dearle\inst{1}
+  \and Graham Kirby\inst{1} \and Peter Christen\inst{2}}
 
-%  Hidden For Blind Review
-\iffalse
-
-\author{{\"O}zg{\"u}r Akg{\"u}n\inst{1} \and Alan Dearle\inst{1} \and\\
-Graham Kirby\inst{1} \and Peter Christen\inst{2}}
-
-\institute{School of Computer Science, University of St Andrews,\\
-St Andrews, Scotland. Contact: \email{ozgur.akgun@st-andrews.ac.uk}
+\institute{School of Computer Science, University of St Andrews,
+St Andrews, Scotland. Contact: \email{\{ozgur.akgun, alan.dearle, graham.kirby\}@st-andrews.ac.uk}
 \and
@@
 
-\fi
 %----------------------------------------------------------------------
@@
 common attributes, with every pair being classified as a link or
-non-link depending on their degree of similarity. Record linkage is
-usually performed in a three-step process: first groups of similar
-candidate records are identified using indexing, pairs within the same
-group are then compared in more detail, and finally classified.
-%
+non-link depending on their similarity. Linkage is
+usually performed in a three-step process: first, groups of similar
+candidate records are identified using indexing, then pairs within the same
+group are compared in more detail, and finally classified.
 Even state-of-the-art indexing techniques, such as Locality Sensitive
 Hashing, have potential drawbacks. They may fail to group together some
-true matching records with high similarity. Conversely, they may group
+true matching records with high similarity, or they may group
 records with low similarity, leading to high computational overhead.
-%
-We propose using metric space indexing to perform \emph{complete} record
-linkage, which results in a parameter-free record linkage process
+We propose using \emph{metric space indexing} (MSI) to perform \emph{complete}
+linkage, resulting in a parameter-free process
 combining indexing, comparison and classification into a single step
-delivering complete and efficient record linkage. Our experimental
-evaluation on real-world datasets from several domains shows that
-linkage using metric space indexing can yield better quality than
+delivering complete and efficient record linkage. An
+evaluation on real-world data from several domains shows that
+linkage using MSI can yield better quality than
 current indexing techniques, with similar execution cost, without the
@@
 requires the comparison of the common attributes (or fields) that are
-available within the datasets. For datasets that contain information
-about individuals, these attributes include the names, addresses,
-dates of birth, and so on, of individuals.
+available within the datasets, for example the names, addresses and
+dates of birth of individuals.
 
 To overcome data quality issues such as typographical errors and
-variations (which are common in name and address values~\cite{Chr12}),
-approximate string comparison functions (such as edit distance, the
+variations (common in name and address values~\cite{Chr12}),
+approximate string comparison functions (e.g. edit distance, the
 Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
-compare pairs of records, leading to a vector of similarities (one
-similarity per attribute compared) for each pair. These similarity
-vectors are then used to classify the compared record pairs into
-\emph{links} (where it is assumed both records in a pair correspond to
-the same real-world entity) and \emph{non-links} (where the records are
-assumed to correspond to two different entities). Various
+compare record pairs, leading to a vector of similarities (one
+similarity per attribute compared) for each pair. These
+are used to classify the record pairs into
+\emph{links} (where it is assumed both records correspond to
+the same real-world entity) and \emph{non-links} (where they are
+assumed to correspond to different entities). Various
 classification methods have been employed in record
@@
 
-Besides the issues of a lack of unique entity identifiers and data
-quality (which will affect linkage quality), record linkage is also
-challenged by the increasing sizes of the datasets to be
-linked~\cite{Don15}. To avoid full pair-wise comparison of all possible
-record pairs (quadratic in the sizes of the datasets to be linked),
+Besides a lack of unique entity identifiers, and data
+quality issues, linkage is also
+challenged by dataset scale~\cite{Don15}. To avoid full pair-wise comparison of all possible
+record pairs (quadratic in the dataset sizes),
 blocking techniques, commonly known as \emph{indexing}~\cite{Chr12b},
-are used. These split the datasets into smaller blocks in a
-computationally efficient way, such that records that are likely to
-correspond to the same entity are grouped into the same block. Only
-records within the same block are then compared in more detail.
+are used. These split the datasets into smaller blocks in an efficient way, grouping together records that are likely to
+correspond to the same entity. Only
+records within blocks are then compared in detail.
 
-While indexing techniques facilitate efficient linkage of very large
-datasets~\cite{Don15}, they generally achieve scalability at the cost
-of reduced linkage quality, because potentially true matching record
-pairs are removed in the indexing step, leading to a reduction in
-recall~\cite{Chr12}. A variety of indexing
-techniques, discussed in more detail in the following section, have
-been proposed, ranging from simple phonetic based blocking~\cite{Chr12}
-and sorting of the datasets~\cite{Dra12} to locality sensitive hashing
-based techniques~\cite{Kim10,Steorts2014}. Techniques for
-unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06}
-learning of optimal blocking schemes have also been proposed.
+While indexing allows efficient linkage of large
+datasets~\cite{Don15}, scalability is at the cost of
+reduced linkage quality, because potentially matching record pairs
+are ignored, leading to lower
+recall~\cite{Chr12}. Indexing techniques, discussed in more
+detail later, range from simple
+phonetic based blocking~\cite{Chr12} and sorting of the
+datasets~\cite{Dra12} to locality sensitive hashing based
+techniques~\cite{Kim10,Steorts2014}, and unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06}
+learning of optimal blocking schemes.
 
@@
 
-Traditional record linkage systems that perform indexing prior to
-comparison and classification, as illustrated on the left in
-\Cref{fig-rl-process}, add a further practical complexity to the
-process. Indexing, comparison and classification are often conducted
-using algorithms and parameters selected based on technical and
-domain expertise of the user of the record linkage system, followed
-by a manual assessment (auditing) of the linkage
-outcomes~\cite{Chr12}. If the quality of the resulting links is not
-good enough for a certain application, the linkage process needs to
-be repeated with different parameter settings and possibly also
-alternative algorithms. This can result in a time-consuming and
-labour-intensive iterative process. A major challenge is often the
-heuristic nature of indexing, where the choice of a certain indexing
-technique and its parameters (including which attributes to use in indexing) will determine the final outcome of a linkage.
+Traditional linkage systems that perform indexing prior to
+comparison and classification (on the left in
+\Cref{fig-rl-process}) add a further complexity. Indexing, comparison and classification are often conducted
+using algorithms and parameters selected using
+domain expertise, followed
+by manual assessment of the linkage
+outcomes~\cite{Chr12}. If the resulting link quality is too low for a certain application, the process is
+repeated with different parameter settings or algorithms, giving a time-consuming iterative process. The choice of an appropriate indexing
+technique and parameters (including which attributes to use in
+indexing) will significantly affect the final linkage outcome.
 
-This paper focuses on approaches that use a similarity threshold to
-classify links. Such techniques are fundamentally limited by the extent
+We focus on approaches using a similarity threshold to
+classify links. These are fundamentally limited by the extent
 to which true matching records are similar, and true non-matches are
 dissimilar---this is dataset-dependent. Within this domain, we define
-a technique to be \emph{complete} if it is guaranteed to identify all
-record pairs within the specified similarity threshold. Commonly used
+a technique to be \emph{complete} if it guarantees to find all
+record pairs within the specified threshold. Many
 indexing techniques are incomplete, since they reduce computational
@@
 definition, incomplete techniques yield lower recall than complete
-techniques. Conversely, and perhaps counter-intuitively, complete
+ones. Conversely, and counter-intuitively, complete
 techniques can yield lower precision with some datasets. This is
@@
 
-\emph{Metric space indexing} (MSI) is a complete technique with much
-lower computational overheads than a brute force approach. It allows
+Metric space indexing (MSI) is a complete technique with
+lower computational cost than a brute force approach. It allows
 indexing, comparison and classification to be combined into a single
-step, as illustrated in the right-hand side in
-Fig.~\ref{fig-rl-process}, making the overall record linkage process
-simpler, more efficient and more effective than incomplete indexing
+step (on the right in
+\Cref{fig-rl-process}), making the process
+simpler, more efficient and more effective than incomplete
 approaches.
 
-\smallskip
-\textbf{Contribution:} The primary contribution of this paper is the
+\textbf{Contribution:} Our primary contribution is the
 novel application of MSI to achieve complete and efficient record
 linkage, without the need for complex parameter tuning. We evaluate our
-approach on several real-world datasets from diverse domains and
+approach on several real-world datasets and
 demonstrate its advantages over existing indexing techniques for
-record linkage.
-
-%----------------------------------------------------------------------
-
-% For final paper:
-% The motivation for this work is the Digitising Scotland project
-% (Dibben 2012), which is in the process of transcribing and linking all
-% the vital events recorded in Scotland between 1856 and 1977. This data
-% set will, when complete, include around 14 million birth records, 11
-% million death records and 4 million marriage records. As part of the
-% work, certain data fields (locations, occupations and causes of death)
-% are also being classified to the relevant standard coding schemes.
-
-%----------------------------------------------------------------------
+record linkage~\footnote{The motivation for this work is the Digitising Scotland
+project~\cite{Dibben2012}, which aims to transcribe and
+link all civil registration events recorded in Scotland between
+1856 and 1973. This data set will include around 14
+million birth records, 11 million death records and 4 million marriage
+records.}.
 
@@
 
-We now briefly review relevant work in the areas of indexing for record
+We review relevant work in the areas of indexing for record
 linkage (for recent surveys see~\cite{Chr12b,Pap16}), and metric space
-indexing~\cite{Zezula2010}.
-
-Techniques to link records across datasets have been investigated for
-over five decades~\cite{Fel69,New59}, with the scalability of linking
+indexing~\cite{Zezula2010}. Techniques to link records have been investigated for
+over five decades~\cite{Fel69,New59}, with scalability
 being an ongoing challenge as datasets grow in size and complexity.
-Traditional blocking~\cite{Chr12b} uses a set of attributes (known as a
-\emph{blocking key}) to insert records that share the same value(s) in
+Traditional blocking~\cite{Chr12b} uses a set of attributes (a
+\emph{blocking key}) to insert records with the same value(s) in
 their blocking key into the same block. Only records within the same
 block are  compared to each other. To overcome variations and
-misspellings, the values used in blocking keys can be phonetically
+misspellings, the values can be phonetically
 encoded using functions such as Soundex, NYSIIS, or
@@
 sounding names (such as `Gail' and `Gayle'). Multiple blocking keys
-may also be used to deal with the problem of missing attribute values.
+may also be used to deal with missing attribute values.
 
-A different approach to indexing is the sorted neighbourhood
-method~\cite{Mon96}, where the datasets to be linked are sorted
-according to a \emph{sorting key} (usually a concatenation of the values
-from several attributes). A sliding window is then moved over the
+A different approach uses sorted neighbourhoods~\cite{Mon96}, where the datasets are sorted
+according to a \emph{sorting key} (usually a concatenation of several attribute values). A sliding window is moved over the
 datasets and only records within the window are compared. Techniques
@@
 characteristics of the sorting key values have been shown to improve
-both linkage efficiency and quality~\cite{Dra12} over
-%other sliding window
-approaches that use fixed size windows.
+both linkage efficiency and quality~\cite{Dra12}.
 
-These blocking techniques are heuristics, commonly requiring domain
+These techniques are heuristics, requiring domain
 knowledge, such as the choice of appropriate blocking or sorting keys.
 Poor choices of blocking attributes result in records being inserted
-into inappropriate blocks, and thus true matches being missed. As a
-result, such techniques can lead to \emph{incomplete} linkage.
-Conversely, many of the record pairs compared in a block may turn out
-to have low similarity, corresponding to non-matches, resulting in
+into inappropriate blocks, and thus true matches being missed, giving \emph{incomplete} linkage.
+Conversely, many pairs compared in a block may have low similarity, being non-matches, giving
 \emph{inefficient} linkage.
 
-Locality sensitive hashing (LSH), originally proposed to allow efficient
+Locality sensitive hashing (LSH), proposed for efficient
 nearest-neighbour search in high-dimensional spaces~\cite{Ind98}, has
-been employed in record linkage as an indexing technique where attribute
+been used for record linkage indexing. Attribute
 values are hashed multiple times, and blocks are created from those
 records that share some hash values. \emph{HARRA}~\cite{Kim10} is a
-record linkage approach based on MinHash~\cite{Broder1997} and LSH
-which blocks, compares, and then merges linked records in an iterative
-fashion, where merged records are re-hashed to improve overall linkage
-quality. \cite{Steorts2014} evaluates two LSH variations, concluding
-that in order to get good results, LSH methods must be tuned to the
-particular datasets being linked. This requires good quality ground
+linkage approach based on MinHash~\cite{Broder1997} and LSH
+which blocks, compares, and then merges linked records iteratively. \cite{Steorts2014} evaluates two LSH variations, concluding
+that to get good results, they must be tuned to the
+datasets. This requires good ground
 truth data which may be unavailable or expensive to obtain.
 
-Metric space indexing (MSI) techniques~\cite{Zezula2010} provide
-indexing structures to support the comparison of records in one set
-with those in another, typically also offering similarity search
-operations. To create an MSI data structure, it is necessary to define
+Metric space indexing (MSI) techniques~\cite{Zezula2010} support similarity search. 
+They require
 a distance measure between records, with certain properties including
 the \emph{triangle inequality}~\cite{Zezula2010}. Similarity search
-operations include \emph{range-search(\textbf{q},d)}, where all
-records $\mathbf{r}$ in a database $\mathbf{D}$ within a distance $d$
-of a query record $\mathbf{q}$ are identified;
-\emph{nearest-neighbour(\textbf{q})}, returning the record $\mathbf{r}
-\in \mathbf{D}$ with smallest distance to $\mathbf{q}$; and
-\emph{nearest-n(\textbf{q},n)}, returning the $n$ closest records in
-$\mathbf{D}$ to $\mathbf{q}$.
-%\cite{Zezula2010} gives an overview of MSI techniques.
-Here we choose one metric space indexing structure, the
-M-tree~\cite{Ciaccia97indexingmetric}, and investigate its efficacy
-for record linkage.
-
-The M-tree is a dynamically balanced tree structure. Every node contains
+operations include \emph{range-search(\textbf{q},d)}, identifying all
+records within a distance $d$
+of a query record $\mathbf{q}$;
+\emph{nearest-neighbour(\textbf{q})}, returning the record with smallest distance to $\mathbf{q}$; and
+\emph{nearest-n(\textbf{q},n)}, returning the $n$ closest records to $\mathbf{q}$. Here we choose one MSI
+structure, the M-tree~\cite{Ciaccia97indexingmetric}, and investigate
+its efficacy for record linkage.
+The M-tree is dynamically balanced. Every node contains
 a reference to a record being indexed, a pointer to its parent, the
@@
 
-A threshold-based linkage method using R-trees~\cite{Hjaltason1998}
-was described in~\cite{Li2006}. It demonstrates that high linkage
-quality can be achieved using Jaccard similarity over selected record
-attributes. \cite{Ciaccia97indexingmetric} shows that M-trees are
-almost always more efficient than R-trees, hence their use in the
-experiments described in this paper.
+A linkage method using R-trees~\cite{Hjaltason1998}
+was described in~\cite{Li2006}, demonstrating that high linkage
+quality can be achieved using Jaccard similarity. \cite{Ciaccia97indexingmetric} shows that M-trees are
+almost always more efficient than R-trees, hence their use here.
 
@@
 
-We address the following general record linkage problem: for two
+We address the following general linkage problem: for two
 datasets $\mathbf{D}_A$ and $\mathbf{D}_B$, we wish to find, for each
 record in $\mathbf{D}_A$, all the records in $\mathbf{D}_B$ that match
-it with regard to a certain distance threshold, $d$ (i.e.\ have a
-distance of $d$ or less). We compare several algorithms for record
-linkage: traditional blocking, one incomplete similarity search method,
-LSH-MinHash, and one complete method, M-tree. We also use a
-simple complete brute force technique as a baseline for comparison,
-although this can only feasibly be applied to the smallest of our
-datasets. All experiments have a number of parameters to configure
-the search space and algorithm behaviour, including a
-\emph{distance threshold}, $d$, specifying a maximum distance,
-equivalent to a minimum similarity, for two records to be classified
-as a link (i.e.\ referring to the same entity).
+it with regard to a certain distance threshold $d$ (i.e.\ have a
+distance of $d$ or less). We compare several linkage algorithms: traditional blocking, an incomplete similarity search method,
+LSH-MinHash, and a complete method, M-tree. We also use a
+complete brute force technique as a baseline, though
+this can only feasibly be applied to our smallest dataset. All
+experiments have a number of parameters to configure the search space
+and algorithm behaviour, including the distance function and the threshold, $d$,
+specifying the maximum distance for
+two records to be classified as a link (i.e.\ referring to the same
+entity).
+We focus on a single distance function in these experiments, to constrain the experimental space. In
+Sect.~\ref{sec-concl} we return to the selection of alternative distance functions.
 
-\smallskip
 \textbf{Brute force:}
-Two nested loops are used to compare every record in
-$\mathbf{D}_A$ with every record in $\mathbf{D}_B$. Each pair is
-classified as a link if the distance between the records is equal to
-or less than the threshold $d$. This approach is guaranteed to identify
+Every record in
+$\mathbf{D}_A$ is compared with every record in $\mathbf{D}_B$. Each pair is
+classified as a link if the distance between the records is less than or equal to
+the threshold. This always finds
 all links, with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|)$.
 
-\smallskip
 \textbf{Traditional blocking:}
-The parameters of traditional blocking are the set of blocking keys
-and the (optionally) phonetic
-% Peter: normalisation?
-encodings applied to each attribute. These are selected as described
-in~\cite{Chr12b}, exploiting knowledge of the domain and of the
-records being linked, and chosen with the intention of giving the
-best possible results for the datasets. Each record in $\mathbf{D}_A$
-is placed into the appropriate block based on its blocking key value.
-The algorithm then iterates over the records in $\mathbf{D}_B$, and
-for each one, compares it with each of the records from $\mathbf{D}_A$
-in the block with the same blocking key value.
+The parameters are the set of blocking keys and
+(optionally) the phonetic encodings applied to each attribute. These are
+selected as described in~\cite{Chr12b}, exploiting knowledge of the
+domain and of the data, and chosen with the intention of
+giving the best possible results. Each record in
+$\mathbf{D}_A$ is placed into the appropriate block based on its
+blocking key value. The algorithm then iterates over the records in
+$\mathbf{D}_B$, and for each one, compares it with each of the records
+from $\mathbf{D}_A$ in the block with the same blocking key value.
 
-\smallskip
 \textbf{LSH-MinHash:}
@@
 size} ($l_{ss}$), \emph{band size} ($l_{bs}$) and \emph{number of
-bands} ($l_{nb}$). Each record in $\mathbf{D}_A$ is placed in an
-\emph{LSH-MinHash} data structure. First, all the attributes of the
-record are concatenated, and the result \emph{shingled} into a set of
+bands} ($l_{nb}$). First, the attributes of each record in $\mathbf{D}_A$
+are concatenated, and the result \emph{shingled} into a set of
 n-grams with $n = l_{ss}$. Next, a set of deterministically generated
@@
 map associated with each of the keys. 
-
 To perform linkage, the algorithm iterates over the records in
-$\mathbf{D}_B$. Each record is hashed using the procedure described
-above, to obtain a set of keys. For each of these keys, the key is
-looked up in the LSH-MinHash data structure, and the records from
-$\mathbf{D}_A$ associated with it added to the result set. Finally,
+$\mathbf{D}_B$. Each record is hashed as described
+above, to obtain a set of keys. Each key is
+looked up in the data structure, and the associated records from
+$\mathbf{D}_A$ added to the result set. Finally,
 the record from $\mathbf{D}_B$ is compared in turn with each record
-in the result set using the same distance function and threshold, $d$,
-as with the other approaches, with the pair being classified as a link
+in the result set, with the pair being classified as a link
 or non-link based on their distance.
@@
 blocking and LSH-MinHash can yield higher precision than complete
-techniques. This can occur when the datasets contain a significant
-number of non-matches that nonetheless have high similarity. In this
+techniques. This can occur when a significant
+number of non-matches nonetheless have high similarity. In this
 situation, the fact that an incomplete technique omits consideration
@@
 
-\smallskip
-\textbf{M-tree:} Besides the distance threshold, $d$, the M-tree
-linkage algorithm has no additional parameters. In a similar manner
-to the \emph{LSH-MinHash} approach, each record in $\mathbf{D}_A$ is inserted into an M-tree. To perform linkage, the algorithm iterates
-over each record $\mathbf{b} \in \mathbf{D}_B$. A
-\emph{range-search(\textbf{b},d)} operation is performed on the
-M-tree, passing the distance threshold $d$ as the second parameter.
-All the returned records are directly classified as links.
+\textbf{M-tree:} The linkage
+algorithm has no additional parameters. As with
+\emph{LSH-MinHash}, each record in $\mathbf{D}_A$ is inserted
+into an M-tree. To perform linkage, the algorithm iterates over each
+record $\mathbf{b} \in \mathbf{D}_B$. A
+\emph{range-search(\textbf{b},d)} operation is performed on the M-tree,
+passing the distance threshold $d$ as the second parameter. All the
+returned records are directly classified as links.
 
@@
 
-We now describe the datasets we use in our evaluation, the set-up
-employed to evaluate our proposed metric indexing approach and compare
-it with traditional blocking techniques, and we then present and discuss
-our results~\footnote{The raw data, additional figures and the source
+We now describe the datasets and method used in our evaluation~\footnote{The raw data, additional figures and the source
 code required to run these experiments can be downloaded from:
 (link removed for double blind review).}.
+\label{sec-data}
+We used three datasets from two domains in our experiments, as
+summarised in Table~\ref{table-datasets}. The first is
+\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer
+to 112 machine learning publications. Cora is commonly used as a
+benchmark dataset in the literature for assessing linkage algorithms.
+Ground truth is provided via a unique \emph{paper\_id} identifier of
+the form ``blum1993". In this experiment linkage is performed over
+the same set of records (i.e.\ a de-duplication~\cite{Chr12}).
 
@@
   \begin{scriptsize}
-  %\addtolength{\tabcolsep}{-0.5pt}
   \begin{tabular}{ccccc}
@@
 
-%\smallskip
-%\textbf{Datasets:}
-\label{sec-data}
-We used three datasets from two domains in our experiments, as
-summarised in Table~\ref{table-datasets}. The first is
-\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer
-to 112 machine learning publications. Cora is commonly used as a
-benchmark dataset in the literature for assessing linkage algorithms.
-Ground truth is provided via a unique \emph{paper\_id} identifier of
-the form ``blum1993". In this experiment linkage is performed over
-the same set of records (i.e.\ a de-duplication~\cite{Chr12}).
-
 The other two datasets are historical Scottish records of vital events
-(birth, marriages and deaths) one registered on the \emph{Isle of
+(birth, marriages and deaths), one registered on the \emph{Isle of
 Skye}, a rural district, and the other records from
-\emph{Kilmarnock}, an industrial town. These datasets were created by
-historical demographers who extensively curated and linked both
-datasets~\cite{reid2006,reid2002}. Both datasets include the names,
-gender, addresses of individuals and their parents. Ground truth was
+\emph{Kilmarnock}, an industrial town. These datasets were created, curated and linked by
+historical demographers~\cite{reid2006,reid2002}. Both include the names and
+genders of individuals and their parents. Ground truth was
 generated by the demographers based on their extensive domain
@@
 
-In all of our experiments we use a single distance metric which is the
-sum of the Levenshtein~\cite{Levenshtein66} edit distances between the
-set of attributes being compared.
+In all of our experiments we use a single distance metric: the
+sum of the attribute-level Levenshtein~\cite{Levenshtein66} edit distances.
 
@@
 
-In this section, we perform linkage on the Cora dataset using all four
+We perform linkage on the Cora dataset using all
 approaches presented in this paper: brute force, traditional blocking,
-LSH and M-tree, using several selected configurations for the blocking
-and LSH approaches. The distance threshold is varied between $0 \le d
-\le 250$~\footnote{Relatively high Levenshtein edit distances are
+LSH and M-tree, using several selected configurations for blocking
+and LSH. The distance threshold is varied between 0 and 250~\footnote{Relatively high Levenshtein edit distances are
 included since Cora includes a number of low-similarity true
-matches.}. For traditional blocking, the following Cora attributes are
+matches.}. For traditional blocking, the following attributes are
 used individually as blocking keys: \emph{author}, \emph{title},
@@
 
-\Cref{cora-quality} shows the precision, recall, and F-measure~\cite{Chr12} for various thresholds. As expected, low 
+\Cref{cora-quality} shows the precision, recall, and
+F-measure~\cite{Chr12} for various thresholds. As expected, low
 thresholds give high precision and low recall, and the reverse for high
-threshold values. Brute force and M-tree give identical results, as
-expected. The best linkage quality, with an F-measure of around 0.7,
-is achieved by a number of linkers, including brute force, M-tree, blocking on \emph{authors} and blocking on all attributes, and two
-of the LSH configurations. All of these give similar overall results,
-apart from blocking on \emph{authors}, which gives much better quality
-at very high distance thresholds. This is due to the incomplete
-nature of the approach, avoiding comparisons of significant numbers
-of high-similarity non-matches and thus avoiding these becoming false
-positives and keeping precision high.
+thresholds. Brute force and M-tree give identical results, as
+expected. The best linkage quality, with an F-measure of around 0.7, is
+achieved by several linkers, including brute force, M-tree, blocking
+on \emph{authors}, blocking on all attributes, and two of the LSH
+configurations. All of these give similar overall results, apart from
+blocking on \emph{authors}, which gives much better quality at very high
+distance thresholds. This is due to the incomplete nature of the
+approach, avoiding comparisons of significant numbers of high-similarity
+non-matches and thus avoiding these becoming false positives and keeping
+precision high.
 
@@
   ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
-%\parbox{0.2\textwidth}{\centering Linker} &
-%\parbox{0.2\textwidth}{\centering ~~~Comparisons} &
-%\parbox{0.17\textwidth}{\centering ~~~Precision} &
-%\parbox{0.17\textwidth}{\centering ~~~Recall} &
-%\parbox{0.17\textwidth}{\centering ~~~F1 Measure} \\
   \noalign{\smallskip} \hline \noalign{\smallskip}
@@
   ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
-%\parbox{0.2\textwidth}{\centering Linker} &
-%\parbox{0.2\textwidth}{\centering Comparisons} &
-%\parbox{0.17\textwidth}{\centering Precision} &
-%\parbox{0.17\textwidth}{\centering Recall} &
-%\parbox{0.17\textwidth}{\centering F1 Measure} \\
 \noalign{\smallskip} \hline \noalign{\smallskip}
@@
 
-The observation that LSH can yield results quickly at the cost of some
-recall suggests that hybrid algorithms may be a potentially profitable
-future direction for exploration. We hypothesise that it may be possible
-to use an incomplete technique like LSH to form indices into a data
-structure such as an M-tree. In this manner, M-tree nodes could be used
-as the starting point for range-search queries and, in so doing, prune
-the search space considerably to yield high quality results quickly.
-We plan to develop such hybrid indexing approaches with the aim to
-achieve fully automated, complete and efficient record linkage.
+We have focused on a single distance function: the sum of attribute-level Levenshtein distances.
+This gives a straightforward intuition of record-level distances, but it is more common to normalise metrics to the range 0-1.
+
+Many distance functions, such as Jaro-Winkler, are not metric, and therefore cannot be used with MSI techniques.
+Care must be taken to preserve metric properties when combining metrics over individual fields, as is highlighted in~\cite{Yujian2007}, since this may yield a function that is not metric.
+It is then possible for MSI techniques to yield results that are subtly incorrect.
+
+Different metrics can give different distributions of inter-record distances, which can affect both the linkage results and the number of comparisons made, and hence the efficiency of the algorithm.
+Datasets with low variation in inter-record distances are said to have high \textit{intrinsic dimensionality}, and tend to require high numbers of comparisons.
+The Scottish vital event datasets, combined with Levenshtein-based metrics, appear to have high intrinsic dimensionality.
+We are now, therefore, investigating the application of various different metrics to this domain, including Jensen-Shannon, Cosine and Structured Entropic Distance, as described in~\cite{Connor2016}.
+We are also investigating the applicability of a novel technique for dimensionality reduction~\cite{Connor2017}.
+
+%----------------------------------------------------------------------
+
+\section{Acknowledgements}
+
+This work was supported by ESRC grants ES/K00574X/2 ``Digitising Scotland''
+and ES/L007487/1 ``Administrative Data Research Centre---Scotland''.
+
+We thank Alice Reid of the University of
+Cambridge and her colleagues, especially Ros Davies and Eilidh Garrett,
+for the work undertaken on the Kilmarnock and Isle of Skye databases.
 
@@
 %----------------------------------------------------------------------
-
-%text cut from earlier versions:
-
-%The MIFile \cite{amato2014mi} approach is an inexact technique for performing similarity search based on metric spaces.
-Using  MI\_file,  each record is represented by the ordering of distances from a collection of reference objects.
-% Peter: are these reference object also records? 
-It is based on the idea that two records close to each other will have similar neighbours and will thus be represented by similar of distances
-% Peter: unclear what you mean with: similar of distances
-from the chosen reference objects. 
-In order to create the indexes, for each value to be added to the
-% peter: value -> record ?
-data-structure, first the k nearest reference objects (governed by a fixed parameter) are found, each is given a score (drawn from 1,2,3... based on the reference object's position from the value).
-% peter: this is unclear .. I think I understand k-NN: for each
-% record you find the k nearest reference objects and you enumerate
-% them
-Next for each score, a  mapping is created in an inverted index mapping
-% peter: unclear..? a mapping in an inverted index mapping?
-from the value to the score of each reference object and a representation of the data associated with the reference object and the value inserted.
-% peter: unclear: what is "the data" associated with the reference
-% object.
-In order to perform a \textit{nearest-neighbour} lookup, first the k nearest reference objects (bounded by a second fixed parameter).
-% peter: this sentence reads incomplete..?
-Next the objects closest to these reference objects are extracted from the inverted map by choosing the neighbours with the highest score.
-% Peter: sorry, this is unclear as well. "the objects closest" ->
-% records closest? And why highest scores..? Is the closest record
-% to a reference object given the highest score (not the lowest
-% score - i.e. not 1-closest, 2-closest etc)
-Much of the search space may be eliminated by  keeping track of the scores that  can contribute to the nearest neighbours.
-% peter: so I guess you use the triangular inequality and the reference
-% objects to prune the search space.
-
-\cite{Yu2016} surveys the use of string similarity in record linkage.
-Algorithms used to compare two records are characterised as being
-character-based or token-based. The first category, including approaches
-such as Levenshtein~\cite{Levenshtein66}, uses character differences in
-the strings being compared. The second, including techniques such as
-Jaccard similarity, treats records as sets of tokens, and utilises set
-similarity. There are also hybrids of the two approaches.
-% Do we need the rest of this paragraph?
-\cite{Yu2016} categorises linkage algorithms as being based on filtering,
-verification, or thresholds.
-% This is missing some context as to what is going on.
-Filtering algorithms generate signatures for each record, construct
-inverted lists containing records with common signatures, and prune
-records with no common signatures.
-% Distance of prefix to what? What is the data in this context?
-Verification algorithms calculate the exact distances of a prefix of the
-data, and use another technique to estimate the edit distance between
-records. If the estimated distance is larger than the desired similarity
-threshold, the candidate may be discarded.
-% This seems a bit throw-away.
-% Also, these aren't themselves linkage algorithms are they?
-Threshold-based algorithms include M-trees, Trie-Join, Pass-Join and
-Partenum.
-
-%We will evaluate the performance of LSH based blocking with regard to the three parameters $l_{ss}$, $l_{bs}$ and $l_{nb}$ in
-Sect.~\ref{sec-exp}.
-
-% \begin{table}[t]
-% \caption{Parameter settings used for the different datasets used in
-%    the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
-%    shingle size, and the size and number of bands, respectively.}
-%  \label{table-parameters}
-%   \centering
-%   \begin{scriptsize}
-%   %\addtolength{\tabcolsep}{-0.5pt}
-%   \begin{tabular}{ccccc}
-%   \hline\noalign{\smallskip}
-%   Dataset~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
-%   name(s) & & & &  \\
-%   \noalign{\smallskip} \hline \noalign{\smallskip}
-%   CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
-%     ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
-%   Isle of Skye &  \\
-%   Kilmarnock  &  \\
-%   \noalign{\smallskip} \hline
-%   \end{tabular}
-%   \end{scriptsize}
-% \end{table}
-
-% \textbf{Experimental Setup:}
-% %
-% We implemented all techniques described in Sect.~\ref{sec-approach}
-% using Java (version 1.8.0\_144) and ran an extensive set of experiments on
-% a compute server with 24 Intel(R) Xeon(R) processors running at 2.30GHz, 
-% 128 GBytes of main memory and running Red Hat 4.8.5-11.

### 2018-02-16 paper.tex
Update on Overleaf.

@@
 common attributes, with every pair being classified as a link or
-non-link depending on their similarity. Linkage is
-usually performed in a three-step process: first, groups of similar
-candidate records are identified using indexing, then pairs within the same
-group are compared in more detail, and finally classified.
-Even state-of-the-art indexing techniques, such as Locality Sensitive
-Hashing, have potential drawbacks. They may fail to group together some
-true matching records with high similarity, or they may group
-records with low similarity, leading to high computational overhead.
-We propose using \emph{metric space indexing} (MSI) to perform \emph{complete}
-linkage, resulting in a parameter-free process
-combining indexing, comparison and classification into a single step
-delivering complete and efficient record linkage. An
-evaluation on real-world data from several domains shows that
-linkage using MSI can yield better quality than
-current indexing techniques, with similar execution cost, without the
-need for domain knowledge or trial and error to configure the process.
+non-link depending on their similarity. Linkage is usually performed in
+a three-step process: first, groups of similar candidate records are
+identified using indexing, then pairs within the same group are compared
+in more detail, and finally classified. Even state-of-the-art indexing
+techniques, such as Locality Sensitive Hashing, have potential
+drawbacks. They may fail to group together some true matching records
+with high similarity, or they may group records with low similarity,
+leading to high computational overhead. We propose using \emph{metric
+space indexing} (MSI) to perform \emph{complete} linkage, resulting in a
+parameter-free process combining indexing, comparison and classification
+into a single step delivering complete and efficient record linkage. An
+evaluation on real-world data from several domains shows that linkage
+using MSI can yield better quality than current indexing techniques,
+with similar execution cost, without the need for domain knowledge or
+trial and error to configure the process.
 
@@
 To overcome data quality issues such as typographical errors and
-variations (common in name and address values~\cite{Chr12}),
-approximate string comparison functions (e.g. edit distance, the
-Jaro-Winkler comparator, or Jaccard similarity~\cite{Chr12}) are used to
-compare record pairs, leading to a vector of similarities (one
-similarity per attribute compared) for each pair. These
-are used to classify the record pairs into
-\emph{links} (where it is assumed both records correspond to
-the same real-world entity) and \emph{non-links} (where they are
-assumed to correspond to different entities). Various
-classification methods have been employed in record
-linkage~\cite{Chr12,Don15}, ranging from simple threshold-based to
-sophisticated clustering, supervised classification techniques, and
-active learning approaches~\cite{Wan15}.
+variations (common in name and address values~\cite{Chr12}), approximate
+string comparison functions (e.g. edit distance, the Jaro-Winkler
+comparator, or Jaccard similarity~\cite{Chr12}) are used to compare
+record pairs, leading to a vector of similarities (one similarity per
+attribute compared) for each pair. These are used to classify the record
+pairs into \emph{links} (where it is assumed both records correspond to
+the same real-world entity) and \emph{non-links} (where they are assumed
+to correspond to different entities). Various classification methods
+have been employed in record linkage~\cite{Chr12,Don15}, ranging from
+simple threshold-based to sophisticated clustering, supervised
+classification techniques, and active learning approaches~\cite{Wan15}.
 
-Besides a lack of unique entity identifiers, and data
-quality issues, linkage is also
-challenged by dataset scale~\cite{Don15}. To avoid full pair-wise comparison of all possible
-record pairs (quadratic in the dataset sizes),
-blocking techniques, commonly known as \emph{indexing}~\cite{Chr12b},
-are used. These split the datasets into smaller blocks in an efficient way, grouping together records that are likely to
-correspond to the same entity. Only
-records within blocks are then compared in detail.
+Besides a lack of unique entity identifiers, and data quality issues,
+linkage is also challenged by dataset scale~\cite{Don15}. To avoid full
+pair-wise comparison of all possible record pairs (quadratic in the
+dataset sizes), blocking techniques, commonly known as
+\emph{indexing}~\cite{Chr12b}, are used. These split the datasets into
+smaller blocks in an efficient way, grouping together records that are
+likely to correspond to the same entity. Only records within blocks are
+then compared in detail.
 
-While indexing allows efficient linkage of large
-datasets~\cite{Don15}, scalability is at the cost of
-reduced linkage quality, because potentially matching record pairs
-are ignored, leading to lower
-recall~\cite{Chr12}. Indexing techniques, discussed in more
-detail later, range from simple
-phonetic based blocking~\cite{Chr12} and sorting of the
-datasets~\cite{Dra12} to locality sensitive hashing based
-techniques~\cite{Kim10,Steorts2014}, and unsupervised~\cite{Kej13,Ram15} and supervised~\cite{Bil06,Mic06}
-learning of optimal blocking schemes.
+While indexing allows efficient linkage of large datasets~\cite{Don15},
+scalability is at the cost of reduced linkage quality, because
+potentially matching record pairs are ignored, leading to lower
+recall~\cite{Chr12}. Indexing techniques, discussed in more detail
+later, range from simple phonetic based blocking~\cite{Chr12} and
+sorting of the datasets~\cite{Dra12} to locality sensitive hashing based
+techniques~\cite{Kim10,Steorts2014}, and unsupervised~\cite{Kej13,Ram15}
+and supervised~\cite{Bil06,Mic06} learning of optimal blocking schemes.
 
@@
 
-Traditional linkage systems that perform indexing prior to
-comparison and classification (on the left in
-\Cref{fig-rl-process}) add a further complexity. Indexing, comparison and classification are often conducted
-using algorithms and parameters selected using
-domain expertise, followed
-by manual assessment of the linkage
-outcomes~\cite{Chr12}. If the resulting link quality is too low for a certain application, the process is
-repeated with different parameter settings or algorithms, giving a time-consuming iterative process. The choice of an appropriate indexing
-technique and parameters (including which attributes to use in
+Traditional linkage systems that perform indexing prior to comparison
+and classification (on the left in \Cref{fig-rl-process}) add a further
+complexity. Indexing, comparison and classification are often conducted
+using algorithms and parameters selected using domain expertise,
+followed by manual assessment of the linkage outcomes~\cite{Chr12}. If
+the resulting link quality is too low for a certain application, the
+process is repeated with different parameter settings or algorithms,
+giving a time-consuming iterative process. The choice of an appropriate
+indexing technique and parameters (including which attributes to use in
 indexing) will significantly affect the final linkage outcome.
 
-We focus on approaches using a similarity threshold to
-classify links. These are fundamentally limited by the extent
-to which true matching records are similar, and true non-matches are
-dissimilar---this is dataset-dependent. Within this domain, we define
-a technique to be \emph{complete} if it guarantees to find all
-record pairs within the specified threshold. Many
-indexing techniques are incomplete, since they reduce computational
-cost at the expense of potentially overlooking some true matches. By
-definition, incomplete techniques yield lower recall than complete
-ones. Conversely, and counter-intuitively, complete
+We focus on approaches using a similarity threshold to classify links.
+These are fundamentally limited by the extent to which true matching
+records are similar, and true non-matches are dissimilar---this is
+dataset-dependent. Within this domain, we define a technique to be
+\emph{complete} if it guarantees to find all record pairs within the
+specified threshold. Many indexing techniques are incomplete, since they
+reduce computational cost at the expense of potentially overlooking some
+true matches. By definition, incomplete techniques yield lower recall
+than complete ones. Conversely, and counter-intuitively, complete
 techniques can yield lower precision with some datasets. This is
@@
 
-Metric space indexing (MSI) is a complete technique with
-lower computational cost than a brute force approach. It allows
-indexing, comparison and classification to be combined into a single
-step (on the right in
-\Cref{fig-rl-process}), making the process
-simpler, more efficient and more effective than incomplete
-approaches.
+Metric space indexing (MSI) is a complete technique with lower
+computational cost than a brute force approach. It allows indexing,
+comparison and classification to be combined into a single step (on the
+right in \Cref{fig-rl-process}), making the process simpler, more
+efficient and more effective than incomplete approaches.
 
-\textbf{Contribution:} Our primary contribution is the
-novel application of MSI to achieve complete and efficient record
-linkage, without the need for complex parameter tuning. We evaluate our
-approach on several real-world datasets and
-demonstrate its advantages over existing indexing techniques for
-record linkage~\footnote{The motivation for this work is the Digitising Scotland
-project~\cite{Dibben2012}, which aims to transcribe and
-link all civil registration events recorded in Scotland between
-1856 and 1973. This data set will include around 14
-million birth records, 11 million death records and 4 million marriage
-records.}.
+The motivation for this work is
+the Digitising Scotland project~\cite{Dibben2012}, which aims to
+transcribe and link all civil registration events recorded in Scotland
+between 1856 and 1973. This data set will include around 14 million
+birth records, 11 million death records and 4 million marriage records.
+
+\smallskip
+\textbf{Contribution:} Our primary contribution is the novel application
+of MSI to achieve complete and efficient record linkage, without the
+need for complex parameter tuning. We evaluate our approach on several
+real-world datasets and demonstrate its advantages over existing
+indexing techniques for record linkage.
 
@@
 
-We review relevant work in the areas of indexing for record
-linkage (for recent surveys see~\cite{Chr12b,Pap16}), and metric space
-indexing~\cite{Zezula2010}. Techniques to link records have been investigated for
-over five decades~\cite{Fel69,New59}, with scalability
+We review relevant work in the areas of indexing for record linkage (for
+recent surveys see~\cite{Chr12b,Pap16}), and metric space
+indexing~\cite{Zezula2010}. Techniques to link records have been
+investigated for over five decades~\cite{Fel69,New59}, with scalability
 being an ongoing challenge as datasets grow in size and complexity.
 Traditional blocking~\cite{Chr12b} uses a set of attributes (a
-\emph{blocking key}) to insert records with the same value(s) in
-their blocking key into the same block. Only records within the same
-block are  compared to each other. To overcome variations and
-misspellings, the values can be phonetically
-encoded using functions such as Soundex, NYSIIS, or
-Double-Metaphone~\cite{Chr12}. These convert a string into a code
-according to its pronunciation, assigning the same code to similar
-sounding names (such as `Gail' and `Gayle'). Multiple blocking keys
-may also be used to deal with missing attribute values.
+\emph{blocking key}) to insert records with the same value(s) in their
+blocking key into the same block. Only records within the same block are
+ compared to each other. To overcome variations and misspellings, the
+values can be phonetically encoded using functions such as Soundex,
+NYSIIS, or Double-Metaphone~\cite{Chr12}. These convert a string into a
+code according to its pronunciation, assigning the same code to similar
+sounding names (such as `Gail' and `Gayle'). Multiple blocking keys may
+also be used to deal with missing attribute values.
 
-A different approach uses sorted neighbourhoods~\cite{Mon96}, where the datasets are sorted
-according to a \emph{sorting key} (usually a concatenation of several attribute values). A sliding window is moved over the
-datasets and only records within the window are compared. Techniques
-that adaptively shrink or expand the window size based on the
+A different approach uses sorted neighbourhoods~\cite{Mon96}, where the
+datasets are sorted according to a \emph{sorting key} (usually a
+concatenation of several attribute values). A sliding window is moved
+over the datasets and only records within the window are compared.
+Techniques that adaptively shrink or expand the window size based on the
 characteristics of the sorting key values have been shown to improve
@@
 
-These techniques are heuristics, requiring domain
-knowledge, such as the choice of appropriate blocking or sorting keys.
-Poor choices of blocking attributes result in records being inserted
-into inappropriate blocks, and thus true matches being missed, giving \emph{incomplete} linkage.
-Conversely, many pairs compared in a block may have low similarity, being non-matches, giving
-\emph{inefficient} linkage.
+These techniques are heuristics, requiring domain knowledge, such as the
+choice of appropriate blocking or sorting keys. Poor choices of blocking
+attributes result in records being inserted into inappropriate blocks,
+and thus true matches being missed, giving \emph{incomplete} linkage.
+Conversely, many pairs compared in a block may have low similarity,
+being non-matches, giving \emph{inefficient} linkage.
 
@@
 nearest-neighbour search in high-dimensional spaces~\cite{Ind98}, has
-been used for record linkage indexing. Attribute
-values are hashed multiple times, and blocks are created from those
-records that share some hash values. \emph{HARRA}~\cite{Kim10} is a
-linkage approach based on MinHash~\cite{Broder1997} and LSH
-which blocks, compares, and then merges linked records iteratively. \cite{Steorts2014} evaluates two LSH variations, concluding
-that to get good results, they must be tuned to the
-datasets. This requires good ground
-truth data which may be unavailable or expensive to obtain.
+been used for record linkage indexing. Attribute values are hashed
+multiple times, and blocks are created from those records that share
+some hash values. \emph{HARRA}~\cite{Kim10} is a linkage approach based
+on MinHash~\cite{Broder1997} and LSH which blocks, compares, and then
+merges linked records iteratively. \cite{Steorts2014} evaluates two LSH
+variations, concluding that to get good results, they must be tuned to
+the datasets. This requires good ground truth data which may be
+unavailable or expensive to obtain.
 
-Metric space indexing (MSI) techniques~\cite{Zezula2010} support similarity search. 
-They require
-a distance measure between records, with certain properties including
-the \emph{triangle inequality}~\cite{Zezula2010}. Similarity search
-operations include \emph{range-search(\textbf{q},d)}, identifying all
-records within a distance $d$
-of a query record $\mathbf{q}$;
-\emph{nearest-neighbour(\textbf{q})}, returning the record with smallest distance to $\mathbf{q}$; and
-\emph{nearest-n(\textbf{q},n)}, returning the $n$ closest records to $\mathbf{q}$. Here we choose one MSI
+Metric space indexing (MSI) techniques~\cite{Zezula2010} support
+similarity search. They require a distance measure between records, with
+certain properties including the \emph{triangle
+inequality}~\cite{Zezula2010}. Similarity search operations include
+\emph{range-search(\textbf{q},d)}, identifying all records within a
+distance $d$ of a query record $\mathbf{q}$;
+\emph{nearest-neighbour(\textbf{q})}, returning the record with smallest
+distance to $\mathbf{q}$; and \emph{nearest-n(\textbf{q},n)}, returning
+the $n$ closest records to $\mathbf{q}$. Here we choose one MSI
 structure, the M-tree~\cite{Ciaccia97indexingmetric}, and investigate
-its efficacy for record linkage.
-The M-tree is dynamically balanced. Every node contains
-a reference to a record being indexed, a pointer to its parent, the
-distance to its parent, and the node's radius. The radius of a node is
-the distance from it to its furthest child. For a parent node with
-radius \textit{r}, all its children may be visualised as being contained
-within a ball of radius \textit{r} from it.
+its efficacy for record linkage. The M-tree is dynamically balanced.
+Every node contains a reference to a record being indexed, a pointer to
+its parent, the distance to its parent, and the node's radius. The
+radius of a node is the distance from it to its furthest child. For a
+parent node with radius \textit{r}, all its children may be visualised
+as being contained within a ball of radius \textit{r} from it.
 
-A linkage method using R-trees~\cite{Hjaltason1998}
-was described in~\cite{Li2006}, demonstrating that high linkage
-quality can be achieved using Jaccard similarity. \cite{Ciaccia97indexingmetric} shows that M-trees are
-almost always more efficient than R-trees, hence their use here.
+A linkage method using R-trees~\cite{Hjaltason1998} was described
+in~\cite{Li2006}, demonstrating that high linkage quality can be
+achieved using Jaccard similarity. \cite{Ciaccia97indexingmetric} shows
+that M-trees are almost always more efficient than R-trees, hence their
+use here.
 
@@
 
-We address the following general linkage problem: for two
-datasets $\mathbf{D}_A$ and $\mathbf{D}_B$, we wish to find, for each
-record in $\mathbf{D}_A$, all the records in $\mathbf{D}_B$ that match
-it with regard to a certain distance threshold $d$ (i.e.\ have a
-distance of $d$ or less). We compare several linkage algorithms: traditional blocking, an incomplete similarity search method,
-LSH-MinHash, and a complete method, M-tree. We also use a
-complete brute force technique as a baseline, though
-this can only feasibly be applied to our smallest dataset. All
-experiments have a number of parameters to configure the search space
-and algorithm behaviour, including the distance function and the threshold, $d$,
-specifying the maximum distance for
-two records to be classified as a link (i.e.\ referring to the same
-entity).
-We focus on a single distance function in these experiments, to constrain the experimental space. In
-Sect.~\ref{sec-concl} we return to the selection of alternative distance functions.
+We address the following general linkage problem: for two datasets
+$\mathbf{D}_A$ and $\mathbf{D}_B$, we wish to find, for each record in
+$\mathbf{D}_A$, all the records in $\mathbf{D}_B$ that match it with
+regard to a certain distance threshold $d$ (i.e.\ have a distance of $d$
+or less). We compare several linkage algorithms: traditional blocking,
+an incomplete similarity search method, LSH-MinHash, and a complete
+method, M-tree. We also use a complete brute force technique as a
+baseline, though this can only feasibly be applied to our smallest
+dataset. All experiments have a number of parameters to configure the
+search space and algorithm behaviour, including the distance function
+and the threshold, $d$, specifying the maximum distance for two records
+to be classified as a link (i.e.\ referring to the same entity). We
+focus on a single distance function in these experiments, to constrain
+the experimental space. In Sect.~\ref{sec-concl} we return to the
+selection of alternative distance functions.
 
-\textbf{Brute force:}
-Every record in
-$\mathbf{D}_A$ is compared with every record in $\mathbf{D}_B$. Each pair is
-classified as a link if the distance between the records is less than or equal to
-the threshold. This always finds
-all links, with complexity $ O(|\mathbf{D}_A| \times |\mathbf{D}_B|)$.
+\textbf{Brute force:} Every record in $\mathbf{D}_A$ is compared with
+every record in $\mathbf{D}_B$. Each pair is classified as a link if the
+distance between the records is less than or equal to the threshold.
+This always finds all links, with complexity $ O(|\mathbf{D}_A| \times
+|\mathbf{D}_B|)$.
 
-\textbf{Traditional blocking:}
-The parameters are the set of blocking keys and
-(optionally) the phonetic encodings applied to each attribute. These are
-selected as described in~\cite{Chr12b}, exploiting knowledge of the
-domain and of the data, and chosen with the intention of
-giving the best possible results. Each record in
-$\mathbf{D}_A$ is placed into the appropriate block based on its
-blocking key value. The algorithm then iterates over the records in
-$\mathbf{D}_B$, and for each one, compares it with each of the records
-from $\mathbf{D}_A$ in the block with the same blocking key value.
+\textbf{Traditional blocking:} The parameters are the set of blocking
+keys and (optionally) the phonetic encodings applied to each attribute.
+These are selected as described in~\cite{Chr12b}, exploiting knowledge
+of the domain and of the data, and chosen with the intention of giving
+the best possible results. Each record in $\mathbf{D}_A$ is placed into
+the appropriate block based on its blocking key value. The algorithm
+then iterates over the records in $\mathbf{D}_B$, and for each one,
+compares it with each of the records from $\mathbf{D}_A$ in the block
+with the same blocking key value.
 
-\textbf{LSH-MinHash:}
-The parameters for LSH-Minhash are~\cite{Broder1997} \emph{shingle
-size} ($l_{ss}$), \emph{band size} ($l_{bs}$) and \emph{number of
-bands} ($l_{nb}$). First, the attributes of each record in $\mathbf{D}_A$
-are concatenated, and the result \emph{shingled} into a set of
-n-grams with $n = l_{ss}$. Next, a set of deterministically generated
-hash functions are applied to each n-gram in the set and the smallest
-result (the MinHash) of each hash application is added to a signature
-for the record. The number of hashes used, and thus the size of the
-signature, is set to $l_{nb} \times l_{bs}$. Finally, the signature is
-split into $l_{nb}$ bands and the values from each band are hashed
-again to create a number of keys. The original record is added to a
-map associated with each of the keys. 
-To perform linkage, the algorithm iterates over the records in
-$\mathbf{D}_B$. Each record is hashed as described
-above, to obtain a set of keys. Each key is
-looked up in the data structure, and the associated records from
-$\mathbf{D}_A$ added to the result set. Finally,
-the record from $\mathbf{D}_B$ is compared in turn with each record
-in the result set, with the pair being classified as a link
-or non-link based on their distance.
-
-In some circumstances, incomplete approaches such as traditional
-blocking and LSH-MinHash can yield higher precision than complete
-techniques. This can occur when a significant
-number of non-matches nonetheless have high similarity. In this
-situation, the fact that an incomplete technique omits consideration
-of some potential links can serve to improve precision, since a
-classification decision based on a certain similarity threshold is
-incorrect for high-similarity non-matches. By definition, recall can
-never be higher for incomplete techniques.
-
-\textbf{M-tree:} The linkage
-algorithm has no additional parameters. As with
-\emph{LSH-MinHash}, each record in $\mathbf{D}_A$ is inserted
-into an M-tree. To perform linkage, the algorithm iterates over each
-record $\mathbf{b} \in \mathbf{D}_B$. A
-\emph{range-search(\textbf{b},d)} operation is performed on the M-tree,
-passing the distance threshold $d$ as the second parameter. All the
-returned records are directly classified as links.
-
-%----------------------------------------------------------------------
-
-\section{Experiments and Results}
-\label{sec-exp}
-
-We now describe the datasets and method used in our evaluation~\footnote{The raw data, additional figures and the source
-code required to run these experiments can be downloaded from:
-(link removed for double blind review).}.
-\label{sec-data}
-We used three datasets from two domains in our experiments, as
-summarised in Table~\ref{table-datasets}. The first is
-\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer
-to 112 machine learning publications. Cora is commonly used as a
-benchmark dataset in the literature for assessing linkage algorithms.
-Ground truth is provided via a unique \emph{paper\_id} identifier of
-the form ``blum1993". In this experiment linkage is performed over
-the same set of records (i.e.\ a de-duplication~\cite{Chr12}).
-
-%----------------------------------------------------------------------
+\textbf{LSH-MinHash:} The parameters for LSH-Minhash
+are~\cite{Broder1997} \emph{shingle size} ($l_{ss}$), \emph{band size}
+($l_{bs}$) and \emph{number of bands} ($l_{nb}$). First, the attributes
+of each record in $\mathbf{D}_A$ are concatenated, and the result
+\emph{shingled} into a set of n-grams with $n = l_{ss}$. Next, a set of
+deterministically generated hash functions are applied to each n-gram in
+the set and the smallest result (the MinHash) of each hash application
+is added to a signature for the record. The number of hashes used, and
+thus the size of the signature, is set to $l_{nb} \times l_{bs}$.
+Finally, the signature is split into $l_{nb}$ bands and the values from
+each band are hashed again to create a number of keys. The original
+record is added to a map associated with each of the keys. To perform
+linkage, the algorithm iterates over the records in $\mathbf{D}_B$. Each
+record is hashed as described above, to obtain a set of keys. Each key
+is looked up in the data structure, and the associated records from
+$\mathbf{D}_A$ added to the result set. Finally, the record from
+$\mathbf{D}_B$ is compared in turn with each record in the result set,
+with the pair being classified as a link or non-link based on their
+distance.
 
@@
 
+In some circumstances, incomplete approaches such as traditional
+blocking and LSH-MinHash can yield higher precision than complete
+techniques. This can occur when a significant number of non-matches
+nonetheless have high similarity. In this situation, the fact that an
+incomplete technique omits consideration of some potential links can
+serve to improve precision, since a classification decision based on a
+certain similarity threshold is incorrect for high-similarity
+non-matches. By definition, recall can never be higher for incomplete
+techniques.
+
+\textbf{M-tree:} The linkage algorithm has no additional parameters. As
+with \emph{LSH-MinHash}, each record in $\mathbf{D}_A$ is inserted into
+an M-tree. To perform linkage, the algorithm iterates over each record
+$\mathbf{b} \in \mathbf{D}_B$. A \emph{range-search(\textbf{b},d)}
+operation is performed on the M-tree, passing the distance threshold $d$
+as the second parameter. All the returned records are directly
+classified as links.
+
+%----------------------------------------------------------------------
+
+\section{Experiments and Results}
+\label{sec-exp}
+
+We now describe the datasets and method used in our
+evaluation, as available from: \url{ADD\_URL\_TO\_GITHUB?}
+%~\footnote{The raw data, additional figures and the source
+%code required to run these experiments can be downloaded from: (link
+%removed for double blind review).}.
+\label{sec-data} We used three
+datasets from two domains in our experiments, as summarised in
+Table~\ref{table-datasets}. The first is \emph{Cora}~\cite{Cora2017},
+which contains 1,295 records that refer to 112 machine learning
+publications. Cora is commonly used as a benchmark dataset in the
+literature for assessing linkage algorithms. Ground truth is provided
+via a unique \emph{paper\_id} identifier of the form ``blum1993". In
+this experiment linkage is performed over the same set of records (i.e.\
+a de-duplication~\cite{Chr12}).
+
 %----------------------------------------------------------------------
@@
 (birth, marriages and deaths), one registered on the \emph{Isle of
-Skye}, a rural district, and the other records from
-\emph{Kilmarnock}, an industrial town. These datasets were created, curated and linked by
-historical demographers~\cite{reid2006,reid2002}. Both include the names and
-genders of individuals and their parents. Ground truth was
-generated by the demographers based on their extensive domain
-knowledge.
+Skye}, a rural district, and the other records from \emph{Kilmarnock},
+an industrial town. These datasets were created, curated and linked by
+historical demographers~\cite{reid2006,reid2002}. Both include the names
+and genders of individuals and their parents. Ground truth was generated
+by the demographers based on their extensive domain knowledge.
 
-In all of our experiments we use a single distance metric: the
-sum of the attribute-level Levenshtein~\cite{Levenshtein66} edit distances.
+In all of our experiments we use a single distance metric: the sum of
+the attribute-level Levenshtein~\cite{Levenshtein66} edit distances.
 
@@
 
-We perform linkage on the Cora dataset using all
-approaches presented in this paper: brute force, traditional blocking,
-LSH and M-tree, using several selected configurations for blocking
-and LSH. The distance threshold is varied between 0 and 250~\footnote{Relatively high Levenshtein edit distances are
-included since Cora includes a number of low-similarity true
-matches.}. For traditional blocking, the following attributes are
-used individually as blocking keys: \emph{author}, \emph{title},
-\emph{venue}, \emph{location}, \emph{publisher} and \emph{year}. We
-also use a combined blocking key comprising all attributes.
+We perform linkage on the Cora dataset using all approaches presented in
+this paper: brute force, traditional blocking, LSH and M-tree, using
+several selected configurations for blocking and LSH. The distance
+threshold is varied between 0 and 250~\footnote{Relatively high
+Levenshtein edit distances are included since Cora includes a number of
+low-similarity true matches.}. For traditional blocking, the following
+attributes are used individually as blocking keys: \emph{author},
+\emph{title}, \emph{venue}, \emph{location}, \emph{publisher} and
+\emph{year}. We also use a combined blocking key comprising all
+attributes.
 
@@
 thresholds give high precision and low recall, and the reverse for high
-thresholds. Brute force and M-tree give identical results, as
-expected. The best linkage quality, with an F-measure of around 0.7, is
-achieved by several linkers, including brute force, M-tree, blocking
-on \emph{authors}, blocking on all attributes, and two of the LSH
+thresholds. Brute force and M-tree give identical results, as expected.
+The best linkage quality, with an F-measure of around 0.7, is achieved
+by several linkers, including brute force, M-tree, blocking on
+\emph{authors}, blocking on all attributes, and two of the LSH
 configurations. All of these give similar overall results, apart from
@@
 used for blocking. The number of distance comparisons is reported as a
-machine-independent proxy for execution cost, since code profiling
-shows that distance calculations are dominant.
+machine-independent proxy for execution cost, since code profiling shows
+that distance calculations are dominant.
 
@@
 
+% Peter: Figure is too high - higher than page hight! ************
+% So I changed the \vspace{5mm} to \vspace{4mm} and made the plots less % wide ************
+
 \begin{figure}
 \centering
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+~~
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 \caption{M-tree}
-\end{subfigure} \vspace{5mm}
+\end{subfigure} \vspace{4mm}
 
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 \caption{Blocking on `title' attribute}
-\end{subfigure} \vspace{5mm}
+\end{subfigure} \vspace{4mm}
 
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 \caption{Blocking using the union of all attrs.}
-\end{subfigure} \vspace{5mm}
+\end{subfigure} \vspace{4mm}
 
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 \caption{LSH using 2 bands of size 10}
-\end{subfigure} \vspace{5mm}
+\end{subfigure} \vspace{4mm}
 
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 
-\begin{table}[ht]
+\begin{table}[t]
 \caption{Linkage quality on Cora dataset with distance threshold
@@
 \centering
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 \vspace{-6mm}
-\caption{M-tree linkage on Kilmarnock dataset\label{kilmarnock-quality-mtree}}
-\end{subfigure} \vspace{5mm}
+\caption{M-tree linkage on Kilmarnock dataset \label{kilmarnock-quality-mtree}}
+\end{subfigure} \vspace{4mm}
 
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 by a complete method such as M-tree. This is due to the incomplete
-linker masking the inability of a classifier based solely on
-record similarity to correctly classify high-similarity non-matches or
+linker masking the inability of a classifier based solely on record
+similarity to correctly classify high-similarity non-matches or
 low-similarity matches.
 
-We have focused on a single distance function: the sum of attribute-level Levenshtein distances.
-This gives a straightforward intuition of record-level distances, but it is more common to normalise metrics to the range 0-1.
+We have focused on a single distance function: the sum of
+attribute-level Levenshtein distances. This gives a straightforward
+intuition of record-level distances, but it is more common to normalise
+metrics to the range 0-1.
 
-Many distance functions, such as Jaro-Winkler, are not metric, and therefore cannot be used with MSI techniques.
-Care must be taken to preserve metric properties when combining metrics over individual fields, as is highlighted in~\cite{Yujian2007}, since this may yield a function that is not metric.
-It is then possible for MSI techniques to yield results that are subtly incorrect.
+Many distance functions, such as Jaro-Winkler, are not metric, and
+therefore cannot be used with MSI techniques. Care must be taken to
+preserve metric properties when combining metrics over individual
+fields, as is highlighted in~\cite{Yujian2007}, since this may yield a
+function that is not metric. It is then possible for MSI techniques to
+yield results that are subtly incorrect.
 
-Different metrics can give different distributions of inter-record distances, which can affect both the linkage results and the number of comparisons made, and hence the efficiency of the algorithm.
-Datasets with low variation in inter-record distances are said to have high \textit{intrinsic dimensionality}, and tend to require high numbers of comparisons.
-The Scottish vital event datasets, combined with Levenshtein-based metrics, appear to have high intrinsic dimensionality.
-We are now, therefore, investigating the application of various different metrics to this domain, including Jensen-Shannon, Cosine and Structured Entropic Distance, as described in~\cite{Connor2016}.
-We are also investigating the applicability of a novel technique for dimensionality reduction~\cite{Connor2017}.
+Different metrics can give different distributions of inter-record
+distances, which can affect both the linkage results and the number of
+comparisons made, and hence the efficiency of the algorithm. Datasets
+with low variation in inter-record distances are said to have high
+\textit{intrinsic dimensionality}, and tend to require high numbers of
+comparisons. The Scottish vital event datasets, combined with
+Levenshtein-based metrics, appear to have high intrinsic dimensionality.
+We are now, therefore, investigating the application of various
+different metrics to this domain, including Jensen-Shannon, Cosine and
+Structured Entropic Distance, as described in~\cite{Connor2016}. We are
+also investigating the applicability of a novel technique for
+dimensionality reduction~\cite{Connor2017}.
 
@@
 
-We thank Alice Reid of the University of
-Cambridge and her colleagues, especially Ros Davies and Eilidh Garrett,
-for the work undertaken on the Kilmarnock and Isle of Skye databases.
+We thank Alice Reid of the University of Cambridge and her colleagues,
+especially Ros Davies and Eilidh Garrett, for the work undertaken on the
+Kilmarnock and Isle of Skye databases.
 

### 2018-02-19 paper.tex

@@
 Record linkage is the process of identifying records that refer to the
-same real-world entities, in situations where entity identifiers are
+same real-world entities in situations where entity identifiers are
 unavailable. Records are linked on the basis of similarity between
@@
 in more detail, and finally classified. Even state-of-the-art indexing
-techniques, such as Locality Sensitive Hashing, have potential
+techniques, such as locality sensitive hashing, have potential
 drawbacks. They may fail to group together some true matching records
@@
 simple threshold-based to sophisticated clustering, supervised
-classification techniques, and active learning approaches~\cite{Wan15}.
+classification, and active learning approaches~\cite{Wan15}.
 
@@
 Traditional linkage systems that perform indexing prior to comparison
-and classification (on the left in \Cref{fig-rl-process}) add a further
-complexity. Indexing, comparison and classification are often conducted
-using algorithms and parameters selected using domain expertise,
-followed by manual assessment of the linkage outcomes~\cite{Chr12}. If
-the resulting link quality is too low for a certain application, the
-process is repeated with different parameter settings or algorithms,
-giving a time-consuming iterative process. The choice of an appropriate
-indexing technique and parameters (including which attributes to use in
-indexing) will significantly affect the final linkage outcome.
+and classification (on the left in Fig.~\ref{fig-rl-process}) add a
+further complexity. Indexing, comparison and classification are often
+conducted using algorithms and parameters selected using domain
+expertise, followed by manual assessment of the linkage
+outcomes~\cite{Chr12}. If the resulting link quality is too low for a
+certain application, the process is repeated with different parameter
+settings or algorithms, giving a time-consuming iterative
+process~\cite{Fis15}. The choice of an appropriate indexing technique
+as well as suitable parameter settings (including which attributes to
+use in indexing) will significantly affect the final linkage outcome.
 
@@
 comparison and classification to be combined into a single step (on the
-right in \Cref{fig-rl-process}), making the process simpler, more
+right in Fig.~\ref{fig-rl-process}), making the process simpler, more
 efficient and more effective than incomplete approaches.
 
-The motivation for this work is
-the Digitising Scotland project~\cite{Dibben2012}, which aims to
-transcribe and link all civil registration events recorded in Scotland
-between 1856 and 1973. This data set will include around 14 million
-birth records, 11 million death records and 4 million marriage records.
+The motivation for this work is the Digitising Scotland
+project~\cite{Dibben2012}, which aims to transcribe and link all civil
+registration events recorded in Scotland between 1856 and 1973. This
+dataset will include around 14 million birth records, 11 million death
+records and 4 million marriage records.
 
@@
 the datasets. This requires good ground truth data which may be
-unavailable or expensive to obtain.
+unavailable in real-world applications or expensive to obtain.
 
@@
 \textbf{Brute force:} Every record in $\mathbf{D}_A$ is compared with
-every record in $\mathbf{D}_B$. Each pair is classified as a link if the
-distance between the records is less than or equal to the threshold.
-This always finds all links, with complexity $ O(|\mathbf{D}_A| \times
-|\mathbf{D}_B|)$.
+every record in $\mathbf{D}_B$. Each pair is classified as a link if
+the distance between the records is less than or equal to the
+threshold $d$. This always finds all links, with complexity
+$ O(|\mathbf{D}_A| \cdot |\mathbf{D}_B|)$.
 
@@
 the appropriate block based on its blocking key value. The algorithm
-then iterates over the records in $\mathbf{D}_B$, and for each one,
+then iterates over the records in $\mathbf{D}_B$, and for each one
 compares it with each of the records from $\mathbf{D}_A$ in the block
@@
 \emph{shingled} into a set of n-grams with $n = l_{ss}$. Next, a set of
-deterministically generated hash functions are applied to each n-gram in
+deterministically generated hash functions is applied to each n-gram in
 the set and the smallest result (the MinHash) of each hash application
@@
 We now describe the datasets and method used in our
-evaluation, as available from: \url{ADD\_URL\_TO\_GITHUB?}
-%~\footnote{The raw data, additional figures and the source
-%code required to run these experiments can be downloaded from: (link
-%removed for double blind review).}.
-\label{sec-data} We used three
-datasets from two domains in our experiments, as summarised in
-Table~\ref{table-datasets}. The first is \emph{Cora}~\cite{Cora2017},
-which contains 1,295 records that refer to 112 machine learning
-publications. Cora is commonly used as a benchmark dataset in the
-literature for assessing linkage algorithms. Ground truth is provided
-via a unique \emph{paper\_id} identifier of the form ``blum1993". In
-this experiment linkage is performed over the same set of records (i.e.\
-a de-duplication~\cite{Chr12}).
-
-%----------------------------------------------------------------------
+evaluation~\footnote{The raw experimental data, additional figures and
+the source code required to run these experiments can be downloaded
+from the following repository. \\
+\url{http://github.com/digitisingscotland/pakdd2018-metric-linkage}.}.
+We used three datasets from two
+domains in our experiments, as summarised in Table~\ref{table-datasets}.
+The first is \emph{Cora}~\cite{Cora2017}, which contains 1,295 records
+that refer to 112 machine learning publications. Cora is commonly used
+as a benchmark dataset in the literature for assessing linkage
+algorithms. Ground truth is provided via a unique \emph{paper\_id}
+identifier of the form ``blum1993". In this experiment linkage is
+performed over the same set of records (i.e.\ a
+de-duplication~\cite{Chr12}).
 
@@
 
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
 \subsection{Cora Results}
+
+\begin{figure}
+\centering
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
+\vspace{-6mm}
+\caption{Brute Force}
+\end{subfigure}%
+~~
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
+\vspace{-6mm}
+\caption{M-tree}
+\end{subfigure} \vspace{3mm}
+
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
+\vspace{-6mm}
+\caption{Blocking on `authors' attribute}
+\end{subfigure}%
+~~
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
+\vspace{-6mm}
+\caption{Blocking on `title' attribute}
+\end{subfigure} \vspace{3mm}
+
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
+\vspace{-6mm}
+\caption{Blocking on `publication year' attr.}
+\end{subfigure}%
+~~
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
+\vspace{-6mm}
+\caption{Blocking using the union of all attrs.}
+\end{subfigure} \vspace{3mm}
+
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
+\vspace{-6mm}
+\caption{LSH using 2 bands of size 2}
+\end{subfigure}%
+~~
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
+\vspace{-6mm}
+\caption{LSH using 2 bands of size 10}
+\end{subfigure} \vspace{3mm}
+
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
+\vspace{-6mm}
+\caption{LSH using 10 bands of size 2}
+\end{subfigure}%
+~~
+\begin{subfigure}{.47\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
+\vspace{-6mm}
+\caption{LSH using 10 bands of size 10}
+\end{subfigure}
+\caption{Linkage results on the Cora dataset.}
+\label{cora-quality}
+\end{figure}
 
@@
 threshold is varied between 0 and 250~\footnote{Relatively high
-Levenshtein edit distances are included since Cora includes a number of
+Levenshtein edit distances are included since Cora contains a number of
 low-similarity true matches.}. For traditional blocking, the following
@@
 
-\Cref{cora-quality} shows the precision, recall, and
-F-measure~\cite{Chr12} for various thresholds. As expected, low
-thresholds give high precision and low recall, and the reverse for high
-thresholds. Brute force and M-tree give identical results, as expected.
-The best linkage quality, with an F-measure of around 0.7, is achieved
-by several linkers, including brute force, M-tree, blocking on
-\emph{authors}, blocking on all attributes, and two of the LSH
-configurations. All of these give similar overall results, apart from
-blocking on \emph{authors}, which gives much better quality at very high
-distance thresholds. This is due to the incomplete nature of the
-approach, avoiding comparisons of significant numbers of high-similarity
+Figure~\ref{cora-quality} shows the precision, recall, and
+F-measure~\cite{Chr12} for various thresholds~\footnote{Noting that
+recent research identifies some problematic aspects with using the
+F-measure to compare record linkage procedures at different similarity
+thresholds~\cite{Han18}.}. As expected, low thresholds give high
+precision and low recall, and the reverse for high thresholds. Brute
+force and M-tree give identical results, as expected. The best linkage
+quality, with an F-measure of around 0.7, is achieved by several
+linkers, including brute force, M-tree, blocking on \emph{authors},
+blocking on all attributes, and two of the LSH configurations. All of
+these give similar overall results, apart from blocking on
+\emph{authors}, which gives much better quality at very high distance
+thresholds. This is due to the incomplete nature of the approach,
+avoiding comparisons of significant numbers of high-similarity
 non-matches and thus avoiding these becoming false positives and keeping
@@
 used for blocking. The number of distance comparisons is reported as a
-machine-independent proxy for execution cost, since code profiling shows
-that distance calculations are dominant.
+machine-independent proxy for execution cost, since code profiling
+shows that distance calculations are dominant.
 
@@
 Sect.~\ref{sec-approach}. Although several of the incomplete linkers
-give as good quality as M-tree, and in some cases at lower cost, this is
-offset by the need to select appropriate configuration parameters. Some
-other linkers give very poor results.
-
-%----------------------------------------------------------------------
-
-% Peter: Figure is too high - higher than page hight! ************
-% So I changed the \vspace{5mm} to \vspace{4mm} and made the plots less % wide ************
-
-\begin{figure}
-\centering
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
-\vspace{-6mm}
-\caption{Brute Force}
-\end{subfigure}%
-~~
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
-\vspace{-6mm}
-\caption{M-tree}
-\end{subfigure} \vspace{4mm}
-
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
-\vspace{-6mm}
-\caption{Blocking on `authors' attribute}
-\end{subfigure}%
-~~
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
-\vspace{-6mm}
-\caption{Blocking on `title' attribute}
-\end{subfigure} \vspace{4mm}
-
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
-\vspace{-6mm}
-\caption{Blocking on `publication year' attr.}
-\end{subfigure}%
-~~
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
-\vspace{-6mm}
-\caption{Blocking using the union of all attrs.}
-\end{subfigure} \vspace{4mm}
-
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
-\vspace{-6mm}
-\caption{LSH using 2 bands of size 2}
-\end{subfigure}%
-~~
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
-\vspace{-6mm}
-\caption{LSH using 2 bands of size 10}
-\end{subfigure} \vspace{4mm}
-
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
-\vspace{-6mm}
-\caption{LSH using 10 bands of size 2}
-\end{subfigure}%
-~~
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
-\vspace{-6mm}
-\caption{LSH using 10 bands of size 10}
-\end{subfigure}
-
-\caption{Linkage results on the Cora dataset.}
-\label{cora-quality}
-\end{figure}
+give as good quality as M-tree, and in some cases at lower cost, this
+is offset by the need to select appropriate configuration parameters.
+Some other linkers give very poor results.
 
@@
 \begin{table}[t]
-\caption{Linkage quality on Cora dataset with distance threshold
-  $d= 70$.}
+\caption{Linkage quality on Cora dataset with distance threshold $d= 70$.}
 \label{comparison-of-results-cora}
@@
 \begin{tabular}{lrrrr} \hline\noalign{\smallskip}
-  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
-  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
+Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & ~~~~~~~~~~~Recall & ~~~~~~F-measure \\
 \noalign{\smallskip} \hline \noalign{\smallskip}
@@
 LSH-5-5         &    22,057 & 0.99 & 0.36 & 0.53 \\
-LSH-10-5           &  26,167	& 0.98	& 0.4	& 0.57      \\
+LSH-10-5        &    26,167 & 0.98 & 0.40 & 0.57 \\
 LSH-2-10        &     4,711 &    1 & 0.15 & 0.27 \\
 LSH-5-10        &     6,501 &    1 & 0.19 & 0.32 \\
-LSH-10-10          & 10,627  	& 0.99	& 0.27& 	0.43       \\
+LSH-10-10       &    10,627 & 0.99 & 0.27 & 0.43 \\
 \noalign{\smallskip} \hline \noalign{\smallskip}
@@
 
+\begin{figure}[t]
+\centering
+\begin{subfigure}{.46\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
+\vspace{-6mm}
+\caption{M-tree linkage on Isle of Skye \\ dataset\label{skye-quality-mtree}}
+\end{subfigure}%
+~~
+\begin{subfigure}{.46\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
+\vspace{-6mm}
+\caption{M-tree linkage on Kilmarnock \\ dataset \label{kilmarnock-quality-mtree}}
+\end{subfigure} \vspace{1mm}
+
+\begin{subfigure}{.46\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
+\vspace{-6mm}
+\caption{F-measure on Isle of Skye dataset \\ with M-tree and all
+LSH configurations}
+\end{subfigure}%
+~~
+\begin{subfigure}{.46\textwidth}
+  \centering
+\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
+\vspace{-6mm}
+\caption{F-measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
+\end{subfigure}
+\caption{Linkage results on the demographic datasets.}
+\label{demography-quality}
+\end{figure}
+
 Birth records were linked to death records, separately for the Skye and
@@
 
-\Cref{demography-quality} plots (a) and (b) show the M-tree precision,
+Figure~\ref{demography-quality} plots (a) and (b) show the M-tree precision,
 recall, and F-measure for various thresholds. In both datasets, the best
@@
 
-\begin{figure}[t]
-\centering
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
-\vspace{-6mm}
-\caption{M-tree linkage on Skye dataset\label{skye-quality-mtree}}
-\end{subfigure}%
-~~
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
-\vspace{-6mm}
-\caption{M-tree linkage on Kilmarnock dataset \label{kilmarnock-quality-mtree}}
-\end{subfigure} \vspace{4mm}
-
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotFs-skye-f}
-\vspace{-6mm}
-\caption{F1 Measure on Skye dataset \\ with M-tree and all LSH configurations}
-\end{subfigure}%
-~~
-\begin{subfigure}{.47\textwidth}
-  \centering
-\includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
-\vspace{-6mm}
-\caption{F1 Measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
-\end{subfigure}
-\caption{Linkage results on the demographic datasets.}
-\label{demography-quality}
-\end{figure}
-
-%----------------------------------------------------------------------
-
 \begin{table}[t]
-\caption{Linkage quality on Skye dataset with distance threshold
-$d= 2$.}
+\caption{Linkage quality on Isle of Skye dataset with distance threshold $d=2$.}
 \label{comparison-of-results-demography-skye}
@@
 LSH-2-2  &   3,109,250 & 0.63 & 0.03 & 0.06 \\
-
 LSH-5-2  &  10,412,496 & 0.64 & 0.11 & 0.19 \\
-
 LSH-10-2 &  53,874,127 & 0.68 & 0.36 & 0.47 \\
-
 LSH-5-5  &      36,566 & 0.76 & 0.01 & 0.01 \\
-
 LSH-10-5 &     129,873 & 0.72 & 0.01 & 0.02 \\
@@
 \begin{table}[t]
-\caption{Linkage quality on Kilmarnock dataset with distance
-threshold $d= 2$.}
+\caption{Linkage quality on Kilmarnock dataset with distance threshold $d=2$.}
 \label{comparison-of-results-demography-kili}
@@
 \begin{tabular}{lrrrrrr} \hline\noalign{\smallskip}
-  Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & 
-  ~~~~~~~~~~~Recall & ~~~~~~F1 Measure \\
+Linker~~~~~~~~~~ & ~~~Comparisons & ~~~~~Precision & ~~~~~~~~~~~Recall & ~~~~~~F-measure \\
 \noalign{\smallskip} \hline \noalign{\smallskip}
@@
 
-We note that the good results yielded by both traditional blocking and
+We note that the good results obtained by both traditional blocking and
 LSH are partly due to the fact that (in the limit) they tend towards
-brute force as the number of records in the block increase. A second,
+brute force as the number of records in the blocks increase. A second,
 unexpected, result is that illustrated in
-\Cref{comparison-of-results-cora}, namely that incomplete approaches
+Table~\ref{comparison-of-results-cora}, namely that incomplete approaches
 such as LSH can in some cases yield higher precision than that achieved
@@
 \textit{intrinsic dimensionality}, and tend to require high numbers of
-comparisons. The Scottish vital event datasets, combined with
-Levenshtein-based metrics, appear to have high intrinsic dimensionality.
-We are now, therefore, investigating the application of various
-different metrics to this domain, including Jensen-Shannon, Cosine and
-Structured Entropic Distance, as described in~\cite{Connor2016}. We are
-also investigating the applicability of a novel technique for
-dimensionality reduction~\cite{Connor2017}.
+comparisons. The Scottish vital event datasets~\cite{reid2002}, combined
+with Levenshtein-based metrics, appear to have high intrinsic
+dimensionality. We are now, therefore, investigating the application of
+various different metrics to this domain, including Jensen-Shannon,
+Cosine and Structured Entropic Distance, as described
+in~\cite{Connor2016}. We are also investigating the applicability of a
+novel technique for dimensionality reduction~\cite{Connor2017}.
 

### 2018-02-19 paper.tex
Update on Overleaf.

@@
 We now describe the datasets and method used in our
-evaluation~\footnote{The raw experimental data, additional figures and
-the source code required to run these experiments can be downloaded
-from the following repository. \\
+evaluation~\footnote{Experimental data, additional figures and
+source code can be downloaded from:
 \url{http://github.com/digitisingscotland/pakdd2018-metric-linkage}.}.

### 2018-02-19 scripts/plots.py
Use 'F-measure' consistently in the plots

@@ -111,7 +111,7 @@ def plotLQ(title, predicate):
             , [ line["F1 Measure"]          for line in data ]
             , color="red"
             , marker="o"
-            , label="F1 Measure"
+            , label="F-measure"
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
@@ -438,7 +438,7 @@ def plotFixedThresholdTraditional1(title, blockingMethods, predicate):
                                                         for line in data if line["Blocking Method"] == blockingMethod ]
             , color="red"
             , marker="o"
-            , label="F1 Measure"
+            , label="F-measure"
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
@@ -542,7 +542,7 @@ def plotFixedThresholdLSH(title, predicate):
             , [ line["F1 Measure"]                      for line in data ]
             , color="red"
             , marker="o"
-            , label="F1 Measure"
+            , label="F-measure"
             )
 
     plt.plot( range(len(xRange))
@@ -667,7 +667,7 @@ title = " - ".join([dataset])
 predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and toFloat(line["Distance Threshold"]) <= 30
 plotFs(title, "Precision", predicate)
 plotFs(title, "Recall", predicate)
-plotFs(title, "F1 Measure", predicate)
+plotFs(title, "F-measure", predicate)
 
 
 dataset = "Skye"
@@ -675,7 +675,7 @@ title = " - ".join([dataset])
 predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and toFloat(line["Distance Threshold"]) <= 30
 plotFs(title, "Precision", predicate)
 plotFs(title, "Recall", predicate)
-plotFs(title, "F1 Measure", predicate)
+plotFs(title, "F-measure", predicate)
 
 
 dataset = "Kilmarnock"
@@ -683,5 +683,5 @@ title = " - ".join([dataset])
 predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and toFloat(line["Distance Threshold"]) <= 30
 plotFs(title, "Precision", predicate)
 plotFs(title, "Recall", predicate)
-plotFs(title, "F1 Measure", predicate)
+plotFs(title, "F-measure", predicate)
 

### 2018-02-19 scripts/copy.sh
Update the plots with B/W friendly versions

@@ -1,0 +1,34 @@
+
+cp "plotLQ -- Cora - Brute Force.eps" ../figures/plotLQ-cora-brute.eps
+
+
+cp "plotLQ -- Cora - LSH - 2 - 2 - 2.eps"  ../figures/plotLQ-cora-lsh-2-2.eps
+cp "plotLQ -- Cora - LSH - 2 - 2 - 5.eps"  ../figures/plotLQ-cora-lsh-2-5.eps
+cp "plotLQ -- Cora - LSH - 2 - 2 - 10.eps" ../figures/plotLQ-cora-lsh-2-10.eps
+
+cp "plotLQ -- Cora - LSH - 2 - 5 - 2.eps"  ../figures/plotLQ-cora-lsh-5-2.eps
+cp "plotLQ -- Cora - LSH - 2 - 5 - 5.eps"  ../figures/plotLQ-cora-lsh-5-5.eps
+cp "plotLQ -- Cora - LSH - 2 - 5 - 10.eps" ../figures/plotLQ-cora-lsh-5-10.eps
+
+cp "plotLQ -- Cora - LSH - 2 - 10 - 2.eps"  ../figures/plotLQ-cora-lsh-10-2.eps
+cp "plotLQ -- Cora - LSH - 2 - 10 - 5.eps"  ../figures/plotLQ-cora-lsh-10-5.eps
+cp "plotLQ -- Cora - LSH - 2 - 10 - 10.eps" ../figures/plotLQ-cora-lsh-10-10.eps
+
+
+cp "plotLQ -- Cora - MTree.eps"           ../figures/plotLQ-cora-mtree.eps
+cp "plotLQ -- Kilmarnock - MTree.eps"     ../figures/plotLQ-kilmarnock-mtree.eps
+cp "plotLQ -- Skye - MTree.eps"           ../figures/plotLQ-skye-mtree.eps
+
+
+cp "plotLQ -- Cora - TradBlocking - 3.eps"   ../figures/plotLQ-cora-trad-year.eps
+cp "plotLQ -- Cora - TradBlocking - 5.eps"   ../figures/plotLQ-cora-trad-authors.eps
+cp "plotLQ -- Cora - TradBlocking - 7.eps"   ../figures/plotLQ-cora-trad-title.eps
+cp "plotLQ -- Cora - TradBlocking - 8.eps"   ../figures/plotLQ-cora-trad-location.eps
[…]

### 2018-02-19 scripts/plots.py

@@ -61,6 +61,9 @@ def toFloat(s):
 
 (fields, stats) = loadCSV("stats.tsv", delimiter="\t")
 
+for line in stats:
+    line["F-measure"] = line["F1 Measure"]
+
 for f in fields:
     print("Field: " + f)
 
@@ -69,7 +72,7 @@ for f in fields:
 # Plotting Linkage Quality
 ################################################################################
 
-def plotLQ(title, predicate):
+def plotLQ(title, predicate, withlegend=False):
     plt.clf()
 
     w,h = plt.figaspect(0.5)
@@ -96,26 +99,30 @@ def plotLQ(title, predicate):
     plt.plot( [ line["Distance Threshold"]  for line in data ]
             , [ line["Precision"]           for line in data ]
             , color="blue"
-            , marker="o"
+            # , marker="o"
+            , linestyle="-."
             , label="Precision"
             )
 
     plt.plot( [ line["Distance Threshold"]  for line in data ]
             , [ line["Recall"]              for line in data ]
             , color="green"
-            , marker="o"
+            # , marker="o"
+            , linestyle=":"
             , label="Recall"
             )
 
     plt.plot( [ line["Distance Threshold"]  for line in data ]
             , [ line["F1 Measure"]          for line in data ]
             , color="red"
-            , marker="o"
+            # , marker="o"
+            , linestyle="-"
             , label="F-measure"
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
-    plt.legend()
+    if withlegend:
+        plt.legend()
 
     plt.savefig("plotLQ -- %s.eps" % title, bbox_inches="tight")
     plt.close()
@@ -127,7 +134,7 @@ dataset = "Cora"
 linker = "Brute Force"
 title = " - ".join([dataset, linker])
 predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and line["Linker"] == linker
-plotLQ(title, predicate)
+plotLQ(title, predicate, withlegend=True)
 
 
 linker = "TradBlocking"

### 2018-02-19 scripts/plots.py
Add legends to Skye&Kilmarnock MTree plots

@@ -170,7 +170,7 @@ dataset = "Skye"
 linker = "MTree"
 title = " - ".join([dataset, linker])
 predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and line["Linker"] == linker
-plotLQ(title, predicate)
+plotLQ(title, predicate, withlegend=True)
 
 
 linker = "LSH"
@@ -191,7 +191,7 @@ dataset = "Kilmarnock"
 linker = "MTree"
 title = " - ".join([dataset, linker])
 predicate = lambda line: line["Data Set (Source)"].startswith(dataset) and line["Linker"] == linker
-plotLQ(title, predicate)
+plotLQ(title, predicate, withlegend=True)
 
 
 linker = "LSH"

### 2018-02-19 scripts/copy.sh
Add legends to Skye&Kilmarnock MTree plots

@@ -29,6 +29,6 @@ cp "plotLQ -- Cora - TradBlocking - 10.eps"  ../figures/plotLQ-cora-trad-publish
 cp "plotLQ -- Cora - TradBlocking - all.eps" ../figures/plotLQ-cora-trad-combined.eps
 
 
-cp "plotFs -- Cora - F-measure.eps" ../figures/plotFs-cora-f.eps
-cp "plotFs -- Skye - F-measure.eps" ../figures/plotFs-skye-f.eps
-cp "plotFs -- Kilmarnock - F-measure.eps" ../figures/plotFs-kilmarnock-f.eps
+cp "plotFs -- Cora - F-measure.eps"             ../figures/plotFs-cora-f.eps
+cp "plotFs -- Skye - F-measure.eps"             ../figures/plotFs-skye-f.eps
+cp "plotFs -- Kilmarnock - F-measure.eps"       ../figures/plotFs-kilmarnock-f.eps

### 2018-02-19 paper.tex
Update on Overleaf.

@@
 We now describe the datasets and method used in our
-evaluation~\footnote{Experimental data, additional figures and
-source code can be downloaded from:
+evaluation~\footnote{Experimental data, additional figures and source
+code can be downloaded from: \\
 \url{http://github.com/digitisingscotland/pakdd2018-metric-linkage}.}.
-We used three datasets from two
-domains in our experiments, as summarised in Table~\ref{table-datasets}.
-The first is \emph{Cora}~\cite{Cora2017}, which contains 1,295 records
-that refer to 112 machine learning publications. Cora is commonly used
-as a benchmark dataset in the literature for assessing linkage
-algorithms. Ground truth is provided via a unique \emph{paper\_id}
-identifier of the form ``blum1993". In this experiment linkage is
-performed over the same set of records (i.e.\ a
-de-duplication~\cite{Chr12}).
+We used three datasets from two domains in our experiments, as
+summarised in Table~\ref{table-datasets}. The first is
+\emph{Cora}~\cite{Cora2017}, which contains 1,295 records that refer to
+112 machine learning publications. Cora is commonly used as a benchmark
+dataset in the literature for assessing linkage algorithms. Ground truth
+is provided via a unique \emph{paper\_id} identifier of the form
+``blum1993". In this experiment linkage is performed over the same set
+of records (i.e.\ a de-duplication~\cite{Chr12}).
 
@@
 LSH-10-5        &    26,167 & 0.98 & 0.40 & 0.57 \\
-LSH-2-10        &     4,711 &    1 & 0.15 & 0.27 \\
-LSH-5-10        &     6,501 &    1 & 0.19 & 0.32 \\
+LSH-2-10        &     4,711 & 1.00 & 0.15 & 0.27 \\
+LSH-5-10        &     6,501 & 1.00 & 0.19 & 0.32 \\
 LSH-10-10       &    10,627 & 0.99 & 0.27 & 0.43 \\
@@
 \centering
-\begin{subfigure}{.46\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 \vspace{-6mm}
-\caption{M-tree linkage on Isle of Skye \\ dataset\label{skye-quality-mtree}}
+\caption{M-tree on Isle of Skye dataset\label{skye-quality-mtree}}
 \end{subfigure}%
 ~~
-\begin{subfigure}{.46\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 \vspace{-6mm}
-\caption{M-tree linkage on Kilmarnock \\ dataset \label{kilmarnock-quality-mtree}}
-\end{subfigure} \vspace{1mm}
+\caption{M-tree on Kilmarnock dataset \label{kilmarnock-quality-mtree}}
+\end{subfigure} \vspace{5mm}
 
-\begin{subfigure}{.46\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering
@@
 ~~
-\begin{subfigure}{.46\textwidth}
+\begin{subfigure}{.47\textwidth}
   \centering

### 2018-02-20 scripts/plots.py
linewidth=2

@@ -102,6 +102,7 @@ def plotLQ(title, predicate, withlegend=False):
             # , marker="o"
             , linestyle="-."
             , label="Precision"
+            , linewidth=2
             )
 
     plt.plot( [ line["Distance Threshold"]  for line in data ]
@@ -110,6 +111,7 @@ def plotLQ(title, predicate, withlegend=False):
             # , marker="o"
             , linestyle=":"
             , label="Recall"
+            , linewidth=2
             )
 
     plt.plot( [ line["Distance Threshold"]  for line in data ]
@@ -118,6 +120,7 @@ def plotLQ(title, predicate, withlegend=False):
             # , marker="o"
             , linestyle="-"
             , label="F-measure"
+            , linewidth=2
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
@@ -244,6 +247,7 @@ def plotBQ(title, predicate):
             , color="green"
             , marker="o"
             , label="Block Quality"
+            , linewidth=2
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
@@ -284,6 +288,7 @@ def plotPQC(title, predicate):
             , color="green"
             , marker="o"
             , label="Pairs Quality"
+            , linewidth=2
             )
 
     plt.plot( [         line["Distance Threshold"]          for line in data ]
@@ -291,6 +296,7 @@ def plotPQC(title, predicate):
             , color="red"
             , marker="o"
             , label="Pairs Completeness"
+            , linewidth=2
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
@@ -432,6 +438,7 @@ def plotFixedThresholdTraditional1(title, blockingMethods, predicate):
             , color="blue"
             , marker="o"
             , label="Precision"
+            , linewidth=2
             )
     plt.plot( range(len(xRange))
             , [ line["Recall"]                          for blockingMethod in blockingMethods
@@ -439,6 +446,7 @@ def plotFixedThresholdTraditional1(title, blockingMethods, predicate):
             , color="green"
             , marker="o"
             , label="Recall"
+            , linewidth=2
             )
     plt.plot( range(len(xRange))
             , [ line["F1 Measure"]                      for blockingMethod in blockingMethods
@@ -446,6 +454,7 @@ def plotFixedThresholdTraditional1(title, blockingMethods, predicate):
             , color="red"
             , marker="o"
[…]

### 2018-02-20 scripts/plots.py
linewidth=3

@@ -102,7 +102,7 @@ def plotLQ(title, predicate, withlegend=False):
             # , marker="o"
             , linestyle="-."
             , label="Precision"
-            , linewidth=2
+            , linewidth=3.0
             )
 
     plt.plot( [ line["Distance Threshold"]  for line in data ]
@@ -111,7 +111,7 @@ def plotLQ(title, predicate, withlegend=False):
             # , marker="o"
             , linestyle=":"
             , label="Recall"
-            , linewidth=2
+            , linewidth=3.0
             )
 
     plt.plot( [ line["Distance Threshold"]  for line in data ]
@@ -120,7 +120,7 @@ def plotLQ(title, predicate, withlegend=False):
             # , marker="o"
             , linestyle="-"
             , label="F-measure"
-            , linewidth=2
+            , linewidth=3.0
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
@@ -247,7 +247,7 @@ def plotBQ(title, predicate):
             , color="green"
             , marker="o"
             , label="Block Quality"
-            , linewidth=2
+            , linewidth=3.0
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
@@ -288,7 +288,7 @@ def plotPQC(title, predicate):
             , color="green"
             , marker="o"
             , label="Pairs Quality"
-            , linewidth=2
+            , linewidth=3.0
             )
 
     plt.plot( [         line["Distance Threshold"]          for line in data ]
@@ -296,7 +296,7 @@ def plotPQC(title, predicate):
             , color="red"
             , marker="o"
             , label="Pairs Completeness"
-            , linewidth=2
+            , linewidth=3.0
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)
@@ -438,7 +438,7 @@ def plotFixedThresholdTraditional1(title, blockingMethods, predicate):
             , color="blue"
             , marker="o"
             , label="Precision"
-            , linewidth=2
+            , linewidth=3.0
             )
     plt.plot( range(len(xRange))
             , [ line["Recall"]                          for blockingMethod in blockingMethods
@@ -446,7 +446,7 @@ def plotFixedThresholdTraditional1(title, blockingMethods, predicate):
             , color="green"
             , marker="o"
             , label="Recall"
-            , linewidth=2
+            , linewidth=3.0
             )
     plt.plot( range(len(xRange))
[…]

### 2018-02-21 scripts/plots.py
figure updates

@@ -1,4 +1,4 @@
-#!/usr/bin/env python3
+#   #!/usr/bin/env python3
 
 import matplotlib
 matplotlib.use("cairo")
@@ -662,6 +662,7 @@ def plotFs(title, measure, predicate):
                         , marker="o"
                         , label="LSH"
                         , linewidth=3.0
+                        , linestyle="-"
                         )
 
     dataMTree = [ line for line in data if line["Linker"] == "MTree" ]
@@ -672,6 +673,7 @@ def plotFs(title, measure, predicate):
             , marker="o"
             , label="M-tree"
             , linewidth=3.0
+            , linestyle=":"
             )
 
     # plt.legend(bbox_to_anchor=(1.05, 1), loc=2, borderaxespad=0.)

### 2018-04-20 paper.tex
CRC

@@
 
-\institute{School of Computer Science, University of St Andrews,
-St Andrews, Scotland. Contact: \email{\{ozgur.akgun, alan.dearle, graham.kirby\}@st-andrews.ac.uk}
+\institute{School of Computer Science, University of St Andrews, St Andrews, Scotland.
+\\ \email{\{ozgur.akgun, alan.dearle, graham.kirby\}@st-andrews.ac.uk}
 \and
-Research School of Computer Science, The Australian National University,\\
-Canberra, Australia. Contact: \email{peter.christen@anu.edu.au}}
+Research School of Computer Science, The Australian National University, Canberra, Australia.
+\\ \email{peter.christen@anu.edu.au}}
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-brute}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{Brute Force}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-mtree}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{M-tree}
-\end{subfigure} \vspace{3mm}
+\end{subfigure} % \vspace{3mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-authors}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{Blocking on `authors' attribute}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-title}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{Blocking on `title' attribute}
-\end{subfigure} \vspace{3mm}
+\end{subfigure} % \vspace{3mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-year}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{Blocking on `publication year' attr.}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-trad-combined}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{Blocking using the union of all attrs.}
-\end{subfigure} \vspace{3mm}
+\end{subfigure} % \vspace{3mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-2}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{LSH using 2 bands of size 2}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-2-10}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{LSH using 2 bands of size 10}
-\end{subfigure} \vspace{3mm}
+\end{subfigure} % \vspace{3mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-2}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{LSH using 10 bands of size 2}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-cora-lsh-10-10}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{LSH using 10 bands of size 10}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-skye-mtree}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{M-tree on Isle of Skye dataset\label{skye-quality-mtree}}
@@
 \includegraphics[width=\textwidth]{figures/plotLQ-kilmarnock-mtree}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{M-tree on Kilmarnock dataset \label{kilmarnock-quality-mtree}}
-\end{subfigure} \vspace{5mm}
+\end{subfigure} % \vspace{5mm}
 
@@
 \includegraphics[width=\textwidth]{figures/plotFs-skye-f}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{F-measure on Isle of Skye dataset \\ with M-tree and all
@@
 \includegraphics[width=\textwidth]{figures/plotFs-kilmarnock-f}
-\vspace{-6mm}
+% \vspace{-6mm}
 \caption{F-measure on Kilmarnock dataset \\ with M-tree and all LSH configurations}
@@
 
-\bibliographystyle{splncs03}
+\bibliographystyle{splncs04}
 \bibliography{paper.bib}
